### 2019-03-05 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
first

---
title: "Growth model analysis 3spp WB"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r functions}

riverOrderedIn <- factor(c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),labels = c("WB","OL","OS","IL"), ordered = T)

```

Plots for the manuscript

```{r Load data, message=FALSE}
library(arm)
library(relaimpo)
library(linkedModels)
library(arrayhelpers)
library(tidyverse)
library(gridExtra)

load("D:/projects/linkedModels/data/out/ddD_bktbntats_1997.RData")
load("D:/projects/linkedModels/data/out/ddddD_bktbntats_1997.RData")
load("D:/projects/linkedModels/data/out/rawCountsAndMasses.RData")

mcmc <- list()
dGIn <- list()
ddGIn2 <- list()

load("D:/projects/linkedModels/data/out/dG_1530971738_crossValFALSE_bkt1997_Nimble.RData")

# input data
d1 <- ddG[[1]][[1]]
d2 <- ddG[[1]][[2]]
meanSampleIntervalMean <- data.frame( meanSampleInterval = colMeans(d1$sampleIntervalMean), season = 1:d1$nSeasons )

#RECALCULATE gr for dataset. Filtering by maxSampleInterval in the creat ddddG step leaves in trailing lengths from the long samp intervals
d2 <- d2 %>%
  group_by(tag) %>%
  # arrange(tag,sampleNumber) %>%
  mutate( lagObservedWeight = lead(observedWeight),
          lagObservedLength = lead(observedLength),
          grWeight = exp(lagObservedWeight - observedWeight)/as.numeric((lagDetectionDate - detectionDate)),
          grLength = (lagObservedLength - observedLength)/as.numeric((lagDetectionDate - detectionDate))
  )  %>%
  ungroup()

#d2 %>% filter(tag=='00088cc059') %>% select(tag,riverOrdered,riverN,lagRiverN,sampleNumber,observedLength,lagObservedLength)

#input data frame
ddGIn2[['bkt']] <- ddG[[1]][[2]]
dGIn[['bkt']] <- dG
# output data
mcmc[['bkt']] <- mcmcProcessed

load("D:/projects/linkedModels/data/out/dG_1530971744_crossValFALSE_bnt1997_Nimble.RData")
ddGIn2[['bnt']] <- ddG[[1]][[2]]
dGIn[['bnt']] <- dG
mcmc[['bnt']]  <- mcmcProcessed

load("D:/projects/linkedModels/data/out/dG_1530993703_crossValFALSE_ats1997_Nimble.RData")
ddGIn2[['ats']] <- ddG[[1]][[2]]
dGIn[['ats']] <- dG
mcmc[['ats']]  <- mcmcProcessed

rm(ddG,dG,mcmcProcessed)

load("D:/projects/linkedModels/data/cd_west_1997.RData")
```

```{r mean lengths}

# Has mean size changed over time?
means <- 
  cd %>% mutate( age = year - cohort ) %>%
    filter( ageInSamples %in% 1:10, age < 3, year < 2016, area %in% c('inside','trib') ) %>%
    group_by(species,riverOrdered,ageInSamples,year,season,age) %>%
    summarize(meanLen = mean(observedLength,na.rm = TRUE),
              sdLen = sd(observedLength,na.rm = TRUE),
              meanGR = mean(grLength,na.rm = TRUE),
              sdGR = sd(grLength,na.rm = TRUE)
              )

means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

lenSlopes <- means %>%
  group_by( species, riverOrdered, ageInSamples, season, age ) %>%
  nest()

lengthModel <- function(df){
  lm( meanLen ~ year, data = df )
}

grModel <- function(df){
  lm( meanGR ~ year, data = df )
}

lenSlopes <- lenSlopes %>%
  mutate( model = map(lenSlopes$data, lengthModel),
          modelGR = map(lenSlopes$data, grModel))

###################################
######### len here, gr below ######
lenSlopesTidy <- lenSlopes %>%
  mutate(glance = map(model, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter(term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
  mutate( sig = ifelse(p.value < 0.0001,"***",sig))  %>%
  filter(!is.nan(statistic))

lenSlopesTidy$seasonGG <- factor(lenSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
lenSlopesTidy$speciesGG <- factor(lenSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

save(lenSlopesTidy,file = 'D:/projects/westBrook-patterns/data/out/lenSlopesTidy.RData')
##

ggplot( means %>% filter(year >= 0), aes(year,meanLen, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 2) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean body length (mm)" ) +
  # geom_text(data = lenSlopesTidy %>% filter(species == 'bkt'), aes(x=1999,y=220, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = lenSlopesTidy %>% filter(species == 'bnt'), aes(x=1999,y=200, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = lenSlopesTidy %>% filter(species == 'ats'), aes(x=1999,y=180, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age<2), aes(x=2000,y=225, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age<2), aes(x=2000,y=203, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age<2), aes(x=2000,y=181, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +

    geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2009,y=104, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2009,y=82, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ riverOrdered + age)

ggsave(paste0('figures/meanLengths.png'), width = 11, height = 11, units='in')

# mean increases 
lenSlopesTidy %>%
  filter() %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# mean increases by species
lenSlopesTidy %>%
  group_by(species) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
ggplot(lenSlopesTidy %>% filter(), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point( size = 3) +
  geom_smooth(method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_wrap(~riverOrdered)

ggsave(paste0('figures/meanLengthsOverAge.png'), width = 6, height = 6, units='in')

summary(lm(estimate~ageInSamples, data = lenSlopesTidy %>% filter(species=='bkt')))
summary(lm(estimate~ageInSamples, data = lenSlopesTidy %>% filter(species=='bnt')))

#########################
#### gr here, len above

grSlopesTidy <- lenSlopes %>%
  mutate(glance = map(modelGR, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter(term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
  mutate( sig = ifelse(p.value < 0.0001,"***",sig)) %>%
  filter(!is.nan(statistic))

save(grSlopesTidy,file = 'D:/projects/westBrook-patterns/data/out/grSlopesTidy.RData')
##
means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot( means %>% filter(), aes(year,meanGR, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean growth rate (mm/d)" ) +
  # geom_text(data = grSlopesTidy %>% filter(species == 'bkt'), aes(x=1999,y=220, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = grSlopesTidy %>% filter(species == 'bnt'), aes(x=1999,y=200, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = grSlopesTidy %>% filter(species == 'ats'), aes(x=1999,y=180, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age<2), aes(x=2000,y=0.7, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age<2), aes(x=2000,y=0.6, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age<2), aes(x=2000,y=181, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +

    geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2011.5,y=0.66, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2011.5,y=0.56, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ riverOrdered + age)

ggsave(paste0('figures/meanGRs.png'), width = 11, height = 11, units='in')

# mean increases 
grSlopesTidy %>%
  filter() %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# mean increases by species
grSlopesTidy %>%
  group_by(species) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
ggplot(grSlopesTidy %>% filter(), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point(size = 3) +
  geom_smooth(method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean growth rate over years" ) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_wrap(~riverOrdered)

ggsave(paste0('figures/meanGRsOverAge.png'), width = 6, height = 6, units='in')

summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bkt')))
summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bnt')))

```

```{r where are the fish?}

allFishBySpeciesYOY$seasonGG <- factor(allFishBySpeciesYOY$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)

allFishBySpeciesYOY$speciesGG <- factor(allFishBySpeciesYOY$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

allFishBySpeciesYOY$riverGG <- factor(allFishBySpeciesYOY$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("WB","OL","OS","IL"), ordered = T)

allFishBySpeciesYOY %>% filter(!is.na(isYOY)) %>%
  ggplot(aes(year,nAllFishBySpeciesYOY, color = speciesGG)) +
    geom_point(size = 2) +
    geom_smooth(method = 'lm', se=F) +
    theme_publication() +
    labs( x = "Year", y = "Estimated abundance" ) +
    scale_color_grey(start = 0.2, end = 0.8) + 
    facet_grid(seasonGG ~ riverGG + isYOY, scales = 'free')

cd2 <- 
  cd %>% mutate(age = year - cohort) %>%
    filter(ageInSamples %in% 1:10, age < 3, year < 2016, enc == 1, area %in% c('inside','trib')) %>%
   # group_by(species,riverOrdered,ageInSamples,year,season,age)
     mutate(riverArea = ifelse( river == "west brook" & section > 20 & section < 32, 2, 1)) %>%
     mutate(riverArea = ifelse( river == "west brook" & section >= 32, 3, riverArea))

counts <- cd2 %>%
  group_by(riverOrdered, riverArea)

```

# below from MS_Analysis_2spp_WB

```{r count summaries for MS}

ddGInAll <- rbind(ddGIn2[[1]],ddGIn2[[2]],ddGIn2[[3]]) %>%
  filter(year %in% 2000:2015, river == 'west brook') %>%
  mutate( consecOcc = ifelse( is.na(grLength), 0,1 ) )

ddGInAll %>%
  group_by(river,species) %>%
  distinct(tag) %>%
  summarize( n = n() )

#   species     n
#   <chr>   <int>
# 1 ats      5032
# 2 bkt      8556
# 3 bnt      6389

ddGInAll %>%
  group_by(river,species) %>%
  summarize( n = n() )

#   species     n
#   <chr>   <int>
# 1 ats     15791
# 2 bkt     18412
# 3 bnt     16510

ddGInAll %>%
  group_by(river,species,enc) %>%
  summarize( n = n() )

#   species   enc     n
#   <chr>   <dbl> <int>
# 1 ats         0  3515
# 2 ats         1 12276
# 3 bkt         0  3464
# 4 bkt         1 14948
# 5 bnt         0  3855
# 6 bnt         1 12655

ddGInAll %>%
  group_by(river,species,consecOcc) %>%
  summarize( n = n() ) %>%
  spread(consecOcc,n, sep="_") %>%
  mutate( prop0 = consecOcc_0/(consecOcc_0+consecOcc_1),
          prop1 = 1 - prop0)

#   river       species consecOcc_0 consecOcc_1 prop0 prop1
# 5 west brook  ats           10203        5588 0.646 0.354
# 6 west brook  bkt           13631        4781 0.740 0.260
# 7 west brook  bnt           12420        4090 0.752 0.248

length(unique(ddGInAll$sampleNumber))

```

```{r table of estimated abundances}

ddGIn2[['bkt']] %>%
  filter(river == 'west brook') %>%
  distinct(river,year,season,nAllFishBySpeciesPBKT,nAllFishBySpeciesPBNT,nAllFishBySpeciesPATS) %>%
  arrange(year, season)

```

```{r bnt density}

ggplot(d2, aes(cBNTStd)) +
  geom_histogram() +
  facet_grid(season~riverOrdered)

ggplot(d2, aes(cBNTStd, grLength)) +
  geom_point(alpha=0.1) +
  geom_smooth(method='lm') +
  facet_grid(isYOY+season~riverOrdered)

ggplot(d2, aes(cBKTStd, grLength)) +
  geom_point(alpha=0.1) +
  geom_smooth(method='lm') +
  facet_grid(isYOY+season~riverOrdered)
```

```{r check lengthExp}
len <- list()
for( spp in c("bkt","bnt","ats") ){

#  len2 <- array2df(mcmc[[spp]]$sims.list$lengthExp, label.x = "est") %>%
#           rename(iter=d1,obs=d2)
  offset <- 1
  if(spp == 'ats') offset <- 0 # not sure why this is needed. REs seem meesed up in the graphs below
  inds <- data.frame(ind = dGIn[[spp]][[1]]$constants$ind[1:(length(dGIn[[spp]][[1]]$constants$ind) - offset)],
                     observedLength = dGIn[[spp]][[1]]$data$lengthDATA,
                     sampleNumber = ddGIn2[[spp]]$sampleNumber[1:(length(ddGIn2[[spp]]$sampleNumber) - offset)],
                     cohort = ddGIn2[[spp]]$cohort[1:(length(ddGIn2[[spp]]$cohort) - offset)],
                     ageInSamples = ddGIn2[[spp]]$ageInSamples[1:(length(ddGIn2[[spp]]$ageInSamples) - offset)],
                     riverOrdered = ddGIn2[[spp]]$riverOrdered[1:(length(ddGIn2[[spp]]$riverOrdered) - offset)],
                     lagObservedLength = ddGIn2[[spp]]$lagObservedLength[1:(length(ddGIn2[[spp]]$lagObservedLength) - offset)],
                     detectionDate = ddGIn2[[spp]]$detectionDate[1:(length(ddGIn2[[spp]]$detectionDate) - offset)],
                     lagDetectionDate = ddGIn2[[spp]]$lagDetectionDate[1:(length(ddGIn2[[spp]]$lagDetectionDate) - offset)])
  inds$obs <- 1:(length(dGIn[[spp]][[1]]$constants$ind) - offset)

  inds <-  inds %>%
    group_by(ind) %>%
    mutate(first = min(obs),
           firstObs = obs == first,
           observed = !is.na(observedLength),
           sampleNumber = sampleNumber,
           nObs = n(),
           grLength = (lagObservedLength - observedLength)/as.numeric((lagDetectionDate - detectionDate))
           ) %>%
    ungroup()

#  len1 <- left_join(len2,inds)
#  rm(len2,inds)

  # means
  lenMean1 <- array2df(mcmc[[spp]]$mean$lengthExp, label.x = "estMean") %>%
                rename(obs=d1)
  len[[spp]] <- left_join(inds,lenMean1)  %>%
                  mutate( estMean = ifelse( firstObs, observedLength, estMean ) )

  grIndREMean1 <- array2df(mcmc[[spp]]$mean$grIndRE, label.x = "grIndRE") %>%
                rename(ind=d1)

  len[[spp]] <- left_join(len[[spp]],grIndREMean1)

#  len <- left_join(len1,lenMean1)
#  rm(len1,lenMean1)
  rm(inds,lenMean1)
  gc()
}

len[['bkt']] %>% filter( nObs == 7 ) %>%
  ggplot(aes(ageInSamples,observedLength)) +
  geom_point(aes(color = factor(ind)), size=3) +
  geom_point(aes(ageInSamples, estMean, color=factor(ind))) +
  geom_line(aes(ageInSamples, estMean, color = factor(ind))) +
  geom_point(aes(ageInSamples, estMean, color=factor(ind)), size=2) +
  geom_point(aes(ageInSamples, estMean, color=factor(ind)), shape = 1, size = 2, color = 'black') +
  theme(legend.position="none") +
  facet_wrap(~cohort)

focalFish <- 60

len[['bkt']] %>% filter( cohort == 2001 ) %>%
  ggplot(aes(ageInSamples,observedLength)) +
  geom_point(aes(color = riverOrdered), size = 3) +
  geom_point(color = 'black', shape = 1, size = 3) +
  geom_point(aes(ageInSamples, estMean, color = riverOrdered)) +
  geom_line( aes(ageInSamples, estMean, group = factor(ind), color = riverOrdered)) +
  geom_point(aes(ageInSamples, estMean, color = riverOrdered), size = 2) +
  geom_point(aes(ageInSamples, estMean, color = riverOrdered), shape = 1, size = 2, color = 'black') +
 # theme(legend.position="none") +
  facet_wrap(~nObs)

len[['bkt']] %>% filter( cohort == 2001 ) %>%
  ggplot(aes(ageInSamples,observedLength)) +
  geom_point(aes(color = grIndRE), size = 3) +
#  geom_point(color = 'black', shape = 1, size = 3) +
  geom_point(aes(ageInSamples, estMean, color = grIndRE)) +
  geom_line( aes(ageInSamples, estMean, group = factor(ind), color = grIndRE)) +
  geom_point(aes(ageInSamples, estMean, color = grIndRE), size = 2) +
#  geom_point(aes(ageInSamples, estMean, color = grIndRE), shape = 1, size = 2, color = 'black') +
 # scale_colour_distiller(palette = "Spectral") +
   scale_colour_gradient(low = "white", high = "black") +
 # theme(legend.position="none") +
  facet_wrap(~nObs)

len[['bkt']] %>% #filter( cohort == 2001 ) %>%
  ggplot(aes(ageInSamples,observedLength)) +
  geom_point(aes(color = grIndRE), size = 3) +
#  geom_point(color = 'black', shape = 1, size = 3) +
  geom_point(aes(ageInSamples, estMean, color = grIndRE)) +
  geom_line( aes(ageInSamples, estMean, group = factor(ind), color = grIndRE)) +
  geom_point(aes(ageInSamples, estMean, color = grIndRE), size = 2) +
#  geom_point(aes(ageInSamples, estMean, color = grIndRE), shape = 1, size = 2, color = 'black') +
 # scale_colour_distiller(palette = "Spectral") +
   scale_colour_gradient(low = "white", high = "black") +
 # theme(legend.position="none") +
  facet_grid(riverOrdered~cohort)

len[['ats']] %>% filter( nObs > 0 ) %>%
  group_by(ind) %>%
  summarize( meanGRLength = mean(grLength,na.rm=TRUE),
             meanGRIndRE = mean(grIndRE,na.rm=TRUE),
             meanNObs = mean(nObs,na.rm=TRUE)) %>%
  ggplot(aes(meanGRLength,meanGRIndRE)) +
  geom_point() +
  geom_smooth(method='lm') +
#  scale_colour_gradient(low = "yellow", high = "red") +
  facet_wrap(~meanNObs)

```

```{r winter growth rates}

ggplot(ddG[[1]][[2]], aes(observedLength,grLength)) + 
  geom_point() +
  facet_grid(river + isYOY ~ season+species)

```

```{r adjusted counts}
allFishBySpeciesYOY$riverOrderedGG <- factor(allFishBySpeciesYOY$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
allFishBySpeciesYOY$seasonGG <- factor(allFishBySpeciesYOY$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
allFishBySpeciesYOY$speciesGG <- factor(allFishBySpeciesYOY$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

#sum across YOY
allFishBySpecies <- allFishBySpeciesYOY %>%
   group_by(river,season,year,species) %>%
   summarize( nAllFishBySpecies = sum(nAllFishBySpeciesYOY, na.rm=T),
                massAllFishBySpecies = sum(massAllFishBySpeciesYOY, na.rm=T))
allFishBySpecies$seasonGG <- factor(allFishBySpecies$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
allFishBySpeciesYOY <- left_join(allFishBySpeciesYOY,allFishBySpecies)

# sum across YOY and species
   allFish <- allFishBySpeciesYOY %>%
     group_by(river,season,year) %>%
     summarize( nAllFish = sum(nAllFishBySpeciesYOY, na.rm=T),
                massAllFish = sum(massAllFishBySpeciesYOY, na.rm=T))
   cd <- left_join(cd,allFish)

allFish$riverOrderedGG <- factor(allFish$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
allFish$seasonGG <- factor(allFish$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
#########

ggplot(allFishBySpeciesYOY %>% filter(riverOrderedGG == "West Brook"), aes(year,nAllFishBySpecies)) +
  geom_point(aes(color = speciesGG)) +
  geom_line(aes(color = speciesGG)) +
  geom_line( data = allFish %>% filter(riverOrderedGG == "West Brook"), aes(year,nAllFish) ) +
  theme_publication() +
  labs( x = "Year", y = "Count" ) +
  facet_wrap( ~ seasonGG )#, scales = "free_y" )

ggsave('figures/3Spp_WB/counts.png')

ggplot(allFishBySpeciesYOY %>% filter(riverOrderedGG == "West Brook", year >=2000), aes(year,nAllFishBySpecies)) +
  geom_point(aes(color = speciesGG), size=3) +
#  geom_line(aes(color = speciesGG)) +
  geom_smooth(aes(color = speciesGG), se=FALSE,method='lm') +
  geom_line( data = allFish %>% filter(riverOrderedGG == "West Brook", year >=2000), aes(year,nAllFish), size=1.25, color='darkgrey', linetype = 5 ) +
  scale_y_continuous(limits = c(0,NA)) +
  theme_publication() +
  labs( x = "Year", y = "Estimated abundance" ) +
  scale_color_grey(start = 0, end = 0.9) +
  facet_wrap( ~ seasonGG , scales = "free_y" )

ggsave('figures/3Spp_WB/countsWSlope.png', width = 11, height = 9, units='in')

nSlopes <- allFishBySpeciesYOY %>%
  filter(riverOrderedGG == "West Brook", year >=2000) %>%
  group_by( species,  season ) %>%
  nest()

nModel <- function(df){
  lm( nAllFishBySpecies ~ year, data = df )
}

nSlopes <- nSlopes %>%
  mutate( model = map(nSlopes$data, nModel) )

nSlopesTidy <- nSlopes %>%
  mutate( glance = map(model, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter( term == 'year')

save(nSlopesTidy,file = 'D:/projects/linkedModels/data/out/nSlopesTidy.RData')
########################
# add in sampleNumber
tmp <- d2 %>% distinct(season,year,sampleNumber)
allFishBySpeciesYOY <- left_join(allFishBySpeciesYOY,tmp)
allFish <- left_join(allFish,tmp)

ggplot(allFishBySpeciesYOY, aes(sampleNumber,nAllFishBySpeciesYOY)) +
  geom_point(aes(color = speciesGG)) +
  geom_line(aes(color = speciesGG)) +
  geom_line( data = allFish, aes(sampleNumber,nAllFish) ) +
  theme_publication() +
  labs( x = "SampleNumber", y = "Count" ) +
  facet_wrap(  ~riverOrderedGG )#, scales = "free_y" )

```

```{r Table 1 - sample summary}

# sample summary table
summaryTable <- allFishBySpecies %>% 
          select(-massAllFishBySpecies) %>% 
          filter(river == 'west brook', year %in% 2000:2015) %>% 
          spread(species,nAllFishBySpecies) %>% 
          arrange(year,seasonGG)

write.csv(summaryTable, file = 'D:/projects/linkedModels/tables/rawCounts.csv')

```

```{r}
# Correlation between age classes
# justification for not using separate age class abundances
# interesting that it is less for ATS
# From the bottom of adjustCounts()
# 
#                               nAllFishBySpeciesPStdBKT_yoy1 nAllFishBySpeciesPStdBKT_yoy2 nAllFishBySpeciesPStdBNT_yoy1 nAllFishBySpeciesPStdBNT_yoy2
# nAllFishBySpeciesPStdBKT_yoy1                          1.00                          0.96                         -0.46                         -0.43
# nAllFishBySpeciesPStdBKT_yoy2                        **0.96**                        1.00                         -0.44                         -0.38
# nAllFishBySpeciesPStdBNT_yoy1                         -0.46                         -0.44                          1.00                          0.94
# nAllFishBySpeciesPStdBNT_yoy2                         -0.43                         -0.38                       * *0.94**                       1.00
# nAllFishBySpeciesPStdATS_yoy1                         -0.13                         -0.17                         -0.21                          0.00
# nAllFishBySpeciesPStdATS_yoy2                          0.06                         -0.03                         -0.05                          0.15
#                               nAllFishBySpeciesPStdATS_yoy1 nAllFishBySpeciesPStdATS_yoy2
# nAllFishBySpeciesPStdBKT_yoy1                         -0.13                          0.06
# nAllFishBySpeciesPStdBKT_yoy2                         -0.17                         -0.03
# nAllFishBySpeciesPStdBNT_yoy1                         -0.21                         -0.05
# nAllFishBySpeciesPStdBNT_yoy2                          0.00                          0.15
# nAllFishBySpeciesPStdATS_yoy1                          1.00                          0.47
# nAllFishBySpeciesPStdATS_yoy2                         *0.47*                         1.00
```

```{r mean lengths}

# Has mean size changed over time?
means <- 
  cd %>% mutate( age = year - cohort ) %>%
    filter( riverOrdered == 'west brook', ageInSamples %in% 1:10, age < 3, year < 2016 ) %>%
    group_by(species,ageInSamples,year,season,age) %>%
    summarize(meanLen = mean(observedLength,na.rm = TRUE),
              sdLen = sd(observedLength,na.rm = TRUE),
              meanGR = mean(grLength,na.rm = TRUE),
              sdGR = sd(grLength,na.rm = TRUE)
              )

lenSlopes <- means %>%
  group_by( species, ageInSamples, season, age ) %>%
  nest()

lengthModel <- function(df){
  lm( meanLen ~ year, data = df )
}

grModel <- function(df){
  lm( meanGR ~ year, data = df )
}

lenSlopes <- lenSlopes %>%
  mutate( model = map(lenSlopes$data, lengthModel),
          modelGR = map(lenSlopes$data, grModel))

###################################
######### len here, gr below ######
lenSlopesTidy <- lenSlopes %>%
  mutate(glance = map(model, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter(term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
  mutate( sig = ifelse(p.value < 0.0001,"***",sig))

save(lenSlopesTidy,file = 'D:/projects/linkedModels/data/out/lenSlopesTidy.RData')
##
means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
lenSlopesTidy$seasonGG <- factor(lenSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
lenSlopesTidy$speciesGG <- factor(lenSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanLen, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean body length (mm)" ) +
  # geom_text(data = lenSlopesTidy %>% filter(species == 'bkt'), aes(x=1999,y=220, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = lenSlopesTidy %>% filter(species == 'bnt'), aes(x=1999,y=200, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = lenSlopesTidy %>% filter(species == 'ats'), aes(x=1999,y=180, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age<2), aes(x=2000,y=225, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age<2), aes(x=2000,y=203, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age<2), aes(x=2000,y=181, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +

    geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2012.5,y=104, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2012.5,y=82, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG~age)

ggsave(paste0('figures/3Spp_WB/meanLengths.png'), width = 11, height = 11, units='in')

# mean increases 
lenSlopesTidy %>%
  filter(species != 'ats') %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# mean increases by species
lenSlopesTidy %>%
  group_by(species) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
ggplot(lenSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point(aes( shape = factor(seasonGG)), size = 3) +
  geom_smooth(method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75)

ggsave(paste0('figures/3Spp_WB/meanLengthsOverAge.png'), width = 6, height = 6, units='in')

summary(lm(estimate~ageInSamples, data = lenSlopesTidy %>% filter(species=='bkt')))
summary(lm(estimate~ageInSamples, data = lenSlopesTidy %>% filter(species=='bnt')))

#########################
#### gr here, len above

grSlopesTidy <- lenSlopes %>%
  mutate(glance = map(modelGR, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter(term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
  mutate( sig = ifelse(p.value < 0.0001,"***",sig))

save(grSlopesTidy,file = 'D:/projects/linkedModels/data/out/grSlopesTidy.RData')
##
means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanGR, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean growth rate (mm/d)" ) +
  # geom_text(data = grSlopesTidy %>% filter(species == 'bkt'), aes(x=1999,y=220, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = grSlopesTidy %>% filter(species == 'bnt'), aes(x=1999,y=200, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  # geom_text(data = grSlopesTidy %>% filter(species == 'ats'), aes(x=1999,y=180, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age<2), aes(x=2000,y=0.7, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age<2), aes(x=2000,y=0.6, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age<2), aes(x=2000,y=181, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +

    geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2011.5,y=0.66, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2011.5,y=0.56, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG~age)

ggsave(paste0('figures/3Spp_WB/meanGRs.png'), width = 11, height = 11, units='in')

# mean increases 
grSlopesTidy %>%
  filter(species != 'ats') %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# mean increases by species
grSlopesTidy %>%
  group_by(species) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
ggplot(grSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point(aes( shape = factor(seasonGG)), size = 3) +
  geom_smooth(method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean growth rate over years" ) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75)

ggsave(paste0('figures/3Spp_WB/meanGRsOverAge.png'), width = 6, height = 6, units='in')

summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bkt')))
summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bnt')))

```

```{r figs 3,4, flow and temp}
d2$riverOrderedGG <- factor(d2$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
d2$seasonGG <- factor(d2$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)

d2 %>% group_by(riverOrderedGG,seasonGG,year) %>%
  filter( riverOrderedGG == "West Brook" ) %>%
  summarize(meanF = mean(meanFlow, na.rm=T),
            sdF = sd(meanFlow, na.rm=T)) %>%
  ggplot( aes(year,meanF) ) +
  geom_point( size = 3 ) +
  geom_line() +
 # geom_smooth(method = 'lm', se=F) +
  geom_errorbar( aes(ymin = meanF - sdF, ymax = meanF + sdF, width = 0.33 ) ) +
  labs( x = "Year", y = "Stream flow (x/x)" ) +
  theme_publication() +
  facet_wrap(~seasonGG)

ggsave('figures/3Spp_WB/flow.png')

lmFunctionF <- function(df){ lm( meanF ~ year, data=df ) }

flowMods <- d2 %>%
  group_by(riverOrderedGG,seasonGG,year) %>%
  summarize(meanF = mean(meanFlow, na.rm=T),
            sdF = sd(meanFlow, na.rm=T)) %>%
  group_by( riverOrderedGG, seasonGG ) %>%
  nest() %>%
  mutate( model = map(data,lmFunctionF) )

flowSlopesTidy <- flowMods %>%
  mutate(tidy = map(model,broom::tidy)) %>%
  dplyr::select(-data,-model) %>%
  unnest() %>%
  filter( term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*",""))

save(flowSlopesTidy,file = 'D:/projects/linkedModels/data/out/flowSlopesTidy.RData')

#################

lmFunction <- function(df){ lm( meanT ~ year, data=df ) }

tempMods <- d2 %>%
  group_by(riverOrderedGG,seasonGG,year) %>%
  summarize(meanT = mean(meanTemperature, na.rm=T),
            sdT = sd(meanTemperature, na.rm=T)) %>%
  group_by( riverOrderedGG, seasonGG ) %>%
  nest() %>%
  mutate( model = map(data,lmFunction) )

tempSlopesTidy <- tempMods %>%
  mutate(tidy = map(model,broom::tidy)) %>%
  dplyr::select(-data,-model) %>%
  unnest() %>%
  filter( term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*",""))

d2 %>% group_by(riverOrderedGG,seasonGG,year) %>%
  summarize(meanT = mean(meanTemperature, na.rm=T),
            sdT = sd(meanTemperature, na.rm=T)) %>%
  filter(riverOrderedGG == "West Brook") %>%
  ggplot( aes(year,meanT) ) +
  geom_smooth(method = 'lm', se=F, color = 'darkgrey') +
  geom_point( size = 3 ) +
  geom_line() +
  geom_errorbar( aes(ymin = meanT - sdT, ymax = meanT + sdT, width = 0.33 ) ) +
  labs( x = "Year", y = "Stream temperature (C)" ) +
  geom_text(data=tempSlopesTidy %>% filter(riverOrderedGG == "West Brook"),aes(x=2000.5,y=17.5, label=paste(round(estimate,2),sig))) +
  theme_publication() +
  facet_wrap(~seasonGG)

ggsave('figures/3Spp_WB/temp.png')

save(tempSlopesTidy,file = 'D:/projects/linkedModels/data/out/tempSlopesTidy.RData')
```

```{r Carys Jones analysis w season}
######
# for Carys Jones analysis
dJonesT <- 
  d2 %>% group_by(riverOrderedGG,seasonGG,year,season) %>%
    summarize(meanT = mean(meanTemperature, na.rm=T),
              sdT = sd(meanTemperature, na.rm=T)) %>%
    filter(riverOrderedGG == "West Brook") %>%
    ungroup() %>%
    dplyr::select(year,season,meanT,sdT)

####
# daily temps

#####
# Lengths
# from 'r mean lengths' chunk above

means <- 
  cd %>% mutate( age = year - cohort ) %>%
    #filter( riverOrdered == 'west brook', ageInSamples %in% 1:10, age < 3, year < 2016 ) %>%
    filter( riverOrdered == 'west brook', year < 2016 ) %>%
    group_by(species,ageInSamples,year,season,age,isYOY) %>%
    summarize(meanLen = mean(observedLength,na.rm = TRUE),
              sdLen = sd(observedLength,na.rm = TRUE),
              meanGR = mean(grLength,na.rm = TRUE),
              sdGR = sd(grLength,na.rm = TRUE),
              nLen = n()
              )

dJonesL <- 
  means %>%
   # filter(season == 3, age == 1) %>%
  #filter(age == 1) %>%
    ungroup() #%>%
    #dplyr::select(species,year,season,age,meanLen,sdLen)

######
# abundance
# run 'r adjusted counts' above

dJonesA <- 
  allFishBySpeciesYOY %>% 
  #  filter(riverOrderedGG == "West Brook", year >=2000, season == 3, isYOY == 2) %>%
  filter(riverOrderedGG == "West Brook") %>% #, year >=2000, isYOY == 2) %>%
    ungroup() %>%
    dplyr::select(species,year,isYOY,season,nAllFishBySpeciesYOY,nAllFishBySpecies)

######
# phi
# run 'r p(detection)' below
# only has summer survival, may want annual survival

dJonesP <- phi %>%
  #filter(season == 2, isYOYGG == 'Adult') %>%
  #filter(isYOYGG == 'Adult') %>%
  dplyr::select(speciesGG,yearGG,season,isYOYGG,isYOY,phi,phiSD,phi25,phi975) %>%
  rename(species = speciesGG, year = yearGG) %>%
  mutate(species = as.character(species))

dJonesAL <- left_join(dJonesA,dJonesL) 
dJonesALT <-  left_join(dJonesAL,dJonesT)
dJonesALTP <-  left_join(dJonesALT,dJonesP)

##
write.csv(dJonesALTP, file = 'D:/projects/linkedModels/dataForCarysJones/dJonesALTP_allSeasonsSpeciesAges.csv')
```

```{r mean gr graphs}
cd$riverOrderedGG <- factor(cd$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
cd$seasonGG <- factor(cd$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
cd$speciesGG <- factor(cd$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

pd <- position_dodge(0.65)

growthYOY <- 
cd %>% group_by(speciesGG,species,riverOrderedGG,seasonGG,isYOY) %>%
  filter( riverOrdered == 'west brook', year < 2016, species != 'ats' )  %>%
  summarize(meanGR = mean(grLength, na.rm=T),
            sdGR = sd(grLength, na.rm=T))

ggplot(growthYOY%>%filter(isYOY == 1), aes(seasonGG,meanGR, color=speciesGG, shape=speciesGG) ) +
  geom_point( position = pd, size = 3 ) +
  geom_errorbar( aes(ymin = meanGR - sdGR, ymax = meanGR + sdGR), width = 0.33, position=pd ) +
  ggtitle( "isYOY = 1" ) +
  labs( x = "Season", y = "Growth rate (mm/d)" ) +
  #scale_x_continuous( breaks = 1997:2015 ) +
  scale_color_grey(start = 0.0, end = 0.5) +
  theme_publication() +
  theme( panel.grid.minor = element_blank() ) 
#  facet_wrap( ~ riverOrderedGG)

ggsave('figures/3Spp_WB/meanGRYOY1.png', width = 4, height = 4, units='in')

ggplot(growthYOY%>%filter(isYOY == 2), aes(seasonGG,meanGR, color=speciesGG, shape=speciesGG) ) +
  geom_point( position = pd, size = 3 ) +
  geom_errorbar( aes(ymin = meanGR - sdGR, ymax = meanGR + sdGR), width = 0.33, position=pd ) +
  ggtitle( "isYOY = 2" ) +
  labs( x = "Season", y = "Growth rate (mm/d)" ) +
  #scale_x_continuous( breaks = 1997:2015 ) +
  scale_color_grey(start = 0.0, end = 0.5) +
  theme_publication() +
  theme( panel.grid.minor = element_blank() )

ggsave('figures/3Spp_WB/meanGRYOY2.png', width = 4, height = 4, units='in')

# get ratios of growth rates for the 2 species
growthYOYWide <- growthYOY %>% ungroup() %>% select(-sdGR,-speciesGG) %>% spread(key=species,value=meanGR) %>%
  mutate(ratio = bnt/bkt)

mean(growthYOYWide$ratio)
# 1.107745

```

```{r mean growth rate ratios}

cd %>% group_by(species) %>% filter(riverOrderedGG == "West Brook",species != 'ats') %>%
  summarize(meanGR = mean(grLength, na.rm=T),
            sdGR = sd(grLength, na.rm=T))
.197/.182#bnt/bkt

cd %>% group_by(isYOY) %>% filter(riverOrderedGG == "West Brook",species != 'ats') %>%
  summarize(meanGR = mean(grLength, na.rm=T),
            sdGR = sd(grLength, na.rm=T))
.145/.244 #adult/yoy
.244/.145

cd %>% group_by(season) %>% filter(riverOrderedGG == "West Brook",species != 'ats') %>%
  summarize(meanGR = mean(grLength, na.rm=T),
            sdGR = sd(grLength, na.rm=T))
.434/((.114+.124+.067)/3)
```

```{r - why bad convergence for bnt in winter?}

cd %>% filter() %>%
  ggplot(aes(observedLength,grLength)) +
  geom_point() +
  facet_grid(season~species)

```

```{r questions}

```

```{r p(detection)}
  # ggPInt <- array2df(ddD$sims.list$pBetaInt, label.x = "est")
  # 
  # ggPInt$chain <- rep(1:ddD$mcmc.info$n.chains, each = ddD$mcmc.info$n.samples/ddD$mcmc.info$n.chains)
  # ggPInt$iter <- 1:as.numeric(ddD$mcmc.info$n.samples/ddD$mcmc.info$n.chains)
  # 
  # ggplot(filter(ggPInt), aes(iter,est)) + 
  #   geom_point( aes(color=factor(chain)), size = 0.1 ) + 
  #   ylim(-1.5,1.5) + 
  #   facet_grid(d2+d4~d5+d3)

  p <- array2df(ddD$q50$pBetaInt, label.x = "est") %>%
    rename(est=est,isYOY=d1,species=d2,season=d3,river=d4,year=d5) %>% filter(river == 1, year > 3)

  p$speciesGG <- factor(p$species, levels = 1:3, labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
  p$seasonGG <- factor(p$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
  p$isYOYGG <- factor(p$isYOY, levels = 1:2, labels = c("Juvenile","Adult"), ordered = T)
  p$yearGG <- p$year + 1996

  ggplot(p, aes(yearGG,invlogit(est),color=factor(speciesGG))) + 
    geom_point(size = 2) + 
    labs( x = 'Year', y = 'Probability of detection') +
    scale_color_grey(start = 0.0, end = 0.85) +
    theme_publication() +
    facet_grid( seasonGG ~ isYOYGG)

  # seem to need to run the ggsave outside of markdown by pasting into console
   ggsave('figures/3Spp_WB/p.png', height = 11, width = 9, units = 'in')

  p %>% group_by(species) %>%
    summarize(meanP = mean(invlogit(est)),
              sdP = sd(invlogit(est)))

  p %>% group_by(season) %>%
    summarize(meanP = mean(invlogit(est)))

  ##################### 
  phi <- array2df(ddD$q50$phiBetaInt, label.x = "est") %>%
    rename(est=est,isYOY=d1,species=d2,season=d3,river=d4,year=d5) %>% filter(river == 1)#, year > 3)

  phi$speciesGG <- factor(phi$species, levels = 1:3, labels = c("bkt", "bnt", "ats"), ordered = T)
  phi$seasonGG <- factor(phi$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
  phi$isYOYGG <- factor(phi$isYOY, levels = 1:2, labels = c("Juvenile","Adult"), ordered = T)
  phi$yearGG <- phi$year + 1996
  phi$phi <- invlogit(phi$est)

  # add in sd of phiBetaInt
  phiSD <- array2df(ddD$sd$phiBetaInt, label.x = "estSD") %>%
    rename(estSD=estSD,isYOY=d1,species=d2,season=d3,river=d4,year=d5) %>% filter(river == 1)#, year > 3)  
  phiSD$phiSD <- invlogit(phiSD$estSD)

    # add in q2.5 of phiBetaInt
  phi25 <- array2df(ddD$q2.5$phiBetaInt, label.x = "est25") %>%
    rename(est25=est25,isYOY=d1,species=d2,season=d3,river=d4,year=d5) %>% filter(river == 1)#, year > 3)  
  phi25$phi25<- invlogit(phi25$est25)

    # add in 97.5 of phiBetaInt
  phi975 <- array2df(ddD$q97.5$phiBetaInt, label.x = "est975") %>%
    rename(est975=est975,isYOY=d1,species=d2,season=d3,river=d4,year=d5) %>% filter(river == 1)#, year > 3)  
  phi975$phi975 <- invlogit(phi975$est975)

  phi <- left_join(phi, phiSD)
  phi <- left_join(phi, phi25)
  phi <- left_join(phi, phi975) 
  phi <- phi %>% select(-estSD,-est25,-est975)

  ggplot(phi %>% filter(year > 3), aes(yearGG,phi,color=factor(speciesGG))) +geom_point() + facet_grid(seasonGG ~ isYOYGG)

  ggsave('figures/3Spp_WB/phi.png')

  phi %>% group_by(species) %>%
    summarize(meanPhi = mean(invlogit(est)),
              sdPhi = sd(invlogit(est)))

  phi %>% group_by(season) %>%
    summarize(meanPhi = mean(invlogit(est)))

# # of skunked sections    
  tmp <- dddD[[2]] %>%
           group_by(riverOrdered,year,season,section) %>%
           summarize( s = sum(enc) ) %>%
           filter( s == 0 )
  table(tmp$riverOrdered,tmp$year,tmp$season)

```

```{r r-hat}

rBkt <- unlist(mcmc[['bkt']]$Rhat)
sort(rBkt[rBkt > 1.1])

rBnt <- unlist(mcmc[['bnt']]$Rhat)
sort(rBnt[rBnt > 1.1])
mcmc[['bnt']]$Rhat$grBeta[,,,1]
mcmc[['bnt']]$Rhat$grInt
mcmc[['bnt']]$Rhat$grIntMu

rHat <- list()
rHat[['bkt']] <- array2df(mcmc[['bkt']]$Rhat$grBeta, levels = list(beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="rHat")
rHat[['bnt']] <- array2df(mcmc[['bnt']]$Rhat$grBeta, levels = list(beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="rHat")
rHat[['ats']] <- array2df(mcmc[['ats']]$Rhat$grBeta, levels = list(beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="rHat")

# rHat[['bkt']] <- array2df(mcmc[['bkt']]$Rhat$grBeta, levels = list(beta=1:8,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="rHat")
# rHat[['bnt']] <- array2df(mcmc[['bnt']]$Rhat$grBeta, levels = list(beta=1:8,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="rHat")
# rHat[['ats']] <- array2df(mcmc[['ats']]$Rhat$grBeta, levels = list(beta=1:8,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="rHat")

hist(rHat[["bkt"]]$rHat)
hist(rHat[["bnt"]]$rHat)
rHat[["bnt"]] %>% filter( rHat >1.1 )

ggplot(rHat[["bkt"]], aes(beta,rHat)) +
  geom_point() +
  facet_grid(river+isYOY~season)

```

```{r mcmc output analysis-grInt}

grInt <- list()
grInt[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$grInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int") %>% mutate(species = 'bkt')
grInt[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$grInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int") %>%   mutate(species = 'bnt')
grInt[['ats']] <- array2df(mcmc[['ats']]$sims.list$grInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int") %>% mutate(species = 'ats')

ggplot( grInt[['bkt']] %>% filter(river %in% c('west brook','wb jimmy')), aes(int)) +
  geom_freqpoly() +
  geom_freqpoly(aes((int)),color = 'brown', data = grInt[['bnt']]) +
  geom_freqpoly(aes((int)),color = 'lightblue', data = grInt[['ats']]) +
 # xlim(-0.5,1.5) +
  theme_publication() +
  facet_grid(season ~ river + isYOY)

#grIntDF <- rbind(grInt[['bkt']],grInt[['bnt']],grInt[['ats']])
#tmp <- grIntDF %>% filter(river == 'west brook')
#save(tmp , file= "data/out/grIntDFAllIters.RData") #for Ron

##################################################
# as range plot

plotHPD <- function(sppToPlot){

  nRivers <- list()
  nRivers[['bkt']] <- 4; nRivers[['bnt']] <- 2; nRivers[['ats']] <- 1

  probs <- c(0.025,0.1,0.5,0.9,0.975)
  q <- grInt[[sppToPlot]] %>%
    group_by(isYOY,season,river) %>%
    summarize(
      p = list(probs),
      q = list(quantile(int,probs))
    ) %>%
    unnest() %>%
    spread(key = p, value = q)

  p <- array2df(mcmc[[sppToPlot]]$mean$grInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="int") %>%
    mutate( season = as.numeric(season) ) %>%
    left_join(.,meanSampleIntervalMean, by = 'season')
  p$intByDay <- p$int#/p$meanSampleInterval
  p$intByDay <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intByDay)
  p$seasonGG <- factor(p$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
  p$parameter <- p$seasonGG

  pUp <- array2df(mcmc[[sppToPlot]]$q97.5$grInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="intUp") %>%
    mutate( season = as.numeric(season) ) 
  pDown <- array2df(mcmc[[sppToPlot]]$q2.5$grInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="intDown") %>%
    mutate( season = as.numeric(season) )

  p <- left_join(p,pUp); p <- left_join(p,pDown)

  p$intByDayUp <- p$intUp#/p$meanSampleInterval; 
  p$intByDayDown <- p$intDown#/p$meanSampleInterval
  p$intByDay <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intByDay)
  p$intByDayUp <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intByDayUp)
  p$intByDayDown <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intByDayDown)

  p$seasonGG <- factor(p$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
  #print(c(min(p$intByDayDown,na.rm=T),max(p$intByDayDown,na.rm=T)))

  #save(p, file= paste0("data/out/allInts_",sppToPlot,".RData"))  #for projections and tables

    gg <- ggplot(p %>% filter(river=='west brook'), aes(x = intByDay, y = reorder(parameter,intByDay,na.rm=TRUE), color = factor(isYOY), shape = factor(isYOY))) + 
        theme_publication() +  
        geom_point(size = 4) +
        geom_segment(aes(x = intByDay, xend = intByDayUp, yend = reorder(parameter, intByDay)), 
        size = 0.5, arrow = arrow(angle = 90, length = unit(0.05, "inches"))) + 
        geom_segment(aes(x = intByDay, xend = intByDayDown, yend = reorder(parameter, intByDay)), 
        size = 0.5,arrow = arrow(angle = 90, length = unit(0.05, "inches"))) +
        xlim(c(-0.00,0.55)) +
        xlab("Mean daily growth (mm/d)") + ylab("Season") +
        scale_color_grey(start = 0, end = 0.65) 
       # ggtitle(sppToPlot)

    return(gg)
}

plotHPD('bkt'); ggsave('figures/3Spp_WB/HDP_bkt.png', height = 5, width = 5, scale = 1)
plotHPD('bnt'); ggsave('figures/3Spp_WB/HDP_bnt.png', height = 5, width = 5, scale = 1)
plotHPD('ats'); ggsave('figures/3Spp_WB/HDP_ats.png', height = 5, width = 5, scale = 1)
```

```{r sigmaInt}

sigmaInt <- list()
sigmaInt[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$sigmaInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int")
sigmaInt[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$sigmaInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int")
sigmaInt[['ats']] <- array2df(mcmc[['ats']]$sims.list$sigmaInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int")

bw <- 0.05
ggplot( sigmaInt[['bkt']] %>% filter(river %in% c('west brook','wb jimmy')), aes(int)) +
  geom_freqpoly(binwidth = bw) +
  geom_freqpoly(aes(int),binwidth = bw,color = 'brown', data = sigmaInt[['bnt']]) +
  geom_freqpoly(aes(int),binwidth = bw,color = 'lightblue', data = sigmaInt[['ats']]) +
  #xlim(-4.5,-1) +
  xlim(-5,2.5) +
  theme_publication() +
  facet_grid(season ~ river + isYOY)

##################################################
# as range plot

plotHPDSigma <- function(sppToPlot){

  nRivers <- list()
  nRivers[['bkt']] <- 4; nRivers[['bnt']] <- 2; nRivers[['ats']] <- 1

  probs <- c(0.025,0.1,0.5,0.9,0.975)
  q <- sigmaInt[[sppToPlot]] %>%
    group_by(isYOY,season,river) %>%
    summarize(
      p = list(probs),
      q = list(quantile(int,probs))
    ) %>%
    unnest() %>%
    spread(key = p, value = q)

  p <- array2df(mcmc[[sppToPlot]]$mean$sigmaInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="int") %>%
    mutate( season = as.numeric(season) ) %>%
    left_join(.,meanSampleIntervalMean, by = 'season')
 # p$intByDay <- p$int/p$meanSampleInterval
#  p$intByDay <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intByDay)
  p$seasonGG <- factor(p$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
  p$parameter <- p$seasonGG

  pUp <- array2df(mcmc[[sppToPlot]]$q97.5$sigmaInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="intUp") %>%
    mutate( season = as.numeric(season) ) 
  pDown <- array2df(mcmc[[sppToPlot]]$q2.5$sigmaInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="intDown") %>%
    mutate( season = as.numeric(season) )

  p <- left_join(p,pUp); p <- left_join(p,pDown)

 # p$intByDayUp <- p$intUp/p$meanSampleInterval; p$intByDayDown <- p$intDown/p$meanSampleInterval
  p$int <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$int)
  p$intUp <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intUp)
  p$intDown <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intDown)

  #print(c(min(p$intDown,na.rm=T),max(p$intDown,na.rm=T)))

    gg <- ggplot(p %>% filter(river == 'west brook'), aes(x = exp(int), y = reorder(parameter,int,na.rm=TRUE), color = factor(isYOY), shape = factor(isYOY))) + 
        theme_publication() +  
        geom_point(size = 3) +
        geom_segment(aes(x = exp(int), xend = exp(intUp), yend = reorder(parameter, int)), 
        size = 0.5, arrow = arrow(angle = 90, length = unit(0.05, "inches"))) + 
        geom_segment(aes(x = exp(int), xend = exp(intDown), yend = reorder(parameter, int)), 
        size = 0.5,arrow = arrow(angle = 90, length = unit(0.05, "inches"))) +
        #xlim(c(-0.0015,0.01)) +
        xlab("Standard deviation of daily growth (mm/d)") + ylab("Season") +
        scale_color_grey(start = 0, end = 0.65) 
       # ggtitle(sppToPlot)

    return(gg)
}

plotHPDSigma('bkt'); ggsave('figures/3Spp_WB/HDPSigma_bkt.png', height = 5, width = 5, scale = 1)
plotHPDSigma('bnt'); ggsave('figures/3Spp_WB/HDPSigma_bnt.png', height = 5, width = 5, scale = 1)
plotHPDSigma('ats'); ggsave('figures/3Spp_WB/HDPsigma_ats.png', height = 5, width = 5, scale = 1)
```

```{r grInt}
#riverOrderedIn <- factor(c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),labels = c("west brook","wb jimmy","wb mitchell","wb obear"), ordered = T)

grIntMu <- list()
grIntMu[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$grIntMu, levels = list(iter=NA,season=1:4), label.x="int")
grIntMu[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$grIntMu, levels = list(iter=NA,season=1:4), label.x="int")
grIntMu[['ats']] <- array2df(mcmc[['ats']]$sims.list$grIntMu, levels = list(iter=NA,season=1:4), label.x="int")

bw <- 0.05
ggplot( grIntMu[['bkt']] , aes(int)) +
  geom_freqpoly(binwidth = bw, size=1.2) +
  geom_freqpoly(aes(int),binwidth = bw,color = 'brown', data = grIntMu[['bnt']]) +
  geom_freqpoly(aes(int),binwidth = bw,color = 'lightblue', data = grIntMu[['ats']]) +
  #xlim(-4.5,-1) +
  xlim(-0.25,0.75) +
 # ylim(0,2500) +
  theme_publication() +
  facet_wrap(~season )

```

```{r mcmc output analysis-grBeta}

# get all betas in a DF
grBeta <- list()
grBeta[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$grBeta, levels = list(iter=NA,beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int") %>%
  mutate( species = 'bkt')
grBeta[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$grBeta, levels = list(iter=NA,beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int") %>%
  mutate( species = 'bnt')
grBeta[['ats']] <- array2df(mcmc[['ats']]$sims.list$grBeta, levels = list(iter=NA,beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int") %>%
  mutate( species = 'ats')

grBetaDF1 <- rbind(grBeta[['bkt']],grBeta[['bnt']],grBeta[['ats']])

grBetaBNT <- list()
grBetaBNT[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$grBetaBNT, levels = list(iter=NA,beta=8:9,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int") %>%
  mutate( species = 'bkt')
grBetaBNT[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$grBetaBNT, levels = list(iter=NA,beta=8:9,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int") %>%
  mutate( species = 'bnt')
grBetaBNT[['ats']] <- array2df(mcmc[['ats']]$sims.list$grBetaBNT, levels = list(iter=NA,beta=8:9,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int") %>%
  mutate( species = 'ats')

grBetaBNTDF <- rbind(grBetaBNT[['bkt']],grBetaBNT[['bnt']],grBetaBNT[['ats']])

grBetaATS <- list()
grBetaATS[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$grBetaATS, levels = list(iter=NA,beta=10,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int") %>%
  mutate( species = 'bkt')
grBetaATS[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$grBetaATS, levels = list(iter=NA,beta=10,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int") %>%
  mutate( species = 'bnt')
grBetaATS[['ats']] <- array2df(mcmc[['ats']]$sims.list$grBetaATS, levels = list(iter=NA,beta=10,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int") %>%
  mutate( species = 'ats')

grBetaATSDF <- rbind(grBetaATS[['bkt']],grBetaATS[['bnt']],grBetaATS[['ats']])

grBetaDF <- rbind(grBetaDF1,grBetaBNTDF,grBetaATSDF)

#tmp <- grBetaDF %>% filter(river == 'west brook')
#save(tmp , file= "data/out/grBetaDFAllIters.RData") #for Ron
##################################################
# as range plot

  nRivers <- list()
  nRivers[['bkt']] <- 4; nRivers[['bnt']] <- 2; nRivers[['ats']] <- 1
  speciesList <- list('bkt','bnt','ats')

  probs <- c(0.025,0.1,0.5,0.9,0.975)
  nSpecies <- 3

  q <- grBetaDF %>%
    group_by(beta,isYOY,season,river,species) %>%
    summarize(
      p = list(probs),
      q = list(quantile(int,probs))
    ) %>%
    unnest() %>%
    spread(key = p, value = q) %>%
    data.frame() %>%
    rename( median = X0.5, lo = X0.025, hi = X0.975 )

  q <- q %>% filter( !(season == 2 & isYOY == 0))

q$riverGG <- factor(q$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("WB","OL","OS","IL"), ordered = T)
q$seasonGG <- factor(q$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
q$speciesGG <- factor(q$species, levels = c('bkt','bnt','ats'), labels = c("BKT", "BNT", "ATS"), ordered = T)
q$betaGG <- factor(q$beta, levels = 1:10, 
                   #levels = c("Temp","Flow","Temp*Flow","Temp^2","Flow^2","BKT","Length","BNT","BNT*BKT","ATS"), # make sure this order matches the nimble code
                   labels = c("Temp","Flow","Temp*Flow","Temp^2","Flow^2","BKT","Length","BNT","BNT*BKT","ATS"),ordered = T)
q$parameter = q$speciesGG
q$parameterOrd <- factor(q$parameter, 
                         levels = c( "ATS", "BNT", "BKT"), 
                         labels = c( "WB_ATS", "WB_BNT", "WB_BKT"), ordered = T)

  q <- q %>% filter( !(betaGG == 'ATS' & (riverGG == "OS" | riverGG == "IL" | riverGG == "OL")) )
  q <- q %>% filter( !(betaGG == 'BNT' & (riverGG == "OS" | riverGG == "IL")) )
  q <- q %>% filter( !(betaGG == 'BNT*BKT' & (riverGG == "OS" | riverGG == "IL")) )
#  q <- q %>% filter( !((riverGG == "OS" | riverGG == "IL")) )

  save(q, file= "data/out/allBetas.RData")  #for projections and tables

  #reorder(parameter,median,na.rm=TRUE)
(min(q$median))
  (max(q$median))
#  q[q$median < -0.13,]

    ggplot(q %>% filter(river == 'west brook', species != 'ats'), aes(x = median, y = parameterOrd, color = factor(species), shape = factor(isYOY))) + 
        geom_vline(xintercept = 0,color='grey') +
        theme_publication() +  
        geom_point(size = 4) +
        geom_segment(aes(x = median, xend = hi, yend = parameterOrd), 
        size = 0.5, arrow = arrow(angle = 90, length = unit(0.1, "inches"))) + 
        geom_segment(aes(x = median, xend = lo, yend = parameterOrd), 
        size = 0.5,arrow = arrow(angle = 90, length = unit(0.1, "inches"))) +
     #   xlim(c(-0.15,0.11)) +
        scale_x_continuous( breaks=c(-0.1,0,0.1)) +
        xlab("Parameter estimate") + theme(axis.text.y = element_blank(), axis.title.y = element_blank()) +
        scale_color_grey(start = 0, end = 0.5)  +
        facet_grid(seasonGG~betaGG)

ggsave('figures/3Spp_WB/grBetas.png', height = 8, width = 10)

save(betas,file = 'D:/projects/linkedModels/data/out/nSlopesTidy.RData')

```

```{r table of Ints}
load(file = "data/out/allInts_bkt.RData") #loads p
pBKT <- p %>% mutate(species = 'bkt')

load(file = "data/out/allInts_bnt.RData") #loads p
pBNT <- p %>% mutate(species = 'bnt')

pp <- rbind(pBKT,pBNT)

intsForTable <- pp %>% 
  filter(river == 'west brook', species %in% c('bkt','bnt')) %>%
  arrange(species,isYOY,seasonGG) %>%
  mutate(lo = round(intDown,3),mean = round(int,3),hi = round(intUp,3)) %>%
  select(species,isYOY,seasonGG,intDown,int,intUp)

write.csv(intsForTable,file = 'D:/projects/linkedModels/data/out/intsForTableAgeGroup.csv')
```

```{r table of betas}

load(file= "data/out/allBetas.RData") #loads q

betasForTable <- q %>% 
  filter(river == 'west brook', species %in% c('bkt','bnt')) %>%
  arrange(species,betaGG,isYOY,seasonGG,betaGG) %>%
  mutate(lo = round(lo,3),median = round(median,3),hi = round(hi,3)) %>%
  select(species,betaGG,isYOY,seasonGG,betaGG,lo,median,hi)

write.csv(betasForTable,file = 'D:/projects/linkedModels/data/out/betasForTableAgeGroup.csv')
```

```{r mcmc output analysis-grBetaSigma}

# get all betas in a DF
grBetaSigma <- list()
grBetaSigma[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$sigmaBeta, levels = list(iter=NA,beta=1:3,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int") %>%
  mutate( species = 'bkt')
grBetaSigma[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$sigmaBeta, levels = list(iter=NA,beta=1:3,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int") %>%
  mutate( species = 'bnt')
grBetaSigma[['ats']] <- array2df(mcmc[['ats']]$sims.list$sigmaBeta, levels = list(iter=NA,beta=1:3,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int") %>%
  mutate( species = 'ats')

grBetaSigmaDF <- rbind(grBetaSigma[['bkt']],grBetaSigma[['bnt']],grBetaSigma[['ats']])

##################################################
# as range plot

  nRivers <- list()
  nRivers[['bkt']] <- 4; nRivers[['bnt']] <- 2; nRivers[['ats']] <- 1
  speciesList <- list('bkt','bnt')

  probs <- c(0.025,0.1,0.5,0.9,0.975)
  nSpecies <- 2

  q <- grBetaSigmaDF %>%
    group_by(beta,isYOY,season,river,species) %>%
    summarize(
      p = list(probs),
      q = list(quantile(int,probs))
    ) %>%
    unnest() %>%
    spread(key = p, value = q) %>%
    data.frame() %>%
    rename( median = X0.5, lo = X0.025, hi = X0.975 )

  q <- q %>% filter( !(season == 2 & isYOY == 0))

q$riverGG <- factor(q$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("WB","OL","OS","IL"), ordered = T)
q$seasonGG <- factor(q$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
q$speciesGG <- factor(q$species, levels = c('bkt','bnt','ats'), labels = c("BKT", "BNT", "ATS"), ordered = T)
q$betaGG <- factor(q$beta, levels = 1:10, labels = c("Temp", "Flow", "Temp*Flow","Temp^2","Flow^2","BKT","Length","BNT","BNT*BKT","ATS"), ordered = T)
q$parameter = q$speciesGG
q$parameterOrd <- factor(q$parameter, 
                         levels = c( "ATS", "BNT", "BKT"), 
                         labels = c( "WB_ATS", "WB_BNT", "WB_BKT"), ordered = T)

  q <- q %>% filter( !(betaGG == 'ATS' & (riverGG == "OS" | riverGG == "IL" | riverGG == "OL")) )
  q <- q %>% filter( !(betaGG == 'BNT' & (riverGG == "OS" | riverGG == "IL")) )
  q <- q %>% filter( !(betaGG == 'BNT*BKT' & (riverGG == "OS" | riverGG == "IL")) )
#  q <- q %>% filter( !((riverGG == "OS" | riverGG == "IL")) )

#reorder(parameter,median,na.rm=TRUE)
(min(q$median))
  (max(q$median))
#  q[q$median < -0.13,]

    ggplot(q %>% filter(river == 'west brook', species != 'ats'), aes(x = median, y = parameterOrd, color = factor(speciesGG), shape = factor(isYOY))) + 
        geom_vline(xintercept = 0,color='black') +
        theme_publication() +  
        geom_point(size = 3) +
        geom_segment(aes(x = median, xend = hi, yend = parameterOrd), 
        size = 0.5, arrow = arrow(angle = 90, length = unit(0.05, "inches"))) + 
        geom_segment(aes(x = median, xend = lo, yend = parameterOrd), 
        size = 0.5,arrow = arrow(angle = 90, length = unit(0.05, "inches"))) +
     #   xlim(c(-0.35,0.5)) +
        xlab("Parameter estimate") + theme(axis.text.y = element_blank(), axis.title.y = element_blank()) +
        scale_color_grey(start = 0, end = 0.5)  +
        facet_grid(seasonGG~betaGG)

ggsave('figures/3Spp_WB/grBetaSigmas.png', height = 8, width = 8)

```

```{r ind RE among spp}

grIndRE <- list()
grIndRE[['bkt']] <- as.data.frame(mcmc[['bkt']]$q50$grIndRE); names(grIndRE[['bkt']]) <- 'value'
grIndRE[['bnt']] <- as.data.frame(mcmc[['bnt']]$q50$grIndRE); names(grIndRE[['bnt']]) <- 'value'
grIndRE[['ats']] <- as.data.frame(mcmc[['ats']]$q50$grIndRE); names(grIndRE[['ats']]) <- 'value'

table(mcmc[['bkt']]$q50$grIndRE==0)/length(mcmc[['bkt']]$q50$grIndRE)
table(mcmc[['bnt']]$q50$grIndRE==0)/length(mcmc[['bnt']]$q50$grIndRE)
table(mcmc[['ats']]$q50$grIndRE==0)/length(mcmc[['ats']]$q50$grIndRE)

ggplot( grIndRE[[1]], aes(value,..density..) ) +
  geom_freqpoly(binwidth = 0.00025) +
  geom_freqpoly(aes(value), color = "brown", binwidth = 0.00025, data = grIndRE[[2]]) +
  geom_freqpoly(aes(value), color = "lightblue", binwidth = 0.00025, data = grIndRE[[3]]) +
  ylim(c(0,100))

```

```{r ind trajectories}

```

```{r predictions}
# 
#####run if get new predictions - takes a while to run ####

 # load('D:/projects/linkedModels/data/out/P_ForMike_bkt.RData')
 # pBoth$species <- 'bkt'
 # pAll <- pBoth; rm(pBoth)
 # 
 # load('D:/projects/linkedModels/data/out/P_ForMike_bnt.RData')
 # pBoth$species <- 'bnt'
 # pAll <- rbind(pAll,pBoth); rm(pBoth)
 # 
 # load('D:/projects/linkedModels/data/out/P_ForMike_ats.RData')
 # pBoth$species <- 'ats'
 # pAll <- rbind(pAll,pBoth); rm(pBoth)
 # 
 # pred <- pAll %>% group_by(species,len,flow,temp,cBKT,cBNT,cATS,isYOY,season,river) %>%
 #   summarize(meanPredGR = mean(predGr, na.rm=T),
 #             sdPredGR =     sd(predGr, na.rm=T),
 #             meanPredGRSigma = mean(predGrSigma, na.rm=T),
 #             sdPredGRSigma =     sd(predGrSigma, na.rm=T),
 #             meanPredGRCV = mean(predCV, na.rm=T),
 #             sdPredGRCV =     sd(predCV, na.rm=T)) %>%
 #    ungroup()
 # 
 # rm(pAll)
 # save(pred,file = 'D:/projects/linkedModels/data/out/P_ForMike_allMeans.RData')

load('D:/projects/linkedModels/data/out/P_ForMike_allMeans.RData')

pred <- pred %>% filter( !(isYOY == 0 & season == 2) ) 
#pred <- pred %>% filter( !(isYOY == 0 & river == 'wb jimmy' & season == 2) )
#graph function, in analyzeOutputFunctions.R

pred$riverGG <- factor(pred$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
pred$seasonGG <- factor(pred$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
pred$speciesGG <- factor(pred$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

plotPredMeans(pred, "meanPredGR", "len", 0)
plotPredMeans(pred %>% filter(river == 'west brook'), "meanPredGR", "len", 1)

plotPredMeans(pred, "meanPredGR", "flow", 0)
plotPredMeans(pred, "meanPredGR", "flow", 1)

plotPredMeans(pred, "meanPredGR", "temp", 0)
plotPredMeans(pred, "meanPredGR", "temp", 1)

plotPredMeans(pred, "meanPredGR", "cBKT", 0)
plotPredMeans(pred, "meanPredGR", "cBKT", 1)

plotPredMeans(pred, "meanPredGR", "cBNT", 0)
plotPredMeans(pred, "meanPredGR", "cBNT", 1)

plotPredMeans(pred, "meanPredGR", "cATS", 0)
plotPredMeans(pred, "meanPredGR", "cATS", 1)

plotPredMeans(pred, "meanPredGR", c("temp","flow"), 0)
plotPredMeans(pred, "meanPredGR", c("temp","flow"), 1)

plotPredMeans(pred, "meanPredGR", c("flow","temp"), 0)
plotPredMeans(pred, "meanPredGR", c("flow","temp"), 1)

ggsave('figures/3Spp_WBtmp.png', width = 9, height = 7, units='in')
```

```{r mean prediction graphs for meanPredGR}

all <- c('len','temp','flow','cBKT','cBNT','cATS')
allLabels <- c('Length (mm)','Temperature (C)','Flow (m3/s)','Brook Trout abundance','Brown Trout abundance','Atlantic Salmon abundance')

all <- c('len','cBKT','temp','cBNT','flow','cATS')
allLabels <- c('Length (mm)','Brook Trout abundance','Stream temperature (C)','Brown Trout abundance','Stream flow (m3/s)','Atlantic Salmon abundance')

remFacetLabel <- c(3,4,5,6)

selectSingleVar <- function(p, all, depVar, varsToPlot, isYOYGG){

  notPlot <- NA
  notPlot[1] <- all[!(all %in% varsToPlot)][1]
  notPlot[2] <- all[!(all %in% varsToPlot)][2]
  notPlot[3] <- all[!(all %in% varsToPlot)][3]
  notPlot[4] <- all[!(all %in% varsToPlot)][4]
  notPlot[5] <- all[!(all %in% varsToPlot)][5]

  pGG <- p %>% filter(isYOY == isYOYGG, 
                        eval(as.name(notPlot[1])) == 0, 
                        eval(as.name(notPlot[2])) == 0, 
                        eval(as.name(notPlot[3])) == 0,                                         
                        eval(as.name(notPlot[4])) == 0, 
                        eval(as.name(notPlot[5])) == 0
                      )#%>%
              # distinct(species, isYOY, riverGG, seasonGG, speciesGG, .keep_all = TRUE)

  return(pGG)  
}

##### by species

isYOYGG <- 0
riverGGToPlot <- "West Brook" #"Open Large"
outputToPlot <- "meanPredGR"

ggSingle <- list()
singleDat <- list()
for(i in 1:length(all))
  #https://stackoverflow.com/questions/31993704/storing-ggplot-objects-in-a-list-from-within-loop-in-r
      local({ i <- i
  singleDat[[i]] <- selectSingleVar(pred, all, outputToPlot, all[i], isYOYGG)
  print(c(i,all[i]))
 # print(str(singleDat[[i]]))

  tmp <- singleDat[[i]] %>% filter(riverGG == riverGGToPlot)
  ggSingle[[i]] <<- ggplot( tmp %>% filter(species != 'ats'), aes( eval(as.name(all[i])), eval(as.name(outputToPlot)), color=speciesGG ) ) +
      geom_line( size = 1.225 ) +
      theme_bw() +
      theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), 
            axis.title.y = element_blank(),
            legend.position="none") +
      geom_hline(yintercept = 0) +
      ylim(-0.15,0.8) +
      scale_x_continuous(allLabels[i]) +
      scale_color_grey(start = 0, end = 0.5)  +
      facet_grid(~seasonGG)

  if( i %in% remFacetLabel ) {
    ggSingle[[i]] <<- ggSingle[[i]] +
    theme(
      strip.background = element_blank(),
      strip.text.x = element_blank()
    )
  }
})

grid.arrange(grobs=ggSingle, left = "Growth rate (mm/d)")
# save from rStudio as export, width=900, height=500
# ggsave(gridPred,'figures/3Spp_WB/predGRYOY1.png', width = 9, height = 7, units='in')

```

```{r mean prediction graphs for meanPredGRSigma}

all <- c('temp','flow')
allLabels <- c('Stream temperature (C)','Stream flow (m3/s)')

remFacetLabel <- c(2)

selectSingleVarSigma <- function(p, all, depVar, varsToPlot, isYOYGG){

  notPlot <- NA
  notPlot[1] <- all[!(all %in% varsToPlot)][1]

  pGG <- p %>% filter(isYOY == isYOYGG, 
                        eval(as.name(notPlot[1])) == 0,
                      len == 0, cBKT == 0, cBNT == 0, cATS == 0
                      )#%>%
              # distinct(species, isYOY, riverGG, seasonGG, speciesGG, .keep_all = TRUE)

  return(pGG)  
}

##### by species

isYOYGG <- 1
riverGGToPlot <- "West Brook" #
outputToPlot <- "meanPredGRSigma"

ggSingle <- list()
singleDat <- list()
for(i in 1:length(all))
  #https://stackoverflow.com/questions/31993704/storing-ggplot-objects-in-a-list-from-within-loop-in-r
      local({ i <- i
  singleDat[[i]] <- selectSingleVarSigma(pred, all, outputToPlot, all[i], isYOYGG)
  print(c(i,all[i]))
 # print(str(singleDat[[i]]))

  tmp <- singleDat[[i]] %>% filter(riverGG == riverGGToPlot, species != 'ats')
  ggSingle[[i]] <<- ggplot( tmp, aes( eval(as.name(all[i])), eval(as.name(outputToPlot)), color=speciesGG ) ) +
      geom_line( size = 1.225 ) +
      theme_bw() +
      theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), 
            axis.title.y = element_blank(),
            legend.position="none") +
      geom_hline(yintercept = 0) +
    #  ylim(-0.15,0.8) +
      scale_x_continuous(allLabels[i]) +
      scale_color_grey(start = 0, end = 0.5)  +
      facet_grid(~seasonGG)

  if( i %in% remFacetLabel ) {
    ggSingle[[i]] <<- ggSingle[[i]] +
    theme(
      strip.background = element_blank(),
      strip.text.x = element_blank()
    )
  }
})

# add thrid row to graph for interaction of tem and flow?

# run this and export
grid.arrange(grobs=ggSingle, left = "Standard deviation of growth rate (mm/d)")

# 0 = 875,600
# 1 = 875,500

#ggsave(ggCV, paste0('figures/3Spp_WBpredGRCV_YOY',isYOYGG,'_',riverGGToPlot,'.png'), width = 9, height = 7, units='in')

```

```{r flow*temp interaction plots}

###############################
# flow*temp interaction plots
isYOYGG <- 0
riverGGToPlot <- "West Brook" #"Open Large" #
outputToPlot <- "meanPredGRSigma" #"meanPredGR" #

interact <- pred %>% filter(isYOY == isYOYGG,
                            riverGG == riverGGToPlot,
                            len == 0, cBKT == 0, cBNT == 0, cATS == 0 ) %>%
                     mutate( temp_spp = paste0(temp,speciesGG))

    ggplot( interact %>% filter(species != 'ats'), aes( flow, eval(as.name(outputToPlot)), group=temp_spp, linetype=speciesGG ) ) +
      geom_line( aes(alpha = temp + 3), size = 1.225 ) +
      theme_bw() +
      theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), 
          #  axis.title.y = element_blank()
            legend.position="none"
            ) +
      geom_hline(yintercept = 0) +
    #  ylim(-0.15,0.8) +
      scale_x_continuous('Stream flow') +
      scale_y_continuous("Standard deviation of growth rate (mm/d)") +
      #scale_color_manual(values=c('blue','darkgreen'))  +
      facet_grid(~seasonGG)

#yoy0 = 875,350
#yoy1 = 875,280
```

```{r bkt*bnt interaction plots}

###############################
# flow*temp interaction plots
isYOYGG <- 1
riverGGToPlot <- "West Brook" #
outputToPlot <- "meanPredGR" #"meanPredGRSigma"

interact <- pred %>% filter(isYOY == isYOYGG,
                            riverGG == riverGGToPlot,
                            len == 0, temp == 0, flow == 0, cATS == 0 ) %>%
                     mutate( bnt_spp = paste0(cBNT,speciesGG))

    ggplot( interact %>% filter(species != 'ats'), aes( cBKT, eval(as.name(outputToPlot)), group=bnt_spp, linetype=speciesGG ) ) +
      geom_line( aes(alpha = cBNT + 3), size = 1.225 ) +
      theme_bw() +
      theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), 
           # axis.title.y = element_blank()
            legend.position="none"
            ) +
      geom_hline(yintercept = 0) +
    #  ylim(-0.15,0.8) +
      scale_x_continuous('Brook Trout abundance') +
      scale_y_continuous("Growth rate (mm/d)") +
      facet_grid(~seasonGG)

# for saving from export
#yoy0 = 875,350
#yoy1 = 875,275

```

```{r projections}
# run 'Load data' chunk at top to get the 'mcmc' object

#data for running the simulation
simData <- list(
  nReps = 100, #num of mcmc iterations to grab
  numInd = 500 , # number of individual to simulate
  nOccYOY = 3,
  nOccAdult = 7,
  riverIndex = 1, #Levels: west brook < wb jimmy < wb mitchell < wb obear
  minSD = -1.5, #range and step for independent variable values in the sim
  maxSD = 1.5, 
  step = 1.5
)
simData$nOcc <- simData$nOccYOY + simData$nOccAdult
simData$indVarSeq = seq(simData$minSD,simData$maxSD,simData$step)

stats3Spp <- grStats3Spp <- anov <- elapsed <- lm <- relImp <- list()
spp <- c('bkt','bnt') #,'ats')

for( speciesForProj in spp ){

  # get mean on mean length by cohort for the fall yoy sample
  obsLenStats <-  
    ddGIn2[[speciesForProj]] %>%
      group_by(cohort,ageInSamples) %>%
      summarise(lenMean = mean(observedLength, na.rm=TRUE),
                lenSD = sd(observedLength, na.rm=TRUE)) %>%
      filter(ageInSamples == 1) %>%
      ungroup() %>%
      summarise(lenMeanMean = mean(lenMean, na.rm=TRUE),
                lenSDMean = mean(lenSD, na.rm=TRUE))

  simData$species = speciesForProj
  simData$meanInit = obsLenStats$lenMeanMean
  simData$sdInit = obsLenStats$lenSDMean
  simData$nSamplesAvail = mcmcInfo$nChains * mcmcInfo$nSamples

  simData <- updateSimData(simData) # add #season, itersForSim etc data to simData
  nB <- getNBetas() # get number of betas from mcmc[]
  m <- createMatrices()

  indVars <- getIndVars() # get independent variables table, -1.5, 0, 1.5

  startTime <- Sys.time()
  sizeStats <- sizeStats2 <- grStats <- grStats2 <- NULL
  #loop over conditions
  for(testRow in indVars$rowNum){
    project <- projectSizes(m,simData,testRow)

    sizeMat <- project$sizeMat
    grMat <- project$grMat

    sizeStats1 <- summarizeLens(sizeMat,testRow)
    sizeStats <- rbind(sizeStats,sizeStats1[[1]]) #means over individuals (grouped by iter, occ)
    sizeStats2 <- rbind(sizeStats2,sizeStats1[[2]]) #means by occ of means over individuals (grouped by occ)

    grStats1 <- summarizeGRs(grMat,testRow)
    grStats <- rbind(grStats,grStats1[[1]]) #means over individuals (grouped by iter, occ)
    grStats2 <- rbind(grStats2,grStats1[[2]]) #means by occ of means over individuals (grouped by occ)
  }
  endTime <- Sys.time()
  (elapsed[[speciesForProj]] <- endTime - startTime)

  stats3Spp[[speciesForProj]] <- list( sizeStats,sizeStats2 )
  grStats3Spp[[speciesForProj]] <- list( grStats,grStats2 )

  # lm[[speciesForProj]] <- lm( lenMean ~ temp + flow + bkt + bnt + ats + temp * flow + temp^2 + flow^2 + bnt*bkt, data = sizeStats %>% filter(occ == (simData$nOcc)))
  # 
  # relImp[[speciesForProj]] <- calc.relimp(lm[[speciesForProj]], type = c("lmg"), rela = TRUE)
  #   
  # anov[[speciesForProj]] <- anova( lm[[speciesForProj]] )

}

#projOut <- list(stats3Spp, lm, relImp, anov, elapsed, simData)
save( stats3Spp, grStats3Spp, elapsed, simData, file = paste0('D:/projects/linkedModels/data/out/projectionOut_N',simData$nReps,'.RData'))

# if not re-running the sim
# load( file = 'D:/projects/linkedModels/data/out/projectionOut.RData')
# load( file = paste0('D:/projects/linkedModels/data/out/projectionOut_N',simData$nReps,'.RData'))
# spp <- c('bkt','bnt','ats')

# raw sim data figs - size
ggplot(stats3Spp[['bkt']][[1]] %>% filter(ats == 0, bkt == 0, bnt == 0), aes(as.numeric(occ),lenMean)) +
  geom_jitter(aes(color = factor(ats))) +
  #geom_smooth(aes(color = factor(temp), linetype = factor(flow)))+#(bnt + 2)/4 )) +
  geom_smooth()+#(bnt + 2)/4 )) +
  geom_smooth(color = 'green', data = stats3Spp[['bnt']][[1]] %>% filter(ats == 0, bkt == 0, bnt == 0)) +
  facet_grid(temp ~ flow)

########################
# raw sim data figs - gr
ggplot(grStats3Spp[['bkt']][[1]] %>% filter(ats == 0, bkt == 0, bnt == 0), aes(as.numeric(occ),grMean)) +
  geom_jitter(aes(color = factor(ats)), width = 0.2, size = 0.5) +
  #geom_smooth(aes(color = factor(temp), linetype = factor(flow)))+#(bnt + 2)/4 )) +
  #geom_smooth()+#(bnt + 2)/4 )) +
  geom_jitter(color = 'darkblue', width = 0.1, size = 0.5, data = grStats3Spp[['bnt']][[1]] %>% filter(ats == 0, bkt == 0, bnt == 0)) +
  facet_grid(temp ~ flow)

```

```{r RI across all - use this for paper GROWTH RATE}

# relative proportion stats with occ in the lm 
grStats3Spp[['bkt']][[1]]$season <- as.factor(c(3,4,1,2,3,4,1,2,3,4))
grStats3Spp[['bkt']][[1]]$temp2 <- grStats3Spp[['bkt']][[1]]$temp^2
grStats3Spp[['bkt']][[1]]$flow2 <- grStats3Spp[['bkt']][[1]]$flow^2

indepVarVals <- c(-1.5,0,1.5) # make sure symetrical around 0
indepVars <- c('temp', 'flow', 'temp:flow', 'bkt', 'bnt', 'ats', 'bkt:bnt')
rIBKTBNTAll <- NULL
# loop over occs
for(rep in 1:simData$nReps){

  print(rep)

  datOccBKTAll <- grStats3Spp[['bkt']][[1]] %>% filter(iter == rep, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
  lmAllBKT <- lm( grMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBKTAll) 
  relImpAllBKT <- calc.relimp(lmAllBKT, type = c("lmg"), rela = TRUE)

  datOccBNTAll <- grStats3Spp[['bnt']][[1]] %>% filter(iter == rep, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
  lmAllBNT <- lm( grMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBNTAll)
  relImpAllBNT <- calc.relimp(lmAllBNT, type = c("lmg"), rela = TRUE)

  lmgBKT <- relImpAllBKT$lmg[(names(relImpAllBKT$lmg) %in% indepVars)]
  rBKT <- lmgBKT / sum(lmgBKT)

  lmgBNT <- relImpAllBNT$lmg[(names(relImpAllBNT$lmg) %in% indepVars)]
  rBNT <- lmgBNT / sum(lmgBNT)

  rIBKTBNTAll <- rbind(rIBKTBNTAll, data.frame(cbind(rBKT,rBNT)) %>% 
                      rownames_to_column() %>% 
                      gather(key = rowname) %>% 
                      mutate(var = rep(1:length(indepVars),2),
                             varText = rep(names(rBKT),2),
                             iter = rep)
                    )
}

rIBKTBNTAllMeans <- rIBKTBNTAll %>%
  group_by(rowname, var) %>%
  summarise(rIMean = mean(value),
            rISd = sd(value),
            rIUpperCI = quantile(value, 0.95))

dodge <- position_dodge(width=0.9)
ggplot(rIBKTBNTAllMeans, aes(factor(var),rIMean, fill = factor(rowname, levels = c('rBKT','rBNT'), ordered = TRUE))) +
  geom_errorbar(aes(ymin = rIMean - 0.001, ymax = rIUpperCI), width = 0.25, position=dodge) +
  geom_col(position=dodge) +
  scale_x_discrete('Parameter',
                   labels = c('1' = 'Temperature','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")) +
  scale_fill_grey() +
  scale_y_continuous('Relative proportion of variance explained') +
  theme_publication() +
  geom_hline(yintercept = 0) +
  theme( legend.position = 'none')

ggsave(paste0('figures/3Spp_WB/relImpOccMeanModAll_GR_N',simData$nReps,'.png'), width = 9, height = 7, units='in')

# for MS
rIBKTBNTAllMeans
#    rowname   var  rIMean    rISd rIUpperCI
#    <chr>   <int>   <dbl>   <dbl>     <dbl>
#  1 rBKT        1 0.132   0.0251     0.170 
#  2 rBKT        2 0.670   0.0369     0.731 
#  3 rBKT        3 0.0116  0.0118     0.0330
#  4 rBKT        4 0.0346  0.0256     0.0703
#  5 rBKT        5 0.0951  0.0284     0.140 
#  6 rBKT        6 0.0441  0.0263     0.0919
#  7 rBKT        7 0.0121  0.0155     0.0410
#  8 rBNT        1 0.106   0.0212     0.141 
#  9 rBNT        2 0.568   0.0364     0.623 
# 10 rBNT        3 0.0885  0.0223     0.124 
# 11 rBNT        4 0.00518 0.00556    0.0163
# 12 rBNT        5 0.122   0.0285     0.169 
# 13 rBNT        6 0.0314  0.0176     0.0624
# 14 rBNT        7 0.0782  0.0270     0.122

rIBKTBNTAll %>%
  mutate(isEnvVar = ifelse(varText %in% c('temp','flow','temp:flow'), 1,0)) %>%
  group_by(rowname,isEnvVar) %>%
  summarize(meanMean = mean(value),
            sumMean = sum(value),
            sdMean = sd(value))

#   rowname isEnvVar meanMean sumMean sdMean
#   <chr>      <dbl>    <dbl>   <dbl>  <dbl>
# 1 rBKT           0   0.0383    15.3 0.0402
# 2 rBKT           1   0.282     84.7 0.279 
# 3 rBNT           0   0.0736    29.4 0.0484
# 4 rBNT           1   0.235     70.6 0.239

#bkt
.282/(.282+.038) # 0.88125
#bnt
.235/(.235+.074) # 0.7605178
.881/.76 # 1.159211
```

```{r RI across yoy/adults - use this for paper GROWTH RATE}

# relative proportion stats with occ in the lm 
grStats3Spp[['bkt']][[1]]$season <- as.factor(c(3,4,1,2,3,4,1,2,3,4))
grStats3Spp[['bkt']][[1]]$temp2 <- grStats3Spp[['bkt']][[1]]$temp^2
grStats3Spp[['bkt']][[1]]$flow2 <- grStats3Spp[['bkt']][[1]]$flow^2

ageForRI <-  'yoy' #'yoy' or 'adult'
occs <- if (ageForRI == 'yoy') 1:3 else 4:10
indepVarVals <- c(-1.5,0,1.5) # make sure symetrical around 0
indepVars <- c('temp', 'flow', 'temp:flow', 'bkt', 'bnt', 'ats', 'bkt:bnt')
rIBKTBNT <- NULL
# loop over occs
for(rep in 1:simData$nReps){

  print(rep)

  datOccBKT <- grStats3Spp[['bkt']][[1]] %>% filter(iter == rep, occ %in% occs, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
  lmAllBKT <- lm( grMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBKT) 
  relImpAllBKT <- calc.relimp(lmAllBKT, type = c("lmg"), rela = TRUE)

  datOccBNT <- grStats3Spp[['bnt']][[1]] %>% filter(iter == rep, occ %in% occs, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
  lmAllBNT <- lm( grMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBNT)
  relImpAllBNT <- calc.relimp(lmAllBNT, type = c("lmg"), rela = TRUE)

  lmgBKT <- relImpAllBKT$lmg[(names(relImpAllBKT$lmg) %in% indepVars)]
  rBKT <- lmgBKT / sum(lmgBKT)

  lmgBNT <- relImpAllBNT$lmg[(names(relImpAllBNT$lmg) %in% indepVars)]
  rBNT <- lmgBNT / sum(lmgBNT)

  rIBKTBNT <- rbind(rIBKTBNT, data.frame(cbind(rBKT,rBNT)) %>% 
                      rownames_to_column() %>% 
                      gather(key = rowname) %>% 
                      mutate(var = rep(1:length(indepVars),2),
                             varText = rep(names(rBKT),2),
                             iter = rep)
                    )
}

rIBKTBNTMeans <- rIBKTBNT %>%
  group_by(rowname, var) %>%
  summarise(rIMean = mean(value),
            rISd = sd(value),
            rIUpperCI = quantile(value, 0.95))

dodge <- position_dodge(width=0.9)
ggplot(rIBKTBNTMeans, aes(factor(var),rIMean, fill = factor(rowname, levels = c('rBKT','rBNT'), ordered = TRUE))) +
  geom_errorbar(aes(ymin = rIMean - 0.001, ymax = rIUpperCI), width = 0.25, position=dodge) +
  geom_col(position=dodge) +
  scale_x_discrete('Parameter',
                   labels = c('1' = 'Temperature','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")) +
  scale_fill_grey() +
  scale_y_continuous('Relative proportion of variance explained') +
  theme_publication() +
  geom_hline(yintercept = 0) +
  theme( legend.position = 'none')

ggsave(paste0('figures/3Spp_WB/relImpOccMeanModWOcc_GR',simData$nReps,'_',ageForRI,'.png'), width = 9, height = 7, units='in')

#    rowname   var  rIMean    rISd rIUpperCI
#    <chr>   <int>   <dbl>   <dbl>     <dbl>
#  1 rBKT        1 0.516   0.0639     0.619 
#  2 rBKT        2 0.123   0.0332     0.175 
#  3 rBKT        3 0.0209  0.0222     0.0568
#  4 rBKT        4 0.0579  0.0307     0.100 
#  5 rBKT        5 0.193   0.0487     0.273 
#  6 rBKT        6 0.0539  0.0374     0.126 
#  7 rBKT        7 0.0362  0.0303     0.0889
#  8 rBNT        1 0.154   0.0566     0.253 
#  9 rBNT        2 0.0106  0.00999    0.0281
# 10 rBNT        3 0.140   0.0479     0.208 
# 11 rBNT        4 0.00421 0.00467    0.0135
# 12 rBNT        5 0.463   0.0780     0.598 
# 13 rBNT        6 0.173   0.0676     0.294 
# 14 rBNT        7 0.0547  0.0340     0.121

# get isEnvVar sens for yoy and adults - need to rerun the above code for each age
rIBKTBNT %>%
  mutate(isEnvVar = ifelse(varText %in% c('temp','flow','temp:flow'), 1,0)) %>%
  group_by(rowname,isEnvVar) %>%
  summarize(meanMean = mean(value),
            sumMean = sum(value),
            sdMean = sd(value)) %>%
  mutate(age = ageForRI)

#   rowname isEnvVar meanMean sumMean sdMean age  
#   <chr>      <dbl>    <dbl>   <dbl>  <dbl> <chr>
# 1 rBKT           0   0.0769    30.8 0.0763 yoy  
# 2 rBKT           1   0.231     69.2 0.209  yoy  
# 3 rBNT           0   0.165     66.2 0.185  yoy  
# 4 rBNT           1   0.113     33.8 0.0888 yoy

#   rowname isEnvVar meanMean sumMean sdMean age  
#   <chr>      <dbl>    <dbl>   <dbl>  <dbl> <chr>
# 1 rBKT           0   0.0287    11.5 0.0272 adult
# 2 rBKT           1   0.295     88.5 0.377  adult
# 3 rBNT           0   0.0411    16.4 0.0285 adult
# 4 rBNT           1   0.279     83.6 0.343  adult

# ratio of env/(env+spp) means - yoy
.231/(.231 + .0769) # bkt 0.7502436
.165/(.165 + .113) # bnt 0.5935252

# ratio of env/(env+spp) means - adult
.295/(.295 + .0287) # bkt 0.9113377
.279/(.279 + .0411) # bnt 0.8716026

.75/.593 # 1.26
.911/.872 # 1.04
##############################################################
############################################################

# projOut <- list(stats3Spp, lm, relImp, anov, elapsed)
load(paste0('D:/projects/linkedModels/data/out/projectionOut_N',simData$nReps,'.RData'))

#spp <- c('bkt','bnt','ats')
# scenarios
# Plot subsets of the simulations
#means for each iter
stats1 <- rbind(data.frame(grStats3Spp[['bkt']][[1]]), 
                data.frame(grStats3Spp[['bnt']][[1]]), 
                data.frame(grStats3Spp[['ats']][[1]])) %>%
          mutate(spp = rep(spp, each = nrow(grStats3Spp[['bkt']][[1]])))

#means for each occ
stats2 <- rbind(data.frame(grStats3Spp[['bkt']][[2]]), 
                data.frame(grStats3Spp[['bnt']][[2]]), 
                data.frame(grStats3Spp[['ats']][[2]])) %>%
          mutate(spp = rep(spp, each = nrow(grStats3Spp[['bkt']][[2]])))

ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,grMeanMean, group = spp)) +
  geom_point(aes(color = spp) ) +
  geom_line( aes(color = spp) ) +
  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 1.5, flow == 0, bkt == 0, bnt == 0, ats == 0) ) +
  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 1.5, flow == 0, bkt == 0, bnt == 0, ats == 0), linetype = 2 )

ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,grMeanMean, group = spp)) +
  geom_point(aes(color = spp) ) +
  geom_line( aes(color = spp) ) +
  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 1.5, bkt == 0, bnt == 0, ats == 0) ) +
  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 1.5, bkt == 0, bnt == 0, ats == 0), linetype = 2 )

ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,grMeanMean, group = spp)) +
  geom_point(aes(color = spp) ) +
  geom_line( aes(color = spp) ) +
  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 0, bkt == 1.5, bnt == 0, ats == 0) ) +
  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 0, bkt == 1.5, bnt == 0, ats == 0), linetype = 2 )

# trajectories across indep var values for different combinations. Trying to relate to relImp results
#temp
p1 <- ggplot( stats2 %>% filter(flow == 0, bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(temp))) +
  geom_line(aes(color = factor(temp))) +
  geom_point(aes(color = factor(temp))) +
  facet_wrap(~spp)

#flow
p2 <- ggplot( stats2 %>% filter(temp == 0, bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(flow))) +
  geom_line(aes(color = factor(flow))) +
  geom_point(aes(color = factor(flow))) +
  facet_wrap(~spp)

#bkt
p3 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(bkt))) +
  geom_line(aes(color = factor(bkt))) +
  geom_point(aes(color = factor(bkt))) +
  facet_wrap(~spp)

#bnt
p4 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(bnt))) +
  geom_line(aes(color = factor(bnt))) +
  geom_point(aes(color = factor(bnt))) +
  facet_wrap(~spp)

#ats
p5 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(ats))) +
  geom_line(aes(color = factor(ats))) +
  geom_point(aes(color = factor(ats))) +
  facet_wrap(~spp)

multiplot(p1, p2, p3, p4, p5, cols=2) # multiplot function is below

#temp*flow
ggplot( stats2 %>% filter(bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(flow))) +
  geom_line(aes(color = factor(flow))) +
  geom_point(aes(color = factor(flow))) +
  facet_grid(temp~spp)

#bkt*bnt
ggplot( stats2 %>% filter(temp == 0, flow == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(bnt))) +
  geom_line(aes(color = factor(bnt))) +
  geom_point(aes(color = factor(bnt))) +
  facet_grid(bkt~spp)

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

# size=100
# gr=10
# sd=1
# mean(rnorm(10000,size+gr,sd))
# mean(size+rnorm(10000,gr,sd))

```

```{r RI by occ - use for paper GROWTH}

# make figures of relative importance for each sampling occasion
rIAll <- cAll <- NULL

for(ageGroupLoop in 0:1){
  for(seasonLoop in 1:4){
#for( occSelected in 1:(simData$nOcc - 1)){
  if(ageGroupLoop == 0 & seasonLoop == 2) next

  lmAll <- ggRIAll <- relImpAll <- coefsAll <- list()

  for(spp1 in spp){
    print(c(ageGroupLoop, seasonLoop, spp1))

    datAgeGroupSeason <- grStats3Spp[[spp1]][[1]] %>%
         mutate(ageGroup = ifelse(occ %in% 1:3, 0, 1),
         season = rep(c(3,4,1,2,3,4,1,2,3,4), nrow(grStats3Spp[[spp1]][[1]])/simData$nOcc))

      lmAll[[spp1]] <- lm( grMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt, data = datAgeGroupSeason %>% filter(ageGroup == ageGroupLoop, season == seasonLoop, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals))

      relImpAll[[spp1]] <- calc.relimp(lmAll[[spp1]], type = c("lmg"), rela = TRUE)

      coefsAll[[spp1]] <- lmAll[[spp1]]$coefficients
  }

  # make bar chart of relative importance of each variable for each spp
  rI <- cbind(data.frame(bktRI = relImpAll[['bkt']]$lmg), 
              data.frame(bntRI = relImpAll[['bnt']]$lmg)) %>%
        rownames_to_column() %>% 
        gather(key=rowname) %>% 
        mutate(var = rep(1:length(relImpAll[['bkt']]$lmg),length(spp)),
               ageGroup = ageGroupLoop,
               season = seasonLoop
               )
  rI$varText <- rep(names(relImpAll[['bkt']]$lmg),2)

  rIAll <- rbind(rIAll,rI)

  #rIAll$age <- ifelse(rIAll$occ %in% 1:2, 0, ifelse(rIAll$occ %in% 3:6, 1, 2))

  }  
}

######################################################
# relImp over occasions for each spp and parameter
labels <- c('1' = 'Temp','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")

# ggplot(rIAll %>% filter(rowname %in% c('bktRI','bntRI')), aes((occ), value, color = factor(rowname))) +
#   geom_point() +
#   geom_line() +
#   geom_vline(xintercept = 2.5) +
#   scale_x_continuous('Sampling occasion',
#                     breaks = 1:9,
#                     labels = c('Autumn','Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn')
#                     ) +
#   scale_y_continuous('Relative proportion of variance explained') +
#   theme_publication() +
#   theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10)) +
#   facet_wrap(~ var, labeller=labeller(var = labels))
# 
# ggsave(paste0('figures/3Spp_WB/relImpOccAll_GR.png'), width = 11, height = 9, units='in')

rIAllMinus <- rIAll %>% mutate(value2 = ifelse(rowname == 'bktRI',value * -1, value))

labelsSeason <- c('1'='Autumn','2'='Winter','3'='Spring','4'='Summer','5'='Autumn','6'='Winter','7'='Spring','8'='Summer','9'='Autumn')
labelsSeason2 <- c('1'='Spring','2'='Summer','3'='Autumn','4'='Winter')
labelAge <- c('0' = 'Growth year 0/1', '1' = 'Growth Year 1+')

### get prop env
summ1 <- rIAll %>%
  mutate(isEnvVar = ifelse(varText %in% c('temp','flow','temp:flow'), 1,0)) %>%
  group_by(rowname,isEnvVar,ageGroup, season) %>%
  summarize(mean=mean(value)) %>%
  arrange(ageGroup,season,rowname,isEnvVar)

summ2 <- summ1 %>%
  group_by(rowname,ageGroup,season) %>%
  summarize(sum=sum(mean))

summ1a <- left_join(summ1,summ2) %>%
  mutate(prop = mean/sum,
         xAxisLoc = ifelse(rowname == 'bktRI',-0.85, 0.85)) %>%
  filter(isEnvVar == 1) 
###

####
# add in betas to rIAllMinus
load("D:/projects/linkedModels/data/out/allBetas.RData") #loads df 'q'
qFig1 <- q %>% filter(riverGG == 'WB', species %in% c('bkt','bnt'))
qFig1$season <- as.numeric(as.character(qFig1$season))
qFig1$isYOY <- as.numeric(as.character(qFig1$isYOY))
qFig <- qFig1 %>%
  mutate( varText = tolower(betaGG)) %>%
  mutate( varText = ifelse(varText == 'temp*flow','temp:flow', varText)) %>%
  mutate( varText = ifelse(varText == 'bnt*bkt','bkt:bnt', varText)) %>%
  select(isYOY,season,species,median,varText)

# prepare rIAllMinus
rIAllMinusQ <- rIAllMinus %>%
  mutate(species = str_sub(rowname,1,3),
         isYOY = ageGroup) %>%
  left_join(qFig) %>%
  mutate(median01 = ifelse(median > 0, 1, 0),
         medianBreaks = cut(median, breaks = c(-1, -0.025, 0 , 0.025, 1)))

#

ggplot(rIAllMinus, aes(var, y = value2, fill = rowname)) +
  geom_col() +
  coord_flip() +
  scale_x_reverse('Variable', 
                     breaks = 1:7,
                     labels = c('Temp', 'Flow', 'Bkt', 'Bnt','Ats','Temp*Flow',  "Bkt*Bnt")) +
  scale_y_continuous('Relative proportion of variance explained',
                     breaks = c(-1.0, -0.5, 0.0, 0.5),
                     labels = c('1.0', '0.5', '0.0', '0.5')) +
  geom_text(data = summ1a, aes(y=xAxisLoc, x=7, label=paste(round(prop,2)), color=rowname),size = 5) +
  scale_color_grey(start = 0.15, end = 0.75) +
  theme_publication() +
  scale_fill_grey(start = 0.15, end = 0.75) +
  facet_grid(season ~ ageGroup, labeller=labeller(season = labelsSeason2, ageGroup = labelAge))

ggsave(paste0('figures/3Spp_WB/relImpOccAll_GR_HorizBars.png'), width = 11, height = 9, units='in')

# bars colered by beta
ggplot(rIAllMinusQ, aes(var, y = value2)) +
  geom_col(aes(fill = median)) +
  coord_flip() +
  scale_x_reverse('Variable', 
                     breaks = 1:7,
                     labels = c('Temp', 'Flow', 'Bkt', 'Bnt','Ats','Temp*Flow',  "Bkt*Bnt")) +
  scale_y_continuous('Relative proportion of variance explained',
                     breaks = c(-0.9, -0.25, -0.5, 0.0, 0.25, 0.5, 0.9),
                     labels = c('0.9', '0.25', '0.5\n\nBrook Trout\n', '0.0', '0.25', '0.5\n\nBrown Trout\n', '0.9')) +
  geom_text(data = summ1a, aes(y=xAxisLoc, x=7, label=paste(round(prop,2))),size = 5) +
  geom_hline(yintercept = 0) +
  #scale_color_grey(start = 0.15, end = 0.75) +
  theme_publication() +
 #  scale_fill_continuous() +
  scale_fill_distiller(name="Parameter\nestimate", palette = "RdBu", direction = 1) +
  theme(legend.position="right",             
        #legend.key = element_rect(colour = NA),
        legend.direction = "vertical",
        legend.key.size= unit(0.25, "in"),
        #    legend.margin = unit(0, "cm"),
        legend.title = element_text(face="plain")#element_blank()
        ) +
  facet_grid(season ~ ageGroup, labeller=labeller(season = labelsSeason2, ageGroup = labelAge))

ggsave(paste0('figures/3Spp_WB/relImpOccAll_GR_HorizBarsColor.png'), width = 11, height = 9, units='in')

#######################
# coefs
labelsC <- c('1' = 'intercept', '2' = 'Temp','3' = 'Flow','4' = 'Bkt','5' = 'Bnt','6' = 'Ats','7' = 'Temp*Flow', '8' = "Bkt*Bnt")

ggplot(cAll %>% filter(var %in% 2:8, rowname %in% c('bktC','bntC')), aes((occ), value, color = factor(rowname))) +
  geom_point(size=3) +
  geom_line() +
  geom_vline(xintercept = 3.5) +
  scale_x_continuous('Sampling occasion',
                    breaks = 2:10,
                    labels = c('Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn','Winter')
                    ) +
  scale_y_continuous('Coefficients') +
  theme_publication() +
  theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10)) +
  facet_wrap(~ var, labeller=labeller(var = labelsC))#, scales = 'free')

ggsave(paste0('figures/3Spp_WB/coefsOccAll.png'), width = 11, height = 9, units='in')

# intercept only
cAll2 <- rbind(cAll,data.frame(rowname=c('bktC','bntC','atsC'),value=simData$meanInit,var=1,occ=1,season=3))
ggplot(cAll2 %>% filter(var == 1), aes((occ), value, color = factor(rowname))) +
  geom_point(size=4) +
  geom_line() +
  geom_vline(xintercept = 3.5) +
  scale_x_continuous('Sampling occasion',
                    breaks = 1:10,
                    labels = c('Autumn','Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn','Winter')
                    ) +
  scale_y_continuous('Intercept') +
  theme_publication() +
  theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10))

ggsave(paste0('figures/3Spp_WB/coefsInterceptOccAll.png'), width = 11, height = 9, units='in')

#####################################
# no need to start sims with diff sizes by species
yoyMeans <- 
cd %>% filter( riverOrdered == 'west brook', ageInSamples == 1 ) %>%
  group_by(species) %>%
  summarize(meanLen = mean(observedLength,na.rm = TRUE),
           sdLen = sd(observedLength,na.rm = TRUE))

#   species meanLen sdLen
#   <chr>     <dbl> <dbl>
# 1 ats        69.0  5.86
# 2 bkt        72.7  9.94
# 3 bnt        72.9  8.73
#####################################

```

```{r RI by occ - not used for paper SIZE}

# make figures of relative importance for each sampling occasion
rIAll <- cAll <- NULL
for( occSelected in 2:simData$nOcc){
  print(occSelected)
  lmAll <- ggRIAll <- relImpAll <- coefsAll <- list()

  for(spp1 in spp){

      lmAll[[spp1]] <- lm( lenMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt, data = stats3Spp[[spp1]][[1]] %>% filter(occ == occSelected))

      relImpAll[[spp1]] <- calc.relimp(lmAll[[spp1]], type = c("lmg"), rela = TRUE)

      coefsAll[[spp1]] <- lmAll[[spp1]]$coefficients
  }

  # make bar chart of relative importance of each variable for each spp
  rI <- cbind(data.frame(bktRI = relImpAll[['bkt']]$lmg), 
              data.frame(bntRI = relImpAll[['bnt']]$lmg), 
              data.frame(atsRI = relImpAll[['ats']]$lmg)) %>%
        rownames_to_column() %>% 
        gather(key=rowname) %>% 
        mutate(var = rep(1:length(relImpAll[['bkt']]$lmg),length(spp)),
               occ = occSelected,
               season = simData$seasons[occSelected])

  rIAll <- rbind(rIAll,rI)

  ggplot(rI, aes(factor(var),value, fill = factor(rowname, levels = c('bktRI','bntRI','atsRI'), ordered = TRUE))) +
    geom_col(position='dodge') +
    scale_x_discrete('Parameter',
                     labels = c('1' = 'Temp','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")) +
    scale_fill_grey() +
    scale_y_continuous('Relative proportion of variance explained') +
    ggtitle(paste0("occ = ",occSelected,"  Season = ",simData$seasons[occSelected])) +
    theme_publication() +
    theme( legend.position = 'none')

##############################  ggsave(paste0('figures/3Spp_WB/relImpOcc',occSelected,'.png'), width = 9, height = 7, units='in')

  # coefficients
  coefs <- cbind(data.frame(bktC = coefsAll[['bkt']]), 
                 data.frame(bntC = coefsAll[['bnt']]), 
                 data.frame(atsC = coefsAll[['ats']])) %>%
        rownames_to_column() %>% 
        gather(key=rowname) %>% 
        mutate(var = rep(1:length(coefsAll[['bkt']]),length(spp)),
               occ = occSelected,
               season = simData$seasons[occSelected])

  cAll <- rbind(cAll,coefs)
}

######################################################
# relImp over occasions for each spp and parameter
labels <- c('1' = 'Temp','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")

# leaving out bkt*bnt, too small
ggplot(rIAll %>% filter(rowname %in% c('bktRI','bntRI')), aes((occ), value, color = factor(rowname))) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 3.5) +
  scale_x_continuous('Sampling occasion',
                    breaks = 2:10,
                    labels = c('Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn','Winter')
                    ) +
  scale_y_continuous('Relative proportion of variance explained') +
  theme_publication() +
  theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10)) +
  facet_wrap(~ var, labeller=labeller(var = labels))

ggsave(paste0('figures/3Spp_WB/relImpOccAll.png'), width = 11, height = 9, units='in')

rIMeans <- rIAll %>%
  group_by(rowname,var) %>%
  summarize( mean = mean(value),
             sd = sd(value),
             stderr = sd(value)/sqrt(length(value))) %>%
  mutate( up = mean + stderr )

dodge <- position_dodge(width=0.9)
ggplot(rIMeans %>% filter(rowname %in% c('bktRI','bntRI')), aes(factor(var),mean, fill = factor(rowname, levels = c('bktRI','bntRI','atsRI'), ordered = TRUE))) +
  geom_errorbar(aes(ymin = mean-0.01, ymax = up), width = 0.25, position=dodge) +
  geom_col(position=dodge) +
  scale_x_discrete('Parameter',
                   labels = c('1' = 'Temperature','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")) +
  scale_fill_grey() +
  scale_y_continuous('Relative proportion of variance explained') +
  theme_publication() +
  theme( legend.position = 'none')

ggsave(paste0('figures/3Spp_WB/relImpOccMean.png'), width = 9, height = 7, units='in')

# mean rel Imp for env vs spp variables
rIMeans %>%
  mutate(isEnvVar = ifelse(var %in% c(1,2,6), 1,0)) %>%
  group_by(rowname,isEnvVar) %>%
  summarize(meanMean = mean(mean),
            sumMean = sum(mean),
            sdMean = sd(mean))

# ratio of env/(env+spp) means
.247/(.246+.0648) # bkt
.122/(.122+.159) # bnt

# ratio of env/(env+spp) sums - donT need to do ratio, just use sum for env
.635/.259 # den-dep bnt/bkt
.741/.365 # env bkt/bnt

#######################
# coefs
labelsC <- c('1' = 'intercept', '2' = 'Temp','3' = 'Flow','4' = 'Bkt','5' = 'Bnt','6' = 'Ats','7' = 'Temp*Flow', '8' = "Bkt*Bnt")

ggplot(cAll %>% filter(var %in% 2:8, rowname %in% c('bktC','bntC')), aes((occ), value, color = factor(rowname))) +
  geom_point(size=3) +
  geom_line() +
  geom_vline(xintercept = 3.5) +
  scale_x_continuous('Sampling occasion',
                    breaks = 2:10,
                    labels = c('Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn','Winter')
                    ) +
  scale_y_continuous('Coefficients') +
  theme_publication() +
  theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10)) +
  facet_wrap(~ var, labeller=labeller(var = labelsC))#, scales = 'free')

ggsave(paste0('figures/3Spp_WB/coefsOccAll.png'), width = 11, height = 9, units='in')

# intercept only
cAll2 <- rbind(cAll,data.frame(rowname=c('bktC','bntC','atsC'),value=simData$meanInit,var=1,occ=1,season=3))
ggplot(cAll2 %>% filter(var == 1), aes((occ), value, color = factor(rowname))) +
  geom_point(size=4) +
  geom_line() +
  geom_vline(xintercept = 3.5) +
  scale_x_continuous('Sampling occasion',
                    breaks = 1:10,
                    labels = c('Autumn','Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn','Winter')
                    ) +
  scale_y_continuous('Intercept') +
  theme_publication() +
  theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10))

ggsave(paste0('figures/3Spp_WB/coefsInterceptOccAll.png'), width = 11, height = 9, units='in')

#####################################
# no need to start sims with diff sizes by species
yoyMeans <- 
cd %>% filter( riverOrdered == 'west brook', ageInSamples == 1 ) %>%
  group_by(species) %>%
  summarize(meanLen = mean(observedLength,na.rm = TRUE),
           sdLen = sd(observedLength,na.rm = TRUE))

#   species meanLen sdLen
#   <chr>     <dbl> <dbl>
# 1 ats        69.0  5.86
# 2 bkt        72.7  9.94
# 3 bnt        72.9  8.73
#####################################

```

```{r tables}

round(mcmc$summary$all.chains[startsWith(attributes(mcmc$summary$all.chains)$dimnames[[1]], "grBeta["),],2)

```

```{r RI across occ - use this for paper SIZE}

# relative proportion stats with occ in the lm 
stats3Spp[['bkt']][[1]]$season <- as.factor(c(3,4,1,2,3,4,1,2,3,4))
stats3Spp[['bkt']][[1]]$temp2 <- stats3Spp[['bkt']][[1]]$temp^2
stats3Spp[['bkt']][[1]]$flow2 <- stats3Spp[['bkt']][[1]]$flow^2

ageForRI <-  'adult' #'yoy' or 'adult'
occs <- if (ageForRI == 'yoy') 1:3 else 4:10
indepVarVals <- c(-1.5,0,1.5) # make sure symetrical around 0
rIBKTBNT <- NULL
# loop over occs
for(rep in 1:simData$nReps){

  print(rep)

  datOccBKT <- stats3Spp[['bkt']][[1]] %>% filter(iter == rep, occ %in% occs, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
  lmAllBKT <- lm( lenMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBKT) 
  relImpAllBKT <- calc.relimp(lmAllBKT, type = c("lmg"), rela = TRUE)

  datOccBNT <- stats3Spp[['bnt']][[1]] %>% filter(iter == rep, occ %in% occs, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
  lmAllBNT <- lm( lenMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBNT)
  relImpAllBNT <- calc.relimp(lmAllBNT, type = c("lmg"), rela = TRUE)

  rBKT <- relImpAllBKT$lmg[2:8] / sum(relImpAllBKT$lmg[2:8]) 
  rBNT <- relImpAllBNT$lmg[2:8] / sum(relImpAllBNT$lmg[2:8])

  rIBKTBNT <- rbind(rIBKTBNT, data.frame(cbind(rBKT,rBNT)) %>% 
                      rownames_to_column() %>% 
                      gather(key = rowname) %>% 
                      mutate(var = rep(1:7,2),
                             varText = rep(names(rBKT),2),
                             iter = rep)
                    )

}

rIBKTBNTMeans <- rIBKTBNT %>%
  group_by(rowname, var) %>%
  summarise(rIMean = mean(value),
            rISd = sd(value),
            rIUpperCI = quantile(value, 0.95))

dodge <- position_dodge(width=0.9)
ggplot(rIBKTBNTMeans, aes(factor(var),rIMean, fill = factor(rowname, levels = c('rBKT','rBNT'), ordered = TRUE))) +
  geom_errorbar(aes(ymin = rIMean - 0.001, ymax = rIUpperCI), width = 0.25, position=dodge) +
  geom_col(position=dodge) +
  scale_x_discrete('Parameter',
                   labels = c('1' = 'Temperature','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")) +
  scale_fill_grey() +
  scale_y_continuous('Relative proportion of variance explained') +
  theme_publication() +
  geom_hline(yintercept = 0) +
  theme( legend.position = 'none')

ggsave(paste0('figures/3Spp_WB/relImpOccMeanModWOcc_Size',simData$nReps,'_',ageForRI,'.png'), width = 9, height = 7, units='in')

# uses 42 G - does not run
#lmAll <- lm( lenMean ~ (temp + flow +  temp^2 + flow^2 + bkt + bnt + ats )*occ + temp * flow + bnt * bkt, data = stats3Spp[['bkt']][[1]] %>%               filter(iter %in% 1) )

rIBKTBNT %>%
  mutate(isEnvVar = ifelse(var %in% c(1,2,6), 1,0)) %>%
  group_by(rowname,isEnvVar) %>%
  summarize(meanMean = mean(value),
            sumMean = sum(value),
            sdMean = sd(value))
# 25 reps
#   rowname isEnvVar meanMean sumMean sdMean
#   <chr>      <dbl>    <dbl>   <dbl>  <dbl>
# 1 rBKT           0   0.0664   0.265  0.105
# 2 rBKT           1   0.245    0.735  0.249
# 3 rBNT           0   0.147    0.588  0.104
# 4 rBNT           1   0.137    0.412  0.158

# 100 reps
#   rowname isEnvVar meanMean sumMean sdMean
#   <chr>      <dbl>    <dbl>   <dbl>  <dbl>
# 1 rBKT           0   0.0676    27.0 0.0914
# 2 rBKT           1   0.243     73.0 0.200 
# 3 rBNT           0   0.146     58.2 0.0922
# 4 rBNT           1   0.139     41.8 0.131

# ratio of env/(env+spp) means
.243/(.243 + .0676) # bkt 0.7823567
.139/(.139 + .146) # bnt 0.4877193

.782/.488 # 1.60
##############################################################
############################################################

# projOut <- list(stats3Spp, lm, relImp, anov, elapsed)
load(paste0('D:/projects/linkedModels/data/out/projectionOut_N',simData$nReps,'.RData'))

spp <- c('bkt','bnt','ats')
# scenarios
# Plot subsets of the simulations
#means for each iter
stats1 <- rbind(data.frame(stats3Spp[['bkt']][[1]]), 
                data.frame(stats3Spp[['bnt']][[1]]), 
                data.frame(stats3Spp[['ats']][[1]])) %>%
          mutate(spp = rep(spp, each = nrow(stats3Spp[['bkt']][[1]])))

#means for each occ
stats2 <- rbind(data.frame(stats3Spp[['bkt']][[2]]), 
                data.frame(stats3Spp[['bnt']][[2]]), 
                data.frame(stats3Spp[['ats']][[2]])) %>%
          mutate(spp = rep(spp, each = nrow(stats3Spp[['bkt']][[2]])))

ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,lenMeanMean, group = spp)) +
  geom_point(aes(color = spp) ) +
  geom_line( aes(color = spp) ) +
  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 1.5, flow == 0, bkt == 0, bnt == 0, ats == 0) ) +
  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 1.5, flow == 0, bkt == 0, bnt == 0, ats == 0), linetype = 2 )

ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,lenMeanMean, group = spp)) +
  geom_point(aes(color = spp) ) +
  geom_line( aes(color = spp) ) +
  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 1.5, bkt == 0, bnt == 0, ats == 0) ) +
  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 1.5, bkt == 0, bnt == 0, ats == 0), linetype = 2 )

ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,lenMeanMean, group = spp)) +
  geom_point(aes(color = spp) ) +
  geom_line( aes(color = spp) ) +
  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 0, bkt == 1.5, bnt == 0, ats == 0) ) +
  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 0, bkt == 1.5, bnt == 0, ats == 0), linetype = 2 )

# trajectories across indep var values for different combinations. Trying to relate to relImp results
#temp
p1 <- ggplot( stats2 %>% filter(flow == 0, bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(temp))) +
  geom_line(aes(color = factor(temp))) +
  geom_point(aes(color = factor(temp))) +
  facet_wrap(~spp)

#flow
p2 <- ggplot( stats2 %>% filter(temp == 0, bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(flow))) +
  geom_line(aes(color = factor(flow))) +
  geom_point(aes(color = factor(flow))) +
  facet_wrap(~spp)

#bkt
p3 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(bkt))) +
  geom_line(aes(color = factor(bkt))) +
  geom_point(aes(color = factor(bkt))) +
  facet_wrap(~spp)

#bnt
p4 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(bnt))) +
  geom_line(aes(color = factor(bnt))) +
  geom_point(aes(color = factor(bnt))) +
  facet_wrap(~spp)

#ats
p5 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(ats))) +
  geom_line(aes(color = factor(ats))) +
  geom_point(aes(color = factor(ats))) +
  facet_wrap(~spp)

multiplot(p1, p2, p3, p4, p5, cols=2) # multiplot function is below

#temp*flow
ggplot( stats2 %>% filter(bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(flow))) +
  geom_line(aes(color = factor(flow))) +
  geom_point(aes(color = factor(flow))) +
  facet_grid(temp~spp)

#bkt*bnt
ggplot( stats2 %>% filter(temp == 0, flow == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(bnt))) +
  geom_line(aes(color = factor(bnt))) +
  geom_point(aes(color = factor(bnt))) +
  facet_grid(bkt~spp)

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

# size=100
# gr=10
# sd=1
# mean(rnorm(10000,size+gr,sd))
# mean(size+rnorm(10000,gr,sd))

```

```{r overview graphs - Fig Appendix}

pd <- position_dodge(0.65)
d2$speciesGG <- factor(d2$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

# isYOY1
d2 %>% group_by(speciesGG,riverOrderedGG,seasonGG,year) %>%
  filter(isYOYDATA == 1, riverOrderedGG == 'West Brook') %>%
  summarize(meanGR = mean(grLength, na.rm=T),
            sdGR = sd(grLength, na.rm=T)) %>%
ggplot( aes(year,meanGR, color=speciesGG, shape=speciesGG) ) +
  geom_point( position=pd ) +
  geom_errorbar( aes(ymin = meanGR - sdGR, ymax = meanGR + sdGR), width = 0.33, position=pd ) +
  ggtitle( "isYOY = 1" ) +
  labs( x = "Year", y = "Growth rate (mm/d)" ) +
  scale_x_continuous( breaks = 1997:2015 ) +
  theme_publication() +
  theme( panel.grid.minor = element_blank(),
         axis.text.x  = element_text(angle=90, vjust=0.5, size=8) ) +
  facet_wrap(~seasonGG )

ggsave('figures/3Spp_WB/grByYearYOY1.png')

# isYOY2
d2 %>% group_by(speciesGG,riverOrderedGG,seasonGG,year) %>%
  filter(isYOYDATA == 2, riverOrderedGG == 'West Brook') %>%
  summarize(meanGR = mean(grLength, na.rm=T),
            sdGR = sd(grLength, na.rm=T)) %>%
  ggplot( aes(year,meanGR, color=speciesGG, shape=speciesGG) ) +
  geom_point( position=pd ) +
  geom_errorbar( aes(ymin = meanGR - sdGR, ymax = meanGR + sdGR), width = 0.33, position=pd ) +
  ggtitle( "isYOY = 2" ) +
  labs( x = "Year", y = "Growth rate (mm/d)" ) +
  scale_x_continuous( breaks = 1997:2015 ) +
  theme_publication() +
  theme( panel.grid.minor = element_blank(),
         axis.text.x  = element_text(angle=90, vjust=0.5, size=8) ) +
  facet_wrap(~seasonGG )

ggsave('figures/3Spp_WB/grByYearYOY2.png')

```

```{r growth by movement}
d2$isYOY <- ifelse( d2$ageInSamples <= 3, 1, 2 )
d2 <-  d2 %>%
  group_by(tag) %>%
  mutate( lagRiverN = lead(riverN), 
          riverNMove = paste(riverN,lagRiverN, sep="_")
          #gr = (observedLength - lagObservedLength)/as.integer(detectionDate - lagDetectionDate)
          )

#d2 %>% filter(tag=='00088cbf44') %>% select(tag,riverN,lagRiverN,sampleNumber,observedLength,lagObservedLength)
table(d2$riverNMove)

cdMeansByisYOYriverNMove <- d2 %>%
  #mutate( gr = (observedLength - lagObservedLength)/as.integer(detectionDate - lagDetectionDate) ) %>%
  group_by(species,riverN,lagRiverN,season,isYOY) %>%
  summarize( meanGr = mean( grLength, na.rm = TRUE ),
             sdGr = sd( grLength, na.rm = TRUE ),
             n = n()) %>%
  mutate( meanGrUp = meanGr + sdGr,
          meanGrDn = meanGr - sdGr
          )

# 
ggplot( cdMeansByisYOYriverNMove %>% filter(!(season == 2 & isYOY == 1), species == "bkt", isYOY == 2) , aes(season, meanGr, color = factor(lagRiverN)) ) +
  geom_errorbar(aes(ymin = meanGrDn, ymax = meanGrUp), width = 0.25, position = pd) +
  geom_point( size = 2 ) +
  geom_line()+
  facet_wrap(~riverN)

d2 %>% filter(riverN == 1 & lagRiverN == 4) %>% select(tag,river,sampleNumber)
d2 %>% filter(riverN == 2, is.na(lagRiverN), !is.na(grLength)) %>% select(tag,riverN,sampleNumber,lagRiverN,observedLength,lagObservedLength,gr,grLength)
d2 %>% filter(tag=='00088cc059') %>% select(tag,riverOrdered,riverN,lagRiverN,sampleNumber,observedLength,lagObservedLength,gr,grLength)
```

### 2019-03-05 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
updated figures

@@
   cd %>% mutate( age = year - cohort ) %>%
-    filter( ageInSamples %in% 1:10, age < 3, year < 2016, area %in% c('inside','trib') ) %>%
+    filter( ageInSamples %in% 1:10, age < 3, year < 2016, area %in% c('inside','trib'), enc == 1) %>%
     group_by(species,riverOrdered,ageInSamples,year,season,age) %>%
@@
               meanGR = mean(grLength,na.rm = TRUE),
-              sdGR = sd(grLength,na.rm = TRUE)
+              sdGR = sd(grLength,na.rm = TRUE),
+              n = n()
               )
@@
 # do increases (slopes) increase over age? Yes
-ggplot(lenSlopesTidy %>% filter(), aes((ageInSamples),estimate, color=speciesGG)) +
+ggplot(lenSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8)), aes((ageInSamples),estimate, color=speciesGG)) +
   geom_point( size = 3) +
@@
 ggsave(paste0('figures/meanLengthsOverAge.png'), width = 6, height = 6, units='in')
+
+# just bkt
+ggplot(lenSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8), species == 'bkt'), aes((ageInSamples),estimate, color=riverOrdered, shape = riverOrdered)) +
+  geom_point( size = 3) +
+  geom_smooth(method='lm', se=FALSE) +
+  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
+  labs( x = "Age", y = "Slope of mean body length over years" ) +
+  theme_publication()
+
+ggsave(paste0('figures/meanLengthsOverAgeBKT.png'), width = 6, height = 6, units='in')
 
@@
 ggsave(paste0('figures/meanGRsOverAge.png'), width = 6, height = 6, units='in')
+
+# just bkt
+ggplot(grSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8), species == 'bkt'), aes((ageInSamples),estimate, color=riverOrdered, shape = riverOrdered)) +
+  geom_point( size = 3) +
+  geom_smooth(method='lm', se=FALSE) +
+  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
+  labs( x = "Age", y = "Slope of mean growth rate over years" ) +
+  theme_publication()
+
+ggsave(paste0('figures/meanGRsOverAgeBKT.png'), width = 6, height = 6, units='in')
 

### 2019-03-06 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
riverOrdered and metaPop01 analysis in one chunk

--- 2019-03-05 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
@@
 
-```{r mean lengths}
+```{r mean lengths by gropuingVariable}
+
+#####################################################
+#####################################################
+# SElect THIS for metaPop analysis, or
+groupingVariables <- c('species','metaPop01','ageInSamples', 'year','season','age')
+groupingVariablesNoYear <- c('species','metaPop01','ageInSamples','season','age')
+
+# THIS for 4 river analysis
+#groupingVariables <- c('species','riverOrdered','ageInSamples','year','season','age')
+#groupingVariablesNoYear <- c('species','riverOrdered','ageInSamples','season','age')
+
+#####################################################
+#####################################################
+
+riverGroupVar <- groupingVariables[2]
 
@@
 means <- 
-  cd %>% mutate( age = year - cohort ) %>%
-    filter( ageInSamples %in% 1:10, age < 3, year < 2016, area %in% c('inside','trib') ) %>%
-    group_by(species,riverOrdered,ageInSamples,year,season,age) %>%
+  cd %>% mutate( age = year - cohort,
+                 metaPop01 = ifelse(riverOrdered == 'wb obear', 0 ,1)) %>%
+    filter( ageInSamples %in% 1:10, age < 3, year < 2016, area %in% c('inside','trib'), enc == 1
+            #!(riverOrdered == 'wb mitchell' & ageInSamples > 8) #low sample sizes
+            ) %>% 
+    group_by_at(vars(one_of(groupingVariables))) %>%
     summarize(meanLen = mean(observedLength,na.rm = TRUE),
@@
 lenSlopes <- means %>%
-  group_by( species, riverOrdered, ageInSamples, season, age ) %>%
+  group_by_at(vars(one_of(groupingVariablesNoYear))) %>%
+  #group_by(species,metaPop01,ageInSamples,year,season,age) %>%
   nest()
@@
 
-save(lenSlopesTidy,file = 'D:/projects/westBrook-patterns/data/out/lenSlopesTidy.RData')
+save(lenSlopesTidy,file = paste0('D:/projects/westBrook-patterns/data/out/lenSlopesTidy_',riverGroupVar,'.RData'))
 ##
@@
   scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(seasonGG ~ riverOrdered + age)
-
-ggsave(paste0('figures/meanLengths.png'), width = 11, height = 11, units='in')
+ # facet_grid(seasonGG ~ riverOrdered + age)
+  facet_grid(as.formula(paste("seasonGG ~", riverGroupVar, "+ age")))
+
+ggsave(paste0('figures/meanLengths_',riverGroupVar,'.png'), width = 11, height = 11, units='in')
 
@@
 
+lenSlopesTidy %>%
+  group_by(species, metaPop01) %>%
+  summarise( means = mean(estimate),
+             sd = sd(estimate))
+
 # do increases (slopes) increase over age? Yes
-ggplot(lenSlopesTidy %>% filter(), aes((ageInSamples),estimate, color=speciesGG)) +
+#ggplot(lenSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8)), aes((ageInSamples),estimate, color=speciesGG)) +
+ggplot(lenSlopesTidy %>% filter(ageInSamples %in% 1:9), aes((ageInSamples),estimate, color=speciesGG)) +
   geom_point( size = 3) +
@@
   scale_color_grey(start = 0.15, end = 0.75) +
-  facet_wrap(~riverOrdered)
-
-ggsave(paste0('figures/meanLengthsOverAge.png'), width = 6, height = 6, units='in')
+  facet_wrap(as.formula(paste("~", riverGroupVar)))
+
+ggsave(paste0('figures/meanLengthsOverAge_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
+
+# just bkt
+ggplot(lenSlopesTidy %>% filter(ageInSamples %in% 1:9, species == 'bkt'), aes_(~ageInSamples, ~estimate, group = as.name(riverGroupVar))) +
+  geom_point(aes_(color= as.name(riverGroupVar)), size = 3) +
+  geom_smooth(aes_(color= as.name(riverGroupVar)), method='lm', se=FALSE) +
+  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
+  labs( x = "Age", y = "Slope of mean body length over years" ) +
+  theme_publication()
+
+ggsave(paste0('figures/meanLengthsOverAgeBKT_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
 
@@
 
-save(grSlopesTidy,file = 'D:/projects/westBrook-patterns/data/out/grSlopesTidy.RData')
+grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+save(grSlopesTidy,file = paste0('D:/projects/westBrook-patterns/data/out/grSlopesTidy_',riverGroupVar,'.RData'))
+##
+
 ##
@@
   scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(seasonGG ~ riverOrdered + age)
-
-ggsave(paste0('figures/meanGRs.png'), width = 11, height = 11, units='in')
+  facet_grid(as.formula(paste("seasonGG ~", riverGroupVar, "+ age")))
+
+ggsave(paste0('figures/meanGRs_',riverGroupVar,'.png'), width = 11, height = 11, units='in')
 
@@
 # do increases (slopes) increase over age? Yes
-ggplot(grSlopesTidy %>% filter(), aes((ageInSamples),estimate, color=speciesGG)) +
-  geom_point(size = 3) +
+#ggplot(lenSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8)), aes((ageInSamples),estimate, color=speciesGG)) +
+ggplot(grSlopesTidy %>% filter(ageInSamples %in% 1:9), aes((ageInSamples),estimate, color=speciesGG)) +
+  geom_point( size = 3) +
   geom_smooth(method='lm', se=FALSE) +
   scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
-  labs( x = "Age", y = "Slope of mean growth rate over years" ) +
+  labs( x = "Age", y = "Slope of mean body length over years" ) +
   theme_publication() +
   scale_color_grey(start = 0.15, end = 0.75) +
-  facet_wrap(~riverOrdered)
-
-ggsave(paste0('figures/meanGRsOverAge.png'), width = 6, height = 6, units='in')
+  facet_wrap(as.formula(paste("~", riverGroupVar)))
+
+ggsave(paste0('figures/meanLengthsOverAge_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
+
+# just bkt
+ggplot(grSlopesTidy %>% filter(ageInSamples %in% 1:9, species == 'bkt'), aes_(~ageInSamples, ~estimate, group = as.name(riverGroupVar))) +
+  geom_point(aes_(color= as.name(riverGroupVar)), size = 3) +
+  geom_smooth(aes_(color= as.name(riverGroupVar)), method='lm', se=FALSE) +
+  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
+  labs( x = "Age", y = "Slope of mean body length over years" ) +
+  theme_publication()
+
+ggsave(paste0('figures/meanLengthsOverAgeBKT_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
 

### 2019-03-06 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
fix naming error

@@
 # SElect THIS for metaPop analysis, or
-groupingVariables <- c('species','metaPop01','ageInSamples', 'year','season','age')
-groupingVariablesNoYear <- c('species','metaPop01','ageInSamples','season','age')
+#groupingVariables <- c('species','metaPop01','ageInSamples', 'year','season','age')
+#groupingVariablesNoYear <- c('species','metaPop01','ageInSamples','season','age')
 
 # THIS for 4 river analysis
-#groupingVariables <- c('species','riverOrdered','ageInSamples','year','season','age')
-#groupingVariablesNoYear <- c('species','riverOrdered','ageInSamples','season','age')
+groupingVariables <- c('species','riverOrdered','ageInSamples','year','season','age')
+groupingVariablesNoYear <- c('species','riverOrdered','ageInSamples','season','age')
 
@@
 
+############################################################################################################
 #########################
@@
 
-ggsave(paste0('figures/meanLengthsOverAge_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
+ggsave(paste0('figures/meanGRsOverAge_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
 
@@
 
-ggsave(paste0('figures/meanLengthsOverAgeBKT_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
+ggsave(paste0('figures/meanGRsOverAgeBKT_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
 

### 2019-03-13 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
add gams

@@
 load("D:/projects/linkedModels/data/cd_west_1997.RData")
-```
-
-```{r mean lengths by gropuingVariable}
+cd <- cd %>% mutate(observedWeight = ifelse(observedWeight == -9999.9, NA, observedWeight),
+                    lagObservedWeight = ifelse(lagObservedWeight == -9999.9, NA, lagObservedWeight))
+```
+
+```{r von Bert, well, gam}
+
+vB <- 
+  cd %>% mutate( age = year - cohort,
+                 metaPop01 = ifelse(riverOrdered == 'wb obear', 0 ,1),
+                 ydayCumul = yday + 365 * (year - cohort),
+                 cf = observedWeight/observedLength^3 * 10000 ) %>%
+    filter( ageInSamples %in% 1:9,
+            age < 3, 
+            year < 2016, area %in% c('inside','trib'), enc == 1
+            #!(riverOrdered == 'wb mitchell' & ageInSamples > 8) #low sample sizes
+            )
+
+# number of knots from gam 
+kValue <- 9 # from mod6 below, just for graphing
+
+# all fish
+ggplot(vB %>% filter(cohort %in% 2000:2013), aes(ydayCumul,observedLength)) +
+  geom_point(size = 0.5, alpha = 0.1) +
+ # geom_smooth(span = 0.1, size = 1.5, se = F) +
+  geom_smooth(method='gam', alpha = 0.1, formula=y~s(x, k = kValue), se = F) +
+  theme_publication()
+
+# by species
+ggplot(vB %>% filter(cohort %in% 2000:2013), aes(ydayCumul,observedLength, color = species)) +
+  geom_point(size = 0.5, alpha = 0.1) +
+ # geom_smooth(span = 0.1, size = 1.5, se = F) +
+  geom_smooth(method='gam', alpha = 0.1, formula=y~s(x, k = kValue), se = F) +
+ # facet_grid(~species) +
+  theme_publication()
+
+# by species, by cohort gridded
+# ggplot(vB %>% filter(cohort %in% 2000:2013), aes(ydayCumul,observedLength, group = species, color = (species))) +
+#   geom_point(size = 0.5, alpha = 0.1) +
+#  # geom_smooth(span = 0.1, size = 1.5, se = F) +
+#   geom_smooth(method='gam', alpha=0.1, formula=y~s(x, k = kValue), se = F) +
+#   facet_wrap( ~ cohort) +
+#   theme_publication()
+
+# by species (bkt, bnt), by cohort, by river gridded
+ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats')), aes(ydayCumul,cf, group = species, color = (species))) +
+  geom_point(size = 0.5, alpha = 0.1) +
+ # geom_smooth(span = 0.1, size = 1.5, se = F) +
+  geom_smooth(method='gam', alpha=0.1, formula=y~s(x, k = kValue), se = F) +
+  facet_grid(riverOrdered ~ cohort) +
+  theme_publication()
+
+# gam models
+library(mgcv)
+mod_bnt <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species == 'bnt'))
+summary(mod1_bnt)
+
+mod_bkt <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species == 'bkt'))
+summary(mod_bkt)
+
+mod_ats <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr'), data = vB %>% filter(species == 'ats'))
+summary(mod_ats)
+
+#######################
+
+# 1. Do we need a RE on year in s()? YES
+mod0a <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod1)
+
+mod0 <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod1)
+
+AIC(mod0,mod0a) %>% rownames_to_column() %>% arrange(AIC)
+
+# 2. Best structure for year,species,river
+mod1 <- gam(observedLength ~ species + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod2 <- gam(observedLength ~ river + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+# river expains the most for 1-variable models
+AIC(mod0,mod1,mod2) %>% rownames_to_column() %>% arrange(AIC)
+
+# 2-way models
+mod3 <- gam(observedLength ~ year * species + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod4 <- gam(observedLength ~ year * riverOrdered * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod4)
+
+mod5 <- gam(observedLength ~ species * riverOrdered * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod4)
+
+# mod5 is best
+AIC(mod0,mod1,mod2,mod3,mod4,mod5) %>% rownames_to_column() %>% arrange(AIC)
+
+# 3-way models
+mod6 <- gam(observedLength ~ year * species * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+# https://www.fromthebottomoftheheap.net/2017/12/14/difference-splines-ii/
+# calculates differences from overall spline for each year
+# a lot slower
+vB$yearOrdered <- factor(vB$year, levels = 2000:2013)
+mod6a <- gam(observedLength ~ year * species * riverOrdered + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+# try with river instead of riverOrdered
+mod6b <- gam(observedLength ~ year * species * river + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod6b)
+
+# 3-way is best
+AIC(mod1,mod2,mod3,mod4,mod5,mod6,mod6a) %>% rownames_to_column() %>% arrange(AIC)
+
+# Any additive 3-way better? NO
+mod7 <- gam(observedLength ~ year + species * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
+
+mod8 <- gam(observedLength ~ year * species + riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
+
+mod9 <- gam(observedLength ~ year + species + riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
+
+AIC(mod1,mod2,mod3,mod4,mod5,mod6,mod7,mod8,mod9) %>% rownames_to_column() %>% arrange(AIC)
+
+summary(mod6)
+plot(mod6)
+
+hist(resid(mod6))
+gam.check(mod6)
+
+# try spline on year - no real advantage, except we can see the squiggle
+mod6a <- gam(cf ~ species * riverOrdered + 
+                              s(ydayCumul, bs = 'cr') + 
+                              s(year), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), 
+             method = "REML")
+plot(mod6a, scale = 0)
+summary(mod6a)
+
+```
+
+```{r mean lengths by grouping Variable}
 
@@
 #####################################################
-# SElect THIS for metaPop analysis, or
+
+# Select THIS for whole pop analysis, or
+groupingVariables <- c('species','ageInSamples','year','season','age')
+groupingVariablesNoYear <- c('species','ageInSamples','season','age')
+riverGroupVar <- NULL
+
+# Select THIS for metaPop analysis, or
 #groupingVariables <- c('species','metaPop01','ageInSamples', 'year','season','age')
 #groupingVariablesNoYear <- c('species','metaPop01','ageInSamples','season','age')
+#riverGroupVar <- groupingVariables[2]
 
 # THIS for 4 river analysis
-groupingVariables <- c('species','riverOrdered','ageInSamples','year','season','age')
-groupingVariablesNoYear <- c('species','riverOrdered','ageInSamples','season','age')
-
+#groupingVariables <- c('species','riverOrdered','ageInSamples','year','season','age')
+#groupingVariablesNoYear <- c('species','riverOrdered','ageInSamples','season','age')
+#riverGroupVar <- groupingVariables[2]
 #####################################################
 #####################################################
-
-riverGroupVar <- groupingVariables[2]
 
@@
                  metaPop01 = ifelse(riverOrdered == 'wb obear', 0 ,1)) %>%
-    filter( ageInSamples %in% 1:10, age < 3, year < 2016, area %in% c('inside','trib'), enc == 1
+    filter( ageInSamples %in% 1:9, age < 3, year < 2016, area %in% c('inside','trib'), enc == 1
             #!(riverOrdered == 'wb mitchell' & ageInSamples > 8) #low sample sizes
@@
 means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+means$metaPop01GG <- factor(means$metaPop01, levels = c(0,1), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
 

### 2019-03-14 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
switched to bam(), many tests

@@
 ```{r Load data, message=FALSE}
+library(itsadug)
 library(arm)
@@
 #######################
-
-# 1. Do we need a RE on year in s()? YES
-mod0a <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod1)
-
-mod0 <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod1)
-
-AIC(mod0,mod0a) %>% rownames_to_column() %>% arrange(AIC)
+vB$yearFactor <- factor(vB$year)
+vB$yearOrdered <- factor(vB$year, levels = 2000:2013)
+vB$yearOrderedSpecies <- factor(paste(vB$yearOrdered,vB$species, sep = '_'))
+
+# 1. Do we need a RE on year in s()? YES, yearOrdered is best
+# 
+mod0_0 <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod0_1 <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod0_2 <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr', by = (yearFactor)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod0_3 <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod0_4 <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr', by = (yearOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+AIC(mod0_0,mod0_1,mod0_2,mod0_3,mod0_4) %>% rownames_to_column() %>% arrange(AIC)
+
+# 2. structure of smooths
+# with ordered factor, need to include an overall smooth, http://www.sfs.uni-tuebingen.de/~jvanrij/Tutorial/GAMM.html 
+mod0_3a <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod0_3b <- gam(observedLength ~ yearOrdered + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+# mod0_3b is best, but we want to estimate a trend as in mod0_3a
+# add + s(ydayCumul, bs = 'cr') to models
+AIC(mod0_3,mod0_3a,mod0_3b) %>% rownames_to_column() %>% arrange(AIC)
+
+# 3. gam vs. bam for this big dataset? bam() is MUCH faster!
+system.time( mod0_3a_1 <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
+# 316 sec
+summary(mod0_3a_1)
+
+system.time( mod0_3a_2 <- bam(observedLength ~ year + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
+# 8 sec
+summary(mod0_3a_2)
+
+# 4. bs = 'cr', vs default 'tp'. AIC for 'cr' is ~ 300 lower. Stick with 'cr'
+mod0_3a_2a <- bam(observedLength ~ year + s(ydayCumul) + s(ydayCumul, by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+AIC(mod0_3a_2,mod0_3a_2a)
+
+# 5. Include Individual as a RE??
+# Taking a long time, maybe run overnight
+vB$tagIndexFactor <- factor(vB$tagIndex)
+system.time( mod0_3a_3 <- bam(observedLength ~ year + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (yearOrdered)) + s(ydayCumul, tagIndexFactor, bs="fs", m=1), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
+# sec
+compareML(mod0_3a_3,mod0_3a_2)
 
 # 2. Best structure for year,species,river
-mod1 <- gam(observedLength ~ species + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-
-mod2 <- gam(observedLength ~ river + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+mod0 <- bam(observedLength ~ year + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = yearOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod1 <- bam(observedLength ~ species + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = yearOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod2 <- bam(observedLength ~ river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = yearOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
 
@@
 # 2-way models
-mod3 <- gam(observedLength ~ year * species + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-
-mod4 <- gam(observedLength ~ year * riverOrdered * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod4)
-
-mod5 <- gam(observedLength ~ species * riverOrdered * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod4)
-
-# mod5 is best
+mod3 <- bam(observedLength ~ year * species + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = yearOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod4 <- bam(observedLength ~ year * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = yearOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod5 <- bam(observedLength ~ species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = yearOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+# mod4 is best
 AIC(mod0,mod1,mod2,mod3,mod4,mod5) %>% rownames_to_column() %>% arrange(AIC)
@@
 # 3-way models
-mod6 <- gam(observedLength ~ year * species * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-
-# https://www.fromthebottomoftheheap.net/2017/12/14/difference-splines-ii/
-# calculates differences from overall spline for each year
-# a lot slower
-vB$yearOrdered <- factor(vB$year, levels = 2000:2013)
-mod6a <- gam(observedLength ~ year * species * riverOrdered + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-
-# try with river instead of riverOrdered
-mod6b <- gam(observedLength ~ year * species * river + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod6b)
-
-# 3-way is best
-AIC(mod1,mod2,mod3,mod4,mod5,mod6,mod6a) %>% rownames_to_column() %>% arrange(AIC)
+mod6 <- bam(observedLength ~ year * species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = yearOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod6)
 
 # Any additive 3-way better? NO
-mod7 <- gam(observedLength ~ year + species * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
-
-mod8 <- gam(observedLength ~ year * species + riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
-
-mod9 <- gam(observedLength ~ year + species + riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
+mod7 <- bam(observedLength ~ year + species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = yearOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
+
+mod8 <- bam(observedLength ~ year * species + river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = yearOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
+
+mod9 <- bam(observedLength ~ year + species + river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = yearOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
 

### 2019-03-14 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
added speciesRiver model

--- 2019-03-13 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
@@
 # by species (bkt, bnt), by cohort, by river gridded
-ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats')), aes(ydayCumul,cf, group = species, color = (species))) +
+ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats')), aes(ydayCumul,observedLength, group = species, color = (species))) +
   geom_point(size = 0.5, alpha = 0.1) +
@@
 #######################
-
-# 1. Do we need a RE on year in s()? YES
-mod0a <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod1)
-
-mod0 <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod1)
-
-AIC(mod0,mod0a) %>% rownames_to_column() %>% arrange(AIC)
+vB$yearFactor <- factor(vB$year)
+vB$yearOrdered <- factor(vB$year, levels = 2000:2013, ordered = T)
+vB$yearOrderedSpecies <- factor(paste(vB$yearOrdered,vB$species, sep = '_'), ordered = T)
+
+vB$cohortFactor <- factor(vB$cohort)
+vB$cohortOrdered <- factor(vB$cohort, levels = 2000:2013, ordered = T)
+vB$cohortOrderedSpecies <- factor(paste(vB$cohortOrdered,vB$species, sep = '_'), ordered = T)
+
+vB$speciesOrdered <- factor(vB$species, levels = c('bkt','bnt'), ordered = T)
+vB$cohortSpeciesRiver <- factor(paste(vB$cohortOrdered,vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
+
+vB$speciesRiver <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
+
+# 1. Do we need separate splines on cohort in s()? YES, cohortOrdered is best
+# 
+mod0_0 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod0_1 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr', by = (cohort)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod0_2 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr', by = (cohortFactor)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod0_3 <- bam(observedLength ~ 
+                cohort + 
+                s(ydayCumul, bs = 'cr') + 
+                s(ydayCumul, bs = 'cr', by = (cohortOrdered)), 
+              data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod0_4 <- bam(observedLength ~ cohort + 
+                s(ydayCumul, bs = 'cr') + 
+                s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)),
+                data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+#mod0_4 is best
+AIC(mod0_0,mod0_1,mod0_2,mod0_3,mod0_4) %>% rownames_to_column() %>% arrange(AIC)
+
+# 2. structure of smooths
+# with ordered factor, need to include an overall smooth, http://www.sfs.uni-tuebingen.de/~jvanrij/Tutorial/GAMM.html 
+mod0_4a <- gam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod0_4b <- gam(observedLength ~ cohortOrdered + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+# mod0_3b is best, but we want to estimate a trend as in mod0_3a
+# add + s(ydayCumul, bs = 'cr') to models
+AIC(mod0_4,mod0_4a,mod0_4b) %>% rownames_to_column() %>% arrange(AIC)
+
+# 3. gam vs. bam for this big dataset? bam() is MUCH faster!
+system.time( mod0_4a_1 <- gam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
+# 316 sec
+summary(mod0_3a_1)
+
+system.time( mod0_4a_2 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
+# 8 sec
+summary(mod0_4a_2)
+
+# 4. bs = 'cr', vs default 'tp'. AIC for 'cr' is ~ 300 lower. Stick with 'cr'
+mod0_4a_2a <- bam(observedLength ~ cohort + s(ydayCumul) + s(ydayCumul, by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+AIC(mod0_4a_2,mod0_4a_2a)
+
+# 5. Include Individual as a RE??
+# Taking a long time, maybe run overnight
+vB$tagIndexFactor <- factor(vB$tagIndex)
+system.time( mod0_3a_3 <- bam(observedLength ~ year + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (yearOrdered)) + s(ydayCumul, tagIndexFactor, bs="fs", m=1), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
+# sec
+compareML(mod0_3a_3,mod0_3a_2)
 
 # 2. Best structure for year,species,river
-mod1 <- gam(observedLength ~ species + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-
-mod2 <- gam(observedLength ~ river + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+mod0 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod1 <- bam(observedLength ~ species + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod2 <- bam(observedLength ~ river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
 
@@
 # 2-way models
-mod3 <- gam(observedLength ~ year * species + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-
-mod4 <- gam(observedLength ~ year * riverOrdered * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod4)
-
-mod5 <- gam(observedLength ~ species * riverOrdered * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod4)
-
-# mod5 is best
+mod3 <- bam(observedLength ~ cohort * species + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod4 <- bam(observedLength ~ cohort * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+mod5 <- bam(observedLength ~ species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+# mod4 is best
 AIC(mod0,mod1,mod2,mod3,mod4,mod5) %>% rownames_to_column() %>% arrange(AIC)
@@
 # 3-way models
-mod6 <- gam(observedLength ~ year * species * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-
-# https://www.fromthebottomoftheheap.net/2017/12/14/difference-splines-ii/
-# calculates differences from overall spline for each year
-# a lot slower
-vB$yearOrdered <- factor(vB$year, levels = 2000:2013)
-mod6a <- gam(observedLength ~ year * species * riverOrdered + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-
-# try with river instead of riverOrdered
-mod6b <- gam(observedLength ~ year * species * river + s(ydayCumul, bs = 'cr', by = (yearOrdered)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+mod6 <- bam(observedLength ~ cohort * species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod6)
+
+mod6a <- bam(observedLength ~ cohort * species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrderedSpecies), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod6a)
+
+mod6b <- bam(observedLength ~ cohort * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrderedSpecies), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
 summary(mod6b)
 
-# 3-way is best
-AIC(mod1,mod2,mod3,mod4,mod5,mod6,mod6a) %>% rownames_to_column() %>% arrange(AIC)
+#mod6a is best
+AIC(mod6,mod6a,mod6b) %>% rownames_to_column() %>% arrange(AIC)
 
 # Any additive 3-way better? NO
-mod7 <- gam(observedLength ~ year + species * riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
-
-mod8 <- gam(observedLength ~ year * species + riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
-
-mod9 <- gam(observedLength ~ year + species + riverOrdered + s(ydayCumul, bs = 'cr', by = (year)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
+mod7 <- bam(observedLength ~ cohort + species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
+
+mod8 <- bam(observedLength ~ cohort * species + river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
+
+mod9 <- bam(observedLength ~ cohort + species + river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))
 
@@
 summary(mod6)
-plot(mod6)
-
-hist(resid(mod6))
-gam.check(mod6)
+plot(mod6a)
+
+hist(resid(mod6a))
+gam.check(mod6a)
+
+plot(mod6a, shade = TRUE, pages = 2,  seWithMean = TRUE)
+
+#################################
+# Predictions
+# https://www.fromthebottomoftheheap.net/2017/10/10/difference-splines-i/
+
+mod6a <- bam(observedLength ~ cohort * species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+
+predGrid <- expand.grid(cohortOrderedSpecies = unique(mod6a$model$cohortOrderedSpecies), 
+                        river = mod6a$xlevels$river,  
+                        ydayCumul = seq(250,1000, by = 30)) %>%
+            mutate(cohort = as.numeric(substr(cohortOrderedSpecies,1,4)),
+                   species = substr(cohortOrderedSpecies,6,8))
+
+pred <- predict(mod6a, newdata = predGrid)#, type = 'lpmatrix')
+
+predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
+
+ggplot(predGrid2, aes(ydayCumul,pred, color = species)) +
+  geom_point() + 
+  geom_line() +
+  facet_grid(river ~ cohort)
+
+mod6csr <- bam(observedLength ~ cohort * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiver), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod6csr)
+plot(mod6csr)
+
+predGrid <- expand.grid(cohortSpeciesRiver = unique(mod6csr$model$cohortSpeciesRiver),
+                        ydayCumul = seq(250,1000, by = 30)) %>%
+                 mutate(cohort = as.numeric(substr(cohortSpeciesRiver,1,4)),
+                        species = substr(cohortSpeciesRiver,6,8),
+                        river = substring(cohortSpeciesRiver,10))
+
+pred <- predict(mod6csr, newdata = predGrid)#, type = 'lpmatrix')
+
+predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
+
+ggplot(predGrid2, aes(ydayCumul,pred, color = species)) +
+  geom_point() + 
+  geom_line() +
+  ylim(0,300) +
+  facet_grid(river ~ cohort)
+
+mod6sr <- bam(observedLength ~ cohort * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', by = speciesRiver), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod6sr)
+
+predGrid <- expand.grid(speciesRiver = unique(mod6sr$model$speciesRiver),
+                        cohort = 2000:2013,
+                        ydayCumul = seq(250,1000, by = 30)) %>%
+                 mutate(species = substr(speciesRiver,1,3),
+                        river = substring(speciesRiver,5))
+
+pred <- predict(mod6sr, newdata = predGrid)#, type = 'lpmatrix')
+
+predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
+
+ggplot(predGrid2, aes(ydayCumul,pred, color = species)) +
+  geom_point() + 
+  geom_line() +
+  facet_grid(river ~ cohort)
 

### 2019-03-15 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
add pairwise by variables for gam

@@
 vB$speciesRiver <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
+vB$speciesCohort <- factor(paste(vB$speciesOrdered,vB$cohortOrdered, sep = '_'), ordered = T)
+vB$cohortRiver <- factor(paste(vB$cohortOrdered,vB$riverOrdered, sep = '_'), ordered = T)
 
@@
 
+# by cohortSpeciesRiver - model is overfit
+
 mod6csr <- bam(observedLength ~ cohort * species * river + 
@@
 
-mod6sr <- bam(observedLength ~ cohort * species * river + 
+# by speciesRiver
+
+mod6SR <- bam(observedLength ~ cohort * species * river + 
                s(ydayCumul, bs = 'cr') +
@@
 
-predGrid <- expand.grid(speciesRiver = unique(mod6sr$model$speciesRiver),
+predGridSR <- expand.grid(speciesRiver = unique(mod6SR$model$speciesRiver),
                         cohort = 2000:2013,
@@
 
-pred <- predict(mod6sr, newdata = predGrid)#, type = 'lpmatrix')
-
-predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
-
-ggplot(predGrid2, aes(ydayCumul,pred, color = species)) +
+predSR <- predict(mod6SR, newdata = predGrid)#, type = 'lpmatrix')
+
+predGrid2SR <- bind_cols(predGridSR,as.data.frame(predSR))
+
+ggplot(predGrid2SR, aes(ydayCumul,pred, color = species)) +
   geom_point() + 
   geom_line() +
+  facet_grid(river ~ cohort)
+
+# by speciesCohort
+
+mod6SC <- bam(observedLength ~ cohort * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', by = speciesCohort), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod6SC)
+
+predGridSC <- expand.grid(speciesCohort = unique(mod6SC$model$speciesCohort),
+                        river = unique(vB$riverOrdered),
+                        ydayCumul = seq(250,1000, by = 30)) %>%
+                 mutate(species = substr(speciesCohort,1,3),
+                        cohort = as.numeric(substring(speciesCohort,5)))
+
+predSC <- predict(mod6SC, newdata = predGridSC)#, type = 'lpmatrix')
+
+predGrid2SC <- bind_cols(predGridSC,as.data.frame(predSC))
+
+ggplot(predGrid2SC, aes(ydayCumul,predSC, color = species)) +
+  geom_point() + 
+  geom_line() +
+  facet_grid(river ~ cohort)
+
+# by cohortRiver
+
+mod6CR <- bam(observedLength ~ cohort * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', by = cohortRiver), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(mod6CR)
+
+predGridCR <- expand.grid(cohortRiver = unique(mod6CR$model$cohortRiver),
+                        species = c('bkt','bnt'),
+                        ydayCumul = seq(250,1000, by = 30)) %>%
+                 mutate(cohort = as.numeric(substr(cohortRiver,1,4)),
+                        river = (substring(cohortRiver,6)))
+
+predCR <- predict(mod6CR, newdata = predGridCR)#, type = 'lpmatrix')
+
+predGrid2CR <- bind_cols(predGridCR,as.data.frame(predCR))
+
+ggplot(predGrid2CR, aes(ydayCumul, predCR, color = species)) +
+  geom_point() + 
+  geom_line() +
+  ylim(50,200) +
   facet_grid(river ~ cohort)

### 2019-03-15 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
add full interactions in s()

--- 2019-03-14 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
@@
 
-```{r functions}
-
-riverOrderedIn <- factor(c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),labels = c("WB","OL","OS","IL"), ordered = T)
-
-```
-
 Plots for the manuscript
@@
 library(arrayhelpers)
+library(mgcv)
 library(tidyverse)
@@
 
-load("D:/projects/linkedModels/data/out/ddD_bktbntats_1997.RData")
-load("D:/projects/linkedModels/data/out/ddddD_bktbntats_1997.RData")
-load("D:/projects/linkedModels/data/out/rawCountsAndMasses.RData")
-
-mcmc <- list()
-dGIn <- list()
-ddGIn2 <- list()
-
-load("D:/projects/linkedModels/data/out/dG_1530971738_crossValFALSE_bkt1997_Nimble.RData")
-
-# input data
-d1 <- ddG[[1]][[1]]
-d2 <- ddG[[1]][[2]]
-meanSampleIntervalMean <- data.frame( meanSampleInterval = colMeans(d1$sampleIntervalMean), season = 1:d1$nSeasons )
-
 #RECALCULATE gr for dataset. Filtering by maxSampleInterval in the creat ddddG step leaves in trailing lengths from the long samp intervals
-d2 <- d2 %>%
-  group_by(tag) %>%
-  # arrange(tag,sampleNumber) %>%
-  mutate( lagObservedWeight = lead(observedWeight),
-          lagObservedLength = lead(observedLength),
-          grWeight = exp(lagObservedWeight - observedWeight)/as.numeric((lagDetectionDate - detectionDate)),
-          grLength = (lagObservedLength - observedLength)/as.numeric((lagDetectionDate - detectionDate))
-  )  %>%
-  ungroup()
-
-#d2 %>% filter(tag=='00088cc059') %>% select(tag,riverOrdered,riverN,lagRiverN,sampleNumber,observedLength,lagObservedLength)
-
-#input data frame
-ddGIn2[['bkt']] <- ddG[[1]][[2]]
-dGIn[['bkt']] <- dG
-# output data
-mcmc[['bkt']] <- mcmcProcessed
-
-load("D:/projects/linkedModels/data/out/dG_1530971744_crossValFALSE_bnt1997_Nimble.RData")
-ddGIn2[['bnt']] <- ddG[[1]][[2]]
-dGIn[['bnt']] <- dG
-mcmc[['bnt']]  <- mcmcProcessed
-
-load("D:/projects/linkedModels/data/out/dG_1530993703_crossValFALSE_ats1997_Nimble.RData")
-ddGIn2[['ats']] <- ddG[[1]][[2]]
-dGIn[['ats']] <- dG
-mcmc[['ats']]  <- mcmcProcessed
-
-rm(ddG,dG,mcmcProcessed)
+# d2 <- d2 %>%
+#   group_by(tag) %>%
+#   # arrange(tag,sampleNumber) %>%
+#   mutate( lagObservedWeight = lead(observedWeight),
+#           lagObservedLength = lead(observedLength),
+#           grWeight = exp(lagObservedWeight - observedWeight)/as.numeric((lagDetectionDate - detectionDate)),
+#           grLength = (lagObservedLength - observedLength)/as.numeric((lagDetectionDate - detectionDate))
+#   )  %>%
+#   ungroup()
 
@@
 
+####################################
 # gam models
-library(mgcv)
-mod_bnt <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species == 'bnt'))
-summary(mod1_bnt)
-
-mod_bkt <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species == 'bkt'))
-summary(mod_bkt)
-
-mod_ats <- gam(observedLength ~ year + s(ydayCumul, bs = 'cr'), data = vB %>% filter(species == 'ats'))
-summary(mod_ats)
-
-#######################
-vB$yearFactor <- factor(vB$year)
-vB$yearOrdered <- factor(vB$year, levels = 2000:2013, ordered = T)
-vB$yearOrderedSpecies <- factor(paste(vB$yearOrdered,vB$species, sep = '_'), ordered = T)
-
+
+cohortOffset <- 2006
+
+# 1-way
 vB$cohortFactor <- factor(vB$cohort)
 vB$cohortOrdered <- factor(vB$cohort, levels = 2000:2013, ordered = T)
-vB$cohortOrderedSpecies <- factor(paste(vB$cohortOrdered,vB$species, sep = '_'), ordered = T)
-
+#vB$cohortFactorCentered <- vB$cohortFactor - 2006
+vB$cohortCentered <- vB$cohort - cohortOffset
+
+vB$speciesFactor <- factor(vB$species, levels = c('bkt','bnt'))
 vB$speciesOrdered <- factor(vB$species, levels = c('bkt','bnt'), ordered = T)
-vB$cohortSpeciesRiver <- factor(paste(vB$cohortOrdered,vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
-
-vB$speciesRiver <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
+
+vB$riverFactor <- factor(vB$river, levels = c('bkt','bnt'))
+#vB$riverOrdered <- riverOrdered exists
+
+# 2-way
+vB$speciesRiverOrdered <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
+vB$speciesCohortOrdered <- factor(paste(vB$speciesOrdered,vB$cohortOrdered, sep = '_'), ordered = T)
+vB$cohortRiverOrdered <- factor(paste(vB$cohortOrdered,vB$riverOrdered, sep = '_'), ordered = T)
+
+# 3-way
+vB$cohortSpeciesRiverOrdered <- factor(paste(vB$cohortOrdered,vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
+
+##########################################################
+# by cohortSpeciesRiver - model is overfit?
+
+modCSR <- bam(observedLength ~ cohortCentered * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered) +
+               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(modCSR)
+#plot(modCSR)
+
+predGridCSR <- expand.grid(cohortSpeciesRiver = unique(modCSR$model$cohortSpeciesRiver),
+                           ydayCumul = seq(250,1000, by = 30)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiver,1,4)),
+                           speciesOrdered = substr(cohortSpeciesRiver,6,8),
+                           riverOrdered = substring(cohortSpeciesRiver,10),
+                           cohortCentered = cohortOrdered - cohortOffset,
+                           species = speciesOrdered,
+                           river = riverOrdered)
+
+predCSR <- predict(modCSR, newdata = predGridCSR)#, type = 'lpmatrix')
+
+predGrid2CSR <- bind_cols(predGridCSR,as.data.frame(predCSR))
+
+ggplot(predGrid2CSR, aes(ydayCumul,pred, color = species)) +
+  geom_point() + 
+  geom_line() +
+  ylim(50,210) +
+  facet_grid(river ~ cohort)
+
+##########################################################
+# by speciesRiver ##########################################################
+
+modSR <- bam(observedLength ~ cohortCentered * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(modSR)
+#plot(modSR, scale = 0)
+
+predGridSR <- expand.grid(speciesRiverOrdered = unique(modSR$model$speciesRiverOrdered),
+                          cohortOrdered = 2000:2013,
+                          ydayCumul = seq(250,1000, by = 30)) %>%
+                   mutate(speciesOrdered = substr(speciesRiverOrdered,1,3),
+                          riverOrdered = substring(speciesRiverOrdered,5),
+                          cohortCentered = cohortOrdered - cohortOffset,
+                          species = speciesOrdered,
+                          river = riverOrdered)
+
+predSR <- predict(modSR, newdata = predGridSR)#, type = 'lpmatrix')
+
+predGrid2SR <- bind_cols(predGridSR,as.data.frame(predSR))
+
+ggplot(predGrid2SR, aes(ydayCumul,predSR, color = species)) +
+  geom_point() + 
+  geom_line() +
+  facet_grid(river ~ cohortOrdered)
+
+# by speciesCohort
+system.time(
+modSC <- bam(observedLength ~ cohortCentered * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+)
+summary(modSC)
+#plot(modSC, scale = 0)
+
+predGridSC <- expand.grid(speciesCohortOrdered = unique(modSC$model$speciesCohortOrdered),
+                          river = unique(vB$riverOrdered),
+                          ydayCumul = seq(250,1000, by = 30)) %>%
+                   mutate(speciesOrdered = substr(speciesCohortOrdered,1,3),
+                          cohortOrdered = as.numeric(substring(speciesCohortOrdered,5)),
+                          cohortCentered = cohortOrdered - cohortOffset,
+                          species = speciesOrdered)
+
+predSC <- predict(modSC, newdata = predGridSC)#, type = 'lpmatrix')
+
+predGrid2SC <- bind_cols(predGridSC,as.data.frame(predSC))
+
+ggplot(predGrid2SC, aes(ydayCumul,predSC, color = species)) +
+  geom_point() + 
+  geom_line() +
+  facet_grid(river ~ cohortOrdered)
+
+# by cohortRiver
+
+modCR <- bam(observedLength ~ cohortCentered * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered), 
+               data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+summary(modCR)
+#plot(modCR, scale = 0)
+
+predGridCR <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
+                          speciesOrdered = c('bkt','bnt'),
+                          ydayCumul = seq(250,1000, by = 30)) %>%
+                   mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
+                          riverOrdered = (substring(cohortRiverOrdered,6)),
+                          cohort = cohortOrdered,
+                          cohortCentered = cohortOrdered - cohortOffset,
+                          river = riverOrdered,
+                          species = speciesOrdered)
+
+predCR <- predict(modCR, newdata = predGridCR)#, type = 'lpmatrix')
+
+predGrid2CR <- bind_cols(predGridCR, as.data.frame(predCR))
+
+ggplot(predGrid2CR, aes(ydayCumul, predCR, color = species)) +
+  geom_point() + 
+  geom_line() +
+  ylim(50,200) +
+  facet_grid(river ~ cohortOrdered)
+
+#############################
+
+# add year to smooths
+
+modCSRY <- bam(observedLength ~ cohortCentered * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered) +
+               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered) +
+               s(year, bs = 'cr', by = speciesOrdered), 
+               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
+               method = "REML")
+summary(modCSRY)
+
+################################################################################
+
+# base model
+
+mod <- bam(observedLength ~ cohortCentered * species * river + 
+               s(ydayCumul, bs = 'cr'), 
+               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
+               method = "REML")
+summary(modCSRY)
+
+AIC(modCSRY,modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
+
+```
+
+```{r early gams}
+# some early exploration
 
@@
   facet_grid(river ~ cohort)
-
-mod6csr <- bam(observedLength ~ cohort * species * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiver), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod6csr)
-plot(mod6csr)
-
-predGrid <- expand.grid(cohortSpeciesRiver = unique(mod6csr$model$cohortSpeciesRiver),
-                        ydayCumul = seq(250,1000, by = 30)) %>%
-                 mutate(cohort = as.numeric(substr(cohortSpeciesRiver,1,4)),
-                        species = substr(cohortSpeciesRiver,6,8),
-                        river = substring(cohortSpeciesRiver,10))
-
-pred <- predict(mod6csr, newdata = predGrid)#, type = 'lpmatrix')
-
-predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
-
-ggplot(predGrid2, aes(ydayCumul,pred, color = species)) +
-  geom_point() + 
-  geom_line() +
-  ylim(0,300) +
-  facet_grid(river ~ cohort)
-
-mod6sr <- bam(observedLength ~ cohort * species * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = speciesRiver), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(mod6sr)
-
-predGrid <- expand.grid(speciesRiver = unique(mod6sr$model$speciesRiver),
-                        cohort = 2000:2013,
-                        ydayCumul = seq(250,1000, by = 30)) %>%
-                 mutate(species = substr(speciesRiver,1,3),
-                        river = substring(speciesRiver,5))
-
-pred <- predict(mod6sr, newdata = predGrid)#, type = 'lpmatrix')
-
-predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
-
-ggplot(predGrid2, aes(ydayCumul,pred, color = species)) +
-  geom_point() + 
-  geom_line() +
-  facet_grid(river ~ cohort)
-
-# try spline on year - no real advantage, except we can see the squiggle
-mod6a <- gam(cf ~ species * riverOrdered + 
-                              s(ydayCumul, bs = 'cr') + 
-                              s(year), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), 
-             method = "REML")
-plot(mod6a, scale = 0)
-summary(mod6a)
-
 ```

### 2019-03-20 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
add modCSR_add

@@
 ##########################################################
-# by cohortSpeciesRiver - model is overfit?
-
-modCSR <- bam(observedLength ~ cohortCentered * species * river + 
+# by cohortSpeciesRiver
+
+modCSR <- bam(observedLength ~ cohort * species * river + 
                s(ydayCumul, bs = 'cr') +
@@
 
-predGridCSR <- expand.grid(cohortSpeciesRiver = unique(modCSR$model$cohortSpeciesRiver),
+predGridCSR <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR$model$cohortSpeciesRiverOrdered),
                            ydayCumul = seq(250,1000, by = 30)) %>%
-                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiver,1,4)),
-                           speciesOrdered = substr(cohortSpeciesRiver,6,8),
-                           riverOrdered = substring(cohortSpeciesRiver,10),
+                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
+                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
+                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
+                           speciesCohortOrdered = paste(speciesOrdered, cohortOrdered, sep = '_'),
+                           speciesRiverOrdered = paste(speciesOrdered, riverOrdered, sep = '_'),
                            cohortCentered = cohortOrdered - cohortOffset,
+                           cohort = cohortOrdered,
                            species = speciesOrdered,
@@
 
-ggplot(predGrid2CSR, aes(ydayCumul,pred, color = species)) +
+ggplot(predGrid2CSR, aes(ydayCumul,predCSR, color = species)) +
   geom_point() + 
@@
 ##########################################################
-# by speciesRiver ##########################################################
-
-modSR <- bam(observedLength ~ cohortCentered * species * river + 
+# by cohortSpeciesRiver - additive splines
+system.time(
+modCSR_add <- bam(observedLength ~ cohort * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
+               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+)
+summary(modCSR_add)
+#plot(modCSR_add)
+
+predGridCSR_add <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR_add$model$cohortSpeciesRiverOrdered),
+                           ydayCumul = seq(250,1000, by = 30)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
+                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
+                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
+                           speciesCohortOrdered = paste(speciesOrdered, cohortOrdered, sep = '_'),
+                           speciesRiverOrdered = paste(speciesOrdered, riverOrdered, sep = '_'),
+                           cohortCentered = cohortOrdered - cohortOffset,
+                           cohort = cohortOrdered,
+                           species = speciesOrdered,
+                           river = riverOrdered)
+
+predCSR_add <- predict(modCSR_add, newdata = predGridCSR_add)#, type = 'lpmatrix')
+
+predGrid2CSR_add <- bind_cols(predGridCSR_add,as.data.frame(predCSR_add))
+
+ggplot(predGrid2CSR_add, aes(ydayCumul,predCSR_add, color = species)) +
+  geom_point() + 
+  geom_line() +
+  ylim(50,210) +
+  facet_grid(river ~ cohort)
+
+##########################################################
+# by speciesRiver
+
+modSR <- bam(observedLength ~ cohort * species * river + 
                s(ydayCumul, bs = 'cr') +
@@
                           cohortCentered = cohortOrdered - cohortOffset,
+                          cohort = cohortOrdered,
                           species = speciesOrdered,
@@
 system.time(
-modSC <- bam(observedLength ~ cohortCentered * species * river + 
+modSC <- bam(observedLength ~ cohort * species * river + 
                s(ydayCumul, bs = 'cr') +
@@
                           cohortCentered = cohortOrdered - cohortOffset,
+                          cohort = cohortOrdered,
                           species = speciesOrdered)
@@
 
-modCR <- bam(observedLength ~ cohortCentered * species * river + 
+modCR <- bam(observedLength ~ cohort * species * river + 
                s(ydayCumul, bs = 'cr') +
@@
 
-# add year to smooths
-
-modCSRY <- bam(observedLength ~ cohortCentered * species * river + 
+# add year to smooths, model seems overfit
+
+# try running without 'speciesOrdered' for year  this does a difference from the main fit, i.e. try s(year, bs = 'cr', by = speciesFactor)
+
+modCSRY <- bam(observedLength ~ cohort * species * river + 
                s(ydayCumul, bs = 'cr') +
@@
                s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered) +
-               s(year, bs = 'cr', by = speciesOrdered), 
+               #s(year, bs = 'cr', by = speciesOrdered), 
+               s(year, bs = 'cr', by = speciesFactor),
                data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
@@
 summary(modCSRY)
+#plot(modCSRY, select = 189)
+
+predGridCSRY <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSRY$model$cohortSpeciesRiverOrdered),
+                            year = unique(modCSRY$model$year),
+                            ydayCumul = seq(250,1000, by = 30)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
+                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
+                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
+                           speciesCohortOrdered = paste(speciesOrdered, cohortOrdered, sep = '_'),
+                           speciesRiverOrdered = paste(speciesOrdered, riverOrdered, sep = '_'),
+                           cohortCentered = cohortOrdered - cohortOffset,
+                           cohort = cohortOrdered,
+                           species = speciesOrdered,
+                           river = riverOrdered)
+
+predCSRY <- predict(modCSRY, newdata = predGridCSRY)#, type = 'lpmatrix')
+
+predGrid2CSRY <- bind_cols(predGridCSRY,as.data.frame(predCSRY))
+
+ggplot(predGrid2CSRY, aes(ydayCumul, predCSRY, color = species)) +
+  geom_point() + 
+  geom_line() +
+  ylim(50,300) +
+  facet_grid(river ~ cohortOrdered)
 
@@
 
-mod <- bam(observedLength ~ cohortCentered * species * river + 
+mod <- bam(observedLength ~ cohort * species * river + 
                s(ydayCumul, bs = 'cr'), 
@@
 
-AIC(modCSRY,modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
+save(mod,modCR,modCSR,modSC,modSR,modCSRY, file = './data/out/mods.RData')
+AIC(modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
 

### 2019-03-20 MS_Analysis_3spp_4rivers_Patterns_from 2spp_BKT only.Rmd
added bkt only file

--- 2019-03-20 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
@@
 
+################################################################################
+
+# base model
+
+mod <- bam(observedLength ~ cohort * river + 
+               s(ydayCumul, bs = 'cr'), 
+               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt')), 
+               method = "REML")
+summary(mod)
+
+predGrid <- expand.grid(cohort = unique(mod$model$cohort),
+                        river = unique(mod$model$river),
+                        ydayCumul = seq(250,1000, by = 30))
+
+pred <- predict(mod, newdata = predGrid)#, type = 'lpmatrix')
+
+predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
+
+ggplot(predGrid2, aes(ydayCumul, pred)) +
+  geom_point() + 
+  geom_line() +
+  facet_grid(river ~ cohort) +
+  ggtitle('mod')
+
 ##########################################################
 # by cohortSpeciesRiver 
-
-modCSR <- bam(observedLength ~ cohort * species * river + 
+kIn <- 4
+system.time(
+modCR <- bam(observedLength ~ cohort * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt')), method = "REML")
+)
+summary(modCR)
+plot(modCR, pages = 8, scale = 0)
+
+predGridCR <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
+                           ydayCumul = seq(250,1000, by = 30)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
+                           riverOrdered = substring(cohortRiverOrdered,6),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
+                           cohortCentered = cohortOrdered - cohortOffset,
+                           cohort = cohortOrdered,
+                           river = riverOrdered)
+
+predCR <- predict(modCR, newdata = predGridCR)#, type = 'lpmatrix')
+
+predGrid2CR <- bind_cols(predGridCR,as.data.frame(predCR))
+
+ggplot(predGrid2CR, aes(ydayCumul,predCR)) +
+  geom_point() + 
+  geom_line() +
+  ylim(50,210) +
+  facet_grid(river ~ cohort)+
+  ggtitle('modCR')
+
+##########################################################
+# by cohortSpeciesRiver, additive in s() 
+kIn <- 4 # 4 looks good
+system.time(
+modC_R <- bam(observedLength ~ cohort * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt')), method = "REML")
+)
+summary(modC_R)
+plot(modC_R, pages = 4, scale = 0)
+
+predGridC_R <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
+                           ydayCumul = seq(250,1000, by = 30)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
+                           riverOrdered = substring(cohortRiverOrdered,6),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
+                           cohortCentered = cohortOrdered - cohortOffset,
+                           cohort = cohortOrdered,
+                           river = riverOrdered)
+
+predC_R <- predict(modC_R, newdata = predGridC_R)#, type = 'lpmatrix')
+
+predGrid2C_R <- bind_cols(predGridC_R,as.data.frame(predC_R))
+
+ggplot(predGrid2C_R, aes(ydayCumul,predC_R)) +
+  geom_point() + 
+  geom_line() +
+  ylim(50,210) +
+  facet_grid(river ~ cohort)+
+  ggtitle('modC_R')
+
+##########################################################
+# by cohortSpeciesRiver - additive splines, no 3-way spline
+system.time(
+modC_S_R <- bam(observedLength ~ cohort * species * river + 
                s(ydayCumul, bs = 'cr') +
@@
                s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered), 
+               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
              data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(modCSR)
-#plot(modCSR)
-
-predGridCSR <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR$model$cohortSpeciesRiverOrdered),
+)
+summary(modC_S_R)
+#plot(modC_S_R)
+
+predGridC_S_R <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR$model$cohortSpeciesRiverOrdered), # use modCSR to get 3-way
                                ydayCumul = seq(250,1000, by = 30)) %>%
@@
 
-predCSR <- predict(modCSR, newdata = predGridCSR)#, type = 'lpmatrix')
-
-predGrid2CSR <- bind_cols(predGridCSR,as.data.frame(predCSR))
-
-ggplot(predGrid2CSR, aes(ydayCumul,predCSR, color = species)) +
+predC_S_R <- predict(modC_S_R, newdata = predGridC_S_R)#, type = 'lpmatrix')
+
+predGrid2C_S_R <- bind_cols(predGridC_S_R,as.data.frame(predC_S_R))
+
+ggplot(predGrid2C_S_R, aes(ydayCumul,predC_S_R, color = species)) +
   geom_point() + 
@@
   ylim(50,210) +
-  facet_grid(river ~ cohort)
-
-##########################################################
-# by cohortSpeciesRiver - additive splines
-system.time(
-modCSR_add <- bam(observedLength ~ cohort * species * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-)
-summary(modCSR_add)
-#plot(modCSR_add)
-
-predGridCSR_add <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR_add$model$cohortSpeciesRiverOrdered),
-                           ydayCumul = seq(250,1000, by = 30)) %>%
-                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
-                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
-                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
-                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
-                           speciesCohortOrdered = paste(speciesOrdered, cohortOrdered, sep = '_'),
-                           speciesRiverOrdered = paste(speciesOrdered, riverOrdered, sep = '_'),
-                           cohortCentered = cohortOrdered - cohortOffset,
-                           cohort = cohortOrdered,
-                           species = speciesOrdered,
-                           river = riverOrdered)
-
-predCSR_add <- predict(modCSR_add, newdata = predGridCSR_add)#, type = 'lpmatrix')
-
-predGrid2CSR_add <- bind_cols(predGridCSR_add,as.data.frame(predCSR_add))
-
-ggplot(predGrid2CSR_add, aes(ydayCumul,predCSR_add, color = species)) +
-  geom_point() + 
-  geom_line() +
-  ylim(50,210) +
-  facet_grid(river ~ cohort)
+  facet_grid(river ~ cohort) +
+  ggtitle('modC_S_R')
 
@@
 summary(modSR)
-#plot(modSR, scale = 0)
+plot(modCR, scale = 0)
 
@@
   geom_line() +
-  facet_grid(river ~ cohortOrdered)
-
+  facet_grid(river ~ cohortOrdered)+
+  ggtitle('modSR')
+
+#######################################
 # by speciesCohort
@@
   geom_line() +
-  facet_grid(river ~ cohortOrdered)
-
+  facet_grid(river ~ cohortOrdered) +
+  ggtitle('modSC')
+
+##################################
 # by cohortRiver
@@
   ylim(50,200) +
-  facet_grid(river ~ cohortOrdered)
+  facet_grid(river ~ cohortOrdered) +
+  ggtitle('modCR')
 
@@
   ylim(50,300) +
-  facet_grid(river ~ cohortOrdered)
-
-################################################################################
-
-# base model
-
-mod <- bam(observedLength ~ cohort * species * river + 
-               s(ydayCumul, bs = 'cr'), 
-               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
-               method = "REML")
-summary(modCSRY)
-
-save(mod,modCR,modCSR,modSC,modSR,modCSRY, file = './data/out/mods.RData')
-AIC(modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
+  facet_grid(river ~ cohortOrdered) +
+  ggtitle('modCSRY')
+
+##########################################
+
+save(mod,modCR,modCSR,modSC,modSR,modCSRY,modC_S_R, file = './data/out/mods.RData')
+AIC(modC_S_R,modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
 

### 2019-03-22 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
added ...bySpp.Rmd

@@
 
+kIn <- 4
+dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt'))
+```
+
+``` {r gam models}
+
 ##########################################################
@@
                s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered), 
+             data = dataIn, method = "REML")
 summary(modCSR)
@@
   ylim(50,210) +
-  facet_grid(river ~ cohort)
+  facet_grid(river ~ cohort)+
+  ggtitle('modCSR')
 
 ##########################################################
-# by cohortSpeciesRiver - additive splines
+# by cohortSpeciesRiver - additive splines, no 3-way spline
 system.time(
-modCSR_add <- bam(observedLength ~ cohort * species * river + 
+modC_S_R <- bam(observedLength ~ cohort * species * river + 
                s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered), 
+             data = dataIn, method = "REML")
 )
-summary(modCSR_add)
-#plot(modCSR_add)
-
-predGridCSR_add <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR_add$model$cohortSpeciesRiverOrdered),
+summary(modC_S_R)
+#plot(modC_S_R)
+
+predGridC_S_R <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR$model$cohortSpeciesRiverOrdered), # use modCSR to get 3-way
                                ydayCumul = seq(250,1000, by = 30)) %>%
@@
 
-predCSR_add <- predict(modCSR_add, newdata = predGridCSR_add)#, type = 'lpmatrix')
-
-predGrid2CSR_add <- bind_cols(predGridCSR_add,as.data.frame(predCSR_add))
-
-ggplot(predGrid2CSR_add, aes(ydayCumul,predCSR_add, color = species)) +
+predC_S_R <- predict(modC_S_R, newdata = predGridC_S_R)#, type = 'lpmatrix')
+
+predGrid2C_S_R <- bind_cols(predGridC_S_R,as.data.frame(predC_S_R))
+
+ggplot(predGrid2C_S_R, aes(ydayCumul,predC_S_R, color = species)) +
   geom_point() + 
@@
   ylim(50,210) +
-  facet_grid(river ~ cohort)
+  facet_grid(river ~ cohort) +
+  ggtitle('modC_S_R')
 
@@
                s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered), 
+             data = dataIn, method = "REML")
 summary(modSR)
@@
                s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered), 
+             data = dataIn, method = "REML")
 )
@@
                s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered), 
-               data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered), 
+               data = dataIn, method = "REML")
 summary(modCR)
@@
                s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
                #s(year, bs = 'cr', by = speciesOrdered), 
-               s(year, bs = 'cr', by = speciesFactor),
-               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
+               s(year, bs = 'cr', k = kIn, by = speciesFactor),
+               data = dataIn, 
                method = "REML")
 summary(modCSRY)
-#plot(modCSRY, select = 189)
+plot(modCSRY, pages=10, scale = 0)
 
@@
   ylim(50,300) +
-  facet_grid(river ~ cohortOrdered)
+  facet_grid(river ~ cohortOrdered) +
+  ggtitle('modCSRY')
+
+# trying a couple of models
+modCSRC <- bam(observedLength ~ cohort * species * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
+               #s(year, bs = 'cr', by = speciesOrdered), 
+               s(cohort, bs = 'cr', k = kIn, by = speciesFactor),
+               data = dataIn, 
+               method = "REML")
+summary(modCSRC)
+plot(modCSRC, pages=8, scale = 0)
+
+modC_S_RC <- bam(observedLength ~ cohort + species + river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn, by = speciesFactor),
+               data = dataIn, 
+               method = "REML")
+summary(modC_S_RC)
+plot(modCSRC, pages=8, scale = 0)
+
+# modCSR_Ind <- bam(observedLength ~ cohort * species * river + 
+#                s(ydayCumul, bs = 'cr') +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
+#                s(ydayCumul, tagIndexFactor, bs="fs", m=1, k = kIn),
+#                data = dataIn, 
+#                method = "REML")
+# summary(modCSR_Ind)
 
@@
                s(ydayCumul, bs = 'cr'), 
-               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
+               data = dataIn, 
                method = "REML")
-summary(modCSRY)
-
-save(mod,modCR,modCSR,modSC,modSR,modCSRY, file = './data/out/mods.RData')
-AIC(modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
+summary(mod)
+
+predGrid <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR$model$cohortSpeciesRiverOrdered),
+                        ydayCumul = seq(250,1000, by = 30)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
+                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
+                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
+                           cohort = cohortOrdered,
+                           species = speciesOrdered,
+                           river = riverOrdered)
+
+pred <- predict(mod, newdata = predGrid)#, type = 'lpmatrix')
+
+predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
+
+ggplot(predGrid2, aes(ydayCumul, pred, color = species)) +
+  geom_point() + 
+  geom_line() +
+  facet_grid(river ~ cohortOrdered) +
+  ggtitle('mod')
+
+##########################################
+
+save(mod,modCR,modCSR,modSC,modSR,modCSRY,modC_S_R, file = './data/out/mods.RData')
+AIC(modC_S_RC,modC_S_R,modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
 

### 2019-03-22 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
added ...bySpp.Rmd

--- 2019-03-15 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
@@
 ---
-title: "Growth model analysis 3spp WB"
+title: "Body size patterns"
 output: html_notebook
@@
 
-```{r von Bert, well, gam}
+```{r gams for body size - model selection}
+speciesIn <- 'bnt'
+depVarIn <- 'observedLength'
+kIn <- 4 # # of knots for the speces and river smooths
 
@@
 
-vB$riverFactor <- factor(vB$river, levels = c('bkt','bnt'))
+vB$riverFactor <- factor(vB$river, levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'))
 #vB$riverOrdered <- riverOrdered exists
@@
 
+vB$tagIndexFactor <- factor(vB$tagIndex)
+
+#vB$riverOrdered <- factor(vB$river,
+#                         levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'),
+#                         ordered = T)
+
+dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn)
+kIn <- 4
+################################################################################
+
+# base model
+
+mod <- bam(observedLength ~ cohort * river + 
+               s(ydayCumul, bs = 'cr'), 
+               data = dataIn, 
+               method = "REML")
+summary(mod)
+
+predGrid <- expand.grid(cohort = unique(mod$model$cohort),
+                        river = unique(mod$model$river),
+                        ydayCumul = seq(250,1000, by = 30))
+
+pred <- predict(mod, newdata = predGrid)#, type = 'lpmatrix')
+
+predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
+
+ggplot(predGrid2, aes(ydayCumul, pred)) +
+  geom_point() + 
+  geom_line() +
+  facet_grid(river ~ cohort) +
+  ggtitle('mod')
+
 ##########################################################
-# by cohortSpeciesRiver - model is overfit?
-
-modCSR <- bam(observedLength ~ cohortCentered * species * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(modCSR)
-#plot(modCSR)
-
-predGridCSR <- expand.grid(cohortSpeciesRiver = unique(modCSR$model$cohortSpeciesRiver),
-                           ydayCumul = seq(250,1000, by = 30)) %>%
-                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiver,1,4)),
-                           speciesOrdered = substr(cohortSpeciesRiver,6,8),
-                           riverOrdered = substring(cohortSpeciesRiver,10),
+# by cohortSpeciesRiver
+
+system.time(
+modCR <- bam(observedLength ~ cohort * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered), 
+             data = dataIn, method = "REML")
+)
+summary(modCR)
+#plot(modCR, pages = 8, scale = 0)
+
+predGridCR <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
+                           ydayCumul = seq(250,1000, by = 30)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
+                           riverOrdered = substring(cohortRiverOrdered,6),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                            cohortCentered = cohortOrdered - cohortOffset,
-                           species = speciesOrdered,
+                           cohort = cohortOrdered,
                            river = riverOrdered)
 
-predCSR <- predict(modCSR, newdata = predGridCSR)#, type = 'lpmatrix')
-
-predGrid2CSR <- bind_cols(predGridCSR,as.data.frame(predCSR))
-
-ggplot(predGrid2CSR, aes(ydayCumul,pred, color = species)) +
+predCR <- predict(modCR, newdata = predGridCR)#, type = 'lpmatrix')
+
+predGrid2CR <- bind_cols(predGridCR, as.data.frame(predCR))
+
+ggplot(predGrid2CR, aes(ydayCumul, predCR)) +
   geom_point() + 
@@
   ylim(50,210) +
-  facet_grid(river ~ cohort)
+  facet_grid(river ~ cohort)+
+  ggtitle('modCR')
 
 ##########################################################
-# by speciesRiver ##########################################################
-
-modSR <- bam(observedLength ~ cohortCentered * species * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(modSR)
-#plot(modSR, scale = 0)
-
-predGridSR <- expand.grid(speciesRiverOrdered = unique(modSR$model$speciesRiverOrdered),
-                          cohortOrdered = 2000:2013,
-                          ydayCumul = seq(250,1000, by = 30)) %>%
-                   mutate(speciesOrdered = substr(speciesRiverOrdered,1,3),
-                          riverOrdered = substring(speciesRiverOrdered,5),
+# by cohortSpeciesRiver with s(cohort)
+
+system.time(
+modCR_c <- bam(observedLength ~ cohort * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn), 
+             data = dataIn, method = "REML")
+)
+summary(modCR_c)
+#plot(modCR_c, pages = 8, scale = 0)
+
+predGridCR_c <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
+                           ydayCumul = seq(250,1000, by = 30)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
+                           riverOrdered = substring(cohortRiverOrdered,6),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                            cohortCentered = cohortOrdered - cohortOffset,
-                          species = speciesOrdered,
+                           cohort = cohortOrdered,
                            river = riverOrdered)
 
-predSR <- predict(modSR, newdata = predGridSR)#, type = 'lpmatrix')
-
-predGrid2SR <- bind_cols(predGridSR,as.data.frame(predSR))
-
-ggplot(predGrid2SR, aes(ydayCumul,predSR, color = species)) +
-  geom_point() + 
-  geom_line() +
-  facet_grid(river ~ cohortOrdered)
-
-# by speciesCohort
+predCR_c <- predict(modCR_c, newdata = predGridCR_c)#, type = 'lpmatrix')
+
+predGrid2CR_c <- bind_cols(predGridCR_c, as.data.frame(predCR_c))
+
+ggplot(predGrid2CR_c, aes(ydayCumul, predCR_c, group = cohort, color = cohort)) +
+  geom_point() + 
+  geom_line() +
+  #ylim(50,210) +
+  facet_wrap(~river)+
+  ggtitle('modCR_c')
+
+##########################################################
+# by cohortRiver - w ind RE
+
+# need more memory
+# system.time(
+# modCR_Ind <- bam(observedLength ~ cohort * river +
+#                s(ydayCumul, bs = 'cr') +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+#                s(ydayCumul, tagIndexFactor, bs="fs", m=1, k = kIn),
+#              data = dataIn, method = "REML")
+# )
+# summary(modCR_Ind)
+# plot(modCR_Ind, pages = 8, scale = 0)
+
+##########################################################
+# by cohortSpeciesRiver, additive in s()
+
 system.time(
-modSC <- bam(observedLength ~ cohortCentered * species * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
+modC_R <- bam(observedLength ~ cohort * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered), 
+             data = dataIn, method = "REML")
 )
-summary(modSC)
-#plot(modSC, scale = 0)
-
-predGridSC <- expand.grid(speciesCohortOrdered = unique(modSC$model$speciesCohortOrdered),
-                          river = unique(vB$riverOrdered),
-                          ydayCumul = seq(250,1000, by = 30)) %>%
-                   mutate(speciesOrdered = substr(speciesCohortOrdered,1,3),
-                          cohortOrdered = as.numeric(substring(speciesCohortOrdered,5)),
+summary(modC_R)
+#plot(modC_R, pages = 4, scale = 0)
+
+predGridC_R <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
+                           ydayCumul = seq(250,1000, by = 30)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
+                           riverOrdered = substring(cohortRiverOrdered,6),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                            cohortCentered = cohortOrdered - cohortOffset,
-                          species = speciesOrdered)
-
-predSC <- predict(modSC, newdata = predGridSC)#, type = 'lpmatrix')
-
-predGrid2SC <- bind_cols(predGridSC,as.data.frame(predSC))
-
-ggplot(predGrid2SC, aes(ydayCumul,predSC, color = species)) +
-  geom_point() + 
-  geom_line() +
-  facet_grid(river ~ cohortOrdered)
-
-# by cohortRiver
-
-modCR <- bam(observedLength ~ cohortCentered * species * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered), 
-               data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
-summary(modCR)
-#plot(modCR, scale = 0)
-
-predGridCR <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
-                          speciesOrdered = c('bkt','bnt'),
+                           cohort = cohortOrdered,
+                           river = riverOrdered)
+
+predC_R <- predict(modC_R, newdata = predGridC_R)#, type = 'lpmatrix')
+
+predGrid2C_R <- bind_cols(predGridC_R,as.data.frame(predC_R))
+
+ggplot(predGrid2C_R, aes(ydayCumul,predC_R)) +
+  geom_point() + 
+  geom_line() +
+  ylim(50,210) +
+  facet_grid(river ~ cohort)+
+  ggtitle('modC_R')
+
+##########################
+
+modC <- bam(observedLength ~ cohort + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered), 
+             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt')), method = "REML")
+
+modR <- bam(observedLength ~ river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered), 
+             data = dataIn, method = "REML")
+
+AIC(mod,modC,modR,modC_R,modCR,modCR_c) %>% rownames_to_column() %>% arrange(AIC)
+save(mod,modC,modR,modC_R,modCR,modCR_c, file = paste0('./data/out/mods_',speciesIn,'.RData'))
+######
+
+##########################################
+
+```
+
+```{r best models}
+
+# Brook trout
+
+modBKT <- bam(observedLength ~ cohort * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn), 
+             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
+
+summary(modBKT)
+#plot(modBKT, pages = 8, scale = 0)
+
+predGridBKT <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
                            ydayCumul = seq(250,1000, by = 30)) %>%
                     mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
-                          riverOrdered = (substring(cohortRiverOrdered,6)),
+                           riverOrdered = substring(cohortRiverOrdered,6),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
+                           cohortCentered = cohortOrdered - cohortOffset,
                            cohort = cohortOrdered,
+                           river = riverOrdered)
+
+predBKT <- predict(modBKT, newdata = predGridBKT)#, type = 'lpmatrix')
+
+predGrid2BKT <- bind_cols(predGridBKT, as.data.frame(predBKT)) %>% mutate(pred = predBKT, species = 'bkt')
+
+#################################
+# Brown trout
+
+modBNT <- bam(observedLength ~ cohort * river + 
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn), 
+             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
+
+summary(modBNT)
+#plot(modBNT, pages = 8, scale = 0)
+
+predGridBNT <- expand.grid(cohortRiverOrdered = unique(modBNT$model$cohortRiverOrdered),
+                           ydayCumul = seq(250,1000, by = 30)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
+                           riverOrdered = substring(cohortRiverOrdered,6),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                            cohortCentered = cohortOrdered - cohortOffset,
-                          river = riverOrdered,
-                          species = speciesOrdered)
-
-predCR <- predict(modCR, newdata = predGridCR)#, type = 'lpmatrix')
-
-predGrid2CR <- bind_cols(predGridCR, as.data.frame(predCR))
-
-ggplot(predGrid2CR, aes(ydayCumul, predCR, color = species)) +
-  geom_point() + 
-  geom_line() +
-  ylim(50,200) +
-  facet_grid(river ~ cohortOrdered)
-
-#############################
-
-# add year to smooths
-
-modCSRY <- bam(observedLength ~ cohortCentered * species * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
-               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered) +
-               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered) +
-               s(year, bs = 'cr', by = speciesOrdered), 
-               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
-               method = "REML")
-summary(modCSRY)
-
-################################################################################
-
-# base model
-
-mod <- bam(observedLength ~ cohortCentered * species * river + 
-               s(ydayCumul, bs = 'cr'), 
-               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
-               method = "REML")
-summary(modCSRY)
-
-AIC(modCSRY,modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
+                           cohort = cohortOrdered,
+                           river = riverOrdered)
+
+predBNT <- predict(modBNT, newdata = predGridBNT)#, type = 'lpmatrix')
+
+predGrid2BNT <- bind_cols(predGridBNT, as.data.frame(predBNT)) %>% mutate(pred = predBNT, species = 'bnt')
+
+##################
+# combine species' predictions
+
+predGrid2Both <- bind_rows(predGrid2BKT, predGrid2BNT) %>% select(-predBKT, -predBNT)
+
+predGrid2Both$riverGG <- factor(predGrid2Both$riverOrdered,
+                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
+                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
+predGrid2Both$speciesGG <- factor(predGrid2Both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot(predGrid2Both, aes(ydayCumul, pred, group = cohort, color = cohort)) +
+  geom_point() + 
+  geom_line() +
+  labs(y = 'Predicted body size (mm)',
+       x = 'Days since age-0 Jan 1',
+       color = 'Cohort') +
+  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
+  theme_publication() +
+  theme(legend.position = "right", 
+        legend.direction = "vertical",
+        legend.title = element_text(face="plain")) +
+  guides(color = guide_colorbar(barheight = 10,
+                                barwidth = 1.5,
+                                ticks = FALSE,
+                                nbin = 100)) +
+#  guide_legend(title = 'Cohort') +
+  facet_grid(speciesGG ~ riverGG)
 
@@
 ```
-
-```{r mean lengths by grouping Variable}
-
-#####################################################
-#####################################################
-
-# Select THIS for whole pop analysis, or
-groupingVariables <- c('species','ageInSamples','year','season','age')
-groupingVariablesNoYear <- c('species','ageInSamples','season','age')
-riverGroupVar <- NULL
-
-# Select THIS for metaPop analysis, or
-#groupingVariables <- c('species','metaPop01','ageInSamples', 'year','season','age')
-#groupingVariablesNoYear <- c('species','metaPop01','ageInSamples','season','age')
-#riverGroupVar <- groupingVariables[2]
-
-# THIS for 4 river analysis
-#groupingVariables <- c('species','riverOrdered','ageInSamples','year','season','age')
-#groupingVariablesNoYear <- c('species','riverOrdered','ageInSamples','season','age')
-#riverGroupVar <- groupingVariables[2]
-#####################################################
-#####################################################
-
-# Has mean size changed over time?
-means <- 
-  cd %>% mutate( age = year - cohort,
-                 metaPop01 = ifelse(riverOrdered == 'wb obear', 0 ,1)) %>%
-    filter( ageInSamples %in% 1:9, age < 3, year < 2016, area %in% c('inside','trib'), enc == 1
-            #!(riverOrdered == 'wb mitchell' & ageInSamples > 8) #low sample sizes
-            ) %>% 
-    group_by_at(vars(one_of(groupingVariables))) %>%
-    summarize(meanLen = mean(observedLength,na.rm = TRUE),
-              sdLen = sd(observedLength,na.rm = TRUE),
-              meanGR = mean(grLength,na.rm = TRUE),
-              sdGR = sd(grLength,na.rm = TRUE),
-              n = n()
-              )
-
-means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-means$metaPop01GG <- factor(means$metaPop01, levels = c(0,1), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-lenSlopes <- means %>%
-  group_by_at(vars(one_of(groupingVariablesNoYear))) %>%
-  #group_by(species,metaPop01,ageInSamples,year,season,age) %>%
-  nest()
-
-lengthModel <- function(df){
-  lm( meanLen ~ year, data = df )
-}
-
-grModel <- function(df){
-  lm( meanGR ~ year, data = df )
-}
-
-lenSlopes <- lenSlopes %>%
-  mutate( model = map(lenSlopes$data, lengthModel),
-          modelGR = map(lenSlopes$data, grModel))
-
-###################################
-######### len here, gr below ######
-lenSlopesTidy <- lenSlopes %>%
-  mutate(glance = map(model, broom::tidy)) %>% 
-  unnest(glance, .drop = TRUE) %>%
-  filter(term == 'year') %>%
-  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
-  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
-  mutate( sig = ifelse(p.value < 0.0001,"***",sig))  %>%
-  filter(!is.nan(statistic))
-
-lenSlopesTidy$seasonGG <- factor(lenSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-lenSlopesTidy$speciesGG <- factor(lenSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-save(lenSlopesTidy,file = paste0('D:/projects/westBrook-patterns/data/out/lenSlopesTidy_',riverGroupVar,'.RData'))
-##
-
-ggplot( means %>% filter(year >= 0), aes(year,meanLen, group= speciesGG)) +
-  geom_point( aes(color = speciesGG), size = 2) +
-  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
-  labs( x = "Year", y = "Mean body length (mm)" ) +
-  # geom_text(data = lenSlopesTidy %>% filter(species == 'bkt'), aes(x=1999,y=220, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  # geom_text(data = lenSlopesTidy %>% filter(species == 'bnt'), aes(x=1999,y=200, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  # geom_text(data = lenSlopesTidy %>% filter(species == 'ats'), aes(x=1999,y=180, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age<2), aes(x=2000,y=225, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age<2), aes(x=2000,y=203, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age<2), aes(x=2000,y=181, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-
-    geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2009,y=104, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2009,y=82, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-  theme_publication() +
-  scale_color_grey(start = 0.15, end = 0.75) +
- # facet_grid(seasonGG ~ riverOrdered + age)
-  facet_grid(as.formula(paste("seasonGG ~", riverGroupVar, "+ age")))
-
-ggsave(paste0('figures/meanLengths_',riverGroupVar,'.png'), width = 11, height = 11, units='in')
-
-# mean increases 
-lenSlopesTidy %>%
-  filter() %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# mean increases by species
-lenSlopesTidy %>%
-  group_by(species) %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-lenSlopesTidy %>%
-  group_by(species, metaPop01) %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# do increases (slopes) increase over age? Yes
-#ggplot(lenSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8)), aes((ageInSamples),estimate, color=speciesGG)) +
-ggplot(lenSlopesTidy %>% filter(ageInSamples %in% 1:9), aes((ageInSamples),estimate, color=speciesGG)) +
-  geom_point( size = 3) +
-  geom_smooth(method='lm', se=FALSE) +
-  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
-  labs( x = "Age", y = "Slope of mean body length over years" ) +
-  theme_publication() +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_wrap(as.formula(paste("~", riverGroupVar)))
-
-ggsave(paste0('figures/meanLengthsOverAge_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
-
-# just bkt
-ggplot(lenSlopesTidy %>% filter(ageInSamples %in% 1:9, species == 'bkt'), aes_(~ageInSamples, ~estimate, group = as.name(riverGroupVar))) +
-  geom_point(aes_(color= as.name(riverGroupVar)), size = 3) +
-  geom_smooth(aes_(color= as.name(riverGroupVar)), method='lm', se=FALSE) +
-  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
-  labs( x = "Age", y = "Slope of mean body length over years" ) +
-  theme_publication()
-
-ggsave(paste0('figures/meanLengthsOverAgeBKT_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
-
-summary(lm(estimate~ageInSamples, data = lenSlopesTidy %>% filter(species=='bkt')))
-summary(lm(estimate~ageInSamples, data = lenSlopesTidy %>% filter(species=='bnt')))
-
-############################################################################################################
-#########################
-#### gr here, len above
-
-grSlopesTidy <- lenSlopes %>%
-  mutate(glance = map(modelGR, broom::tidy)) %>% 
-  unnest(glance, .drop = TRUE) %>%
-  filter(term == 'year') %>%
-  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
-  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
-  mutate( sig = ifelse(p.value < 0.0001,"***",sig))  %>%
-  filter(!is.nan(statistic))
-
-grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-save(grSlopesTidy,file = paste0('D:/projects/westBrook-patterns/data/out/grSlopesTidy_',riverGroupVar,'.RData'))
-##
-
-##
-means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-ggplot( means %>% filter(), aes(year,meanGR, group= speciesGG)) +
-  geom_point( aes(color = speciesGG), size = 4) +
-  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
-  labs( x = "Year", y = "Mean growth rate (mm/d)" ) +
-  # geom_text(data = grSlopesTidy %>% filter(species == 'bkt'), aes(x=1999,y=220, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  # geom_text(data = grSlopesTidy %>% filter(species == 'bnt'), aes(x=1999,y=200, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  # geom_text(data = grSlopesTidy %>% filter(species == 'ats'), aes(x=1999,y=180, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age<2), aes(x=2000,y=0.7, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
-  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age<2), aes(x=2000,y=0.6, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
-#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age<2), aes(x=2000,y=181, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-
-    geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2011.5,y=0.66, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
-  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2011.5,y=0.56, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
-#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-  theme_publication() +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(as.formula(paste("seasonGG ~", riverGroupVar, "+ age")))
-
-ggsave(paste0('figures/meanGRs_',riverGroupVar,'.png'), width = 11, height = 11, units='in')
-
-# mean increases 
-grSlopesTidy %>%
-  filter() %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# mean increases by species
-grSlopesTidy %>%
-  group_by(species) %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# do increases (slopes) increase over age? Yes
-#ggplot(lenSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8)), aes((ageInSamples),estimate, color=speciesGG)) +
-ggplot(grSlopesTidy %>% filter(ageInSamples %in% 1:9), aes((ageInSamples),estimate, color=speciesGG)) +
-  geom_point( size = 3) +
-  geom_smooth(method='lm', se=FALSE) +
-  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
-  labs( x = "Age", y = "Slope of mean body length over years" ) +
-  theme_publication() +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_wrap(as.formula(paste("~", riverGroupVar)))
-
-ggsave(paste0('figures/meanGRsOverAge_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
-
-# just bkt
-ggplot(grSlopesTidy %>% filter(ageInSamples %in% 1:9, species == 'bkt'), aes_(~ageInSamples, ~estimate, group = as.name(riverGroupVar))) +
-  geom_point(aes_(color= as.name(riverGroupVar)), size = 3) +
-  geom_smooth(aes_(color= as.name(riverGroupVar)), method='lm', se=FALSE) +
-  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
-  labs( x = "Age", y = "Slope of mean body length over years" ) +
-  theme_publication()
-
-ggsave(paste0('figures/meanGRsOverAgeBKT_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
-
-summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bkt')))
-summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bnt')))
-
-```
-
-```{r where are the fish?}
-
-allFishBySpeciesYOY$seasonGG <- factor(allFishBySpeciesYOY$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-
-allFishBySpeciesYOY$speciesGG <- factor(allFishBySpeciesYOY$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-allFishBySpeciesYOY$riverGG <- factor(allFishBySpeciesYOY$river,
-                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("WB","OL","OS","IL"), ordered = T)
-
-allFishBySpeciesYOY %>% filter(!is.na(isYOY)) %>%
-  ggplot(aes(year,nAllFishBySpeciesYOY, color = speciesGG)) +
-    geom_point(size = 2) +
-    geom_smooth(method = 'lm', se=F) +
-    theme_publication() +
-    labs( x = "Year", y = "Estimated abundance" ) +
-    scale_color_grey(start = 0.2, end = 0.8) + 
-    facet_grid(seasonGG ~ riverGG + isYOY, scales = 'free')
-
-cd2 <- 
-  cd %>% mutate(age = year - cohort) %>%
-    filter(ageInSamples %in% 1:10, age < 3, year < 2016, enc == 1, area %in% c('inside','trib')) %>%
-   # group_by(species,riverOrdered,ageInSamples,year,season,age)
-     mutate(riverArea = ifelse( river == "west brook" & section > 20 & section < 32, 2, 1)) %>%
-     mutate(riverArea = ifelse( river == "west brook" & section >= 32, 3, riverArea))
-
-counts <- cd2 %>%
-  group_by(riverOrdered, riverArea)
-
-```
-
-# below from MS_Analysis_2spp_WB
-
-```{r count summaries for MS}
-
-ddGInAll <- rbind(ddGIn2[[1]],ddGIn2[[2]],ddGIn2[[3]]) %>%
-  filter(year %in% 2000:2015, river == 'west brook') %>%
-  mutate( consecOcc = ifelse( is.na(grLength), 0,1 ) )
-
-ddGInAll %>%
-  group_by(river,species) %>%
-  distinct(tag) %>%
-  summarize( n = n() )
-
-#   species     n
-#   <chr>   <int>
-# 1 ats      5032
-# 2 bkt      8556
-# 3 bnt      6389
-
-ddGInAll %>%
-  group_by(river,species) %>%
-  summarize( n = n() )
-
-#   species     n
-#   <chr>   <int>
-# 1 ats     15791
-# 2 bkt     18412
-# 3 bnt     16510
-
-ddGInAll %>%
-  group_by(river,species,enc) %>%
-  summarize( n = n() )
-
-#   species   enc     n
-#   <chr>   <dbl> <int>
-# 1 ats         0  3515
-# 2 ats         1 12276
-# 3 bkt         0  3464
-# 4 bkt         1 14948
-# 5 bnt         0  3855
-# 6 bnt         1 12655
-
-ddGInAll %>%
-  group_by(river,species,consecOcc) %>%
-  summarize( n = n() ) %>%
-  spread(consecOcc,n, sep="_") %>%
-  mutate( prop0 = consecOcc_0/(consecOcc_0+consecOcc_1),
-          prop1 = 1 - prop0)
-
-#   river       species consecOcc_0 consecOcc_1 prop0 prop1
-# 5 west brook  ats           10203        5588 0.646 0.354
-# 6 west brook  bkt           13631        4781 0.740 0.260
-# 7 west brook  bnt           12420        4090 0.752 0.248
-
-length(unique(ddGInAll$sampleNumber))
-
-```
-
-```{r table of estimated abundances}
-
-ddGIn2[['bkt']] %>%
-  filter(river == 'west brook') %>%
-  distinct(river,year,season,nAllFishBySpeciesPBKT,nAllFishBySpeciesPBNT,nAllFishBySpeciesPATS) %>%
-  arrange(year, season)
-
-```
-
-```{r bnt density}
-
-ggplot(d2, aes(cBNTStd)) +
-  geom_histogram() +
-  facet_grid(season~riverOrdered)
-
-ggplot(d2, aes(cBNTStd, grLength)) +
-  geom_point(alpha=0.1) +
-  geom_smooth(method='lm') +
-  facet_grid(isYOY+season~riverOrdered)
-
-ggplot(d2, aes(cBKTStd, grLength)) +
-  geom_point(alpha=0.1) +
-  geom_smooth(method='lm') +
-  facet_grid(isYOY+season~riverOrdered)
-```
-
-```{r check lengthExp}
-len <- list()
-for( spp in c("bkt","bnt","ats") ){
-
-#  len2 <- array2df(mcmc[[spp]]$sims.list$lengthExp, label.x = "est") %>%
-#           rename(iter=d1,obs=d2)
-  offset <- 1
-  if(spp == 'ats') offset <- 0 # not sure why this is needed. REs seem meesed up in the graphs below
-  inds <- data.frame(ind = dGIn[[spp]][[1]]$constants$ind[1:(length(dGIn[[spp]][[1]]$constants$ind) - offset)],
-                     observedLength = dGIn[[spp]][[1]]$data$lengthDATA,
-                     sampleNumber = ddGIn2[[spp]]$sampleNumber[1:(length(ddGIn2[[spp]]$sampleNumber) - offset)],
-                     cohort = ddGIn2[[spp]]$cohort[1:(length(ddGIn2[[spp]]$cohort) - offset)],
-                     ageInSamples = ddGIn2[[spp]]$ageInSamples[1:(length(ddGIn2[[spp]]$ageInSamples) - offset)],
-                     riverOrdered = ddGIn2[[spp]]$riverOrdered[1:(length(ddGIn2[[spp]]$riverOrdered) - offset)],
-                     lagObservedLength = ddGIn2[[spp]]$lagObservedLength[1:(length(ddGIn2[[spp]]$lagObservedLength) - offset)],
-                     detectionDate = ddGIn2[[spp]]$detectionDate[1:(length(ddGIn2[[spp]]$detectionDate) - offset)],
-                     lagDetectionDate = ddGIn2[[spp]]$lagDetectionDate[1:(length(ddGIn2[[spp]]$lagDetectionDate) - offset)])
-  inds$obs <- 1:(length(dGIn[[spp]][[1]]$constants$ind) - offset)
-
-  inds <-  inds %>%
-    group_by(ind) %>%
-    mutate(first = min(obs),
-           firstObs = obs == first,
-           observed = !is.na(observedLength),
-           sampleNumber = sampleNumber,
-           nObs = n(),
-           grLength = (lagObservedLength - observedLength)/as.numeric((lagDetectionDate - detectionDate))
-           ) %>%
-    ungroup()
-
-#  len1 <- left_join(len2,inds)
-#  rm(len2,inds)
-
-  # means
-  lenMean1 <- array2df(mcmc[[spp]]$mean$lengthExp, label.x = "estMean") %>%
-                rename(obs=d1)
-  len[[spp]] <- left_join(inds,lenMean1)  %>%
-                  mutate( estMean = ifelse( firstObs, observedLength, estMean ) )
-
-  grIndREMean1 <- array2df(mcmc[[spp]]$mean$grIndRE, label.x = "grIndRE") %>%
-                rename(ind=d1)
-
-  len[[spp]] <- left_join(len[[spp]],grIndREMean1)
-
-#  len <- left_join(len1,lenMean1)
-#  rm(len1,lenMean1)
-  rm(inds,lenMean1)
-  gc()
-}
-
-len[['bkt']] %>% filter( nObs == 7 ) %>%
-  ggplot(aes(ageInSamples,observedLength)) +
-  geom_point(aes(color = factor(ind)), size=3) +
-  geom_point(aes(ageInSamples, estMean, color=factor(ind))) +
-  geom_line(aes(ageInSamples, estMean, color = factor(ind))) +
-  geom_point(aes(ageInSamples, estMean, color=factor(ind)), size=2) +
-  geom_point(aes(ageInSamples, estMean, color=factor(ind)), shape = 1, size = 2, color = 'black') +
-  theme(legend.position="none") +
-  facet_wrap(~cohort)
-
-focalFish <- 60
-
-len[['bkt']] %>% filter( cohort == 2001 ) %>%
-  ggplot(aes(ageInSamples,observedLength)) +
-  geom_point(aes(color = riverOrdered), size = 3) +
-  geom_point(color = 'black', shape = 1, size = 3) +
-  geom_point(aes(ageInSamples, estMean, color = riverOrdered)) +
-  geom_line( aes(ageInSamples, estMean, group = factor(ind), color = riverOrdered)) +
-  geom_point(aes(ageInSamples, estMean, color = riverOrdered), size = 2) +
-  geom_point(aes(ageInSamples, estMean, color = riverOrdered), shape = 1, size = 2, color = 'black') +
- # theme(legend.position="none") +
-  facet_wrap(~nObs)
-
-len[['bkt']] %>% filter( cohort == 2001 ) %>%
-  ggplot(aes(ageInSamples,observedLength)) +
-  geom_point(aes(color = grIndRE), size = 3) +
-#  geom_point(color = 'black', shape = 1, size = 3) +
-  geom_point(aes(ageInSamples, estMean, color = grIndRE)) +
-  geom_line( aes(ageInSamples, estMean, group = factor(ind), color = grIndRE)) +
-  geom_point(aes(ageInSamples, estMean, color = grIndRE), size = 2) +
-#  geom_point(aes(ageInSamples, estMean, color = grIndRE), shape = 1, size = 2, color = 'black') +
- # scale_colour_distiller(palette = "Spectral") +
-   scale_colour_gradient(low = "white", high = "black") +
- # theme(legend.position="none") +
-  facet_wrap(~nObs)
-
-len[['bkt']] %>% #filter( cohort == 2001 ) %>%
-  ggplot(aes(ageInSamples,observedLength)) +
-  geom_point(aes(color = grIndRE), size = 3) +
-#  geom_point(color = 'black', shape = 1, size = 3) +
-  geom_point(aes(ageInSamples, estMean, color = grIndRE)) +
-  geom_line( aes(ageInSamples, estMean, group = factor(ind), color = grIndRE)) +
-  geom_point(aes(ageInSamples, estMean, color = grIndRE), size = 2) +
-#  geom_point(aes(ageInSamples, estMean, color = grIndRE), shape = 1, size = 2, color = 'black') +
- # scale_colour_distiller(palette = "Spectral") +
-   scale_colour_gradient(low = "white", high = "black") +
- # theme(legend.position="none") +
-  facet_grid(riverOrdered~cohort)
-
-len[['ats']] %>% filter( nObs > 0 ) %>%
-  group_by(ind) %>%
-  summarize( meanGRLength = mean(grLength,na.rm=TRUE),
-             meanGRIndRE = mean(grIndRE,na.rm=TRUE),
-             meanNObs = mean(nObs,na.rm=TRUE)) %>%
-  ggplot(aes(meanGRLength,meanGRIndRE)) +
-  geom_point() +
-  geom_smooth(method='lm') +
-#  scale_colour_gradient(low = "yellow", high = "red") +
-  facet_wrap(~meanNObs)
-
-```
-
-```{r winter growth rates}
-
-ggplot(ddG[[1]][[2]], aes(observedLength,grLength)) + 
-  geom_point() +
-  facet_grid(river + isYOY ~ season+species)
-
-```
-
-```{r adjusted counts}
-allFishBySpeciesYOY$riverOrderedGG <- factor(allFishBySpeciesYOY$river,
-                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
-allFishBySpeciesYOY$seasonGG <- factor(allFishBySpeciesYOY$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-allFishBySpeciesYOY$speciesGG <- factor(allFishBySpeciesYOY$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-#sum across YOY
-allFishBySpecies <- allFishBySpeciesYOY %>%
-   group_by(river,season,year,species) %>%
-   summarize( nAllFishBySpecies = sum(nAllFishBySpeciesYOY, na.rm=T),
-                massAllFishBySpecies = sum(massAllFishBySpeciesYOY, na.rm=T))
-allFishBySpecies$seasonGG <- factor(allFishBySpecies$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-allFishBySpeciesYOY <- left_join(allFishBySpeciesYOY,allFishBySpecies)
-
-# sum across YOY and species
-   allFish <- allFishBySpeciesYOY %>%
-     group_by(river,season,year) %>%
-     summarize( nAllFish = sum(nAllFishBySpeciesYOY, na.rm=T),
-                massAllFish = sum(massAllFishBySpeciesYOY, na.rm=T))
-   cd <- left_join(cd,allFish)
-
-allFish$riverOrderedGG <- factor(allFish$river,
-                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
-allFish$seasonGG <- factor(allFish$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-#########
-
-ggplot(allFishBySpeciesYOY %>% filter(riverOrderedGG == "West Brook"), aes(year,nAllFishBySpecies)) +
-  geom_point(aes(color = speciesGG)) +
-  geom_line(aes(color = speciesGG)) +
-  geom_line( data = allFish %>% filter(riverOrderedGG == "West Brook"), aes(year,nAllFish) ) +
-  theme_publication() +
-  labs( x = "Year", y = "Count" ) +
-  facet_wrap( ~ seasonGG )#, scales = "free_y" )
-
-ggsave('figures/3Spp_WB/counts.png')
-
-ggplot(allFishBySpeciesYOY %>% filter(riverOrderedGG == "West Brook", year >=2000), aes(year,nAllFishBySpecies)) +
-  geom_point(aes(color = speciesGG), size=3) +
-#  geom_line(aes(color = speciesGG)) +
-  geom_smooth(aes(color = speciesGG), se=FALSE,method='lm') +
-  geom_line( data = allFish %>% filter(riverOrderedGG == "West Brook", year >=2000), aes(year,nAllFish), size=1.25, color='darkgrey', linetype = 5 ) +
-  scale_y_continuous(limits = c(0,NA)) +
-  theme_publication() +
-  labs( x = "Year", y = "Estimated abundance" ) +
-  scale_color_grey(start = 0, end = 0.9) +
-  facet_wrap( ~ seasonGG , scales = "free_y" )
-
-ggsave('figures/3Spp_WB/countsWSlope.png', width = 11, height = 9, units='in')
-
-nSlopes <- allFishBySpeciesYOY %>%
-  filter(riverOrderedGG == "West Brook", year >=2000) %>%
-  group_by( species,  season ) %>%
-  nest()
-
-nModel <- function(df){
-  lm( nAllFishBySpecies ~ year, data = df )
-}
-
-nSlopes <- nSlopes %>%
-  mutate( model = map(nSlopes$data, nModel) )
-
-nSlopesTidy <- nSlopes %>%
-  mutate( glance = map(model, broom::tidy)) %>% 
-  unnest(glance, .drop = TRUE) %>%
-  filter( term == 'year')
-
-save(nSlopesTidy,file = 'D:/projects/linkedModels/data/out/nSlopesTidy.RData')
-########################
-# add in sampleNumber
-tmp <- d2 %>% distinct(season,year,sampleNumber)
-allFishBySpeciesYOY <- left_join(allFishBySpeciesYOY,tmp)
-allFish <- left_join(allFish,tmp)
-
-ggplot(allFishBySpeciesYOY, aes(sampleNumber,nAllFishBySpeciesYOY)) +
-  geom_point(aes(color = speciesGG)) +
-  geom_line(aes(color = speciesGG)) +
-  geom_line( data = allFish, aes(sampleNumber,nAllFish) ) +
-  theme_publication() +
-  labs( x = "SampleNumber", y = "Count" ) +
-  facet_wrap(  ~riverOrderedGG )#, scales = "free_y" )
-
-```
-
-```{r Table 1 - sample summary}
-
-# sample summary table
-summaryTable <- allFishBySpecies %>% 
-          select(-massAllFishBySpecies) %>% 
-          filter(river == 'west brook', year %in% 2000:2015) %>% 
-          spread(species,nAllFishBySpecies) %>% 
-          arrange(year,seasonGG)
-
-write.csv(summaryTable, file = 'D:/projects/linkedModels/tables/rawCounts.csv')
-
-```
-
-```{r}
-# Correlation between age classes
-# justification for not using separate age class abundances
-# interesting that it is less for ATS
-# From the bottom of adjustCounts()
-# 
-#                               nAllFishBySpeciesPStdBKT_yoy1 nAllFishBySpeciesPStdBKT_yoy2 nAllFishBySpeciesPStdBNT_yoy1 nAllFishBySpeciesPStdBNT_yoy2
-# nAllFishBySpeciesPStdBKT_yoy1                          1.00                          0.96                         -0.46                         -0.43
-# nAllFishBySpeciesPStdBKT_yoy2                        **0.96**                        1.00                         -0.44                         -0.38
-# nAllFishBySpeciesPStdBNT_yoy1                         -0.46                         -0.44                          1.00                          0.94
-# nAllFishBySpeciesPStdBNT_yoy2                         -0.43                         -0.38                       * *0.94**                       1.00
-# nAllFishBySpeciesPStdATS_yoy1                         -0.13                         -0.17                         -0.21                          0.00
-# nAllFishBySpeciesPStdATS_yoy2                          0.06                         -0.03                         -0.05                          0.15
-#                               nAllFishBySpeciesPStdATS_yoy1 nAllFishBySpeciesPStdATS_yoy2
-# nAllFishBySpeciesPStdBKT_yoy1                         -0.13                          0.06
-# nAllFishBySpeciesPStdBKT_yoy2                         -0.17                         -0.03
-# nAllFishBySpeciesPStdBNT_yoy1                         -0.21                         -0.05
-# nAllFishBySpeciesPStdBNT_yoy2                          0.00                          0.15
-# nAllFishBySpeciesPStdATS_yoy1                          1.00                          0.47
-# nAllFishBySpeciesPStdATS_yoy2                         *0.47*                         1.00
-```
-
-```{r mean lengths}
-
-# Has mean size changed over time?
-means <- 
-  cd %>% mutate( age = year - cohort ) %>%
-    filter( riverOrdered == 'west brook', ageInSamples %in% 1:10, age < 3, year < 2016 ) %>%
-    group_by(species,ageInSamples,year,season,age) %>%
-    summarize(meanLen = mean(observedLength,na.rm = TRUE),
-              sdLen = sd(observedLength,na.rm = TRUE),
-              meanGR = mean(grLength,na.rm = TRUE),
-              sdGR = sd(grLength,na.rm = TRUE)
-              )
-
-lenSlopes <- means %>%
-  group_by( species, ageInSamples, season, age ) %>%
-  nest()
-
-lengthModel <- function(df){
-  lm( meanLen ~ year, data = df )
-}
-
-grModel <- function(df){
-  lm( meanGR ~ year, data = df )
-}
-
-lenSlopes <- lenSlopes %>%
-  mutate( model = map(lenSlopes$data, lengthModel),
-          modelGR = map(lenSlopes$data, grModel))
-
-###################################
-######### len here, gr below ######
-lenSlopesTidy <- lenSlopes %>%
-  mutate(glance = map(model, broom::tidy)) %>% 
-  unnest(glance, .drop = TRUE) %>%
-  filter(term == 'year') %>%
-  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
-  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
-  mutate( sig = ifelse(p.value < 0.0001,"***",sig))
-
-save(lenSlopesTidy,file = 'D:/projects/linkedModels/data/out/lenSlopesTidy.RData')
-##
-means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-lenSlopesTidy$seasonGG <- factor(lenSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-lenSlopesTidy$speciesGG <- factor(lenSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanLen, group= speciesGG)) +
-  geom_point( aes(color = speciesGG), size = 4) +
-  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
-  labs( x = "Year", y = "Mean body length (mm)" ) +
-  # geom_text(data = lenSlopesTidy %>% filter(species == 'bkt'), aes(x=1999,y=220, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  # geom_text(data = lenSlopesTidy %>% filter(species == 'bnt'), aes(x=1999,y=200, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  # geom_text(data = lenSlopesTidy %>% filter(species == 'ats'), aes(x=1999,y=180, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age<2), aes(x=2000,y=225, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age<2), aes(x=2000,y=203, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age<2), aes(x=2000,y=181, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-
-    geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2012.5,y=104, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2012.5,y=82, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-  theme_publication() +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(seasonGG~age)
-
-ggsave(paste0('figures/3Spp_WB/meanLengths.png'), width = 11, height = 11, units='in')
-
-# mean increases 
-lenSlopesTidy %>%
-  filter(species != 'ats') %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# mean increases by species
-lenSlopesTidy %>%
-  group_by(species) %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# do increases (slopes) increase over age? Yes
-ggplot(lenSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
-  geom_point(aes( shape = factor(seasonGG)), size = 3) +
-  geom_smooth(method='lm', se=FALSE) +
-  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
-  labs( x = "Age", y = "Slope of mean body length over years" ) +
-  theme_publication() +
-  scale_color_grey(start = 0.15, end = 0.75)
-
-ggsave(paste0('figures/3Spp_WB/meanLengthsOverAge.png'), width = 6, height = 6, units='in')
-
-summary(lm(estimate~ageInSamples, data = lenSlopesTidy %>% filter(species=='bkt')))
-summary(lm(estimate~ageInSamples, data = lenSlopesTidy %>% filter(species=='bnt')))
-
-#########################
-#### gr here, len above
-
-grSlopesTidy <- lenSlopes %>%
-  mutate(glance = map(modelGR, broom::tidy)) %>% 
-  unnest(glance, .drop = TRUE) %>%
-  filter(term == 'year') %>%
-  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
-  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
-  mutate( sig = ifelse(p.value < 0.0001,"***",sig))
-
-save(grSlopesTidy,file = 'D:/projects/linkedModels/data/out/grSlopesTidy.RData')
-##
-means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanGR, group= speciesGG)) +
-  geom_point( aes(color = speciesGG), size = 4) +
-  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
-  labs( x = "Year", y = "Mean growth rate (mm/d)" ) +
-  # geom_text(data = grSlopesTidy %>% filter(species == 'bkt'), aes(x=1999,y=220, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  # geom_text(data = grSlopesTidy %>% filter(species == 'bnt'), aes(x=1999,y=200, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  # geom_text(data = grSlopesTidy %>% filter(species == 'ats'), aes(x=1999,y=180, label=paste(round(estimate,2),round(p.value,4)), color = speciesGG)) +
-  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age<2), aes(x=2000,y=0.7, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
-  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age<2), aes(x=2000,y=0.6, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
-#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age<2), aes(x=2000,y=181, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-
-    geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2011.5,y=0.66, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
-  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2011.5,y=0.56, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
-#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
-  theme_publication() +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(seasonGG~age)
-
-ggsave(paste0('figures/3Spp_WB/meanGRs.png'), width = 11, height = 11, units='in')
-
-# mean increases 
-grSlopesTidy %>%
-  filter(species != 'ats') %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# mean increases by species
-grSlopesTidy %>%
-  group_by(species) %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# do increases (slopes) increase over age? Yes
-ggplot(grSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
-  geom_point(aes( shape = factor(seasonGG)), size = 3) +
-  geom_smooth(method='lm', se=FALSE) +
-  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
-  labs( x = "Age", y = "Slope of mean growth rate over years" ) +
-  theme_publication() +
-  scale_color_grey(start = 0.15, end = 0.75)
-
-ggsave(paste0('figures/3Spp_WB/meanGRsOverAge.png'), width = 6, height = 6, units='in')
-
-summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bkt')))
-summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bnt')))
-
-```
-
-```{r figs 3,4, flow and temp}
-d2$riverOrderedGG <- factor(d2$river,
-                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
-d2$seasonGG <- factor(d2$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-
-d2 %>% group_by(riverOrderedGG,seasonGG,year) %>%
-  filter( riverOrderedGG == "West Brook" ) %>%
-  summarize(meanF = mean(meanFlow, na.rm=T),
-            sdF = sd(meanFlow, na.rm=T)) %>%
-  ggplot( aes(year,meanF) ) +
-  geom_point( size = 3 ) +
-  geom_line() +
- # geom_smooth(method = 'lm', se=F) +
-  geom_errorbar( aes(ymin = meanF - sdF, ymax = meanF + sdF, width = 0.33 ) ) +
-  labs( x = "Year", y = "Stream flow (x/x)" ) +
-  theme_publication() +
-  facet_wrap(~seasonGG)
-
-ggsave('figures/3Spp_WB/flow.png')
-
-lmFunctionF <- function(df){ lm( meanF ~ year, data=df ) }
-
-flowMods <- d2 %>%
-  group_by(riverOrderedGG,seasonGG,year) %>%
-  summarize(meanF = mean(meanFlow, na.rm=T),
-            sdF = sd(meanFlow, na.rm=T)) %>%
-  group_by( riverOrderedGG, seasonGG ) %>%
-  nest() %>%
-  mutate( model = map(data,lmFunctionF) )
-
-flowSlopesTidy <- flowMods %>%
-  mutate(tidy = map(model,broom::tidy)) %>%
-  dplyr::select(-data,-model) %>%
-  unnest() %>%
-  filter( term == 'year') %>%
-  mutate( sig = ifelse(p.value < 0.05,"*",""))
-
-save(flowSlopesTidy,file = 'D:/projects/linkedModels/data/out/flowSlopesTidy.RData')
-
-#################
-
-lmFunction <- function(df){ lm( meanT ~ year, data=df ) }
-
-tempMods <- d2 %>%
-  group_by(riverOrderedGG,seasonGG,year) %>%
-  summarize(meanT = mean(meanTemperature, na.rm=T),
-            sdT = sd(meanTemperature, na.rm=T)) %>%
-  group_by( riverOrderedGG, seasonGG ) %>%
-  nest() %>%
-  mutate( model = map(data,lmFunction) )
-
-tempSlopesTidy <- tempMods %>%
-  mutate(tidy = map(model,broom::tidy)) %>%
-  dplyr::select(-data,-model) %>%
-  unnest() %>%
-  filter( term == 'year') %>%
-  mutate( sig = ifelse(p.value < 0.05,"*",""))
-
-d2 %>% group_by(riverOrderedGG,seasonGG,year) %>%
-  summarize(meanT = mean(meanTemperature, na.rm=T),
-            sdT = sd(meanTemperature, na.rm=T)) %>%
-  filter(riverOrderedGG == "West Brook") %>%
-  ggplot( aes(year,meanT) ) +
-  geom_smooth(method = 'lm', se=F, color = 'darkgrey') +
-  geom_point( size = 3 ) +
-  geom_line() +
-  geom_errorbar( aes(ymin = meanT - sdT, ymax = meanT + sdT, width = 0.33 ) ) +
-  labs( x = "Year", y = "Stream temperature (C)" ) +
-  geom_text(data=tempSlopesTidy %>% filter(riverOrderedGG == "West Brook"),aes(x=2000.5,y=17.5, label=paste(round(estimate,2),sig))) +
-  theme_publication() +
-  facet_wrap(~seasonGG)
-
-ggsave('figures/3Spp_WB/temp.png')
-
-save(tempSlopesTidy,file = 'D:/projects/linkedModels/data/out/tempSlopesTidy.RData')
-```
-
-```{r Carys Jones analysis w season}
-######
-# for Carys Jones analysis
-dJonesT <- 
-  d2 %>% group_by(riverOrderedGG,seasonGG,year,season) %>%
-    summarize(meanT = mean(meanTemperature, na.rm=T),
-              sdT = sd(meanTemperature, na.rm=T)) %>%
-    filter(riverOrderedGG == "West Brook") %>%
-    ungroup() %>%
-    dplyr::select(year,season,meanT,sdT)
-
-####
-# daily temps
-
-#####
-# Lengths
-# from 'r mean lengths' chunk above
-
-means <- 
-  cd %>% mutate( age = year - cohort ) %>%
-    #filter( riverOrdered == 'west brook', ageInSamples %in% 1:10, age < 3, year < 2016 ) %>%
-    filter( riverOrdered == 'west brook', year < 2016 ) %>%
-    group_by(species,ageInSamples,year,season,age,isYOY) %>%
-    summarize(meanLen = mean(observedLength,na.rm = TRUE),
-              sdLen = sd(observedLength,na.rm = TRUE),
-              meanGR = mean(grLength,na.rm = TRUE),
-              sdGR = sd(grLength,na.rm = TRUE),
-              nLen = n()
-              )
-
-dJonesL <- 
-  means %>%
-   # filter(season == 3, age == 1) %>%
-  #filter(age == 1) %>%
-    ungroup() #%>%
-    #dplyr::select(species,year,season,age,meanLen,sdLen)
-
-######
-# abundance
-# run 'r adjusted counts' above
-
-dJonesA <- 
-  allFishBySpeciesYOY %>% 
-  #  filter(riverOrderedGG == "West Brook", year >=2000, season == 3, isYOY == 2) %>%
-  filter(riverOrderedGG == "West Brook") %>% #, year >=2000, isYOY == 2) %>%
-    ungroup() %>%
-    dplyr::select(species,year,isYOY,season,nAllFishBySpeciesYOY,nAllFishBySpecies)
-
-######
-# phi
-# run 'r p(detection)' below
-# only has summer survival, may want annual survival
-
-dJonesP <- phi %>%
-  #filter(season == 2, isYOYGG == 'Adult') %>%
-  #filter(isYOYGG == 'Adult') %>%
-  dplyr::select(speciesGG,yearGG,season,isYOYGG,isYOY,phi,phiSD,phi25,phi975) %>%
-  rename(species = speciesGG, year = yearGG) %>%
-  mutate(species = as.character(species))
-
-dJonesAL <- left_join(dJonesA,dJonesL) 
-dJonesALT <-  left_join(dJonesAL,dJonesT)
-dJonesALTP <-  left_join(dJonesALT,dJonesP)
-
-##
-write.csv(dJonesALTP, file = 'D:/projects/linkedModels/dataForCarysJones/dJonesALTP_allSeasonsSpeciesAges.csv')
-```
-
-```{r mean gr graphs}
-cd$riverOrderedGG <- factor(cd$river,
-                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
-cd$seasonGG <- factor(cd$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-cd$speciesGG <- factor(cd$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-pd <- position_dodge(0.65)
-
-growthYOY <- 
-cd %>% group_by(speciesGG,species,riverOrderedGG,seasonGG,isYOY) %>%
-  filter( riverOrdered == 'west brook', year < 2016, species != 'ats' )  %>%
-  summarize(meanGR = mean(grLength, na.rm=T),
-            sdGR = sd(grLength, na.rm=T))
-
-ggplot(growthYOY%>%filter(isYOY == 1), aes(seasonGG,meanGR, color=speciesGG, shape=speciesGG) ) +
-  geom_point( position = pd, size = 3 ) +
-  geom_errorbar( aes(ymin = meanGR - sdGR, ymax = meanGR + sdGR), width = 0.33, position=pd ) +
-  ggtitle( "isYOY = 1" ) +
-  labs( x = "Season", y = "Growth rate (mm/d)" ) +
-  #scale_x_continuous( breaks = 1997:2015 ) +
-  scale_color_grey(start = 0.0, end = 0.5) +
-  theme_publication() +
-  theme( panel.grid.minor = element_blank() ) 
-#  facet_wrap( ~ riverOrderedGG)
-
-ggsave('figures/3Spp_WB/meanGRYOY1.png', width = 4, height = 4, units='in')
-
-ggplot(growthYOY%>%filter(isYOY == 2), aes(seasonGG,meanGR, color=speciesGG, shape=speciesGG) ) +
-  geom_point( position = pd, size = 3 ) +
-  geom_errorbar( aes(ymin = meanGR - sdGR, ymax = meanGR + sdGR), width = 0.33, position=pd ) +
-  ggtitle( "isYOY = 2" ) +
-  labs( x = "Season", y = "Growth rate (mm/d)" ) +
-  #scale_x_continuous( breaks = 1997:2015 ) +
-  scale_color_grey(start = 0.0, end = 0.5) +
-  theme_publication() +
-  theme( panel.grid.minor = element_blank() )
-
-ggsave('figures/3Spp_WB/meanGRYOY2.png', width = 4, height = 4, units='in')
-
-# get ratios of growth rates for the 2 species
-growthYOYWide <- growthYOY %>% ungroup() %>% select(-sdGR,-speciesGG) %>% spread(key=species,value=meanGR) %>%
-  mutate(ratio = bnt/bkt)
-
-mean(growthYOYWide$ratio)
-# 1.107745
-
-```
-
-```{r mean growth rate ratios}
-
-cd %>% group_by(species) %>% filter(riverOrderedGG == "West Brook",species != 'ats') %>%
-  summarize(meanGR = mean(grLength, na.rm=T),
-            sdGR = sd(grLength, na.rm=T))
-.197/.182#bnt/bkt
-
-cd %>% group_by(isYOY) %>% filter(riverOrderedGG == "West Brook",species != 'ats') %>%
-  summarize(meanGR = mean(grLength, na.rm=T),
-            sdGR = sd(grLength, na.rm=T))
-.145/.244 #adult/yoy
-.244/.145
-
-cd %>% group_by(season) %>% filter(riverOrderedGG == "West Brook",species != 'ats') %>%
-  summarize(meanGR = mean(grLength, na.rm=T),
-            sdGR = sd(grLength, na.rm=T))
-.434/((.114+.124+.067)/3)
-```
-
-```{r - why bad convergence for bnt in winter?}
-
-cd %>% filter() %>%
-  ggplot(aes(observedLength,grLength)) +
-  geom_point() +
-  facet_grid(season~species)
-
-```
-
-```{r questions}
-
-```
-
-```{r p(detection)}
-  # ggPInt <- array2df(ddD$sims.list$pBetaInt, label.x = "est")
-  # 
-  # ggPInt$chain <- rep(1:ddD$mcmc.info$n.chains, each = ddD$mcmc.info$n.samples/ddD$mcmc.info$n.chains)
-  # ggPInt$iter <- 1:as.numeric(ddD$mcmc.info$n.samples/ddD$mcmc.info$n.chains)
-  # 
-  # ggplot(filter(ggPInt), aes(iter,est)) + 
-  #   geom_point( aes(color=factor(chain)), size = 0.1 ) + 
-  #   ylim(-1.5,1.5) + 
-  #   facet_grid(d2+d4~d5+d3)
-
-  p <- array2df(ddD$q50$pBetaInt, label.x = "est") %>%
-    rename(est=est,isYOY=d1,species=d2,season=d3,river=d4,year=d5) %>% filter(river == 1, year > 3)
-
-  p$speciesGG <- factor(p$species, levels = 1:3, labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-  p$seasonGG <- factor(p$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-  p$isYOYGG <- factor(p$isYOY, levels = 1:2, labels = c("Juvenile","Adult"), ordered = T)
-  p$yearGG <- p$year + 1996
-
-  ggplot(p, aes(yearGG,invlogit(est),color=factor(speciesGG))) + 
-    geom_point(size = 2) + 
-    labs( x = 'Year', y = 'Probability of detection') +
-    scale_color_grey(start = 0.0, end = 0.85) +
-    theme_publication() +
-    facet_grid( seasonGG ~ isYOYGG)
-
-  # seem to need to run the ggsave outside of markdown by pasting into console
-   ggsave('figures/3Spp_WB/p.png', height = 11, width = 9, units = 'in')
-
-  p %>% group_by(species) %>%
-    summarize(meanP = mean(invlogit(est)),
-              sdP = sd(invlogit(est)))
-
-  p %>% group_by(season) %>%
-    summarize(meanP = mean(invlogit(est)))
-
-  ##################### 
-  phi <- array2df(ddD$q50$phiBetaInt, label.x = "est") %>%
-    rename(est=est,isYOY=d1,species=d2,season=d3,river=d4,year=d5) %>% filter(river == 1)#, year > 3)
-
-  phi$speciesGG <- factor(phi$species, levels = 1:3, labels = c("bkt", "bnt", "ats"), ordered = T)
-  phi$seasonGG <- factor(phi$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-  phi$isYOYGG <- factor(phi$isYOY, levels = 1:2, labels = c("Juvenile","Adult"), ordered = T)
-  phi$yearGG <- phi$year + 1996
-  phi$phi <- invlogit(phi$est)
-
-  # add in sd of phiBetaInt
-  phiSD <- array2df(ddD$sd$phiBetaInt, label.x = "estSD") %>%
-    rename(estSD=estSD,isYOY=d1,species=d2,season=d3,river=d4,year=d5) %>% filter(river == 1)#, year > 3)  
-  phiSD$phiSD <- invlogit(phiSD$estSD)
-
-    # add in q2.5 of phiBetaInt
-  phi25 <- array2df(ddD$q2.5$phiBetaInt, label.x = "est25") %>%
-    rename(est25=est25,isYOY=d1,species=d2,season=d3,river=d4,year=d5) %>% filter(river == 1)#, year > 3)  
-  phi25$phi25<- invlogit(phi25$est25)
-
-    # add in 97.5 of phiBetaInt
-  phi975 <- array2df(ddD$q97.5$phiBetaInt, label.x = "est975") %>%
-    rename(est975=est975,isYOY=d1,species=d2,season=d3,river=d4,year=d5) %>% filter(river == 1)#, year > 3)  
-  phi975$phi975 <- invlogit(phi975$est975)
-
-  phi <- left_join(phi, phiSD)
-  phi <- left_join(phi, phi25)
-  phi <- left_join(phi, phi975) 
-  phi <- phi %>% select(-estSD,-est25,-est975)
-
-  ggplot(phi %>% filter(year > 3), aes(yearGG,phi,color=factor(speciesGG))) +geom_point() + facet_grid(seasonGG ~ isYOYGG)
-
-  ggsave('figures/3Spp_WB/phi.png')
-
-  phi %>% group_by(species) %>%
-    summarize(meanPhi = mean(invlogit(est)),
-              sdPhi = sd(invlogit(est)))
-
-  phi %>% group_by(season) %>%
-    summarize(meanPhi = mean(invlogit(est)))
-
-# # of skunked sections    
-  tmp <- dddD[[2]] %>%
-           group_by(riverOrdered,year,season,section) %>%
-           summarize( s = sum(enc) ) %>%
-           filter( s == 0 )
-  table(tmp$riverOrdered,tmp$year,tmp$season)
-
-```
-
-```{r r-hat}
-
-rBkt <- unlist(mcmc[['bkt']]$Rhat)
-sort(rBkt[rBkt > 1.1])
-
-rBnt <- unlist(mcmc[['bnt']]$Rhat)
-sort(rBnt[rBnt > 1.1])
-mcmc[['bnt']]$Rhat$grBeta[,,,1]
-mcmc[['bnt']]$Rhat$grInt
-mcmc[['bnt']]$Rhat$grIntMu
-
-rHat <- list()
-rHat[['bkt']] <- array2df(mcmc[['bkt']]$Rhat$grBeta, levels = list(beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="rHat")
-rHat[['bnt']] <- array2df(mcmc[['bnt']]$Rhat$grBeta, levels = list(beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="rHat")
-rHat[['ats']] <- array2df(mcmc[['ats']]$Rhat$grBeta, levels = list(beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="rHat")
-
-# rHat[['bkt']] <- array2df(mcmc[['bkt']]$Rhat$grBeta, levels = list(beta=1:8,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="rHat")
-# rHat[['bnt']] <- array2df(mcmc[['bnt']]$Rhat$grBeta, levels = list(beta=1:8,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="rHat")
-# rHat[['ats']] <- array2df(mcmc[['ats']]$Rhat$grBeta, levels = list(beta=1:8,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="rHat")
-
-hist(rHat[["bkt"]]$rHat)
-hist(rHat[["bnt"]]$rHat)
-rHat[["bnt"]] %>% filter( rHat >1.1 )
-
-ggplot(rHat[["bkt"]], aes(beta,rHat)) +
-  geom_point() +
-  facet_grid(river+isYOY~season)
-
-```
-
-```{r mcmc output analysis-grInt}
-
-grInt <- list()
-grInt[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$grInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int") %>% mutate(species = 'bkt')
-grInt[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$grInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int") %>%   mutate(species = 'bnt')
-grInt[['ats']] <- array2df(mcmc[['ats']]$sims.list$grInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int") %>% mutate(species = 'ats')
-
-ggplot( grInt[['bkt']] %>% filter(river %in% c('west brook','wb jimmy')), aes(int)) +
-  geom_freqpoly() +
-  geom_freqpoly(aes((int)),color = 'brown', data = grInt[['bnt']]) +
-  geom_freqpoly(aes((int)),color = 'lightblue', data = grInt[['ats']]) +
- # xlim(-0.5,1.5) +
-  theme_publication() +
-  facet_grid(season ~ river + isYOY)
-
-#grIntDF <- rbind(grInt[['bkt']],grInt[['bnt']],grInt[['ats']])
-#tmp <- grIntDF %>% filter(river == 'west brook')
-#save(tmp , file= "data/out/grIntDFAllIters.RData") #for Ron
-
-##################################################
-# as range plot
-
-plotHPD <- function(sppToPlot){
-
-  nRivers <- list()
-  nRivers[['bkt']] <- 4; nRivers[['bnt']] <- 2; nRivers[['ats']] <- 1
-
-  probs <- c(0.025,0.1,0.5,0.9,0.975)
-  q <- grInt[[sppToPlot]] %>%
-    group_by(isYOY,season,river) %>%
-    summarize(
-      p = list(probs),
-      q = list(quantile(int,probs))
-    ) %>%
-    unnest() %>%
-    spread(key = p, value = q)
-
-  p <- array2df(mcmc[[sppToPlot]]$mean$grInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="int") %>%
-    mutate( season = as.numeric(season) ) %>%
-    left_join(.,meanSampleIntervalMean, by = 'season')
-  p$intByDay <- p$int#/p$meanSampleInterval
-  p$intByDay <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intByDay)
-  p$seasonGG <- factor(p$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-  p$parameter <- p$seasonGG
-
-  pUp <- array2df(mcmc[[sppToPlot]]$q97.5$grInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="intUp") %>%
-    mutate( season = as.numeric(season) ) 
-  pDown <- array2df(mcmc[[sppToPlot]]$q2.5$grInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="intDown") %>%
-    mutate( season = as.numeric(season) )
-
-  p <- left_join(p,pUp); p <- left_join(p,pDown)
-
-  p$intByDayUp <- p$intUp#/p$meanSampleInterval; 
-  p$intByDayDown <- p$intDown#/p$meanSampleInterval
-  p$intByDay <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intByDay)
-  p$intByDayUp <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intByDayUp)
-  p$intByDayDown <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intByDayDown)
-
-  p$seasonGG <- factor(p$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-  #print(c(min(p$intByDayDown,na.rm=T),max(p$intByDayDown,na.rm=T)))
-
-  #save(p, file= paste0("data/out/allInts_",sppToPlot,".RData"))  #for projections and tables
-
-    gg <- ggplot(p %>% filter(river=='west brook'), aes(x = intByDay, y = reorder(parameter,intByDay,na.rm=TRUE), color = factor(isYOY), shape = factor(isYOY))) + 
-        theme_publication() +  
-        geom_point(size = 4) +
-        geom_segment(aes(x = intByDay, xend = intByDayUp, yend = reorder(parameter, intByDay)), 
-        size = 0.5, arrow = arrow(angle = 90, length = unit(0.05, "inches"))) + 
-        geom_segment(aes(x = intByDay, xend = intByDayDown, yend = reorder(parameter, intByDay)), 
-        size = 0.5,arrow = arrow(angle = 90, length = unit(0.05, "inches"))) +
-        xlim(c(-0.00,0.55)) +
-        xlab("Mean daily growth (mm/d)") + ylab("Season") +
-        scale_color_grey(start = 0, end = 0.65) 
-       # ggtitle(sppToPlot)
-
-    return(gg)
-}
-
-plotHPD('bkt'); ggsave('figures/3Spp_WB/HDP_bkt.png', height = 5, width = 5, scale = 1)
-plotHPD('bnt'); ggsave('figures/3Spp_WB/HDP_bnt.png', height = 5, width = 5, scale = 1)
-plotHPD('ats'); ggsave('figures/3Spp_WB/HDP_ats.png', height = 5, width = 5, scale = 1)
-```
-
-```{r sigmaInt}
-
-sigmaInt <- list()
-sigmaInt[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$sigmaInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int")
-sigmaInt[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$sigmaInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int")
-sigmaInt[['ats']] <- array2df(mcmc[['ats']]$sims.list$sigmaInt, levels = list(iter=NA,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int")
-
-bw <- 0.05
-ggplot( sigmaInt[['bkt']] %>% filter(river %in% c('west brook','wb jimmy')), aes(int)) +
-  geom_freqpoly(binwidth = bw) +
-  geom_freqpoly(aes(int),binwidth = bw,color = 'brown', data = sigmaInt[['bnt']]) +
-  geom_freqpoly(aes(int),binwidth = bw,color = 'lightblue', data = sigmaInt[['ats']]) +
-  #xlim(-4.5,-1) +
-  xlim(-5,2.5) +
-  theme_publication() +
-  facet_grid(season ~ river + isYOY)
-
-##################################################
-# as range plot
-
-plotHPDSigma <- function(sppToPlot){
-
-  nRivers <- list()
-  nRivers[['bkt']] <- 4; nRivers[['bnt']] <- 2; nRivers[['ats']] <- 1
-
-  probs <- c(0.025,0.1,0.5,0.9,0.975)
-  q <- sigmaInt[[sppToPlot]] %>%
-    group_by(isYOY,season,river) %>%
-    summarize(
-      p = list(probs),
-      q = list(quantile(int,probs))
-    ) %>%
-    unnest() %>%
-    spread(key = p, value = q)
-
-  p <- array2df(mcmc[[sppToPlot]]$mean$sigmaInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="int") %>%
-    mutate( season = as.numeric(season) ) %>%
-    left_join(.,meanSampleIntervalMean, by = 'season')
- # p$intByDay <- p$int/p$meanSampleInterval
-#  p$intByDay <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intByDay)
-  p$seasonGG <- factor(p$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-  p$parameter <- p$seasonGG
-
-  pUp <- array2df(mcmc[[sppToPlot]]$q97.5$sigmaInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="intUp") %>%
-    mutate( season = as.numeric(season) ) 
-  pDown <- array2df(mcmc[[sppToPlot]]$q2.5$sigmaInt, levels = list(isYOY=c(1,2),season=1:4,river=riverOrderedIn[1:nRivers[[sppToPlot]]]), label.x="intDown") %>%
-    mutate( season = as.numeric(season) )
-
-  p <- left_join(p,pUp); p <- left_join(p,pDown)
-
- # p$intByDayUp <- p$intUp/p$meanSampleInterval; p$intByDayDown <- p$intDown/p$meanSampleInterval
-  p$int <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$int)
-  p$intUp <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intUp)
-  p$intDown <- ifelse(p$isYOY == 1 & p$season == 2, NA, p$intDown)
-
-  #print(c(min(p$intDown,na.rm=T),max(p$intDown,na.rm=T)))
-
-    gg <- ggplot(p %>% filter(river == 'west brook'), aes(x = exp(int), y = reorder(parameter,int,na.rm=TRUE), color = factor(isYOY), shape = factor(isYOY))) + 
-        theme_publication() +  
-        geom_point(size = 3) +
-        geom_segment(aes(x = exp(int), xend = exp(intUp), yend = reorder(parameter, int)), 
-        size = 0.5, arrow = arrow(angle = 90, length = unit(0.05, "inches"))) + 
-        geom_segment(aes(x = exp(int), xend = exp(intDown), yend = reorder(parameter, int)), 
-        size = 0.5,arrow = arrow(angle = 90, length = unit(0.05, "inches"))) +
-        #xlim(c(-0.0015,0.01)) +
-        xlab("Standard deviation of daily growth (mm/d)") + ylab("Season") +
-        scale_color_grey(start = 0, end = 0.65) 
-       # ggtitle(sppToPlot)
-
-    return(gg)
-}
-
-plotHPDSigma('bkt'); ggsave('figures/3Spp_WB/HDPSigma_bkt.png', height = 5, width = 5, scale = 1)
-plotHPDSigma('bnt'); ggsave('figures/3Spp_WB/HDPSigma_bnt.png', height = 5, width = 5, scale = 1)
-plotHPDSigma('ats'); ggsave('figures/3Spp_WB/HDPsigma_ats.png', height = 5, width = 5, scale = 1)
-```
-
-```{r grInt}
-#riverOrderedIn <- factor(c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),labels = c("west brook","wb jimmy","wb mitchell","wb obear"), ordered = T)
-
-grIntMu <- list()
-grIntMu[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$grIntMu, levels = list(iter=NA,season=1:4), label.x="int")
-grIntMu[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$grIntMu, levels = list(iter=NA,season=1:4), label.x="int")
-grIntMu[['ats']] <- array2df(mcmc[['ats']]$sims.list$grIntMu, levels = list(iter=NA,season=1:4), label.x="int")
-
-bw <- 0.05
-ggplot( grIntMu[['bkt']] , aes(int)) +
-  geom_freqpoly(binwidth = bw, size=1.2) +
-  geom_freqpoly(aes(int),binwidth = bw,color = 'brown', data = grIntMu[['bnt']]) +
-  geom_freqpoly(aes(int),binwidth = bw,color = 'lightblue', data = grIntMu[['ats']]) +
-  #xlim(-4.5,-1) +
-  xlim(-0.25,0.75) +
- # ylim(0,2500) +
-  theme_publication() +
-  facet_wrap(~season )
-
-```
-
-```{r mcmc output analysis-grBeta}
-
-# get all betas in a DF
-grBeta <- list()
-grBeta[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$grBeta, levels = list(iter=NA,beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int") %>%
-  mutate( species = 'bkt')
-grBeta[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$grBeta, levels = list(iter=NA,beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int") %>%
-  mutate( species = 'bnt')
-grBeta[['ats']] <- array2df(mcmc[['ats']]$sims.list$grBeta, levels = list(iter=NA,beta=1:7,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int") %>%
-  mutate( species = 'ats')
-
-grBetaDF1 <- rbind(grBeta[['bkt']],grBeta[['bnt']],grBeta[['ats']])
-
-grBetaBNT <- list()
-grBetaBNT[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$grBetaBNT, levels = list(iter=NA,beta=8:9,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int") %>%
-  mutate( species = 'bkt')
-grBetaBNT[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$grBetaBNT, levels = list(iter=NA,beta=8:9,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int") %>%
-  mutate( species = 'bnt')
-grBetaBNT[['ats']] <- array2df(mcmc[['ats']]$sims.list$grBetaBNT, levels = list(iter=NA,beta=8:9,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int") %>%
-  mutate( species = 'ats')
-
-grBetaBNTDF <- rbind(grBetaBNT[['bkt']],grBetaBNT[['bnt']],grBetaBNT[['ats']])
-
-grBetaATS <- list()
-grBetaATS[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$grBetaATS, levels = list(iter=NA,beta=10,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int") %>%
-  mutate( species = 'bkt')
-grBetaATS[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$grBetaATS, levels = list(iter=NA,beta=10,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int") %>%
-  mutate( species = 'bnt')
-grBetaATS[['ats']] <- array2df(mcmc[['ats']]$sims.list$grBetaATS, levels = list(iter=NA,beta=10,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int") %>%
-  mutate( species = 'ats')
-
-grBetaATSDF <- rbind(grBetaATS[['bkt']],grBetaATS[['bnt']],grBetaATS[['ats']])
-
-grBetaDF <- rbind(grBetaDF1,grBetaBNTDF,grBetaATSDF)
-
-#tmp <- grBetaDF %>% filter(river == 'west brook')
-#save(tmp , file= "data/out/grBetaDFAllIters.RData") #for Ron
-##################################################
-# as range plot
-
-  nRivers <- list()
-  nRivers[['bkt']] <- 4; nRivers[['bnt']] <- 2; nRivers[['ats']] <- 1
-  speciesList <- list('bkt','bnt','ats')
-
-  probs <- c(0.025,0.1,0.5,0.9,0.975)
-  nSpecies <- 3
-
-  q <- grBetaDF %>%
-    group_by(beta,isYOY,season,river,species) %>%
-    summarize(
-      p = list(probs),
-      q = list(quantile(int,probs))
-    ) %>%
-    unnest() %>%
-    spread(key = p, value = q) %>%
-    data.frame() %>%
-    rename( median = X0.5, lo = X0.025, hi = X0.975 )
-
-  q <- q %>% filter( !(season == 2 & isYOY == 0))
-
-q$riverGG <- factor(q$river,
-                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("WB","OL","OS","IL"), ordered = T)
-q$seasonGG <- factor(q$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-q$speciesGG <- factor(q$species, levels = c('bkt','bnt','ats'), labels = c("BKT", "BNT", "ATS"), ordered = T)
-q$betaGG <- factor(q$beta, levels = 1:10, 
-                   #levels = c("Temp","Flow","Temp*Flow","Temp^2","Flow^2","BKT","Length","BNT","BNT*BKT","ATS"), # make sure this order matches the nimble code
-                   labels = c("Temp","Flow","Temp*Flow","Temp^2","Flow^2","BKT","Length","BNT","BNT*BKT","ATS"),ordered = T)
-q$parameter = q$speciesGG
-q$parameterOrd <- factor(q$parameter, 
-                         levels = c( "ATS", "BNT", "BKT"), 
-                         labels = c( "WB_ATS", "WB_BNT", "WB_BKT"), ordered = T)
-
-  q <- q %>% filter( !(betaGG == 'ATS' & (riverGG == "OS" | riverGG == "IL" | riverGG == "OL")) )
-  q <- q %>% filter( !(betaGG == 'BNT' & (riverGG == "OS" | riverGG == "IL")) )
-  q <- q %>% filter( !(betaGG == 'BNT*BKT' & (riverGG == "OS" | riverGG == "IL")) )
-#  q <- q %>% filter( !((riverGG == "OS" | riverGG == "IL")) )
-
-  save(q, file= "data/out/allBetas.RData")  #for projections and tables
-
-  #reorder(parameter,median,na.rm=TRUE)
-(min(q$median))
-  (max(q$median))
-#  q[q$median < -0.13,]
-
-    ggplot(q %>% filter(river == 'west brook', species != 'ats'), aes(x = median, y = parameterOrd, color = factor(species), shape = factor(isYOY))) + 
-        geom_vline(xintercept = 0,color='grey') +
-        theme_publication() +  
-        geom_point(size = 4) +
-        geom_segment(aes(x = median, xend = hi, yend = parameterOrd), 
-        size = 0.5, arrow = arrow(angle = 90, length = unit(0.1, "inches"))) + 
-        geom_segment(aes(x = median, xend = lo, yend = parameterOrd), 
-        size = 0.5,arrow = arrow(angle = 90, length = unit(0.1, "inches"))) +
-     #   xlim(c(-0.15,0.11)) +
-        scale_x_continuous( breaks=c(-0.1,0,0.1)) +
-        xlab("Parameter estimate") + theme(axis.text.y = element_blank(), axis.title.y = element_blank()) +
-        scale_color_grey(start = 0, end = 0.5)  +
-        facet_grid(seasonGG~betaGG)
-
-ggsave('figures/3Spp_WB/grBetas.png', height = 8, width = 10)
-
-save(betas,file = 'D:/projects/linkedModels/data/out/nSlopesTidy.RData')
-
-```
-
-```{r table of Ints}
-load(file = "data/out/allInts_bkt.RData") #loads p
-pBKT <- p %>% mutate(species = 'bkt')
-
-load(file = "data/out/allInts_bnt.RData") #loads p
-pBNT <- p %>% mutate(species = 'bnt')
-
-pp <- rbind(pBKT,pBNT)
-
-intsForTable <- pp %>% 
-  filter(river == 'west brook', species %in% c('bkt','bnt')) %>%
-  arrange(species,isYOY,seasonGG) %>%
-  mutate(lo = round(intDown,3),mean = round(int,3),hi = round(intUp,3)) %>%
-  select(species,isYOY,seasonGG,intDown,int,intUp)
-
-write.csv(intsForTable,file = 'D:/projects/linkedModels/data/out/intsForTableAgeGroup.csv')
-```
-
-```{r table of betas}
-
-load(file= "data/out/allBetas.RData") #loads q
-
-betasForTable <- q %>% 
-  filter(river == 'west brook', species %in% c('bkt','bnt')) %>%
-  arrange(species,betaGG,isYOY,seasonGG,betaGG) %>%
-  mutate(lo = round(lo,3),median = round(median,3),hi = round(hi,3)) %>%
-  select(species,betaGG,isYOY,seasonGG,betaGG,lo,median,hi)
-
-write.csv(betasForTable,file = 'D:/projects/linkedModels/data/out/betasForTableAgeGroup.csv')
-```
-
-```{r mcmc output analysis-grBetaSigma}
-
-# get all betas in a DF
-grBetaSigma <- list()
-grBetaSigma[['bkt']] <- array2df(mcmc[['bkt']]$sims.list$sigmaBeta, levels = list(iter=NA,beta=1:3,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:4]), label.x="int") %>%
-  mutate( species = 'bkt')
-grBetaSigma[['bnt']] <- array2df(mcmc[['bnt']]$sims.list$sigmaBeta, levels = list(iter=NA,beta=1:3,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:2]), label.x="int") %>%
-  mutate( species = 'bnt')
-grBetaSigma[['ats']] <- array2df(mcmc[['ats']]$sims.list$sigmaBeta, levels = list(iter=NA,beta=1:3,isYOY=c(0,1),season=1:4,river=riverOrderedIn[1:1]), label.x="int") %>%
-  mutate( species = 'ats')
-
-grBetaSigmaDF <- rbind(grBetaSigma[['bkt']],grBetaSigma[['bnt']],grBetaSigma[['ats']])
-
-##################################################
-# as range plot
-
-  nRivers <- list()
-  nRivers[['bkt']] <- 4; nRivers[['bnt']] <- 2; nRivers[['ats']] <- 1
-  speciesList <- list('bkt','bnt')
-
-  probs <- c(0.025,0.1,0.5,0.9,0.975)
-  nSpecies <- 2
-
-  q <- grBetaSigmaDF %>%
-    group_by(beta,isYOY,season,river,species) %>%
-    summarize(
-      p = list(probs),
-      q = list(quantile(int,probs))
-    ) %>%
-    unnest() %>%
-    spread(key = p, value = q) %>%
-    data.frame() %>%
-    rename( median = X0.5, lo = X0.025, hi = X0.975 )
-
-  q <- q %>% filter( !(season == 2 & isYOY == 0))
-
-q$riverGG <- factor(q$river,
-                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("WB","OL","OS","IL"), ordered = T)
-q$seasonGG <- factor(q$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-q$speciesGG <- factor(q$species, levels = c('bkt','bnt','ats'), labels = c("BKT", "BNT", "ATS"), ordered = T)
-q$betaGG <- factor(q$beta, levels = 1:10, labels = c("Temp", "Flow", "Temp*Flow","Temp^2","Flow^2","BKT","Length","BNT","BNT*BKT","ATS"), ordered = T)
-q$parameter = q$speciesGG
-q$parameterOrd <- factor(q$parameter, 
-                         levels = c( "ATS", "BNT", "BKT"), 
-                         labels = c( "WB_ATS", "WB_BNT", "WB_BKT"), ordered = T)
-
-  q <- q %>% filter( !(betaGG == 'ATS' & (riverGG == "OS" | riverGG == "IL" | riverGG == "OL")) )
-  q <- q %>% filter( !(betaGG == 'BNT' & (riverGG == "OS" | riverGG == "IL")) )
-  q <- q %>% filter( !(betaGG == 'BNT*BKT' & (riverGG == "OS" | riverGG == "IL")) )
-#  q <- q %>% filter( !((riverGG == "OS" | riverGG == "IL")) )
-
-#reorder(parameter,median,na.rm=TRUE)
-(min(q$median))
-  (max(q$median))
-#  q[q$median < -0.13,]
-
-    ggplot(q %>% filter(river == 'west brook', species != 'ats'), aes(x = median, y = parameterOrd, color = factor(speciesGG), shape = factor(isYOY))) + 
-        geom_vline(xintercept = 0,color='black') +
-        theme_publication() +  
-        geom_point(size = 3) +
-        geom_segment(aes(x = median, xend = hi, yend = parameterOrd), 
-        size = 0.5, arrow = arrow(angle = 90, length = unit(0.05, "inches"))) + 
-        geom_segment(aes(x = median, xend = lo, yend = parameterOrd), 
-        size = 0.5,arrow = arrow(angle = 90, length = unit(0.05, "inches"))) +
-     #   xlim(c(-0.35,0.5)) +
-        xlab("Parameter estimate") + theme(axis.text.y = element_blank(), axis.title.y = element_blank()) +
-        scale_color_grey(start = 0, end = 0.5)  +
-        facet_grid(seasonGG~betaGG)
-
-ggsave('figures/3Spp_WB/grBetaSigmas.png', height = 8, width = 8)
-
-```
-
-```{r ind RE among spp}
-
-grIndRE <- list()
-grIndRE[['bkt']] <- as.data.frame(mcmc[['bkt']]$q50$grIndRE); names(grIndRE[['bkt']]) <- 'value'
-grIndRE[['bnt']] <- as.data.frame(mcmc[['bnt']]$q50$grIndRE); names(grIndRE[['bnt']]) <- 'value'
-grIndRE[['ats']] <- as.data.frame(mcmc[['ats']]$q50$grIndRE); names(grIndRE[['ats']]) <- 'value'
-
-table(mcmc[['bkt']]$q50$grIndRE==0)/length(mcmc[['bkt']]$q50$grIndRE)
-table(mcmc[['bnt']]$q50$grIndRE==0)/length(mcmc[['bnt']]$q50$grIndRE)
-table(mcmc[['ats']]$q50$grIndRE==0)/length(mcmc[['ats']]$q50$grIndRE)
-
-ggplot( grIndRE[[1]], aes(value,..density..) ) +
-  geom_freqpoly(binwidth = 0.00025) +
-  geom_freqpoly(aes(value), color = "brown", binwidth = 0.00025, data = grIndRE[[2]]) +
-  geom_freqpoly(aes(value), color = "lightblue", binwidth = 0.00025, data = grIndRE[[3]]) +
-  ylim(c(0,100))
-
-```
-
-```{r ind trajectories}
-
-```
-
-```{r predictions}
-# 
-#####run if get new predictions - takes a while to run ####
-
- # load('D:/projects/linkedModels/data/out/P_ForMike_bkt.RData')
- # pBoth$species <- 'bkt'
- # pAll <- pBoth; rm(pBoth)
- # 
- # load('D:/projects/linkedModels/data/out/P_ForMike_bnt.RData')
- # pBoth$species <- 'bnt'
- # pAll <- rbind(pAll,pBoth); rm(pBoth)
- # 
- # load('D:/projects/linkedModels/data/out/P_ForMike_ats.RData')
- # pBoth$species <- 'ats'
- # pAll <- rbind(pAll,pBoth); rm(pBoth)
- # 
- # pred <- pAll %>% group_by(species,len,flow,temp,cBKT,cBNT,cATS,isYOY,season,river) %>%
- #   summarize(meanPredGR = mean(predGr, na.rm=T),
- #             sdPredGR =     sd(predGr, na.rm=T),
- #             meanPredGRSigma = mean(predGrSigma, na.rm=T),
- #             sdPredGRSigma =     sd(predGrSigma, na.rm=T),
- #             meanPredGRCV = mean(predCV, na.rm=T),
- #             sdPredGRCV =     sd(predCV, na.rm=T)) %>%
- #    ungroup()
- # 
- # rm(pAll)
- # save(pred,file = 'D:/projects/linkedModels/data/out/P_ForMike_allMeans.RData')
-
-load('D:/projects/linkedModels/data/out/P_ForMike_allMeans.RData')
-
-pred <- pred %>% filter( !(isYOY == 0 & season == 2) ) 
-#pred <- pred %>% filter( !(isYOY == 0 & river == 'wb jimmy' & season == 2) )
-#graph function, in analyzeOutputFunctions.R
-
-pred$riverGG <- factor(pred$river,
-                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
-pred$seasonGG <- factor(pred$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-pred$speciesGG <- factor(pred$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-plotPredMeans(pred, "meanPredGR", "len", 0)
-plotPredMeans(pred %>% filter(river == 'west brook'), "meanPredGR", "len", 1)
-
-plotPredMeans(pred, "meanPredGR", "flow", 0)
-plotPredMeans(pred, "meanPredGR", "flow", 1)
-
-plotPredMeans(pred, "meanPredGR", "temp", 0)
-plotPredMeans(pred, "meanPredGR", "temp", 1)
-
-plotPredMeans(pred, "meanPredGR", "cBKT", 0)
-plotPredMeans(pred, "meanPredGR", "cBKT", 1)
-
-plotPredMeans(pred, "meanPredGR", "cBNT", 0)
-plotPredMeans(pred, "meanPredGR", "cBNT", 1)
-
-plotPredMeans(pred, "meanPredGR", "cATS", 0)
-plotPredMeans(pred, "meanPredGR", "cATS", 1)
-
-plotPredMeans(pred, "meanPredGR", c("temp","flow"), 0)
-plotPredMeans(pred, "meanPredGR", c("temp","flow"), 1)
-
-plotPredMeans(pred, "meanPredGR", c("flow","temp"), 0)
-plotPredMeans(pred, "meanPredGR", c("flow","temp"), 1)
-
-ggsave('figures/3Spp_WBtmp.png', width = 9, height = 7, units='in')
-```
-
-```{r mean prediction graphs for meanPredGR}
-
-all <- c('len','temp','flow','cBKT','cBNT','cATS')
-allLabels <- c('Length (mm)','Temperature (C)','Flow (m3/s)','Brook Trout abundance','Brown Trout abundance','Atlantic Salmon abundance')
-
-all <- c('len','cBKT','temp','cBNT','flow','cATS')
-allLabels <- c('Length (mm)','Brook Trout abundance','Stream temperature (C)','Brown Trout abundance','Stream flow (m3/s)','Atlantic Salmon abundance')
-
-remFacetLabel <- c(3,4,5,6)
-
-selectSingleVar <- function(p, all, depVar, varsToPlot, isYOYGG){
-
-  notPlot <- NA
-  notPlot[1] <- all[!(all %in% varsToPlot)][1]
-  notPlot[2] <- all[!(all %in% varsToPlot)][2]
-  notPlot[3] <- all[!(all %in% varsToPlot)][3]
-  notPlot[4] <- all[!(all %in% varsToPlot)][4]
-  notPlot[5] <- all[!(all %in% varsToPlot)][5]
-
-  pGG <- p %>% filter(isYOY == isYOYGG, 
-                        eval(as.name(notPlot[1])) == 0, 
-                        eval(as.name(notPlot[2])) == 0, 
-                        eval(as.name(notPlot[3])) == 0,                                         
-                        eval(as.name(notPlot[4])) == 0, 
-                        eval(as.name(notPlot[5])) == 0
-                      )#%>%
-              # distinct(species, isYOY, riverGG, seasonGG, speciesGG, .keep_all = TRUE)
-
-  return(pGG)  
-}
-
-##### by species
-
-isYOYGG <- 0
-riverGGToPlot <- "West Brook" #"Open Large"
-outputToPlot <- "meanPredGR"
-
-ggSingle <- list()
-singleDat <- list()
-for(i in 1:length(all))
-  #https://stackoverflow.com/questions/31993704/storing-ggplot-objects-in-a-list-from-within-loop-in-r
-      local({ i <- i
-  singleDat[[i]] <- selectSingleVar(pred, all, outputToPlot, all[i], isYOYGG)
-  print(c(i,all[i]))
- # print(str(singleDat[[i]]))
-
-  tmp <- singleDat[[i]] %>% filter(riverGG == riverGGToPlot)
-  ggSingle[[i]] <<- ggplot( tmp %>% filter(species != 'ats'), aes( eval(as.name(all[i])), eval(as.name(outputToPlot)), color=speciesGG ) ) +
-      geom_line( size = 1.225 ) +
-      theme_bw() +
-      theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), 
-            axis.title.y = element_blank(),
-            legend.position="none") +
-      geom_hline(yintercept = 0) +
-      ylim(-0.15,0.8) +
-      scale_x_continuous(allLabels[i]) +
-      scale_color_grey(start = 0, end = 0.5)  +
-      facet_grid(~seasonGG)
-
-  if( i %in% remFacetLabel ) {
-    ggSingle[[i]] <<- ggSingle[[i]] +
-    theme(
-      strip.background = element_blank(),
-      strip.text.x = element_blank()
-    )
-  }
-})
-
-grid.arrange(grobs=ggSingle, left = "Growth rate (mm/d)")
-# save from rStudio as export, width=900, height=500
-# ggsave(gridPred,'figures/3Spp_WB/predGRYOY1.png', width = 9, height = 7, units='in')
-
-```
-
-```{r mean prediction graphs for meanPredGRSigma}
-
-all <- c('temp','flow')
-allLabels <- c('Stream temperature (C)','Stream flow (m3/s)')
-
-remFacetLabel <- c(2)
-
-selectSingleVarSigma <- function(p, all, depVar, varsToPlot, isYOYGG){
-
-  notPlot <- NA
-  notPlot[1] <- all[!(all %in% varsToPlot)][1]
-
-  pGG <- p %>% filter(isYOY == isYOYGG, 
-                        eval(as.name(notPlot[1])) == 0,
-                      len == 0, cBKT == 0, cBNT == 0, cATS == 0
-                      )#%>%
-              # distinct(species, isYOY, riverGG, seasonGG, speciesGG, .keep_all = TRUE)
-
-  return(pGG)  
-}
-
-##### by species
-
-isYOYGG <- 1
-riverGGToPlot <- "West Brook" #
-outputToPlot <- "meanPredGRSigma"
-
-ggSingle <- list()
-singleDat <- list()
-for(i in 1:length(all))
-  #https://stackoverflow.com/questions/31993704/storing-ggplot-objects-in-a-list-from-within-loop-in-r
-      local({ i <- i
-  singleDat[[i]] <- selectSingleVarSigma(pred, all, outputToPlot, all[i], isYOYGG)
-  print(c(i,all[i]))
- # print(str(singleDat[[i]]))
-
-  tmp <- singleDat[[i]] %>% filter(riverGG == riverGGToPlot, species != 'ats')
-  ggSingle[[i]] <<- ggplot( tmp, aes( eval(as.name(all[i])), eval(as.name(outputToPlot)), color=speciesGG ) ) +
-      geom_line( size = 1.225 ) +
-      theme_bw() +
-      theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), 
-            axis.title.y = element_blank(),
-            legend.position="none") +
-      geom_hline(yintercept = 0) +
-    #  ylim(-0.15,0.8) +
-      scale_x_continuous(allLabels[i]) +
-      scale_color_grey(start = 0, end = 0.5)  +
-      facet_grid(~seasonGG)
-
-  if( i %in% remFacetLabel ) {
-    ggSingle[[i]] <<- ggSingle[[i]] +
-    theme(
-      strip.background = element_blank(),
-      strip.text.x = element_blank()
-    )
-  }
-})
-
-# add thrid row to graph for interaction of tem and flow?
-
-# run this and export
-grid.arrange(grobs=ggSingle, left = "Standard deviation of growth rate (mm/d)")
-
-# 0 = 875,600
-# 1 = 875,500
-
-#ggsave(ggCV, paste0('figures/3Spp_WBpredGRCV_YOY',isYOYGG,'_',riverGGToPlot,'.png'), width = 9, height = 7, units='in')
-
-```
-
-```{r flow*temp interaction plots}
-
-###############################
-# flow*temp interaction plots
-isYOYGG <- 0
-riverGGToPlot <- "West Brook" #"Open Large" #
-outputToPlot <- "meanPredGRSigma" #"meanPredGR" #
-
-interact <- pred %>% filter(isYOY == isYOYGG,
-                            riverGG == riverGGToPlot,
-                            len == 0, cBKT == 0, cBNT == 0, cATS == 0 ) %>%
-                     mutate( temp_spp = paste0(temp,speciesGG))
-
-    ggplot( interact %>% filter(species != 'ats'), aes( flow, eval(as.name(outputToPlot)), group=temp_spp, linetype=speciesGG ) ) +
-      geom_line( aes(alpha = temp + 3), size = 1.225 ) +
-      theme_bw() +
-      theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), 
-          #  axis.title.y = element_blank()
-            legend.position="none"
-            ) +
-      geom_hline(yintercept = 0) +
-    #  ylim(-0.15,0.8) +
-      scale_x_continuous('Stream flow') +
-      scale_y_continuous("Standard deviation of growth rate (mm/d)") +
-      #scale_color_manual(values=c('blue','darkgreen'))  +
-      facet_grid(~seasonGG)
-
-#yoy0 = 875,350
-#yoy1 = 875,280
-```
-
-```{r bkt*bnt interaction plots}
-
-###############################
-# flow*temp interaction plots
-isYOYGG <- 1
-riverGGToPlot <- "West Brook" #
-outputToPlot <- "meanPredGR" #"meanPredGRSigma"
-
-interact <- pred %>% filter(isYOY == isYOYGG,
-                            riverGG == riverGGToPlot,
-                            len == 0, temp == 0, flow == 0, cATS == 0 ) %>%
-                     mutate( bnt_spp = paste0(cBNT,speciesGG))
-
-    ggplot( interact %>% filter(species != 'ats'), aes( cBKT, eval(as.name(outputToPlot)), group=bnt_spp, linetype=speciesGG ) ) +
-      geom_line( aes(alpha = cBNT + 3), size = 1.225 ) +
-      theme_bw() +
-      theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), 
-           # axis.title.y = element_blank()
-            legend.position="none"
-            ) +
-      geom_hline(yintercept = 0) +
-    #  ylim(-0.15,0.8) +
-      scale_x_continuous('Brook Trout abundance') +
-      scale_y_continuous("Growth rate (mm/d)") +
-      facet_grid(~seasonGG)
-
-# for saving from export
-#yoy0 = 875,350
-#yoy1 = 875,275
-
-```
-
-```{r projections}
-# run 'Load data' chunk at top to get the 'mcmc' object
-
-#data for running the simulation
-simData <- list(
-  nReps = 100, #num of mcmc iterations to grab
-  numInd = 500 , # number of individual to simulate
-  nOccYOY = 3,
-  nOccAdult = 7,
-  riverIndex = 1, #Levels: west brook < wb jimmy < wb mitchell < wb obear
-  minSD = -1.5, #range and step for independent variable values in the sim
-  maxSD = 1.5, 
-  step = 1.5
-)
-simData$nOcc <- simData$nOccYOY + simData$nOccAdult
-simData$indVarSeq = seq(simData$minSD,simData$maxSD,simData$step)
-
-stats3Spp <- grStats3Spp <- anov <- elapsed <- lm <- relImp <- list()
-spp <- c('bkt','bnt') #,'ats')
-
-for( speciesForProj in spp ){
-
-  # get mean on mean length by cohort for the fall yoy sample
-  obsLenStats <-  
-    ddGIn2[[speciesForProj]] %>%
-      group_by(cohort,ageInSamples) %>%
-      summarise(lenMean = mean(observedLength, na.rm=TRUE),
-                lenSD = sd(observedLength, na.rm=TRUE)) %>%
-      filter(ageInSamples == 1) %>%
-      ungroup() %>%
-      summarise(lenMeanMean = mean(lenMean, na.rm=TRUE),
-                lenSDMean = mean(lenSD, na.rm=TRUE))
-
-  simData$species = speciesForProj
-  simData$meanInit = obsLenStats$lenMeanMean
-  simData$sdInit = obsLenStats$lenSDMean
-  simData$nSamplesAvail = mcmcInfo$nChains * mcmcInfo$nSamples
-
-  simData <- updateSimData(simData) # add #season, itersForSim etc data to simData
-  nB <- getNBetas() # get number of betas from mcmc[]
-  m <- createMatrices()
-
-  indVars <- getIndVars() # get independent variables table, -1.5, 0, 1.5
-
-  startTime <- Sys.time()
-  sizeStats <- sizeStats2 <- grStats <- grStats2 <- NULL
-  #loop over conditions
-  for(testRow in indVars$rowNum){
-    project <- projectSizes(m,simData,testRow)
-
-    sizeMat <- project$sizeMat
-    grMat <- project$grMat
-
-    sizeStats1 <- summarizeLens(sizeMat,testRow)
-    sizeStats <- rbind(sizeStats,sizeStats1[[1]]) #means over individuals (grouped by iter, occ)
-    sizeStats2 <- rbind(sizeStats2,sizeStats1[[2]]) #means by occ of means over individuals (grouped by occ)
-
-    grStats1 <- summarizeGRs(grMat,testRow)
-    grStats <- rbind(grStats,grStats1[[1]]) #means over individuals (grouped by iter, occ)
-    grStats2 <- rbind(grStats2,grStats1[[2]]) #means by occ of means over individuals (grouped by occ)
-  }
-  endTime <- Sys.time()
-  (elapsed[[speciesForProj]] <- endTime - startTime)
-
-  stats3Spp[[speciesForProj]] <- list( sizeStats,sizeStats2 )
-  grStats3Spp[[speciesForProj]] <- list( grStats,grStats2 )
-
-  # lm[[speciesForProj]] <- lm( lenMean ~ temp + flow + bkt + bnt + ats + temp * flow + temp^2 + flow^2 + bnt*bkt, data = sizeStats %>% filter(occ == (simData$nOcc)))
-  # 
-  # relImp[[speciesForProj]] <- calc.relimp(lm[[speciesForProj]], type = c("lmg"), rela = TRUE)
-  #   
-  # anov[[speciesForProj]] <- anova( lm[[speciesForProj]] )
-
-}
-
-#projOut <- list(stats3Spp, lm, relImp, anov, elapsed, simData)
-save( stats3Spp, grStats3Spp, elapsed, simData, file = paste0('D:/projects/linkedModels/data/out/projectionOut_N',simData$nReps,'.RData'))
-
-# if not re-running the sim
-# load( file = 'D:/projects/linkedModels/data/out/projectionOut.RData')
-# load( file = paste0('D:/projects/linkedModels/data/out/projectionOut_N',simData$nReps,'.RData'))
-# spp <- c('bkt','bnt','ats')
-
-# raw sim data figs - size
-ggplot(stats3Spp[['bkt']][[1]] %>% filter(ats == 0, bkt == 0, bnt == 0), aes(as.numeric(occ),lenMean)) +
-  geom_jitter(aes(color = factor(ats))) +
-  #geom_smooth(aes(color = factor(temp), linetype = factor(flow)))+#(bnt + 2)/4 )) +
-  geom_smooth()+#(bnt + 2)/4 )) +
-  geom_smooth(color = 'green', data = stats3Spp[['bnt']][[1]] %>% filter(ats == 0, bkt == 0, bnt == 0)) +
-  facet_grid(temp ~ flow)
-
-########################
-# raw sim data figs - gr
-ggplot(grStats3Spp[['bkt']][[1]] %>% filter(ats == 0, bkt == 0, bnt == 0), aes(as.numeric(occ),grMean)) +
-  geom_jitter(aes(color = factor(ats)), width = 0.2, size = 0.5) +
-  #geom_smooth(aes(color = factor(temp), linetype = factor(flow)))+#(bnt + 2)/4 )) +
-  #geom_smooth()+#(bnt + 2)/4 )) +
-  geom_jitter(color = 'darkblue', width = 0.1, size = 0.5, data = grStats3Spp[['bnt']][[1]] %>% filter(ats == 0, bkt == 0, bnt == 0)) +
-  facet_grid(temp ~ flow)
-
-```
-
-```{r RI across all - use this for paper GROWTH RATE}
-
-# relative proportion stats with occ in the lm 
-grStats3Spp[['bkt']][[1]]$season <- as.factor(c(3,4,1,2,3,4,1,2,3,4))
-grStats3Spp[['bkt']][[1]]$temp2 <- grStats3Spp[['bkt']][[1]]$temp^2
-grStats3Spp[['bkt']][[1]]$flow2 <- grStats3Spp[['bkt']][[1]]$flow^2
-
-indepVarVals <- c(-1.5,0,1.5) # make sure symetrical around 0
-indepVars <- c('temp', 'flow', 'temp:flow', 'bkt', 'bnt', 'ats', 'bkt:bnt')
-rIBKTBNTAll <- NULL
-# loop over occs
-for(rep in 1:simData$nReps){
-
-  print(rep)
-
-  datOccBKTAll <- grStats3Spp[['bkt']][[1]] %>% filter(iter == rep, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
-  lmAllBKT <- lm( grMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBKTAll) 
-  relImpAllBKT <- calc.relimp(lmAllBKT, type = c("lmg"), rela = TRUE)
-
-  datOccBNTAll <- grStats3Spp[['bnt']][[1]] %>% filter(iter == rep, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
-  lmAllBNT <- lm( grMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBNTAll)
-  relImpAllBNT <- calc.relimp(lmAllBNT, type = c("lmg"), rela = TRUE)
-
-  lmgBKT <- relImpAllBKT$lmg[(names(relImpAllBKT$lmg) %in% indepVars)]
-  rBKT <- lmgBKT / sum(lmgBKT)
-
-  lmgBNT <- relImpAllBNT$lmg[(names(relImpAllBNT$lmg) %in% indepVars)]
-  rBNT <- lmgBNT / sum(lmgBNT)
-
-  rIBKTBNTAll <- rbind(rIBKTBNTAll, data.frame(cbind(rBKT,rBNT)) %>% 
-                      rownames_to_column() %>% 
-                      gather(key = rowname) %>% 
-                      mutate(var = rep(1:length(indepVars),2),
-                             varText = rep(names(rBKT),2),
-                             iter = rep)
-                    )
-}
-
-rIBKTBNTAllMeans <- rIBKTBNTAll %>%
-  group_by(rowname, var) %>%
-  summarise(rIMean = mean(value),
-            rISd = sd(value),
-            rIUpperCI = quantile(value, 0.95))
-
-dodge <- position_dodge(width=0.9)
-ggplot(rIBKTBNTAllMeans, aes(factor(var),rIMean, fill = factor(rowname, levels = c('rBKT','rBNT'), ordered = TRUE))) +
-  geom_errorbar(aes(ymin = rIMean - 0.001, ymax = rIUpperCI), width = 0.25, position=dodge) +
-  geom_col(position=dodge) +
-  scale_x_discrete('Parameter',
-                   labels = c('1' = 'Temperature','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")) +
-  scale_fill_grey() +
-  scale_y_continuous('Relative proportion of variance explained') +
-  theme_publication() +
-  geom_hline(yintercept = 0) +
-  theme( legend.position = 'none')
-
-ggsave(paste0('figures/3Spp_WB/relImpOccMeanModAll_GR_N',simData$nReps,'.png'), width = 9, height = 7, units='in')
-
-# for MS
-rIBKTBNTAllMeans
-#    rowname   var  rIMean    rISd rIUpperCI
-#    <chr>   <int>   <dbl>   <dbl>     <dbl>
-#  1 rBKT        1 0.132   0.0251     0.170 
-#  2 rBKT        2 0.670   0.0369     0.731 
-#  3 rBKT        3 0.0116  0.0118     0.0330
-#  4 rBKT        4 0.0346  0.0256     0.0703
-#  5 rBKT        5 0.0951  0.0284     0.140 
-#  6 rBKT        6 0.0441  0.0263     0.0919
-#  7 rBKT        7 0.0121  0.0155     0.0410
-#  8 rBNT        1 0.106   0.0212     0.141 
-#  9 rBNT        2 0.568   0.0364     0.623 
-# 10 rBNT        3 0.0885  0.0223     0.124 
-# 11 rBNT        4 0.00518 0.00556    0.0163
-# 12 rBNT        5 0.122   0.0285     0.169 
-# 13 rBNT        6 0.0314  0.0176     0.0624
-# 14 rBNT        7 0.0782  0.0270     0.122
-
-rIBKTBNTAll %>%
-  mutate(isEnvVar = ifelse(varText %in% c('temp','flow','temp:flow'), 1,0)) %>%
-  group_by(rowname,isEnvVar) %>%
-  summarize(meanMean = mean(value),
-            sumMean = sum(value),
-            sdMean = sd(value))
-
-#   rowname isEnvVar meanMean sumMean sdMean
-#   <chr>      <dbl>    <dbl>   <dbl>  <dbl>
-# 1 rBKT           0   0.0383    15.3 0.0402
-# 2 rBKT           1   0.282     84.7 0.279 
-# 3 rBNT           0   0.0736    29.4 0.0484
-# 4 rBNT           1   0.235     70.6 0.239
-
-#bkt
-.282/(.282+.038) # 0.88125
-#bnt
-.235/(.235+.074) # 0.7605178
-.881/.76 # 1.159211
-```
-
-```{r RI across yoy/adults - use this for paper GROWTH RATE}
-
-# relative proportion stats with occ in the lm 
-grStats3Spp[['bkt']][[1]]$season <- as.factor(c(3,4,1,2,3,4,1,2,3,4))
-grStats3Spp[['bkt']][[1]]$temp2 <- grStats3Spp[['bkt']][[1]]$temp^2
-grStats3Spp[['bkt']][[1]]$flow2 <- grStats3Spp[['bkt']][[1]]$flow^2
-
-ageForRI <-  'yoy' #'yoy' or 'adult'
-occs <- if (ageForRI == 'yoy') 1:3 else 4:10
-indepVarVals <- c(-1.5,0,1.5) # make sure symetrical around 0
-indepVars <- c('temp', 'flow', 'temp:flow', 'bkt', 'bnt', 'ats', 'bkt:bnt')
-rIBKTBNT <- NULL
-# loop over occs
-for(rep in 1:simData$nReps){
-
-  print(rep)
-
-  datOccBKT <- grStats3Spp[['bkt']][[1]] %>% filter(iter == rep, occ %in% occs, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
-  lmAllBKT <- lm( grMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBKT) 
-  relImpAllBKT <- calc.relimp(lmAllBKT, type = c("lmg"), rela = TRUE)
-
-  datOccBNT <- grStats3Spp[['bnt']][[1]] %>% filter(iter == rep, occ %in% occs, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
-  lmAllBNT <- lm( grMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBNT)
-  relImpAllBNT <- calc.relimp(lmAllBNT, type = c("lmg"), rela = TRUE)
-
-  lmgBKT <- relImpAllBKT$lmg[(names(relImpAllBKT$lmg) %in% indepVars)]
-  rBKT <- lmgBKT / sum(lmgBKT)
-
-  lmgBNT <- relImpAllBNT$lmg[(names(relImpAllBNT$lmg) %in% indepVars)]
-  rBNT <- lmgBNT / sum(lmgBNT)
-
-  rIBKTBNT <- rbind(rIBKTBNT, data.frame(cbind(rBKT,rBNT)) %>% 
-                      rownames_to_column() %>% 
-                      gather(key = rowname) %>% 
-                      mutate(var = rep(1:length(indepVars),2),
-                             varText = rep(names(rBKT),2),
-                             iter = rep)
-                    )
-}
-
-rIBKTBNTMeans <- rIBKTBNT %>%
-  group_by(rowname, var) %>%
-  summarise(rIMean = mean(value),
-            rISd = sd(value),
-            rIUpperCI = quantile(value, 0.95))
-
-dodge <- position_dodge(width=0.9)
-ggplot(rIBKTBNTMeans, aes(factor(var),rIMean, fill = factor(rowname, levels = c('rBKT','rBNT'), ordered = TRUE))) +
-  geom_errorbar(aes(ymin = rIMean - 0.001, ymax = rIUpperCI), width = 0.25, position=dodge) +
-  geom_col(position=dodge) +
-  scale_x_discrete('Parameter',
-                   labels = c('1' = 'Temperature','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")) +
-  scale_fill_grey() +
-  scale_y_continuous('Relative proportion of variance explained') +
-  theme_publication() +
-  geom_hline(yintercept = 0) +
-  theme( legend.position = 'none')
-
-ggsave(paste0('figures/3Spp_WB/relImpOccMeanModWOcc_GR',simData$nReps,'_',ageForRI,'.png'), width = 9, height = 7, units='in')
-
-#    rowname   var  rIMean    rISd rIUpperCI
-#    <chr>   <int>   <dbl>   <dbl>     <dbl>
-#  1 rBKT        1 0.516   0.0639     0.619 
-#  2 rBKT        2 0.123   0.0332     0.175 
-#  3 rBKT        3 0.0209  0.0222     0.0568
-#  4 rBKT        4 0.0579  0.0307     0.100 
-#  5 rBKT        5 0.193   0.0487     0.273 
-#  6 rBKT        6 0.0539  0.0374     0.126 
-#  7 rBKT        7 0.0362  0.0303     0.0889
-#  8 rBNT        1 0.154   0.0566     0.253 
-#  9 rBNT        2 0.0106  0.00999    0.0281
-# 10 rBNT        3 0.140   0.0479     0.208 
-# 11 rBNT        4 0.00421 0.00467    0.0135
-# 12 rBNT        5 0.463   0.0780     0.598 
-# 13 rBNT        6 0.173   0.0676     0.294 
-# 14 rBNT        7 0.0547  0.0340     0.121
-
-# get isEnvVar sens for yoy and adults - need to rerun the above code for each age
-rIBKTBNT %>%
-  mutate(isEnvVar = ifelse(varText %in% c('temp','flow','temp:flow'), 1,0)) %>%
-  group_by(rowname,isEnvVar) %>%
-  summarize(meanMean = mean(value),
-            sumMean = sum(value),
-            sdMean = sd(value)) %>%
-  mutate(age = ageForRI)
-
-#   rowname isEnvVar meanMean sumMean sdMean age  
-#   <chr>      <dbl>    <dbl>   <dbl>  <dbl> <chr>
-# 1 rBKT           0   0.0769    30.8 0.0763 yoy  
-# 2 rBKT           1   0.231     69.2 0.209  yoy  
-# 3 rBNT           0   0.165     66.2 0.185  yoy  
-# 4 rBNT           1   0.113     33.8 0.0888 yoy
-
-#   rowname isEnvVar meanMean sumMean sdMean age  
-#   <chr>      <dbl>    <dbl>   <dbl>  <dbl> <chr>
-# 1 rBKT           0   0.0287    11.5 0.0272 adult
-# 2 rBKT           1   0.295     88.5 0.377  adult
-# 3 rBNT           0   0.0411    16.4 0.0285 adult
-# 4 rBNT           1   0.279     83.6 0.343  adult
-
-# ratio of env/(env+spp) means - yoy
-.231/(.231 + .0769) # bkt 0.7502436
-.165/(.165 + .113) # bnt 0.5935252
-
-# ratio of env/(env+spp) means - adult
-.295/(.295 + .0287) # bkt 0.9113377
-.279/(.279 + .0411) # bnt 0.8716026
-
-.75/.593 # 1.26
-.911/.872 # 1.04
-##############################################################
-############################################################
-
-# projOut <- list(stats3Spp, lm, relImp, anov, elapsed)
-load(paste0('D:/projects/linkedModels/data/out/projectionOut_N',simData$nReps,'.RData'))
-
-#spp <- c('bkt','bnt','ats')
-# scenarios
-# Plot subsets of the simulations
-#means for each iter
-stats1 <- rbind(data.frame(grStats3Spp[['bkt']][[1]]), 
-                data.frame(grStats3Spp[['bnt']][[1]]), 
-                data.frame(grStats3Spp[['ats']][[1]])) %>%
-          mutate(spp = rep(spp, each = nrow(grStats3Spp[['bkt']][[1]])))
-
-#means for each occ
-stats2 <- rbind(data.frame(grStats3Spp[['bkt']][[2]]), 
-                data.frame(grStats3Spp[['bnt']][[2]]), 
-                data.frame(grStats3Spp[['ats']][[2]])) %>%
-          mutate(spp = rep(spp, each = nrow(grStats3Spp[['bkt']][[2]])))
-
-ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,grMeanMean, group = spp)) +
-  geom_point(aes(color = spp) ) +
-  geom_line( aes(color = spp) ) +
-  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 1.5, flow == 0, bkt == 0, bnt == 0, ats == 0) ) +
-  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 1.5, flow == 0, bkt == 0, bnt == 0, ats == 0), linetype = 2 )
-
-ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,grMeanMean, group = spp)) +
-  geom_point(aes(color = spp) ) +
-  geom_line( aes(color = spp) ) +
-  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 1.5, bkt == 0, bnt == 0, ats == 0) ) +
-  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 1.5, bkt == 0, bnt == 0, ats == 0), linetype = 2 )
-
-ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,grMeanMean, group = spp)) +
-  geom_point(aes(color = spp) ) +
-  geom_line( aes(color = spp) ) +
-  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 0, bkt == 1.5, bnt == 0, ats == 0) ) +
-  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 0, bkt == 1.5, bnt == 0, ats == 0), linetype = 2 )
-
-# trajectories across indep var values for different combinations. Trying to relate to relImp results
-#temp
-p1 <- ggplot( stats2 %>% filter(flow == 0, bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(temp))) +
-  geom_line(aes(color = factor(temp))) +
-  geom_point(aes(color = factor(temp))) +
-  facet_wrap(~spp)
-
-#flow
-p2 <- ggplot( stats2 %>% filter(temp == 0, bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(flow))) +
-  geom_line(aes(color = factor(flow))) +
-  geom_point(aes(color = factor(flow))) +
-  facet_wrap(~spp)
-
-#bkt
-p3 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(bkt))) +
-  geom_line(aes(color = factor(bkt))) +
-  geom_point(aes(color = factor(bkt))) +
-  facet_wrap(~spp)
-
-#bnt
-p4 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(bnt))) +
-  geom_line(aes(color = factor(bnt))) +
-  geom_point(aes(color = factor(bnt))) +
-  facet_wrap(~spp)
-
-#ats
-p5 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(ats))) +
-  geom_line(aes(color = factor(ats))) +
-  geom_point(aes(color = factor(ats))) +
-  facet_wrap(~spp)
-
-multiplot(p1, p2, p3, p4, p5, cols=2) # multiplot function is below
-
-#temp*flow
-ggplot( stats2 %>% filter(bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(flow))) +
-  geom_line(aes(color = factor(flow))) +
-  geom_point(aes(color = factor(flow))) +
-  facet_grid(temp~spp)
-
-#bkt*bnt
-ggplot( stats2 %>% filter(temp == 0, flow == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,grMeanMean, group = factor(bnt))) +
-  geom_line(aes(color = factor(bnt))) +
-  geom_point(aes(color = factor(bnt))) +
-  facet_grid(bkt~spp)
-
-# Multiple plot function
-#
-# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
-# - cols:   Number of columns in layout
-# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
-#
-# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
-# then plot 1 will go in the upper left, 2 will go in the upper right, and
-# 3 will go all the way across the bottom.
-#
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-
-  numPlots = length(plots)
-
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-
- if (numPlots==1) {
-    print(plots[[1]])
-
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
-
-# size=100
-# gr=10
-# sd=1
-# mean(rnorm(10000,size+gr,sd))
-# mean(size+rnorm(10000,gr,sd))
-
-```
-
-```{r RI by occ - use for paper GROWTH}
-
-# make figures of relative importance for each sampling occasion
-rIAll <- cAll <- NULL
-
-for(ageGroupLoop in 0:1){
-  for(seasonLoop in 1:4){
-#for( occSelected in 1:(simData$nOcc - 1)){
-  if(ageGroupLoop == 0 & seasonLoop == 2) next
-
-  lmAll <- ggRIAll <- relImpAll <- coefsAll <- list()
-
-  for(spp1 in spp){
-    print(c(ageGroupLoop, seasonLoop, spp1))
-
-    datAgeGroupSeason <- grStats3Spp[[spp1]][[1]] %>%
-         mutate(ageGroup = ifelse(occ %in% 1:3, 0, 1),
-         season = rep(c(3,4,1,2,3,4,1,2,3,4), nrow(grStats3Spp[[spp1]][[1]])/simData$nOcc))
-
-      lmAll[[spp1]] <- lm( grMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt, data = datAgeGroupSeason %>% filter(ageGroup == ageGroupLoop, season == seasonLoop, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals))
-
-      relImpAll[[spp1]] <- calc.relimp(lmAll[[spp1]], type = c("lmg"), rela = TRUE)
-
-      coefsAll[[spp1]] <- lmAll[[spp1]]$coefficients
-  }
-
-  # make bar chart of relative importance of each variable for each spp
-  rI <- cbind(data.frame(bktRI = relImpAll[['bkt']]$lmg), 
-              data.frame(bntRI = relImpAll[['bnt']]$lmg)) %>%
-        rownames_to_column() %>% 
-        gather(key=rowname) %>% 
-        mutate(var = rep(1:length(relImpAll[['bkt']]$lmg),length(spp)),
-               ageGroup = ageGroupLoop,
-               season = seasonLoop
-               )
-  rI$varText <- rep(names(relImpAll[['bkt']]$lmg),2)
-
-  rIAll <- rbind(rIAll,rI)
-
-  #rIAll$age <- ifelse(rIAll$occ %in% 1:2, 0, ifelse(rIAll$occ %in% 3:6, 1, 2))
-
-  }  
-}
-
-######################################################
-# relImp over occasions for each spp and parameter
-labels <- c('1' = 'Temp','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")
-
-# ggplot(rIAll %>% filter(rowname %in% c('bktRI','bntRI')), aes((occ), value, color = factor(rowname))) +
-#   geom_point() +
-#   geom_line() +
-#   geom_vline(xintercept = 2.5) +
-#   scale_x_continuous('Sampling occasion',
-#                     breaks = 1:9,
-#                     labels = c('Autumn','Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn')
-#                     ) +
-#   scale_y_continuous('Relative proportion of variance explained') +
-#   theme_publication() +
-#   theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10)) +
-#   facet_wrap(~ var, labeller=labeller(var = labels))
-# 
-# ggsave(paste0('figures/3Spp_WB/relImpOccAll_GR.png'), width = 11, height = 9, units='in')
-
-rIAllMinus <- rIAll %>% mutate(value2 = ifelse(rowname == 'bktRI',value * -1, value))
-
-labelsSeason <- c('1'='Autumn','2'='Winter','3'='Spring','4'='Summer','5'='Autumn','6'='Winter','7'='Spring','8'='Summer','9'='Autumn')
-labelsSeason2 <- c('1'='Spring','2'='Summer','3'='Autumn','4'='Winter')
-labelAge <- c('0' = 'Growth year 0/1', '1' = 'Growth Year 1+')
-
-### get prop env
-summ1 <- rIAll %>%
-  mutate(isEnvVar = ifelse(varText %in% c('temp','flow','temp:flow'), 1,0)) %>%
-  group_by(rowname,isEnvVar,ageGroup, season) %>%
-  summarize(mean=mean(value)) %>%
-  arrange(ageGroup,season,rowname,isEnvVar)
-
-summ2 <- summ1 %>%
-  group_by(rowname,ageGroup,season) %>%
-  summarize(sum=sum(mean))
-
-summ1a <- left_join(summ1,summ2) %>%
-  mutate(prop = mean/sum,
-         xAxisLoc = ifelse(rowname == 'bktRI',-0.85, 0.85)) %>%
-  filter(isEnvVar == 1) 
-###
-
-####
-# add in betas to rIAllMinus
-load("D:/projects/linkedModels/data/out/allBetas.RData") #loads df 'q'
-qFig1 <- q %>% filter(riverGG == 'WB', species %in% c('bkt','bnt'))
-qFig1$season <- as.numeric(as.character(qFig1$season))
-qFig1$isYOY <- as.numeric(as.character(qFig1$isYOY))
-qFig <- qFig1 %>%
-  mutate( varText = tolower(betaGG)) %>%
-  mutate( varText = ifelse(varText == 'temp*flow','temp:flow', varText)) %>%
-  mutate( varText = ifelse(varText == 'bnt*bkt','bkt:bnt', varText)) %>%
-  select(isYOY,season,species,median,varText)
-
-# prepare rIAllMinus
-rIAllMinusQ <- rIAllMinus %>%
-  mutate(species = str_sub(rowname,1,3),
-         isYOY = ageGroup) %>%
-  left_join(qFig) %>%
-  mutate(median01 = ifelse(median > 0, 1, 0),
-         medianBreaks = cut(median, breaks = c(-1, -0.025, 0 , 0.025, 1)))
-
-#
-
-ggplot(rIAllMinus, aes(var, y = value2, fill = rowname)) +
-  geom_col() +
-  coord_flip() +
-  scale_x_reverse('Variable', 
-                     breaks = 1:7,
-                     labels = c('Temp', 'Flow', 'Bkt', 'Bnt','Ats','Temp*Flow',  "Bkt*Bnt")) +
-  scale_y_continuous('Relative proportion of variance explained',
-                     breaks = c(-1.0, -0.5, 0.0, 0.5),
-                     labels = c('1.0', '0.5', '0.0', '0.5')) +
-  geom_text(data = summ1a, aes(y=xAxisLoc, x=7, label=paste(round(prop,2)), color=rowname),size = 5) +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  theme_publication() +
-  scale_fill_grey(start = 0.15, end = 0.75) +
-  facet_grid(season ~ ageGroup, labeller=labeller(season = labelsSeason2, ageGroup = labelAge))
-
-ggsave(paste0('figures/3Spp_WB/relImpOccAll_GR_HorizBars.png'), width = 11, height = 9, units='in')
-
-# bars colered by beta
-ggplot(rIAllMinusQ, aes(var, y = value2)) +
-  geom_col(aes(fill = median)) +
-  coord_flip() +
-  scale_x_reverse('Variable', 
-                     breaks = 1:7,
-                     labels = c('Temp', 'Flow', 'Bkt', 'Bnt','Ats','Temp*Flow',  "Bkt*Bnt")) +
-  scale_y_continuous('Relative proportion of variance explained',
-                     breaks = c(-0.9, -0.25, -0.5, 0.0, 0.25, 0.5, 0.9),
-                     labels = c('0.9', '0.25', '0.5\n\nBrook Trout\n', '0.0', '0.25', '0.5\n\nBrown Trout\n', '0.9')) +
-  geom_text(data = summ1a, aes(y=xAxisLoc, x=7, label=paste(round(prop,2))),size = 5) +
-  geom_hline(yintercept = 0) +
-  #scale_color_grey(start = 0.15, end = 0.75) +
-  theme_publication() +
- #  scale_fill_continuous() +
-  scale_fill_distiller(name="Parameter\nestimate", palette = "RdBu", direction = 1) +
-  theme(legend.position="right",             
-        #legend.key = element_rect(colour = NA),
-        legend.direction = "vertical",
-        legend.key.size= unit(0.25, "in"),
-        #    legend.margin = unit(0, "cm"),
-        legend.title = element_text(face="plain")#element_blank()
-        ) +
-  facet_grid(season ~ ageGroup, labeller=labeller(season = labelsSeason2, ageGroup = labelAge))
-
-ggsave(paste0('figures/3Spp_WB/relImpOccAll_GR_HorizBarsColor.png'), width = 11, height = 9, units='in')
-
-#######################
-# coefs
-labelsC <- c('1' = 'intercept', '2' = 'Temp','3' = 'Flow','4' = 'Bkt','5' = 'Bnt','6' = 'Ats','7' = 'Temp*Flow', '8' = "Bkt*Bnt")
-
-ggplot(cAll %>% filter(var %in% 2:8, rowname %in% c('bktC','bntC')), aes((occ), value, color = factor(rowname))) +
-  geom_point(size=3) +
-  geom_line() +
-  geom_vline(xintercept = 3.5) +
-  scale_x_continuous('Sampling occasion',
-                    breaks = 2:10,
-                    labels = c('Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn','Winter')
-                    ) +
-  scale_y_continuous('Coefficients') +
-  theme_publication() +
-  theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10)) +
-  facet_wrap(~ var, labeller=labeller(var = labelsC))#, scales = 'free')
-
-ggsave(paste0('figures/3Spp_WB/coefsOccAll.png'), width = 11, height = 9, units='in')
-
-# intercept only
-cAll2 <- rbind(cAll,data.frame(rowname=c('bktC','bntC','atsC'),value=simData$meanInit,var=1,occ=1,season=3))
-ggplot(cAll2 %>% filter(var == 1), aes((occ), value, color = factor(rowname))) +
-  geom_point(size=4) +
-  geom_line() +
-  geom_vline(xintercept = 3.5) +
-  scale_x_continuous('Sampling occasion',
-                    breaks = 1:10,
-                    labels = c('Autumn','Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn','Winter')
-                    ) +
-  scale_y_continuous('Intercept') +
-  theme_publication() +
-  theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10))
-
-ggsave(paste0('figures/3Spp_WB/coefsInterceptOccAll.png'), width = 11, height = 9, units='in')
-
-#####################################
-# no need to start sims with diff sizes by species
-yoyMeans <- 
-cd %>% filter( riverOrdered == 'west brook', ageInSamples == 1 ) %>%
-  group_by(species) %>%
-  summarize(meanLen = mean(observedLength,na.rm = TRUE),
-           sdLen = sd(observedLength,na.rm = TRUE))
-
-#   species meanLen sdLen
-#   <chr>     <dbl> <dbl>
-# 1 ats        69.0  5.86
-# 2 bkt        72.7  9.94
-# 3 bnt        72.9  8.73
-#####################################
-
-```
-
-```{r RI by occ - not used for paper SIZE}
-
-# make figures of relative importance for each sampling occasion
-rIAll <- cAll <- NULL
-for( occSelected in 2:simData$nOcc){
-  print(occSelected)
-  lmAll <- ggRIAll <- relImpAll <- coefsAll <- list()
-
-  for(spp1 in spp){
-
-      lmAll[[spp1]] <- lm( lenMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt, data = stats3Spp[[spp1]][[1]] %>% filter(occ == occSelected))
-
-      relImpAll[[spp1]] <- calc.relimp(lmAll[[spp1]], type = c("lmg"), rela = TRUE)
-
-      coefsAll[[spp1]] <- lmAll[[spp1]]$coefficients
-  }
-
-  # make bar chart of relative importance of each variable for each spp
-  rI <- cbind(data.frame(bktRI = relImpAll[['bkt']]$lmg), 
-              data.frame(bntRI = relImpAll[['bnt']]$lmg), 
-              data.frame(atsRI = relImpAll[['ats']]$lmg)) %>%
-        rownames_to_column() %>% 
-        gather(key=rowname) %>% 
-        mutate(var = rep(1:length(relImpAll[['bkt']]$lmg),length(spp)),
-               occ = occSelected,
-               season = simData$seasons[occSelected])
-
-  rIAll <- rbind(rIAll,rI)
-
-  ggplot(rI, aes(factor(var),value, fill = factor(rowname, levels = c('bktRI','bntRI','atsRI'), ordered = TRUE))) +
-    geom_col(position='dodge') +
-    scale_x_discrete('Parameter',
-                     labels = c('1' = 'Temp','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")) +
-    scale_fill_grey() +
-    scale_y_continuous('Relative proportion of variance explained') +
-    ggtitle(paste0("occ = ",occSelected,"  Season = ",simData$seasons[occSelected])) +
-    theme_publication() +
-    theme( legend.position = 'none')
-
-##############################  ggsave(paste0('figures/3Spp_WB/relImpOcc',occSelected,'.png'), width = 9, height = 7, units='in')
-
-  # coefficients
-  coefs <- cbind(data.frame(bktC = coefsAll[['bkt']]), 
-                 data.frame(bntC = coefsAll[['bnt']]), 
-                 data.frame(atsC = coefsAll[['ats']])) %>%
-        rownames_to_column() %>% 
-        gather(key=rowname) %>% 
-        mutate(var = rep(1:length(coefsAll[['bkt']]),length(spp)),
-               occ = occSelected,
-               season = simData$seasons[occSelected])
-
-  cAll <- rbind(cAll,coefs)
-}
-
-######################################################
-# relImp over occasions for each spp and parameter
-labels <- c('1' = 'Temp','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")
-
-# leaving out bkt*bnt, too small
-ggplot(rIAll %>% filter(rowname %in% c('bktRI','bntRI')), aes((occ), value, color = factor(rowname))) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 3.5) +
-  scale_x_continuous('Sampling occasion',
-                    breaks = 2:10,
-                    labels = c('Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn','Winter')
-                    ) +
-  scale_y_continuous('Relative proportion of variance explained') +
-  theme_publication() +
-  theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10)) +
-  facet_wrap(~ var, labeller=labeller(var = labels))
-
-ggsave(paste0('figures/3Spp_WB/relImpOccAll.png'), width = 11, height = 9, units='in')
-
-rIMeans <- rIAll %>%
-  group_by(rowname,var) %>%
-  summarize( mean = mean(value),
-             sd = sd(value),
-             stderr = sd(value)/sqrt(length(value))) %>%
-  mutate( up = mean + stderr )
-
-dodge <- position_dodge(width=0.9)
-ggplot(rIMeans %>% filter(rowname %in% c('bktRI','bntRI')), aes(factor(var),mean, fill = factor(rowname, levels = c('bktRI','bntRI','atsRI'), ordered = TRUE))) +
-  geom_errorbar(aes(ymin = mean-0.01, ymax = up), width = 0.25, position=dodge) +
-  geom_col(position=dodge) +
-  scale_x_discrete('Parameter',
-                   labels = c('1' = 'Temperature','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")) +
-  scale_fill_grey() +
-  scale_y_continuous('Relative proportion of variance explained') +
-  theme_publication() +
-  theme( legend.position = 'none')
-
-ggsave(paste0('figures/3Spp_WB/relImpOccMean.png'), width = 9, height = 7, units='in')
-
-# mean rel Imp for env vs spp variables
-rIMeans %>%
-  mutate(isEnvVar = ifelse(var %in% c(1,2,6), 1,0)) %>%
-  group_by(rowname,isEnvVar) %>%
-  summarize(meanMean = mean(mean),
-            sumMean = sum(mean),
-            sdMean = sd(mean))
-
-# ratio of env/(env+spp) means
-.247/(.246+.0648) # bkt
-.122/(.122+.159) # bnt
-
-# ratio of env/(env+spp) sums - donT need to do ratio, just use sum for env
-.635/.259 # den-dep bnt/bkt
-.741/.365 # env bkt/bnt
-
-#######################
-# coefs
-labelsC <- c('1' = 'intercept', '2' = 'Temp','3' = 'Flow','4' = 'Bkt','5' = 'Bnt','6' = 'Ats','7' = 'Temp*Flow', '8' = "Bkt*Bnt")
-
-ggplot(cAll %>% filter(var %in% 2:8, rowname %in% c('bktC','bntC')), aes((occ), value, color = factor(rowname))) +
-  geom_point(size=3) +
-  geom_line() +
-  geom_vline(xintercept = 3.5) +
-  scale_x_continuous('Sampling occasion',
-                    breaks = 2:10,
-                    labels = c('Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn','Winter')
-                    ) +
-  scale_y_continuous('Coefficients') +
-  theme_publication() +
-  theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10)) +
-  facet_wrap(~ var, labeller=labeller(var = labelsC))#, scales = 'free')
-
-ggsave(paste0('figures/3Spp_WB/coefsOccAll.png'), width = 11, height = 9, units='in')
-
-# intercept only
-cAll2 <- rbind(cAll,data.frame(rowname=c('bktC','bntC','atsC'),value=simData$meanInit,var=1,occ=1,season=3))
-ggplot(cAll2 %>% filter(var == 1), aes((occ), value, color = factor(rowname))) +
-  geom_point(size=4) +
-  geom_line() +
-  geom_vline(xintercept = 3.5) +
-  scale_x_continuous('Sampling occasion',
-                    breaks = 1:10,
-                    labels = c('Autumn','Winter','Spring','Summer', 'Autumn','Winter','Spring','Summer', 'Autumn','Winter')
-                    ) +
-  scale_y_continuous('Intercept') +
-  theme_publication() +
-  theme(axis.text.x  = element_text(angle=90, vjust=0.25, size=10))
-
-ggsave(paste0('figures/3Spp_WB/coefsInterceptOccAll.png'), width = 11, height = 9, units='in')
-
-#####################################
-# no need to start sims with diff sizes by species
-yoyMeans <- 
-cd %>% filter( riverOrdered == 'west brook', ageInSamples == 1 ) %>%
-  group_by(species) %>%
-  summarize(meanLen = mean(observedLength,na.rm = TRUE),
-           sdLen = sd(observedLength,na.rm = TRUE))
-
-#   species meanLen sdLen
-#   <chr>     <dbl> <dbl>
-# 1 ats        69.0  5.86
-# 2 bkt        72.7  9.94
-# 3 bnt        72.9  8.73
-#####################################
-
-```
-
-```{r tables}
-
-round(mcmc$summary$all.chains[startsWith(attributes(mcmc$summary$all.chains)$dimnames[[1]], "grBeta["),],2)
-
-```
-
-```{r RI across occ - use this for paper SIZE}
-
-# relative proportion stats with occ in the lm 
-stats3Spp[['bkt']][[1]]$season <- as.factor(c(3,4,1,2,3,4,1,2,3,4))
-stats3Spp[['bkt']][[1]]$temp2 <- stats3Spp[['bkt']][[1]]$temp^2
-stats3Spp[['bkt']][[1]]$flow2 <- stats3Spp[['bkt']][[1]]$flow^2
-
-ageForRI <-  'adult' #'yoy' or 'adult'
-occs <- if (ageForRI == 'yoy') 1:3 else 4:10
-indepVarVals <- c(-1.5,0,1.5) # make sure symetrical around 0
-rIBKTBNT <- NULL
-# loop over occs
-for(rep in 1:simData$nReps){
-
-  print(rep)
-
-  datOccBKT <- stats3Spp[['bkt']][[1]] %>% filter(iter == rep, occ %in% occs, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
-  lmAllBKT <- lm( lenMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBKT) 
-  relImpAllBKT <- calc.relimp(lmAllBKT, type = c("lmg"), rela = TRUE)
-
-  datOccBNT <- stats3Spp[['bnt']][[1]] %>% filter(iter == rep, occ %in% occs, temp %in% indepVarVals, flow %in% indepVarVals, bkt %in% indepVarVals, bnt %in% indepVarVals, ats %in% indepVarVals)
-  lmAllBNT <- lm( lenMean ~ temp + flow + temp * flow + temp^2 + flow^2 + bkt + bnt + ats + bnt * bkt + occ, data = datOccBNT)
-  relImpAllBNT <- calc.relimp(lmAllBNT, type = c("lmg"), rela = TRUE)
-
-  rBKT <- relImpAllBKT$lmg[2:8] / sum(relImpAllBKT$lmg[2:8]) 
-  rBNT <- relImpAllBNT$lmg[2:8] / sum(relImpAllBNT$lmg[2:8])
-
-  rIBKTBNT <- rbind(rIBKTBNT, data.frame(cbind(rBKT,rBNT)) %>% 
-                      rownames_to_column() %>% 
-                      gather(key = rowname) %>% 
-                      mutate(var = rep(1:7,2),
-                             varText = rep(names(rBKT),2),
-                             iter = rep)
-                    )
-
-}
-
-rIBKTBNTMeans <- rIBKTBNT %>%
-  group_by(rowname, var) %>%
-  summarise(rIMean = mean(value),
-            rISd = sd(value),
-            rIUpperCI = quantile(value, 0.95))
-
-dodge <- position_dodge(width=0.9)
-ggplot(rIBKTBNTMeans, aes(factor(var),rIMean, fill = factor(rowname, levels = c('rBKT','rBNT'), ordered = TRUE))) +
-  geom_errorbar(aes(ymin = rIMean - 0.001, ymax = rIUpperCI), width = 0.25, position=dodge) +
-  geom_col(position=dodge) +
-  scale_x_discrete('Parameter',
-                   labels = c('1' = 'Temperature','2' = 'Flow','3' = 'Bkt','4' = 'Bnt','5' = 'Ats','6' = 'Temp*Flow', '7' = "Bkt*Bnt")) +
-  scale_fill_grey() +
-  scale_y_continuous('Relative proportion of variance explained') +
-  theme_publication() +
-  geom_hline(yintercept = 0) +
-  theme( legend.position = 'none')
-
-ggsave(paste0('figures/3Spp_WB/relImpOccMeanModWOcc_Size',simData$nReps,'_',ageForRI,'.png'), width = 9, height = 7, units='in')
-
-# uses 42 G - does not run
-#lmAll <- lm( lenMean ~ (temp + flow +  temp^2 + flow^2 + bkt + bnt + ats )*occ + temp * flow + bnt * bkt, data = stats3Spp[['bkt']][[1]] %>%               filter(iter %in% 1) )
-
-rIBKTBNT %>%
-  mutate(isEnvVar = ifelse(var %in% c(1,2,6), 1,0)) %>%
-  group_by(rowname,isEnvVar) %>%
-  summarize(meanMean = mean(value),
-            sumMean = sum(value),
-            sdMean = sd(value))
-# 25 reps
-#   rowname isEnvVar meanMean sumMean sdMean
-#   <chr>      <dbl>    <dbl>   <dbl>  <dbl>
-# 1 rBKT           0   0.0664   0.265  0.105
-# 2 rBKT           1   0.245    0.735  0.249
-# 3 rBNT           0   0.147    0.588  0.104
-# 4 rBNT           1   0.137    0.412  0.158
-
-# 100 reps
-#   rowname isEnvVar meanMean sumMean sdMean
-#   <chr>      <dbl>    <dbl>   <dbl>  <dbl>
-# 1 rBKT           0   0.0676    27.0 0.0914
-# 2 rBKT           1   0.243     73.0 0.200 
-# 3 rBNT           0   0.146     58.2 0.0922
-# 4 rBNT           1   0.139     41.8 0.131
-
-# ratio of env/(env+spp) means
-.243/(.243 + .0676) # bkt 0.7823567
-.139/(.139 + .146) # bnt 0.4877193
-
-.782/.488 # 1.60
-##############################################################
-############################################################
-
-# projOut <- list(stats3Spp, lm, relImp, anov, elapsed)
-load(paste0('D:/projects/linkedModels/data/out/projectionOut_N',simData$nReps,'.RData'))
-
-spp <- c('bkt','bnt','ats')
-# scenarios
-# Plot subsets of the simulations
-#means for each iter
-stats1 <- rbind(data.frame(stats3Spp[['bkt']][[1]]), 
-                data.frame(stats3Spp[['bnt']][[1]]), 
-                data.frame(stats3Spp[['ats']][[1]])) %>%
-          mutate(spp = rep(spp, each = nrow(stats3Spp[['bkt']][[1]])))
-
-#means for each occ
-stats2 <- rbind(data.frame(stats3Spp[['bkt']][[2]]), 
-                data.frame(stats3Spp[['bnt']][[2]]), 
-                data.frame(stats3Spp[['ats']][[2]])) %>%
-          mutate(spp = rep(spp, each = nrow(stats3Spp[['bkt']][[2]])))
-
-ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,lenMeanMean, group = spp)) +
-  geom_point(aes(color = spp) ) +
-  geom_line( aes(color = spp) ) +
-  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 1.5, flow == 0, bkt == 0, bnt == 0, ats == 0) ) +
-  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 1.5, flow == 0, bkt == 0, bnt == 0, ats == 0), linetype = 2 )
-
-ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,lenMeanMean, group = spp)) +
-  geom_point(aes(color = spp) ) +
-  geom_line( aes(color = spp) ) +
-  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 1.5, bkt == 0, bnt == 0, ats == 0) ) +
-  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 1.5, bkt == 0, bnt == 0, ats == 0), linetype = 2 )
-
-ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, ats == 0), aes(occ,lenMeanMean, group = spp)) +
-  geom_point(aes(color = spp) ) +
-  geom_line( aes(color = spp) ) +
-  geom_point(aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 0, bkt == 1.5, bnt == 0, ats == 0) ) +
-  geom_line( aes(color = spp), data = stats2 %>% filter(temp == 0, flow == 0, bkt == 1.5, bnt == 0, ats == 0), linetype = 2 )
-
-# trajectories across indep var values for different combinations. Trying to relate to relImp results
-#temp
-p1 <- ggplot( stats2 %>% filter(flow == 0, bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(temp))) +
-  geom_line(aes(color = factor(temp))) +
-  geom_point(aes(color = factor(temp))) +
-  facet_wrap(~spp)
-
-#flow
-p2 <- ggplot( stats2 %>% filter(temp == 0, bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(flow))) +
-  geom_line(aes(color = factor(flow))) +
-  geom_point(aes(color = factor(flow))) +
-  facet_wrap(~spp)
-
-#bkt
-p3 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(bkt))) +
-  geom_line(aes(color = factor(bkt))) +
-  geom_point(aes(color = factor(bkt))) +
-  facet_wrap(~spp)
-
-#bnt
-p4 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(bnt))) +
-  geom_line(aes(color = factor(bnt))) +
-  geom_point(aes(color = factor(bnt))) +
-  facet_wrap(~spp)
-
-#ats
-p5 <- ggplot( stats2 %>% filter(temp == 0, flow == 0, bkt == 0, bnt == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(ats))) +
-  geom_line(aes(color = factor(ats))) +
-  geom_point(aes(color = factor(ats))) +
-  facet_wrap(~spp)
-
-multiplot(p1, p2, p3, p4, p5, cols=2) # multiplot function is below
-
-#temp*flow
-ggplot( stats2 %>% filter(bkt == 0, bnt == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(flow))) +
-  geom_line(aes(color = factor(flow))) +
-  geom_point(aes(color = factor(flow))) +
-  facet_grid(temp~spp)
-
-#bkt*bnt
-ggplot( stats2 %>% filter(temp == 0, flow == 0, ats == 0, spp %in% c('bkt','bnt')), aes(occ,lenMeanMean, group = factor(bnt))) +
-  geom_line(aes(color = factor(bnt))) +
-  geom_point(aes(color = factor(bnt))) +
-  facet_grid(bkt~spp)
-
-# Multiple plot function
-#
-# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
-# - cols:   Number of columns in layout
-# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
-#
-# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
-# then plot 1 will go in the upper left, 2 will go in the upper right, and
-# 3 will go all the way across the bottom.
-#
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-
-  numPlots = length(plots)
-
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-
- if (numPlots==1) {
-    print(plots[[1]])
-
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
-
-# size=100
-# gr=10
-# sd=1
-# mean(rnorm(10000,size+gr,sd))
-# mean(size+rnorm(10000,gr,sd))
-
-```
-
-```{r overview graphs - Fig Appendix}
-
-pd <- position_dodge(0.65)
-d2$speciesGG <- factor(d2$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-# isYOY1
-d2 %>% group_by(speciesGG,riverOrderedGG,seasonGG,year) %>%
-  filter(isYOYDATA == 1, riverOrderedGG == 'West Brook') %>%
-  summarize(meanGR = mean(grLength, na.rm=T),
-            sdGR = sd(grLength, na.rm=T)) %>%
-ggplot( aes(year,meanGR, color=speciesGG, shape=speciesGG) ) +
-  geom_point( position=pd ) +
-  geom_errorbar( aes(ymin = meanGR - sdGR, ymax = meanGR + sdGR), width = 0.33, position=pd ) +
-  ggtitle( "isYOY = 1" ) +
-  labs( x = "Year", y = "Growth rate (mm/d)" ) +
-  scale_x_continuous( breaks = 1997:2015 ) +
-  theme_publication() +
-  theme( panel.grid.minor = element_blank(),
-         axis.text.x  = element_text(angle=90, vjust=0.5, size=8) ) +
-  facet_wrap(~seasonGG )
-
-ggsave('figures/3Spp_WB/grByYearYOY1.png')
-
-# isYOY2
-d2 %>% group_by(speciesGG,riverOrderedGG,seasonGG,year) %>%
-  filter(isYOYDATA == 2, riverOrderedGG == 'West Brook') %>%
-  summarize(meanGR = mean(grLength, na.rm=T),
-            sdGR = sd(grLength, na.rm=T)) %>%
-  ggplot( aes(year,meanGR, color=speciesGG, shape=speciesGG) ) +
-  geom_point( position=pd ) +
-  geom_errorbar( aes(ymin = meanGR - sdGR, ymax = meanGR + sdGR), width = 0.33, position=pd ) +
-  ggtitle( "isYOY = 2" ) +
-  labs( x = "Year", y = "Growth rate (mm/d)" ) +
-  scale_x_continuous( breaks = 1997:2015 ) +
-  theme_publication() +
-  theme( panel.grid.minor = element_blank(),
-         axis.text.x  = element_text(angle=90, vjust=0.5, size=8) ) +
-  facet_wrap(~seasonGG )
-
-ggsave('figures/3Spp_WB/grByYearYOY2.png')
-
-```
-
-```{r growth by movement}
-d2$isYOY <- ifelse( d2$ageInSamples <= 3, 1, 2 )
-d2 <-  d2 %>%
-  group_by(tag) %>%
-  mutate( lagRiverN = lead(riverN), 
-          riverNMove = paste(riverN,lagRiverN, sep="_")
-          #gr = (observedLength - lagObservedLength)/as.integer(detectionDate - lagDetectionDate)
-          )
-
-#d2 %>% filter(tag=='00088cbf44') %>% select(tag,riverN,lagRiverN,sampleNumber,observedLength,lagObservedLength)
-table(d2$riverNMove)
-
-cdMeansByisYOYriverNMove <- d2 %>%
-  #mutate( gr = (observedLength - lagObservedLength)/as.integer(detectionDate - lagDetectionDate) ) %>%
-  group_by(species,riverN,lagRiverN,season,isYOY) %>%
-  summarize( meanGr = mean( grLength, na.rm = TRUE ),
-             sdGr = sd( grLength, na.rm = TRUE ),
-             n = n()) %>%
-  mutate( meanGrUp = meanGr + sdGr,
-          meanGrDn = meanGr - sdGr
-          )
-
-# 
-ggplot( cdMeansByisYOYriverNMove %>% filter(!(season == 2 & isYOY == 1), species == "bkt", isYOY == 2) , aes(season, meanGr, color = factor(lagRiverN)) ) +
-  geom_errorbar(aes(ymin = meanGrDn, ymax = meanGrUp), width = 0.25, position = pd) +
-  geom_point( size = 2 ) +
-  geom_line()+
-  facet_wrap(~riverN)
-
-d2 %>% filter(riverN == 1 & lagRiverN == 4) %>% select(tag,river,sampleNumber)
-d2 %>% filter(riverN == 2, is.na(lagRiverN), !is.na(grLength)) %>% select(tag,riverN,sampleNumber,lagRiverN,observedLength,lagObservedLength,gr,grLength)
-d2 %>% filter(tag=='00088cc059') %>% select(tag,riverOrdered,riverN,lagRiverN,sampleNumber,observedLength,lagObservedLength,gr,grLength)
-```

### 2019-03-22 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
added graphs

@@
 
-modCSRY <- bam(observedLength ~ cohort * species * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
-               #s(year, bs = 'cr', by = speciesOrdered), 
-               s(year, bs = 'cr', k = kIn, by = speciesFactor),
-               data = dataIn, 
-               method = "REML")
-summary(modCSRY)
-plot(modCSRY, pages=10, scale = 0)
-
-predGridCSRY <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSRY$model$cohortSpeciesRiverOrdered),
-                            year = unique(modCSRY$model$year),
-                            ydayCumul = seq(250,1000, by = 30)) %>%
-                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
-                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
-                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
-                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
-                           speciesCohortOrdered = paste(speciesOrdered, cohortOrdered, sep = '_'),
-                           speciesRiverOrdered = paste(speciesOrdered, riverOrdered, sep = '_'),
-                           cohortCentered = cohortOrdered - cohortOffset,
-                           cohort = cohortOrdered,
-                           species = speciesOrdered,
-                           river = riverOrdered)
-
-predCSRY <- predict(modCSRY, newdata = predGridCSRY)#, type = 'lpmatrix')
-
-predGrid2CSRY <- bind_cols(predGridCSRY,as.data.frame(predCSRY))
-
-ggplot(predGrid2CSRY, aes(ydayCumul, predCSRY, color = species)) +
-  geom_point() + 
-  geom_line() +
-  ylim(50,300) +
-  facet_grid(river ~ cohortOrdered) +
-  ggtitle('modCSRY')
+# modCSRY <- bam(observedLength ~ cohort * species * river + 
+#                s(ydayCumul, bs = 'cr') +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
+#                #s(year, bs = 'cr', by = speciesOrdered), 
+#                s(year, bs = 'cr', k = kIn, by = speciesFactor),
+#                data = dataIn, 
+#                method = "REML")
+# summary(modCSRY)
+# plot(modCSRY, pages=10, scale = 0)
+# 
+# predGridCSRY <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSRY$model$cohortSpeciesRiverOrdered),
+#                             year = unique(modCSRY$model$year),
+#                             ydayCumul = seq(250,1000, by = 30)) %>%
+#                     mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
+#                            speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
+#                            riverOrdered = substring(cohortSpeciesRiverOrdered,10),
+#                            cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
+#                            speciesCohortOrdered = paste(speciesOrdered, cohortOrdered, sep = '_'),
+#                            speciesRiverOrdered = paste(speciesOrdered, riverOrdered, sep = '_'),
+#                            cohortCentered = cohortOrdered - cohortOffset,
+#                            cohort = cohortOrdered,
+#                            species = speciesOrdered,
+#                            river = riverOrdered)
+# 
+# predCSRY <- predict(modCSRY, newdata = predGridCSRY)#, type = 'lpmatrix')
+# 
+# predGrid2CSRY <- bind_cols(predGridCSRY,as.data.frame(predCSRY))
+# 
+# ggplot(predGrid2CSRY, aes(ydayCumul, predCSRY, color = species)) +
+#   geom_point() + 
+#   geom_line() +
+#   ylim(50,300) +
+#   facet_grid(river ~ cohortOrdered) +
+#   ggtitle('modCSRY')
 
 # trying a couple of models
-modCSRC <- bam(observedLength ~ cohort * species * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
-               #s(year, bs = 'cr', by = speciesOrdered), 
-               s(cohort, bs = 'cr', k = kIn, by = speciesFactor),
-               data = dataIn, 
-               method = "REML")
-summary(modCSRC)
-plot(modCSRC, pages=8, scale = 0)
-
-modC_S_RC <- bam(observedLength ~ cohort + species + river + 
+modCSR_C <- bam(observedLength ~ cohort * species * river + 
                s(ydayCumul, bs = 'cr') +
@@
                method = "REML")
-summary(modC_S_RC)
-plot(modCSRC, pages=8, scale = 0)
+summary(modCSR_C)
+plot(modCSR_C, pages=10, scale = 0)
 
@@
 
-save(mod,modCR,modCSR,modSC,modSR,modCSRY,modC_S_R, file = './data/out/mods.RData')
-AIC(modC_S_RC,modC_S_R,modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
+save(mod,modCR,modCSR,modSC,modSR,modC_S_R,modCSR_C, file = './data/out/mods.RData')
+AIC(modCSR_C,modC_S_R,modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
 

### 2019-03-22 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
added graphs

@@
 
-```{r gams for body size - model selection}
+```{r gams for body size - data prep}
 speciesIn <- 'bnt'
@@
 
+cohortOffset <- 2006
+
+# 1-way
+vB$cohortFactor <- factor(vB$cohort)
+vB$cohortOrdered <- factor(vB$cohort, levels = 2000:2013, ordered = T)
+#vB$cohortFactorCentered <- vB$cohortFactor - 2006
+vB$cohortCentered <- vB$cohort - cohortOffset
+
+vB$speciesFactor <- factor(vB$species, levels = c('bkt','bnt'))
+vB$speciesOrdered <- factor(vB$species, levels = c('bkt','bnt'), ordered = T)
+
+vB$riverFactor <- factor(vB$river, levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'))
+#vB$riverOrdered <- riverOrdered exists
+
+# 2-way
+vB$speciesRiverOrdered <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
+vB$speciesCohortOrdered <- factor(paste(vB$speciesOrdered,vB$cohortOrdered, sep = '_'), ordered = T)
+vB$cohortRiverOrdered <- factor(paste(vB$cohortOrdered,vB$riverOrdered, sep = '_'), ordered = T)
+
+# 3-way
+vB$cohortSpeciesRiverOrdered <- factor(paste(vB$cohortOrdered,vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
+
+vB$tagIndexFactor <- factor(vB$tagIndex)
+
+#vB$riverOrdered <- factor(vB$river,
+#                         levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'),
+#                         ordered = T)
+
+dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn)
+kIn <- 4
+```
+
+Run to here to start
+
+```{r raw data graphs}
 # number of knots from gam 
@@
   theme_publication()
+```
 
+```{r gam models}
 ####################################
@@
 
-cohortOffset <- 2006
-
-# 1-way
-vB$cohortFactor <- factor(vB$cohort)
-vB$cohortOrdered <- factor(vB$cohort, levels = 2000:2013, ordered = T)
-#vB$cohortFactorCentered <- vB$cohortFactor - 2006
-vB$cohortCentered <- vB$cohort - cohortOffset
-
-vB$speciesFactor <- factor(vB$species, levels = c('bkt','bnt'))
-vB$speciesOrdered <- factor(vB$species, levels = c('bkt','bnt'), ordered = T)
-
-vB$riverFactor <- factor(vB$river, levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'))
-#vB$riverOrdered <- riverOrdered exists
-
-# 2-way
-vB$speciesRiverOrdered <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
-vB$speciesCohortOrdered <- factor(paste(vB$speciesOrdered,vB$cohortOrdered, sep = '_'), ordered = T)
-vB$cohortRiverOrdered <- factor(paste(vB$cohortOrdered,vB$riverOrdered, sep = '_'), ordered = T)
-
-# 3-way
-vB$cohortSpeciesRiverOrdered <- factor(paste(vB$cohortOrdered,vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
-
-vB$tagIndexFactor <- factor(vB$tagIndex)
-
-#vB$riverOrdered <- factor(vB$river,
-#                         levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'),
-#                         ordered = T)
-
-dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn)
-kIn <- 4
 ################################################################################
@@
 
-```{r best models}
+```{r best models for each species}
 
@@
                s(cohort, bs = 'cr', k = kIn), 
-             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
+             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")
 
@@
 
+save(modBKT,modBNT,predBKT,predBNT,predGrid2Both, file = paste0('./data/out/bestModsBKTBNT.RData'))
+
+```
+
+```{r graphs of best models}
+# load('./data/out/bestModsBKTBNT.RData')
+
+# cohorts by river/species
 ggplot(predGrid2Both, aes(ydayCumul, pred, group = cohort, color = cohort)) +
@@
 #  guide_legend(title = 'Cohort') +
-  facet_grid(speciesGG ~ riverGG)
+  facet_grid(speciesGG ~ riverGG, scales = 'free_y')
+
+# specific splines
+library(gratia)
+cohortBKT <- evaluate_smooth(modBKT, "s(cohort)") %>% mutate(species = 'bkt')
+cohortBNT <- evaluate_smooth(modBNT, "s(cohort)") %>% mutate(species = 'bnt')
+cohortBoth <- bind_rows(cohortBKT,cohortBNT) %>% mutate(upperCI = est + se * 1.96, lowerCI = est - se * 1.96)
+
+cohortBoth$speciesGG <- factor(cohortBoth$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot(cohortBoth, aes(cohort, group = speciesGG)) + 
+  geom_ribbon(aes( ymin = lowerCI, ymax = upperCI), fill  = 'grey70') +
+  geom_line(aes(y = est, linetype = speciesGG), size = 1.5) +
+  labs(y = 'Effect size',
+       x = 'Cohort',
+       linetype = 'Species') +
+  theme_publication() +
+  theme(legend.position = "right", 
+        legend.direction = "vertical",
+        legend.title = element_text(face="plain"))
+
+# main spline over ydayCumul
+plotBKT = plot.gam(modBKT, select = 1)
+plotBNT = plot.gam(modBNT, select = 1)
+
+plotBKTydayCumul <- data.frame(x = plotBKT[[1]]$x, y = plotBKT[[1]]$fit, se = plotBKT[[1]]$se) %>% mutate(species = 'bkt')
+plotBNTydayCumul <- data.frame(x = plotBNT[[1]]$x, y = plotBNT[[1]]$fit, se = plotBNT[[1]]$se) %>% mutate(species = 'bnt')
+plotBothydayCumul <- bind_rows(plotBKTydayCumul,plotBNTydayCumul) %>% mutate(upperCI = y + se * 1.96, lowerCI = y - se * 1.96)
+
+plotBothydayCumul$speciesGG <- factor(plotBothydayCumul$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot(plotBothydayCumul, aes(x, group = speciesGG)) + 
+  geom_ribbon(aes( ymin = lowerCI, ymax = upperCI, fill  = speciesGG), alpha = 0.5) +
+  geom_line(aes(y = y, linetype = speciesGG), size = 1.5) +
+  scale_fill_manual(values = c('grey30', 'grey70')) +
+  labs(y = 'Predicted body size (mm)',
+       x = 'Days since age-0 Jan 1',
+       linetype = 'Species') +
+  theme_publication() +
+  theme(legend.position = "right", 
+        legend.direction = "vertical",
+        legend.title = element_text(face="plain"))
+
+# main
+
+ggplot(tmp2, aes(x,y)) + geom_line()
+
+appraise(modBKT)
 

### 2019-03-25 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
add salmon01, cohortStrength to models

@@
 ```{r gams for body size - data prep}
-speciesIn <- 'bnt'
+speciesIn <- 'bkt'
 depVarIn <- 'observedLength'
@@
 
+vB$salmon01 <- ifelse(vB$cohort <=2004, 1, 0)
+
+```
+
+```{r cohort strength}
+
+cohortStrength <- vB %>%
+  distinct(tagIndex, species, metaPop01, cohort) %>% #ignoring 10 fish that were in both o'bear and WB
+  group_by(species, metaPop01, cohort) %>%
+  summarize(cohortStrength = n()) %>%
+  filter(!(cohort <2000 & species !='ats')) %>% # no trib data
+  mutate(cohortStrengthScaled = scale(cohortStrength))
+
+vB <- left_join(vB, cohortStrength)
+
+ggplot(cohortStrength, aes(cohort, cohortStrengthScaled, group = species, linetype = species, shape = species)) +
+  geom_point(size = 2) +
+  geom_line(size = 1.1, color = 'grey40') +
+  labs(x = 'Cohort',
+       y = 'Cohort strength') +
+  theme_publication() +
+  facet_grid(~metaPop01)
+
+```
+
+```{r dataIn}
 dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn)
-kIn <- 4
 ```
@@
 
-```{r gam models}
+```{r gam models - only need to run for model selection}
 ####################################
@@
 
-AIC(mod,modC,modR,modC_R,modCR,modCR_c) %>% rownames_to_column() %>% arrange(AIC)
-save(mod,modC,modR,modC_R,modCR,modCR_c, file = paste0('./data/out/mods_',speciesIn,'.RData'))
-######
+##########################################
+# add in salmon01 and cohortStrengthScaled
 
-##########################################
+system.time(
+modCR_S_c <- bam(observedLength ~ cohort * river + factor(salmon01) +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn), 
+             data = dataIn, method = "REML")
+)
+system.time(
+modCR_c_cS <- bam(observedLength ~ cohort * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
+             data = dataIn, method = "REML")
+)
+
+system.time(
+modCR_S_c_cS <- bam(observedLength ~ cohort * river + factor(salmon01) +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
+             data = dataIn, method = "REML")
+)
+summary(modCR_S_c_cS)
+
+system.time(
+modCR_SR_c_cS <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
+             data = dataIn, method = "REML")
+)
+summary(modCR_SR_c_cS)
+
+system.time(
+modCR_SR_c_cSR <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered), 
+             data = dataIn, method = "REML")
+)
+summary(modCR_SR_c_cSR)
+#plot(modCR_S_c_cS, scale = 0, pages=8)
+#plotGamVarBySpecies(var = 's(cohortStrengthScaled)')
+
+AIC(mod,modC,modR,modC_R,modCR, modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR) %>% rownames_to_column() %>% arrange(AIC)
+
+save(mod,modC,modR,modC_R,modCR,modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, file = paste0('./data/out/mods_',speciesIn,'.RData'))
 
@@
 
-modBKT <- bam(observedLength ~ cohort * river + 
+modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) +
                s(ydayCumul, bs = 'cr') +
@@
                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn), 
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
@@
 predGridBKT <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
-                           ydayCumul = seq(250,1000, by = 30)) %>%
+                           ydayCumul = seq(250,1000, by = 30),
+                           salmon01 = c(0,1),
+                           cohortStrengthScaled = seq(-2,2, by = 0.1)) %>%
                     mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
@@
 
-modBNT <- bam(observedLength ~ cohort * river + 
+modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) +
                s(ydayCumul, bs = 'cr') +
@@
                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn), 
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")
@@
 predGridBNT <- expand.grid(cohortRiverOrdered = unique(modBNT$model$cohortRiverOrdered),
-                           ydayCumul = seq(250,1000, by = 30)) %>%
+                           ydayCumul = seq(250,1000, by = 30),
+                           salmon01 = c(0,1),
+                           cohortStrengthScaled = seq(-2,2, by = 0.1)) %>%
                     mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
@@
 # load('./data/out/bestModsBKTBNT.RData')
+source('./functions/plotGams.R')
 
 # cohorts by river/species
-ggplot(predGrid2Both, aes(ydayCumul, pred, group = cohort, color = cohort)) +
-  geom_point() + 
-  geom_line() +
+ggplot(predGrid2Both %>% filter(cohortStrengthScaled == 0, salmon01 == 0), aes(ydayCumul, pred, group = cohort, color = cohort)) +
+  #geom_point() + 
+  geom_line(size = 1.1) +
   labs(y = 'Predicted body size (mm)',
@@
 
-# specific splines
-library(gratia)
-cohortBKT <- evaluate_smooth(modBKT, "s(cohort)") %>% mutate(species = 'bkt')
-cohortBNT <- evaluate_smooth(modBNT, "s(cohort)") %>% mutate(species = 'bnt')
-cohortBoth <- bind_rows(cohortBKT,cohortBNT) %>% mutate(upperCI = est + se * 1.96, lowerCI = est - se * 1.96)
-
-cohortBoth$speciesGG <- factor(cohortBoth$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-ggplot(cohortBoth, aes(cohort, group = speciesGG)) + 
-  geom_ribbon(aes( ymin = lowerCI, ymax = upperCI), fill  = 'grey70') +
-  geom_line(aes(y = est, linetype = speciesGG), size = 1.5) +
-  labs(y = 'Effect size',
-       x = 'Cohort',
-       linetype = 'Species') +
-  theme_publication() +
-  theme(legend.position = "right", 
-        legend.direction = "vertical",
-        legend.title = element_text(face="plain"))
+# specific splines by species
+plotCohort <- plotGamVarBySpecies(var = 's(cohort)', xAxisLab = 'Cohort')
 
 # main spline over ydayCumul
-plotBKT = plot.gam(modBKT, select = 1)
-plotBNT = plot.gam(modBNT, select = 1)
+plotBKT = plot.gam(modBKT, select = 1, scale = 0)
+plotBNT = plot.gam(modBNT, select = 1, scale = 0)
 
@@
 
-# main
+# cohort effects
 

### 2019-03-25 functions/plotGams.R

@@ -1,0 +1,48 @@
+
+plotGamVarBySpecies <- function(var = 's(cohort)', xAxisLab = 'Cohort', bkt = modBKT, bnt = modBNT){
+  require(gratia)
+  bkt <- evaluate_smooth(modBKT, var) %>% mutate(species = 'bkt')
+  bnt <- evaluate_smooth(modBNT, var) %>% mutate(species = 'bnt')
+  both <- bind_rows(BKT,BNT) %>% mutate(upperCI = est + se * 1.96, lowerCI = est - se * 1.96)
+  
+  both$speciesGG <- factor(both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+  
+  gg <- 
+    ggplot(both, aes(cohort, group = speciesGG)) + 
+      geom_ribbon(aes( ymin = lowerCI, ymax = upperCI), fill  = 'grey70') +
+      geom_line(aes(y = est, linetype = speciesGG), size = 1.5) +
+      labs(y = 'Effect size',
+           x = xAxisLab,
+           linetype = 'Species') +
+      theme_publication() +
+      theme(legend.position = "right", 
+            legend.direction = "vertical",
+            legend.title = element_text(face="plain"))
+  gg
+} 
+  # 
+  # # main spline over ydayCumul
+  # plotBKT = plot.gam(modBKT, select = 1, scale = 0)
+  # plotBNT = plot.gam(modBNT, select = 1, scale = 0)
+  # 
+  # plotBKTydayCumul <- data.frame(x = plotBKT[[1]]$x, y = plotBKT[[1]]$fit, se = plotBKT[[1]]$se) %>% mutate(species = 'bkt')
+  # plotBNTydayCumul <- data.frame(x = plotBNT[[1]]$x, y = plotBNT[[1]]$fit, se = plotBNT[[1]]$se) %>% mutate(species = 'bnt')
+  # plotBothydayCumul <- bind_rows(plotBKTydayCumul,plotBNTydayCumul) %>% mutate(upperCI = y + se * 1.96, lowerCI = y - se * 1.96)
+  # 
+  # plotBothydayCumul$speciesGG <- factor(plotBothydayCumul$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+  # 
+  # gg <- 
[…]

### 2019-03-28 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
added spp-specific cohortStrength effects

@@
   summarize(cohortStrength = n()) %>%
-  filter(!(cohort <2000 & species !='ats')) %>% # no trib data
+  filter(!(cohort < 2000 & species !='ats')) %>% # no trib data
   mutate(cohortStrengthScaled = scale(cohortStrength))
@@
 vB <- left_join(vB, cohortStrength)
+
+cohortStrengthWide <- spread(cohortStrength, species, cohortStrengthScaled) %>%
+  rename(cohortStrengthScaledBKT = bkt,
+         cohortStrengthScaledBNT = bnt,
+         cohortStrengthScaledATS = ats
+         ) %>%
+  dplyr::select(-cohortStrength)
+
+vB <- left_join(vB, cohortStrengthWide)
+
+vB$cohortStrengthScaledBKT <- ifelse(is.na(vB$cohortStrengthScaledBKT), -2.5, vB$cohortStrengthScaledBKT)
+vB$cohortStrengthScaledBNT <- ifelse(is.na(vB$cohortStrengthScaledBNT), -2.5, vB$cohortStrengthScaledBNT)
+vB$cohortStrengthScaledATS <- ifelse(is.na(vB$cohortStrengthScaledATS), -2.5, vB$cohortStrengthScaledATS)
 
@@
 
-AIC(mod,modC,modR,modC_R,modCR, modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR) %>% rownames_to_column() %>% arrange(AIC)
+# effects of each species together - above is only conspecific
+system.time(
+modCR_SR_c_cSR3 <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
+             data = dataIn, method = "REML")
+)
+summary(modCR_SR_c_cSR3)
+#plot(modCR_SR_c_cSR3, scale = 0, pages=5)
+
+system.time(
+modCR_SR_c_cSR3a <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn) +
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
+             data = dataIn, method = "REML")
+)
+summary(modCR_SR_c_cSR3a)
+#plot(modCR_SR_c_cSR3a, scale = 0, pages=5)
+
+system.time(
+modCR_SR_c_cSR3b <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
+             data = dataIn, method = "REML")
+)
+summary(modCR_SR_c_cSR3b)
+#plot(modCR_SR_c_cSR3b, scale = 0, pages=5)
+
+AIC(mod,modC,modR,modC_R,modCR, modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, modCR_SR_c_cSR3, modCR_SR_c_cSR3a,modCR_SR_c_cSR3b) %>% rownames_to_column() %>% arrange(AIC)
 
 save(mod,modC,modR,modC_R,modCR,modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, file = paste0('./data/out/mods_',speciesIn,'.RData'))
+
+# BKT
+
+# BNT
+#             rowname       df      AIC
+# 1  modCR_SR_c_cSR3a 89.52290 231873.4
+# 2  modCR_SR_c_cSR3b 89.54674 231873.6
+# 3     modCR_SR_c_cS 84.40505 231935.7
+# 4      modCR_S_c_cS 83.47382 231962.5
+# 5   modCR_SR_c_cSR3 88.15840 232100.8
+# 6        modCR_c_cS 80.92171 232142.0
+# 7    modCR_SR_c_cSR 80.89406 232201.5
+# 8         modCR_S_c 79.53048 232281.7
+# 9           modCR_c 77.80808 232420.2
+# 10            modCR 78.58400 233816.0
+# 11           modC_R 53.21199 234214.3
+# 12              mod 13.99759 238544.8
+# 13             modR 14.97078 243472.0
+# 14             modC 48.73277 431440.2
 
@@
 # Brook trout
-
-modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) +
+# using salmonid01 * river because we need diff effects per river - even though AIC is worse for this model 
+modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
                s(ydayCumul, bs = 'cr') +
@@
 
-modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) +
+modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
                s(ydayCumul, bs = 'cr') +
@@
                s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")
@@
 # specific splines by species
-plotCohort <- plotGamVarBySpecies(var = 's(cohort)', xAxisLab = 'Cohort')
+(plotCohort <- plotGamVarBySpecies(var = 's(cohort)', xAxisLab = 'Cohort'))
+(plotCohortStrength <- plotGamVarBySpecies(var = 's(cohortStrengthScaled)', xAxisLab = 'Cohort Strength'))
+
+#
 

### 2019-06-27 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
update model selection for gams

@@
 library(relaimpo)
+library(GGally)
 library(linkedModels)
@@
 # 2-way
+vB$speciesRiverFactor <- factor(paste(vB$speciesFactor,vB$riverFactor, sep = '_'))
+vB$cohortRiverFactor <- factor(paste(vB$cohortFactor,vB$riverFactor, sep = '_'))
+vB$seasonRiverFactor <- factor(paste(vB$season,vB$riverFactor, sep = '_'))
+
 vB$speciesRiverOrdered <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
@@
 
-vB$salmon01 <- ifelse(vB$cohort <=2004, 1, 0)
-
+vB$salmon01 <- ifelse(vB$cohort <= 2004 & vB$river == 'west brook', 1, 0)
+
+vB$speciesGG <- factor(vB$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+vB$seasonGG <- factor(vB$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+
+# scaling temperature and flow
+vB$meanTemperatureScaled <- as.numeric(scale(vB$meanTemperature))
+vB$meanFlowScaled <- as.numeric(scale(vB$meanFlow))
+vB$meanTemperatureFlowScaled <- vB$meanTemperatureScaled * vB$meanFlowScaled
+
+seasonalTF <- vB %>% group_by(season) %>%
+  summarize(meanTBySeason = mean(meanTemperature, na.rm = TRUE),
+            sdTBySeason = sd(meanTemperature, na.rm = TRUE),
+            meanFBySeason = mean(meanFlow, na.rm = TRUE),
+            sdFBySeason = sd(meanFlow, na.rm = TRUE))
+
+vB <- left_join(vB,seasonalTF) %>%
+  mutate(meanTemperatureScaledBySeason = (meanTemperature - meanTBySeason)/sdTBySeason, 
+         meanFlowScaledBySeason = (meanFlow - meanFBySeason)/sdFBySeason)
+
+vB$meanTemperatureFlowScaledBySeason <- vB$meanTemperatureScaledBySeason * vB$meanFlowScaledBySeason
 ```
@@
 
-ggplot(cohortStrength, aes(cohort, cohortStrengthScaled, group = species, linetype = species, shape = species)) +
+ggplot(cohortStrength, aes(cohort, cohortStrengthScaled, group = species, linetype = species, shape = species, color = species)) +
   geom_point(size = 2) +
@@
 
-```{r gam models - only need to run for model selection}
-####################################
-# gam models
-
-################################################################################
-
-# base model
-
-mod <- bam(observedLength ~ cohort * river + 
+```{r raw data histos}
+
+ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt'), river == "west brook"), aes(observedLength)) +
+  geom_freqpoly(aes(color = species)) +
+  theme_publication() +
+  facet_grid(season ~ year)
+
+# season 3
+ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt'), river == "west brook", season == 3), aes(observedLength)) +
+  geom_freqpoly(aes(color = speciesGG), binwidth = 6, size = 1.35) +
+  theme_publication() +
+  theme(strip.text.x = element_text(size = 8), 
+        strip.background = element_blank(),
+        legend.title = element_text("Species"),
+        legend.position = "right",
+        legend.direction = "vertical") + 
+  labs(color='Species') +
+  scale_x_continuous("Fork Length (mm)", limits = c(49,200)) +
+  scale_y_continuous("Count") +
+  #scale_color_manual(values = c("lightblue", "lightgreen")) +
+  scale_colour_brewer(palette = "Paired")+
+  facet_wrap(~ year)
+
+ggsave('./figures/season3freqpoly.png', width = 11, height = 8, units='in')
+
+# 1 cohort
+ggplot(vB %>% filter(cohort == 2009, species %in% c('bkt', 'bnt'), river == "west brook"), aes(observedLength)) +
+  geom_freqpoly(aes(color = speciesGG), binwidth = 6, size = 1.35) +
+  theme_publication() +
+  theme(strip.text.x = element_text(size = 8), 
+        strip.background = element_blank(),
+        legend.title = element_text("Species"),
+        legend.position = "right",
+        legend.direction = "vertical") + 
+  labs(color='Species') +
+  scale_x_continuous("Fork Length (mm)", limits = c(49,200)) +
+  scale_y_continuous("Count") +
+  #scale_color_manual(values = c("lightblue", "lightgreen")) +
+  scale_colour_brewer(palette = "Paired")+
+  facet_grid(season ~ year)
+
+ggsave('./figures/cohort2009freqpoly.png', width = 11, height = 8, units='in')
+
+```
+
+```{r mean lengths: from MS_Analysis_3spp_WB.Rmd. For Grenada talk}
+
+# Has mean size changed over time?
+means <- 
+  cd %>% mutate( age = year - cohort ) %>%
+    filter( riverOrdered == 'west brook', ageInSamples %in% 1:10, age < 3, year < 2016 ) %>%
+    group_by(species,ageInSamples,year,season,age) %>%
+    summarize(meanLen = mean(observedLength,na.rm = TRUE),
+              sdLen = sd(observedLength,na.rm = TRUE),
+              meanGR = mean(grLength,na.rm = TRUE),
+              sdGR = sd(grLength,na.rm = TRUE)
+              )
+
+lenSlopes <- means %>%
+  group_by( species, ageInSamples, season, age ) %>%
+  nest()
+
+lengthModel <- function(df){
+  lm( meanLen ~ year, data = df )
+}
+
+grModel <- function(df){
+  lm( meanGR ~ year, data = df )
+}
+
+lenSlopes <- lenSlopes %>%
+  mutate( model = map(lenSlopes$data, lengthModel),
+          modelGR = map(lenSlopes$data, grModel))
+
+###################################
+######### len here, gr below ######
+lenSlopesTidy <- lenSlopes %>%
+  mutate(glance = map(model, broom::tidy)) %>% 
+  unnest(glance, .drop = TRUE) %>%
+  filter(term == 'year') %>%
+  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
+  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
+  mutate( sig = ifelse(p.value < 0.0001,"***",sig))
+
+means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+lenSlopesTidy$seasonGG <- factor(lenSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+lenSlopesTidy$speciesGG <- factor(lenSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanLen, group= speciesGG)) +
+  geom_point( aes(color = speciesGG), size = 4) +
+  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
+  labs( x = "Year", y = "Mean body length (mm)" ) +
+
+  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age<2), 
+            aes(x=2000,y=225, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age<2), 
+            aes(x=2000,y=195, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+
+  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), 
+            aes(x=2011.8,y=104, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), 
+            aes(x=2011.8,y=74, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+
+
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+      #  legend.position = "right",
+     #   legend.direction = "vertical"
+        legend.position = "bottom",
+        legend.direction = "horizontal") + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG ~ age, scales='free')
+
+ggsave('./figures/meanSizes.png', width = 11, height = 11, units='in')
+
+# one season/year only
+ggplot( means %>% filter(year >= 2000, species != 'ats', ageInSamples == 8), aes(year,meanLen, group= speciesGG)) +
+  geom_point( aes(color = speciesGG), size = 4) +
+  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
+  labs( x = "Year", y = "Mean body length (mm)" ) +
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+      #  legend.position = "right",
+     #   legend.direction = "vertical"
+                legend.position = "bottom",
+        legend.direction = "horizontal") + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG ~ age)
+ggsave('./figures/meanSizesAIS_1.png', width = 11, height = 8, units='in')
+
+# is a model with salmon presence absence better than a year model?
+means$ats01 <- ifelse(means$year <= 2005, 1, 0)
+
+modMeans1 = lm(meanLen ~ year * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
+modMeans2 = lm(meanLen ~ year * factor(ats01) * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
+modMeans3 = lm(meanLen ~ factor(ats01) * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
+modMeans4 = lm(meanLen ~ factor(ats01)*year, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
+AIC(modMeans1, modMeans2, modMeans3, modMeans4)
+
+# one season/year only
+ggplot( means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1), aes(year,meanLen, group= speciesGG)) +
+  geom_point( aes(color = speciesGG), size = 4) +
+  geom_smooth( aes(group = factor(ats01)), color = 'black', se = FALSE, method = 'lm') +
+  labs( x = "Year", y = "Mean body length (mm)" ) +
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+      #  legend.position = "right",
+     #   legend.direction = "vertical"
+        legend.position = "bottom",
+        legend.direction = "horizontal") + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG ~ age)
+
+ggsave('./figures/meanSizesAIS_1_ats01.png', width = 11, height = 8, units='in')
+
+# mean increases 
+lenSlopesTidy %>%
+  filter(species != 'ats') %>%
+  summarise( means = mean(estimate),
+             sd = sd(estimate))
+
+# mean increases by species
+lenSlopesTidy %>%
+  group_by(species) %>%
+  summarise( means = mean(estimate),
+             sd = sd(estimate))
+
+# do increases (slopes) increase over age? Yes
+ggplot(lenSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
+  geom_point(size = 5) +
+  geom_smooth(method='lm', se=FALSE, size = 1.5) +
+  scale_x_continuous(breaks = 1:10, labels = c('0 Fall','0 Winter','1 Spring','1 Summer','1 Fall','1 Winter','2 Spring','2 Summer','2 Fall','2 Winter')) +
+  labs( x = "Age Season", y = "Slope of mean body length over years" ) +
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+        legend.position = "bottom",
+        legend.direction = "horizontal",
+        axis.text.x = element_text(angle = 40, hjust = 1)) + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75)
+
+ggsave(paste0('figures/meanLengthsOverAge.png'), width = 11, height = 10, units='in')
+
+# species diff in increases in length?
+sppEffect <- (lm(estimate ~ ageInSamples * (species), data = lenSlopesTidy))
+summary(sppEffect)
+anova(sppEffect)
+
+# mean diff in slopes btw bkt and bnt
+lenSlopesTidy %>%
+  filter(species != 'ats') %>%
+  group_by(species) %>%
+  summarise(means = mean(estimate))
+
+#####################################  
+#################################
+# growth rate
+grSlopesTidy <- lenSlopes %>%
+  mutate(glance = map(modelGR, broom::tidy)) %>% 
+  unnest(glance, .drop = TRUE) %>%
+  filter(term == 'year') %>%
+  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
+  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
+  mutate( sig = ifelse(p.value < 0.0001,"***",sig))
+
+##
+means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanGR, group= speciesGG)) +
+  geom_point( aes(color = speciesGG), size = 4) +
+  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
+  labs( x = "Year", y = "Mean growth rate (mm/d)" ) +
+  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age<2), 
+            aes(x=2000,y=0.68, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age<2), 
+            aes(x=2000,y=0.55, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+
+  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), 
+            aes(x=2011.5,y=0.64, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), 
+            aes(x=2011.5,y=0.51, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+      #  legend.position = "right",
+     #   legend.direction = "vertical"
+        legend.position = "bottom",
+        legend.direction = "horizontal") + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG~age)
+
+ggsave('./figures/meanGRs.png', width = 11, height = 11, units='in')
+
+# mean increases 
+grSlopesTidy %>%
+  filter(species != 'ats') %>%
+  summarise( means = mean(estimate),
+             sd = sd(estimate))
+
+# mean increases by species
+grSlopesTidy %>%
+  group_by(species) %>%
+  summarise( means = mean(estimate),
+             sd = sd(estimate))
+
+# do increases (slopes) increase over age? Yes
+ggplot(grSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
+  geom_point(size = 5) +
+  geom_smooth(method='lm', se=FALSE) +  
+  geom_hline(yintercept = 0) +
+  scale_x_continuous(breaks = 1:10, labels = c('0 Fall','0 Winter','1 Spring','1 Summer','1 Fall','1 Winter','2 Spring','2 Summer','2 Fall','2 Winter')) +
+  labs( x = "Age Season", y = "Slope of mean growth rate over years" ) +
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+        legend.position = "bottom",
+        legend.direction = "horizontal",
+        axis.text.x = element_text(angle = 40, hjust = 1)) + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75)
+
+ggsave('./figures/meanGRsOverAge.png', width = 11, height = 11, units='in')
+
+summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bkt')))
+summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bnt')))
+
+# species diff in increases in gr?
+sppEffectGR <- (lm(estimate ~ ageInSamples * (species), data = grSlopesTidy))
+summary(sppEffectGR)
+anova(sppEffectGR)
+
+# mean diff in slopes btw bkt and bnt
+grSlopesTidy %>%
+  filter(species != 'ats') %>%
+  group_by(species) %>%
+  summarise(means = mean(estimate))
+
+```
+
+```{r size-dep imm or mort? For Spain talk}
+
+### for losses
+
+cd01 <-  cd %>%
+  filter(knownZ == 1) %>%
+  group_by(tagIndex) %>%
+  mutate(first = min(sampleNumber),
+         firstObs = sampleNumber == first,
+         last = max(sampleNumber),
+         lastObs = sampleNumber == last,
+         notLastObs = !lastObs,
+         nObs = n(),
+         age = year - cohort
+  ) %>%
+  ungroup()
+
+meansByAIS <- cd01 %>%
+  group_by(species, ageInSamples, cohort, river) %>%
+  summarise(meanLen = mean(observedLength, na.rm =T),
+            sdLen = sd(observedLength, na.rm = T),
+            n = n())
+cd01 <- left_join(cd01, meansByAIS) %>%
+  mutate(observedLengthStd = (observedLength - meanLen) / sdLen,
+         cohortStd = cohort - 2007)
+
+#cd %>% select(tagIndex,sampleNumber, firstObs,lastObs, knownZ, observedLength) %>% head(50) %>% as.data.frame()
+cd01Mod <- cd01 %>% filter(species %in% c('bkt', 'bnt'), river == 'west brook', ageInSamples %in% 1:10, cohort %in% 2000:2014)
+
+ggplot(cd01Mod, aes(observedLengthStd, notLastObs*1, color = factor(species), group = factor(species))) +
+  geom_point(alpha = 0.01) +
+  geom_smooth(method = "glm", 
+    method.args = list(family = "binomial"), 
+    se = FALSE) +
+ # scale_color_viridis_c() +
+  facet_wrap( ~ cohort, scales = 'free')
+
+#### by species and ageInSamples
+
+sensSlopes <- cd01Mod %>%
+  group_by( species, ageInSamples ) %>%
+  nest()
+
+survModel <- function(df){
+  glm(notLastObs*1 ~ observedLengthStd + cohortStd, family = 'binomial', data = df )
+}
+enterModel <- function(df){
+  glm(firstObs*1 ~ observedLengthStd + cohortStd, family = 'binomial', data = df )
+}
+
+sensSlopes <- sensSlopes %>%
+  mutate(modelSurv = map(sensSlopes$data, survModel),
+         modelEnter = map(sensSlopes$data, enterModel))
+
+sensSlopesTidy <- sensSlopes %>%
+  mutate(surv = map(modelSurv, broom::tidy),
+         enter = map(modelEnter, broom::tidy)) %>% 
+  unnest(surv, enter, .drop = TRUE) %>%
+  filter(term == 'cohortStd') %>% # cohort effect
+  rename(estimateSurv = estimate, estimateEnter = estimate1) %>%
+  select(species,ageInSamples,estimateSurv, estimateEnter) #%>%
+  #gather("type", "estimate", estimateSurv:estimateEnter)
+
+ggplot(sensSlopesTidy, aes(ageInSamples, estimate, color = type)) +
+  geom_point() +
+  geom_line() +
+  facet_wrap(~species)
+
+```
+
+```{r dummy graphs for Spain talk}
+
+sizesForGraph <- data.frame(year = c(2000,2015), surv = c(100, 150))
+ggplot(sizesForGraph, aes(year,surv)) +
+  geom_line(size = 2) +
+  labs(x = 'Year', y = 'Mean body length (mm)') +
+  ylim(80,200) +
+  theme_publication(base_size = 36)
+ggsave('./figures/sizesDummy.png', width = 11, height = 11, units='in')
+
+survForGraph <- data.frame(size = c(-2,2), surv = c(0.25,0.75))
+ggplot(survForGraph, aes(size,surv)) +
+  geom_line(size = 2) +
+  labs(x = 'Standardized length', y = 'Probability of seen again') +
+  theme_publication(base_size = 36)
+ggsave('./figures/pSurvivalDummy.png', width = 11, height = 11, units='in')
+
+firstForGraph <- data.frame(size = c(-2,2), surv = c(0.35,0.65))
+ggplot(firstForGraph, aes(size,surv)) +
+  geom_line(size = 2) +
+  labs(x = 'Standardized length', y = 'Probability of first observation') +
+  theme_publication(base_size = 36)
+ggsave('./figures/pFirstDummy.png', width = 11, height = 11, units='in')
+
+sizesForGraph <- data.frame(year = c(2000,2015), surv = c(100, 150))
+ggplot(sizesForGraph, aes(year,surv)) +
+  geom_line(size = 3) +
+  labs(x = 'Year', y = 'Mean body length (mm)') +
+  ylim(80,200) +
+  theme_publication(base_size = 36)
+ggsave('./figures/sizesDummy.png', width = 11, height = 11, units='in')
+
+#### show interactions
+
+survForGraph <- data.frame(ind = c(1,1,2,2,3,3), size = c(-2,2), surv = c(0.52,0.55,0.51,0.24,0.54,0.82))
+ggplot(survForGraph, aes(size,surv, group = ind)) +
+  geom_line(size = 2) +
+  labs(x = 'Standardized length', y = 'Probability of seen again') +
+  theme_publication(base_size = 36)
+ggsave('./figures/pSurvivalDummy3.png', width = 11, height = 11, units='in')
+
+firstForGraph <- data.frame(ind = c(1,1,2,2,3,3), size = c(-2,2), surv = c(0.42,0.45,0.41,0.14,0.44,0.72))
+ggplot(firstForGraph, aes(size,surv, group = ind)) +
+  geom_line(size = 2) +
+  labs(x = 'Standardized length', y = 'Probability of first observation') +
+  theme_publication(base_size = 36)
+ggsave('./figures/pFirstDummy3.png', width = 11, height = 11, units='in')
+
+```
+
+```{r get betas from growth model for sensitivities}
+load("D:/projects/linkedModels/data/out/allBetas.RData") #loads df 'q'
+qFig1 <- q %>% filter(riverGG == 'WB', species %in% c('bkt','bnt'))
+qFig1$season <- as.numeric(as.character(qFig1$season))
+qFig1$isYOY <- as.numeric(as.character(qFig1$isYOY))
+qFig <- qFig1 %>%
+  mutate( varText = tolower(betaGG)) %>%
+  mutate( varText = ifelse(varText == 'temp*flow','temp:flow', varText)) %>%
+  mutate( varText = ifelse(varText == 'bnt*bkt','bkt:bnt', varText)) %>%
+  select(isYOY,season,species,median,varText)
+qFig$seasonGG <- factor(qFig$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+#get slopes of env variables
+
+load('D:/projects/linkedModels/data/out/flowSlopesTidy.RData')
+flowSlopesTidy2 <- flowSlopesTidy %>% filter(riverOrderedGG == 'West Brook') %>% select(estimate, seasonGG) %>% mutate(varText = 'flow')
+load('D:/projects/linkedModels/data/out/tempSlopesTidy.RData')
+tempSlopesTidy2 <- tempSlopesTidy %>% filter(riverOrderedGG == 'West Brook') %>% select(estimate, seasonGG) %>% mutate(varText = 'temp')
+load('D:/projects/linkedModels/data/out/nSlopesTidy.RData')
+nSlopesTidy2 <- nSlopesTidy %>% select(estimate, season, species) %>% mutate(varText = species)
+nSlopesTidy2$seasonGG <- factor(nSlopesTidy2$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+nSlopesTidy2 <- nSlopesTidy2 %>% select(estimate,seasonGG,varText)
+
+allSlopesTidy <- bind_rows(nSlopesTidy2, tempSlopesTidy2, flowSlopesTidy2)
+
+allSlopesTidyWBetas <- left_join(qFig, allSlopesTidy, by = c('seasonGG', 'varText')) %>% mutate(sens = median * estimate)
+
+ggplot(allSlopesTidyWBetas%>%filter(!is.na(sens), varText != 'ats'), aes(seasonGG, sens, color = varText)) +
+  geom_point() +
+  facet_grid(species + isYOY ~ varText, scales = 'free')
+```
+
+```{r }
+############# 1st attempt at imm em
+mod1 <- glm(notLastObs*1 ~ observedLengthStd * cohortStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+#mod2 <- glm(notLastObs*1 ~ observedLengthStd * factor(cohortStd) * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+mod3 <- glm(notLastObs*1 ~ observedLengthStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+
+AIC(mod1,mod3)
+
+summary(mod1)
+anova(mod1, test="Chi")
+
+pred <- expand.grid(observedLengthStd = seq(-2,2, length.out = 50), cohortStd = unique(cd01Mod$cohortStd), species = c('bkt', 'bnt'), ageInSamples = unique(cd01Mod$ageInSamples))
+
+pred$pred <- predict.glm(mod1, newdata = pred, type = 'response')
+
+u <- cd01Mod %>% distinct(age, season, ageInSamples) %>% arrange(ageInSamples) %>% filter(!(age==2 & ageInSamples == 6))
+pred <- left_join(pred, u)
+
+pred$seasonGG <- factor(pred$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+pred$cohort <- pred$cohortStd + 2007
+pred$speciesGG <- factor(pred$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot(pred, aes(observedLengthStd, pred, color = (cohortStd), group = (cohortStd))) +
+  geom_point() +
+  scale_color_viridis_c() +
+  scale_y_continuous("Probability of survival") +
+  facet_grid(species ~ ageInSamples, scales = 'free')
+
+ggplot(pred %>% filter(cohortStd %in% 0, age < 3), aes(observedLengthStd, pred, color = speciesGG)) +
+  geom_point(size = 2) +
+  scale_x_continuous("Standardized length") +
+  scale_y_continuous(lim = c(0,1),"Probability of seen again") +
+  theme_publication(base_size = 28) +
+  theme(legend.title = element_text("Species"),
+        legend.position = "bottom",
+        legend.direction = "horizontal"
+        ) + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG ~ age)
+
+ggsave('./figures/pSurvival.png', width = 11, height = 11, units='in')
+
+################################
+# gains - immigration, firstObs
+
+ggplot(cd01Mod, aes(observedLengthStd, firstObs*1, color = factor(species), group = factor(species))) +
+  geom_jitter(alpha = 0.1) +
+  geom_smooth(method = "glm", 
+    method.args = list(family = "binomial"), 
+    se = FALSE) +
+ # scale_color_viridis_c() +
+  xlim(c(-2.75,2.75)) +
+  facet_grid(ageInSamples ~ cohort)
+
+ggplot(cd01Mod%>%filter(ageInSamples == 3), aes(observedLengthStd, firstObs*1, color = factor(species), group = factor(species))) +
+  geom_jitter(alpha = 0.1) +
+  geom_smooth(method = "glm", 
+    method.args = list(family = "binomial"), 
+    se = FALSE) +
+ # scale_color_viridis_c() +
+  xlim(c(-2.75,2.75)) +
+  facet_wrap( ~ cohort)
+
+modi1 <- glm(firstObs*1 ~ observedLengthStd * cohortStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+modi2 <- glm(firstObs*1 ~ observedLengthStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+modi3 <- glm(firstObs*1 ~ observedLengthStd * species , family = 'binomial', data = cd01Mod)
+modi4 <- glm(firstObs*1 ~ observedLengthStd * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+modi5 <- glm(firstObs*1 ~ observedLengthStd, family = 'binomial', data = cd01Mod)
+
+AIC(modi1,modi2,modi3,modi4,modi5) %>% rownames_to_column() %>% arrange(AIC)
+
+summary(modi1)
+anova(modi1, test="Chi")
+###
+predi <- expand.grid(observedLengthStd = seq(-2,2, length.out = 50), cohortStd = unique(cd01Mod$cohortStd), species = c('bkt', 'bnt'), ageInSamples = unique(cd01Mod$ageInSamples))
+
+predi$pred <- predict.glm(modi1, newdata = predi, type = 'response')
+
+u <- cd01Mod %>% distinct(age, season, ageInSamples) %>% arrange(ageInSamples) %>% filter(!(age==2 & ageInSamples == 6))
+predi <- left_join(predi, u)
+
+ggplot(predi, aes(observedLengthStd, pred, color = (cohortStd), group = (cohortStd))) +
+  geom_point() +
+  scale_color_viridis_c() +
+  scale_y_continuous("Probability of first observation") +
+  facet_grid(species ~ ageInSamples, scales = 'free')
+
+predi$seasonGG <- factor(predi$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+predi$cohort <- predi$cohortStd + 2007
+predi$speciesGG <- factor(predi$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot(predi %>% filter(cohortStd %in% c(0), species == 'bkt', age < 3), aes(observedLengthStd, pred)) +
+  geom_point() +
+  scale_color_viridis_c() +
+  scale_x_continuous("Standardized length") +
+  scale_y_continuous(lim = c(0,1),"Probability of first observation") +
+  theme_publication() +
+  facet_grid(seasonGG ~ age)
+
+ggplot(predi %>% filter(cohortStd %in% c(0), age < 3), aes(observedLengthStd, pred, color = speciesGG)) +
+  geom_point(size = 2) +
+  scale_x_continuous("Standardized length") +
+  scale_y_continuous(lim = c(0,1),"Probability of first observation") +
+  theme_publication(base_size = 28) +
+  theme(legend.title = element_text("Species"),
+        legend.position = "bottom",
+        legend.direction = "horizontal"
+        ) + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG ~ age)
+
+ggsave('./figures/pFirstObs.png', width = 11, height = 11, units='in')
+```
+
+```{r gam models - find best lm, before fitting detailed gams - only need to run for model selection}
+# only including river, season, salmon01 becaseu we assume these have categorical effects
+
+mod0 <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) + 
                s(ydayCumul, bs = 'cr'), 
@@
                method = "REML")
-summary(mod)
-
+
+mod1 <- bam(observedLength ~ riverFactor + factor(season) + factor(salmon01) + 
+               s(ydayCumul, bs = 'cr'), 
+               data = dataIn, 
+               method = "REML")
+
+mod2 <- bam(observedLength ~ riverFactor * factor(season) + 
+               s(ydayCumul, bs = 'cr'), 
+               data = dataIn, 
+               method = "REML")
+
+mod3 <- bam(observedLength ~ riverFactor + factor(season) + 
+               s(ydayCumul, bs = 'cr'), 
+               data = dataIn, 
+               method = "REML")
+
+mod4 <- bam(observedLength ~ riverFactor + 
+               s(ydayCumul, bs = 'cr'), 
+               data = dataIn, 
+               method = "REML")
+
+mod5 <- bam(observedLength ~ factor(season) + 
+               s(ydayCumul, bs = 'cr'), 
+               data = dataIn, 
+               method = "REML")
+
+mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +
+               s(ydayCumul, bs = 'cr'), 
+               data = dataIn, 
+               method = "REML")
+
+AIC(mod0,mod1,mod2,mod3,mod4,mod5,mod6) %>% rownames_to_column() %>% arrange(AIC)
+summary(mod0)
+plot(mod0)
+gam.check(mod0)
+
+```
+
+```{r gam structure given best model from above}
+
+# BKT, mod0 is best fixed effects model
+
+system.time(
+mod0_YDCcr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCc_r <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) , 
+               data = dataIn, 
+               method = "REML")
+)
+
+AIC(mod0_YDCcr, mod0_YDCc_r) %>% rownames_to_column() %>% arrange(AIC)
+
+system.time(
+mod0_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_Cr_CSr_Ts_Fs_TFs <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = season) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = season) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = season), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
+summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
+plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)
+
+system.time(
+mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_C_CS_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
+summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
+plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)
+
+system.time(
+mod0_YDCcr_C_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_CR_Cr_CS_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_CR_Cr_CS_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_CR_Cr_CSr_T_F_TF <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr'), 
+               data = dataIn, 
+               method = "REML")
+)
+
+# take out s()'s
+
+system.time(
+mod0_CR_Cr_CSr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered),  # this is conspecific cS only
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_CR_Cr_Tr_Fr_TFr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_CR_CSr_Tr_Fr_TFr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_CSr_Tr_Fr_TFr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
+               data = dataIn, 
+               method = "REML")
+)
+########
+
+system.time(
+mod0_CR_Cr_CS_T_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr'), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_CR_Cr_CS_Tr_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_C_Cr_CS_Tr_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
+               data = dataIn, 
+               method = "REML")
+)
+####
+system.time(
+mod0_C_Cr_CS_T_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+              # s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr'), 
+               data = dataIn, 
+               method = "REML")
+)
+
+tmp <- 
+bam(observedLength ~ factor(river) * factor(season) + factor(salmon01) + 
+      #factor(river) *  + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+              # s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+                s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
+                s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+                s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = season) +
+                s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = season) +
+                s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = season), 
+               data = dataIn, 
+               method = "REML")
+summary(tmp)
+plot(tmp, scale = 0, shade = TRUE)
+
+AIC(mod0, mod0_CR, mod0_CR_Cr_CSr_Tr_Fr_TFr,mod0_CR_C_CSr_Tr_Fr_TFr,mod0_CR_Cr_CS_Tr_Fr_TFr,mod0_CR_Cr_CSr_T_F_TF,
+    mod0_CR_Cr_CSr,mod0_CR_Cr_Tr_Fr_TFr,mod0_CR_CSr_Tr_Fr_TFr,mod0_CSr_Tr_Fr_TFr,mod0_CR_Cr_CS_T_F_TF,mod0_CR_Cr_CS_Tr_F_TF,mod0_C_Cr_CS_Tr_F_TF,mod0_C_Cr_CS_T_F_TF) %>% rownames_to_column() %>% arrange(AIC)
+
+summary(mod0_CR_Cr_CS_T_F_TF)
+gam.check(mod0_CR_Cr_CS_T_F_TF)
+plot(mod0_CR_Cr_CS_T_F_TF, scale = 0, pages=8)
+
+```
+
+```{r}
 predGrid <- expand.grid(cohort = unique(mod$model$cohort),
@@
                s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn),  # this is conspecific cS only
              data = dataIn, method = "REML")
@@
                s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered), 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered),  # this is conspecific cS only
              data = dataIn, method = "REML")
@@
 
+system.time(
+modCR_SR_c_cSMP <- bam(observedLength ~ cohort * river + factor(salmon01) * river + metaPop01 +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = metaPop01),  # this is conspecific cS only
+             data = dataIn, method = "REML")
+)
+summary(modCR_SR_c_cSMP)
+
 # effects of each species together - above is only conspecific
@@
 system.time(
-modCR_SR_c_cSR3b <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+modCR_SR_c_cSR3b <- bam(observedLength ~ cohort * river + factor(salmon01) * river + metaPop01 +
                s(ydayCumul, bs = 'cr') +
@@
 
-AIC(mod,modC,modR,modC_R,modCR, modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, modCR_SR_c_cSR3, modCR_SR_c_cSR3a,modCR_SR_c_cSR3b) %>% rownames_to_column() %>% arrange(AIC)
-
-save(mod,modC,modR,modC_R,modCR,modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, file = paste0('./data/out/mods_',speciesIn,'.RData'))
+# take out salmon because the salmon01 is in cohortStrengthScaledATS
+system.time(
+modCR__c_cSR3b <- bam(observedLength ~ cohort * river + metaPop01 +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
+             data = dataIn, method = "REML")
+)
+summary(modCR_SR_c_cSR3b)
+
+AIC(mod,modC,modR,modC_R,modCR, modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, modCR_SR_c_cSMP, modCR_SR_c_cSR3, modCR_SR_c_cSR3a,modCR_SR_c_cSR3b, modCR__c_cSR3b) %>% rownames_to_column() %>% arrange(AIC)
+
+save(mod,modC,modR,modC_R,modCR,modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, modCR_SR_c_cSMP, modCR_SR_c_cSR3, modCR_SR_c_cSR3a,modCR_SR_c_cSR3b, file = paste0('./data/out/mods_',speciesIn,'.RData'))
 
 # BKT 
+#             rowname        df      AIC
+# 1     modCR_SR_c_cS 162.34756 414267.2
+# 2      modCR_S_c_cS 163.43154 414602.7
+# 3        modCR_c_cS 161.63691 414750.3
+# 4  modCR_SR_c_cSR3a  88.21646 415388.9
+# 5  modCR_SR_c_cSR3b 167.32968 416062.3
+# 6    modCR_SR_c_cSR 163.93920 416587.5
+# 7   modCR_SR_c_cSR3 173.36884 416611.4
+# 8         modCR_S_c 157.23995 418645.9
+# 9           modCR_c 158.50186 418665.1
+# 10            modCR 154.36186 419135.1
+# 11           modC_R  63.10879 421258.1
+# 12              mod  17.99204 427309.6
+# 13             modC  48.73277 431440.2
+# 14             modR  22.80483 432214.7
 
@@
 
-```{r best models for each species}
-
-# Brook trout
-# using salmonid01 * river because we need diff effects per river - even though AIC is worse for this model 
-modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
-             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
-
-summary(modBKT)
-#plot(modBKT, pages = 8, scale = 0)
-
-predGridBKT <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
+```{r models with flow and temp}
+
+################################################################################
+
+# base model
+modCR_SR_c_cS <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn),  # this is conspecific cS only
+             data = dataIn, method = "REML")
+
+modCR_SR_c_cS_T <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaled, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+modCR_SR_c_cS_F <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanFlowScaled, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+modCR_SR_c_cS_T_F <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaled, bs = 'cr', k = kIn) +
+               s(meanFlowScaled, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+modCR_SR_c_cS_TF <- bam(observedLength ~ cohort * riverOrdered + factor(salmon01) * riverOrdered +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaled, bs = 'cr', k = kIn) +
+               s(meanFlowScaled, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaled, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+modCR_SR_c_cS_TFs <- bam(observedLength ~ cohort * river + factor(salmon01) * river + meanTemperatureScaled * meanFlowScaled +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaled, bs = 'cr', k = kIn, by = factor(season)) +
+               s(meanFlowScaled, bs = 'cr', k = kIn, by = factor(season)) +
+               s(meanTemperatureFlowScaled, bs = 'cr', k = kIn, by = factor(season)),
+             data = dataIn, method = "REML")
+
+summary(modCR_SR_c_cS_TFs)
+plot(modCR_SR_c_cS_TFs, scale = 0, pages=8)
+
+modCR_SR_c_cS_TFs2 <- bam(observedLength ~ cohort * river + factor(salmon01) * river + meanTemperatureScaled * meanFlowScaled * factor(season) +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaled, bs = 'cr', k = kIn, by = factor(season)) +
+               s(meanFlowScaled, bs = 'cr', k = kIn, by = factor(season)) +
+               s(meanTemperatureFlowScaled, bs = 'cr', k = kIn, by = factor(season)),
+             data = dataIn, method = "REML")
+
+modCR_SR_c_cS_TFr <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaled, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanFlowScaled, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(meanTemperatureFlowScaled, bs = 'cr', k = kIn, by = riverOrdered),
+             data = dataIn, method = "REML")
+
+AIC(modCR_SR_c_cS,modCR_SR_c_cS_T, modCR_SR_c_cS_F, modCR_SR_c_cS_T_F, modCR_SR_c_cS_TF, modCR_SR_c_cS_TFs, modCR_SR_c_cS_TFs2, modCR_SR_c_cS_TFr) %>% rownames_to_column() %>% arrange(AIC)
+
+summary(modCR_SR_c_cS_TFr)
+plot(modCR_SR_c_cS_TFr, scale = 0, pages=8)
+
+predGridTF <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
                            ydayCumul = seq(250,1000, by = 30),
                            salmon01 = c(0,1),
-                           cohortStrengthScaled = seq(-2,2, by = 0.1)) %>%
+                           cohortStrengthScaledBKT = seq(-2,2, by = 0.1)) %>%
                     mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
@@
 
-predBKT <- predict(modBKT, newdata = predGridBKT)#, type = 'lpmatrix')
+predTF <- predict(modCR_SR_c_cS_TFr, newdata = predGridTF)#, type = 'lpmatrix')
 
@@
 
-#################################
-# Brown trout
-
-modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
-               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
-               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
-             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")
-
-summary(modBNT)
-#plot(modBNT, pages = 8, scale = 0)
-
-predGridBNT <- expand.grid(cohortRiverOrdered = unique(modBNT$model$cohortRiverOrdered),
+predGrid <- expand.grid(cohort = unique(mod$model$cohort),
+                        river = unique(mod$model$river),
+                        ydayCumul = seq(250,1000, by = 30))
+
+pred <- predict(mod, newdata = predGrid)#, type = 'lpmatrix')
+
+predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
+
+ggplot(predGrid2, aes(ydayCumul, pred)) +
+  geom_point() + 
+  geom_line() +
+  facet_grid(river ~ cohort) +
+  ggtitle('mod')
+
+```
+
+```{r best models for each species}
+
+# Brook trout
+
+#using salmon01 * river because we need diff effects per river - even though AIC is worse for this model
+modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) +
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn),
+             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
+
+summary(modBKT)
+
+#plot(modBKT, pages = 8, scale = 0)
+
+predGridBKT <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
                            ydayCumul = seq(250,1000, by = 30),
                            salmon01 = c(0,1),
-                           cohortStrengthScaled = seq(-2,2, by = 0.1)) %>%
+                           cohortStrengthScaledBKT = seq(-2,2, by = 0.1)) %>%
                     mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
@@
 
+predBKT <- predict(modBKT, newdata = predGridBKT)#, type = 'lpmatrix')
+
+predGrid2BKT <- bind_cols(predGridBKT, as.data.frame(predBKT)) %>% mutate(pred = predBKT, species = 'bkt')
+
+#################################
+# Brown trout
+
+modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
+             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")
+
+summary(modBNT)
+#plot(modBNT, pages = 8, scale = 0)
+
+predGridBNT <- expand.grid(cohortRiverOrdered = unique(modBNT$model$cohortRiverOrdered),
+                           ydayCumul = seq(250,1000, by = 30),
+                           salmon01 = c(0,1),
+                           metaPop01 = c(0,1),
+                           cohortStrengthScaledBKT = seq(-2,2, by = 0.25),
+                           cohortStrengthScaledBNT = seq(-2,2, by = 0.25),
+                           cohortStrengthScaledATS = seq(-2,2, by = 0.25)) %>%
+                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
+                           riverOrdered = substring(cohortRiverOrdered,6),
+                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
+                           cohortCentered = cohortOrdered - cohortOffset,
+                           cohort = cohortOrdered,
+                           river = riverOrdered)
+
 predBNT <- predict(modBNT, newdata = predGridBNT)#, type = 'lpmatrix')
@@
 
-predGrid2Both <- bind_rows(predGrid2BKT, predGrid2BNT) %>% select(-predBKT, -predBNT)
+predGrid2Both <- bind_rows(predGrid2BKT, predGrid2BNT) %>% dplyr:::select(-predBKT, -predBNT)
 
@@
 # cohorts by river/species
-ggplot(predGrid2Both %>% filter(cohortStrengthScaled == 0, salmon01 == 0), aes(ydayCumul, pred, group = cohort, color = cohort)) +
+ggplot(predGrid2Both %>% 
+         filter(
+         (cohortStrengthScaledBKT == 0 & salmon01 == 0) | (cohortStrengthScaledBKT == 0 & cohortStrengthScaledBNT == 0 & cohortStrengthScaledATS == 0 & metaPop01 == 1 & salmon01 ==0)),
+#           cohortStrengthScaledBKT == 0, cohortStrengthScaledBNT == 0, cohortStrengthScaledATS == 0, metaPop01 == 1, salmon01 == 0),
+       aes(ydayCumul, pred, group = cohort, color = cohort)) +
   #geom_point() + 
+  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
   geom_line(size = 1.1) +
@@
         legend.direction = "vertical",
-        legend.title = element_text(face="plain")) +
+        legend.title = element_text(face = "plain")) +
   guides(color = guide_colorbar(barheight = 10,
@@
 
+ggsave('./figures/predLengths.png', width = 11, height = 11, units='in')
+
 # specific splines by species
@@
 #
+plot.gam(modBKT, pages = 10)
 

### 2019-06-28 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
new models

@@
 
+# take out s()'s
+
 system.time(
-mod0_YDCcr_C_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
+mod0_YDCcr_Cr_CSr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor), 
                data = dataIn, 
@@
 system.time(
-mod0_CR_Cr_CS_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
+mod0_YDCcr_Cr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               #s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
                data = dataIn, 
@@
 system.time(
-mod0_CR_Cr_CS_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
+mod0_YDCcr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+              # s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
                data = dataIn, 
@@
 system.time(
-mod0_CR_Cr_CSr_T_F_TF <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr'), 
-               data = dataIn, 
-               method = "REML")
-)
-
-# take out s()'s
-
-system.time(
-mod0_CR_Cr_CSr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered),  # this is conspecific cS only
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_CR_Cr_Tr_Fr_TFr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_CR_CSr_Tr_Fr_TFr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_CSr_Tr_Fr_TFr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
+mod0_YDCcr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+             #  s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+             #  s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
                data = dataIn, 
@@
 
-system.time(
-mod0_CR_Cr_CS_T_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr'), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_CR_Cr_CS_Tr_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_C_Cr_CS_Tr_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
-               data = dataIn, 
-               method = "REML")
-)
-####
-system.time(
-mod0_C_Cr_CS_T_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-              # s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr'), 
-               data = dataIn, 
-               method = "REML")
-)
-
-tmp <- 
-bam(observedLength ~ factor(river) * factor(season) + factor(salmon01) + 
-      #factor(river) *  + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-              # s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-                s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
-                s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-                s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = season) +
-                s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = season) +
-                s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = season), 
-               data = dataIn, 
-               method = "REML")
-summary(tmp)
-plot(tmp, scale = 0, shade = TRUE)
-
-AIC(mod0, mod0_CR, mod0_CR_Cr_CSr_Tr_Fr_TFr,mod0_CR_C_CSr_Tr_Fr_TFr,mod0_CR_Cr_CS_Tr_Fr_TFr,mod0_CR_Cr_CSr_T_F_TF,
-    mod0_CR_Cr_CSr,mod0_CR_Cr_Tr_Fr_TFr,mod0_CR_CSr_Tr_Fr_TFr,mod0_CSr_Tr_Fr_TFr,mod0_CR_Cr_CS_T_F_TF,mod0_CR_Cr_CS_Tr_F_TF,mod0_C_Cr_CS_Tr_F_TF,mod0_C_Cr_CS_T_F_TF) %>% rownames_to_column() %>% arrange(AIC)
-
-summary(mod0_CR_Cr_CS_T_F_TF)
-gam.check(mod0_CR_Cr_CS_T_F_TF)
-plot(mod0_CR_Cr_CS_T_F_TF, scale = 0, pages=8)
+AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CSr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_Tsr_Fsr_TFsr,mod0_YDCcr_CSr_Tsr_Fsr_TFsr, mod0_YDCcr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
+
+#                           rowname        df      AIC
+# 1  mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr 277.67003 401304.4
+# 2   mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr 281.22421 401407.3
+# 3   mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr 277.09498 401772.0
+# 4     mod0_YDCcr_CSr_Tsr_Fsr_TFsr 279.84632 402037.9
+# 5    mod0_YDCcr_C_CS_Tsr_Fsr_TFsr 269.69828 402073.1
+# 6     mod0_YDCcr_Cr_CSr_Tr_Fr_TFr 207.97536 402187.1
+# 7     mod0_YDCcr_Cr_CSr_Ts_Fs_TFs 190.64645 402464.7
+# 8      mod0_YDCcr_Cr_Tsr_Fsr_TFsr 289.33427 403529.7
+# 9         mod0_YDCcr_Tsr_Fsr_TFsr 283.93242 404139.1
+# 10              mod0_YDCcr_Cr_CSr 184.24038 413605.9
+# 11                     mod0_YDCcr 168.77693 420503.0
+# 12                    mod0_YDCc_r  71.99108 422677.4
+
+summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
+plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)
+
+```
+
+```{r predictions}
+# BKT
+# system.time(
+# mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+#                s(ydayCumul, bs = 'cr') +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+#                s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+#                s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+#                s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+#                s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+#                s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+#                data = dataIn, 
+#                method = "REML")
+# )
+
+predGridBKT <- expand.grid(cohortRiverFactor = unique(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr$model$cohortRiverFactor),
+                           seasonRiverFactor = unique(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr$model$seasonRiverFactor),
+                           ydayCumul = seq(250,1000, by = 30),
+                           salmon01 = c(0,1),
+                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
+                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
+                           meanFlowScaledBySeason = 0) %>% #seq(-1.5,1.5, by = 0.5)) %>%
+                          # meanTemperatureFlowScaledBySeason = seq(-1.5,1.5, by = 0.5)) %>%
+                    mutate(cohortFactor = as.numeric(substr(cohortRiverFactor,1,4)),
+                           riverFactor = substring(cohortRiverFactor,6),
+                           season = substring(seasonRiverFactor,1),
+                           #cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
+                           #seasonRiverFactor = paste(season, riverFactor, sep = '_'),
+                           #cohortCentered = cohortOrdered - cohortOffset,
+                           cohort = cohortFactor,
+                           #river = riverFactor,
+                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)
+
+ydayCumulSeason <- distinct(dataIn, ydayCumul,season,age) %>% arrange(ydayCumul) %>% mutate(age_season = paste(age,season,sep = '_')) %>%
+  group_by(age_season) %>% summarise(age=min(age),season=min(season),minY=min(ydayCumul),maxY=max(ydayCumul))
+#38 obear fish were captured in january, but get season 4 and are age 1, so have very small ydayCumul -need to manually reset 
+#dataIn %>% filter(age==1,season==4,ydayCumul==367) %>% select(tag,cohort,detectionDate,year,ydayCumul,yday,river,age,season)
+ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '1_4',670,ydayCumulSeason$minY)
+ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '2_4',1035,ydayCumulSeason$minY)
+ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '0_3',0,ydayCumulSeason$minY)
+ydayCumulSeason %>% mutate(minY2 = lead(minY))
+
+yDayAll <- data.frame(y = 250:1000)
+# pick up here
+#### need to do something like this  left_join(yDayAll, ydayCumulSeason, by = c('yDay' = 'minY'))
+
+predGridBKT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
+                           cohortFactor = unique(dataIn$cohort),
+                           ydayCumul = seq(250,1000, by = 30),
+                           salmon01 = c(0,1),
+                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
+                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
+                           meanFlowScaledBySeason = 0) %>%
+                    left_join(., ydayCumulSeason) %>%
+                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
+                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
+                           cohort = cohortFactor,
+                           #river = riverFactor,
+                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)
+
+predBKT <- predict(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, newdata = predGridBKT)#, type = 'lpmatrix')
+
+predGrid2BKT <- bind_cols(predGridBKT, as.data.frame(predBKT)) %>% mutate(pred = predBKT, species = 'bkt')
+
+ggplot(predGrid2BKT %>% filter(salmon01 == 0),
+
+       aes(ydayCumul, pred, group = cohort, color = cohort)) +
+  #geom_point() + 
+  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
+  geom_line(size = 1.1) +
+  labs(y = 'Predicted body size (mm)',
+       x = 'Days since age-0 Jan 1',
+       color = 'Cohort') +
+  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
+  theme_publication() +
+  theme(legend.position = "right", 
+        legend.direction = "vertical",
+        legend.title = element_text(face = "plain")) +
+  guides(color = guide_colorbar(barheight = 10,
+                                barwidth = 1.5,
+                                ticks = FALSE,
+                                nbin = 100)) +
+#  guide_legend(title = 'Cohort') +
+  facet_grid(~riverFactor, scales = 'free_y')
 
@@
 ```{r}
-predGrid <- expand.grid(cohort = unique(mod$model$cohort),
-                        river = unique(mod$model$river),
-                        ydayCumul = seq(250,1000, by = 30))
-
-pred <- predict(mod, newdata = predGrid)#, type = 'lpmatrix')
-
-predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
-
-ggplot(predGrid2, aes(ydayCumul, pred)) +
-  geom_point() + 
-  geom_line() +
-  facet_grid(river ~ cohort) +
-  ggtitle('mod')
-
-##########################################################
-# by cohortSpeciesRiver
-
-system.time(
-modCR <- bam(observedLength ~ cohort * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered), 
-             data = dataIn, method = "REML")
-)
-summary(modCR)
-#plot(modCR, pages = 8, scale = 0)
-
-predGridCR <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
-                           ydayCumul = seq(250,1000, by = 30)) %>%
-                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
-                           riverOrdered = substring(cohortRiverOrdered,6),
-                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
-                           cohortCentered = cohortOrdered - cohortOffset,
-                           cohort = cohortOrdered,
-                           river = riverOrdered)
-
-predCR <- predict(modCR, newdata = predGridCR)#, type = 'lpmatrix')
-
-predGrid2CR <- bind_cols(predGridCR, as.data.frame(predCR))
-
-ggplot(predGrid2CR, aes(ydayCumul, predCR)) +
-  geom_point() + 
-  geom_line() +
-  ylim(50,210) +
-  facet_grid(river ~ cohort)+
-  ggtitle('modCR')
-
-##########################################################
-# by cohortSpeciesRiver with s(cohort)
-
-system.time(
-modCR_c <- bam(observedLength ~ cohort * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn), 
-             data = dataIn, method = "REML")
-)
-summary(modCR_c)
-#plot(modCR_c, pages = 8, scale = 0)
-
-predGridCR_c <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
-                           ydayCumul = seq(250,1000, by = 30)) %>%
-                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
-                           riverOrdered = substring(cohortRiverOrdered,6),
-                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
-                           cohortCentered = cohortOrdered - cohortOffset,
-                           cohort = cohortOrdered,
-                           river = riverOrdered)
-
-predCR_c <- predict(modCR_c, newdata = predGridCR_c)#, type = 'lpmatrix')
-
-predGrid2CR_c <- bind_cols(predGridCR_c, as.data.frame(predCR_c))
-
-ggplot(predGrid2CR_c, aes(ydayCumul, predCR_c, group = cohort, color = cohort)) +
-  geom_point() + 
-  geom_line() +
-  #ylim(50,210) +
-  facet_wrap(~river)+
-  ggtitle('modCR_c')
-
-##########################################################
-# by cohortRiver - w ind RE
-
-# need more memory
-# system.time(
-# modCR_Ind <- bam(observedLength ~ cohort * river +
-#                s(ydayCumul, bs = 'cr') +
-#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-#                s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-#                s(ydayCumul, tagIndexFactor, bs="fs", m=1, k = kIn),
-#              data = dataIn, method = "REML")
-# )
-# summary(modCR_Ind)
-# plot(modCR_Ind, pages = 8, scale = 0)
-
-##########################################################
-# by cohortSpeciesRiver, additive in s()
-
-system.time(
-modC_R <- bam(observedLength ~ cohort * river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered), 
-             data = dataIn, method = "REML")
-)
-summary(modC_R)
-#plot(modC_R, pages = 4, scale = 0)
-
-predGridC_R <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
-                           ydayCumul = seq(250,1000, by = 30)) %>%
-                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
-                           riverOrdered = substring(cohortRiverOrdered,6),
-                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
-                           cohortCentered = cohortOrdered - cohortOffset,
-                           cohort = cohortOrdered,
-                           river = riverOrdered)
-
-predC_R <- predict(modC_R, newdata = predGridC_R)#, type = 'lpmatrix')
-
-predGrid2C_R <- bind_cols(predGridC_R,as.data.frame(predC_R))
-
-ggplot(predGrid2C_R, aes(ydayCumul,predC_R)) +
-  geom_point() + 
-  geom_line() +
-  ylim(50,210) +
-  facet_grid(river ~ cohort)+
-  ggtitle('modC_R')
-
-##########################
-
-modC <- bam(observedLength ~ cohort + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered), 
-             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt')), method = "REML")
-
-modR <- bam(observedLength ~ river + 
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered), 
-             data = dataIn, method = "REML")
-
-##########################################
-# add in salmon01 and cohortStrengthScaled
-
-system.time(
-modCR_S_c <- bam(observedLength ~ cohort * river + factor(salmon01) +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn), 
-             data = dataIn, method = "REML")
-)
-system.time(
-modCR_c_cS <- bam(observedLength ~ cohort * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn),  # this is conspecific cS only
-             data = dataIn, method = "REML")
-)
-
-system.time(
-modCR_S_c_cS <- bam(observedLength ~ cohort * river + factor(salmon01) +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn),  # this is conspecific cS only
-             data = dataIn, method = "REML")
-)
-summary(modCR_S_c_cS)
-
-system.time(
-modCR_SR_c_cS <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn),  # this is conspecific cS only
-             data = dataIn, method = "REML")
-)
-summary(modCR_SR_c_cS)
-
-system.time(
-modCR_SR_c_cSR <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered),  # this is conspecific cS only
-             data = dataIn, method = "REML")
-)
-summary(modCR_SR_c_cSR)
-#plot(modCR_S_c_cS, scale = 0, pages=8)
-#plotGamVarBySpecies(var = 's(cohortStrengthScaled)')
-
-system.time(
-modCR_SR_c_cSMP <- bam(observedLength ~ cohort * river + factor(salmon01) * river + metaPop01 +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = metaPop01),  # this is conspecific cS only
-             data = dataIn, method = "REML")
-)
-summary(modCR_SR_c_cSMP)
-
-# effects of each species together - above is only conspecific
-system.time(
-modCR_SR_c_cSR3 <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
-               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
-             data = dataIn, method = "REML")
-)
-summary(modCR_SR_c_cSR3)
-#plot(modCR_SR_c_cSR3, scale = 0, pages=5)
-
-system.time(
-modCR_SR_c_cSR3a <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn) +
-               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
-               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
-             data = dataIn, method = "REML")
-)
-summary(modCR_SR_c_cSR3a)
-#plot(modCR_SR_c_cSR3a, scale = 0, pages=5)
-
-system.time(
-modCR_SR_c_cSR3b <- bam(observedLength ~ cohort * river + factor(salmon01) * river + metaPop01 +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
-               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
-               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
-             data = dataIn, method = "REML")
-)
-summary(modCR_SR_c_cSR3b)
-#plot(modCR_SR_c_cSR3b, scale = 0, pages=5)
-
-# take out salmon because the salmon01 is in cohortStrengthScaledATS
-system.time(
-modCR__c_cSR3b <- bam(observedLength ~ cohort * river + metaPop01 +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
-               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
-               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
-             data = dataIn, method = "REML")
-)
-summary(modCR_SR_c_cSR3b)
-
-AIC(mod,modC,modR,modC_R,modCR, modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, modCR_SR_c_cSMP, modCR_SR_c_cSR3, modCR_SR_c_cSR3a,modCR_SR_c_cSR3b, modCR__c_cSR3b) %>% rownames_to_column() %>% arrange(AIC)
-
-save(mod,modC,modR,modC_R,modCR,modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, modCR_SR_c_cSMP, modCR_SR_c_cSR3, modCR_SR_c_cSR3a,modCR_SR_c_cSR3b, file = paste0('./data/out/mods_',speciesIn,'.RData'))
-
+#earlier runs before setting fixed effect first. kept forf reference
 # BKT 
@@
 # 14             modC 48.73277 431440.2
-
-```
-
-```{r models with flow and temp}
-
-################################################################################
-
-# base model
-modCR_SR_c_cS <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn),  # this is conspecific cS only
-             data = dataIn, method = "REML")
-
-modCR_SR_c_cS_T <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaled, bs = 'cr', k = kIn),
-             data = dataIn, method = "REML")
-
-modCR_SR_c_cS_F <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanFlowScaled, bs = 'cr', k = kIn),
-             data = dataIn, method = "REML")
-
-modCR_SR_c_cS_T_F <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaled, bs = 'cr', k = kIn) +
-               s(meanFlowScaled, bs = 'cr', k = kIn),
-             data = dataIn, method = "REML")
-
-modCR_SR_c_cS_TF <- bam(observedLength ~ cohort * riverOrdered + factor(salmon01) * riverOrdered +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaled, bs = 'cr', k = kIn) +
-               s(meanFlowScaled, bs = 'cr', k = kIn) +
-               s(meanTemperatureFlowScaled, bs = 'cr', k = kIn),
-             data = dataIn, method = "REML")
-
-modCR_SR_c_cS_TFs <- bam(observedLength ~ cohort * river + factor(salmon01) * river + meanTemperatureScaled * meanFlowScaled +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaled, bs = 'cr', k = kIn, by = factor(season)) +
-               s(meanFlowScaled, bs = 'cr', k = kIn, by = factor(season)) +
-               s(meanTemperatureFlowScaled, bs = 'cr', k = kIn, by = factor(season)),
-             data = dataIn, method = "REML")
-
-summary(modCR_SR_c_cS_TFs)
-plot(modCR_SR_c_cS_TFs, scale = 0, pages=8)
-
-modCR_SR_c_cS_TFs2 <- bam(observedLength ~ cohort * river + factor(salmon01) * river + meanTemperatureScaled * meanFlowScaled * factor(season) +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaled, bs = 'cr', k = kIn, by = factor(season)) +
-               s(meanFlowScaled, bs = 'cr', k = kIn, by = factor(season)) +
-               s(meanTemperatureFlowScaled, bs = 'cr', k = kIn, by = factor(season)),
-             data = dataIn, method = "REML")
-
-modCR_SR_c_cS_TFr <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaled, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanFlowScaled, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(meanTemperatureFlowScaled, bs = 'cr', k = kIn, by = riverOrdered),
-             data = dataIn, method = "REML")
-
-AIC(modCR_SR_c_cS,modCR_SR_c_cS_T, modCR_SR_c_cS_F, modCR_SR_c_cS_T_F, modCR_SR_c_cS_TF, modCR_SR_c_cS_TFs, modCR_SR_c_cS_TFs2, modCR_SR_c_cS_TFr) %>% rownames_to_column() %>% arrange(AIC)
-
-summary(modCR_SR_c_cS_TFr)
-plot(modCR_SR_c_cS_TFr, scale = 0, pages=8)
-
-predGridTF <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
-                           ydayCumul = seq(250,1000, by = 30),
-                           salmon01 = c(0,1),
-                           cohortStrengthScaledBKT = seq(-2,2, by = 0.1)) %>%
-                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
-                           riverOrdered = substring(cohortRiverOrdered,6),
-                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
-                           cohortCentered = cohortOrdered - cohortOffset,
-                           cohort = cohortOrdered,
-                           river = riverOrdered)
-
-predTF <- predict(modCR_SR_c_cS_TFr, newdata = predGridTF)#, type = 'lpmatrix')
-
-predGrid2BKT <- bind_cols(predGridBKT, as.data.frame(predBKT)) %>% mutate(pred = predBKT, species = 'bkt')
-
-predGrid <- expand.grid(cohort = unique(mod$model$cohort),
-                        river = unique(mod$model$river),
-                        ydayCumul = seq(250,1000, by = 30))
-
-pred <- predict(mod, newdata = predGrid)#, type = 'lpmatrix')
-
-predGrid2 <- bind_cols(predGrid,as.data.frame(pred))
-
-ggplot(predGrid2, aes(ydayCumul, pred)) +
-  geom_point() + 
-  geom_line() +
-  facet_grid(river ~ cohort) +
-  ggtitle('mod')
 

### 2019-07-10 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
took season out of main effects and splines

@@
 ```{r gams for body size - data prep}
-speciesIn <- 'bkt'
+speciesIn <- 'bnt'
 depVarIn <- 'observedLength'
@@
 ```{r gam models - find best lm, before fitting detailed gams - only need to run for model selection}
-# only including river, season, salmon01 becaseu we assume these have categorical effects
-
-mod0 <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) + 
-               s(ydayCumul, bs = 'cr'), 
-               data = dataIn, 
-               method = "REML")
-
-mod1 <- bam(observedLength ~ riverFactor + factor(season) + factor(salmon01) + 
-               s(ydayCumul, bs = 'cr'), 
-               data = dataIn, 
-               method = "REML")
+# only including river, salmon01 becaseu we assume these have categorical effects
+
+# not including season
+# models with season have big dip in predictions for season 1. Season is...
 
@@
 
-mod3 <- bam(observedLength ~ riverFactor + factor(season) + 
-               s(ydayCumul, bs = 'cr'), 
-               data = dataIn, 
-               method = "REML")
-
 mod4 <- bam(observedLength ~ riverFactor + 
@@
 
-mod5 <- bam(observedLength ~ factor(season) + 
-               s(ydayCumul, bs = 'cr'), 
-               data = dataIn, 
-               method = "REML")
-
 mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +
@@
 
-AIC(mod0,mod1,mod2,mod3,mod4,mod5,mod6) %>% rownames_to_column() %>% arrange(AIC)
-summary(mod0)
-plot(mod0)
-gam.check(mod0)
-
-```
-
-```{r gam structure given best model from above}
-
-# BKT, mod0 is best fixed effects model
-
-system.time(
-mod0_YDCcr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_YDCc_r <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) , 
-               data = dataIn, 
-               method = "REML")
-)
-
-AIC(mod0_YDCcr, mod0_YDCc_r) %>% rownames_to_column() %>% arrange(AIC)
-
-system.time(
-mod0_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+AIC(mod2,mod4,mod6) %>% rownames_to_column() %>% arrange(AIC)
+summary(mod6)
+plot(mod6)
+gam.check(mod6)
+
+# bkt
+#   rowname       df      AIC
+# 1    mod6 14.99337 429769.2
+# 2    mod2 25.99569 432704.3
+# 3    mod4 13.98558 433764.8
+
+#bnt
+#   rowname       df      AIC
+# 1    mod6 12.99685 238636.2
+# 2    mod2 17.99755 242840.1
+# 3    mod4 11.98667 243594.3
+
+```
+
+```{r models with no fixed season effects or categories for splines}
+
+system.time(
+mod6_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
                s(ydayCumul, bs = 'cr') +
@@
 system.time(
-mod0_YDCcr_Cr_CSr_Ts_Fs_TFs <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+mod6_YDCcr_Cr_CSr_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
                s(ydayCumul, bs = 'cr') +
@@
                s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = season) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = season) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = season), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
-               data = dataIn, 
-               method = "REML")
-)
-
-AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
-summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
-plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)
-
-system.time(
-mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
                s(ydayCumul, bs = 'cr') +
@@
                s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_YDCcr_C_CS_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_C_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
                s(ydayCumul, bs = 'cr') +
@@
                s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
-               data = dataIn, 
-               method = "REML")
-)
-
-AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
-summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
-plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)
-
-# take out s()'s
-
-system.time(
-mod0_YDCcr_Cr_CSr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_YDCcr_Cr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               #s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_YDCcr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-              # s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod0_YDCcr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-             #  s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-             #  s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
-               data = dataIn, 
-               method = "REML")
-)
-########
-
-AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CSr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_Tsr_Fsr_TFsr,mod0_YDCcr_CSr_Tsr_Fsr_TFsr, mod0_YDCcr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
-
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDC_C_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
+               data = dataIn, 
+               method = "REML")
+)
+
+AIC(mod6_YDC_C_CS_T_F_TF,mod6_YDCcr_C_CS_T_F_TF,mod6_YDCcr_Cr_CS_T_F_TF,mod6_YDCcr_Cr_CSr_T_F_TF,mod6_YDCcr_Cr_CSr_Tr_Fr_TFr)  %>% rownames_to_column() %>% arrange(AIC)
+
+summary(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr)
+plot(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr, pages = 8, shade = TRUE, scale = 0)
+
+#BKT
 #                       rowname        df      AIC
-# 1  mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr 277.67003 401304.4
-# 2   mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr 281.22421 401407.3
-# 3   mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr 277.09498 401772.0
-# 4     mod0_YDCcr_CSr_Tsr_Fsr_TFsr 279.84632 402037.9
-# 5    mod0_YDCcr_C_CS_Tsr_Fsr_TFsr 269.69828 402073.1
-# 6     mod0_YDCcr_Cr_CSr_Tr_Fr_TFr 207.97536 402187.1
-# 7     mod0_YDCcr_Cr_CSr_Ts_Fs_TFs 190.64645 402464.7
-# 8      mod0_YDCcr_Cr_Tsr_Fsr_TFsr 289.33427 403529.7
-# 9         mod0_YDCcr_Tsr_Fsr_TFsr 283.93242 404139.1
-# 10              mod0_YDCcr_Cr_CSr 184.24038 413605.9
-# 11                     mod0_YDCcr 168.77693 420503.0
-# 12                    mod0_YDCc_r  71.99108 422677.4
-
-summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
-plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)
+# 1 mod6_YDCcr_Cr_CSr_Tr_Fr_TFr 194.27159 402517.3
+# 2    mod6_YDCcr_Cr_CSr_T_F_TF 178.16547 402696.5
+# 3     mod6_YDCcr_Cr_CS_T_F_TF 170.29804 403228.7
+# 4      mod6_YDCcr_C_CS_T_F_TF 166.30172 404261.2
+# 5        mod6_YDC_C_CS_T_F_TF  29.89126 412576.2
+
+#BNT
+#                       rowname        df      AIC
+# 1 mod6_YDCcr_Cr_CSr_Tr_Fr_TFr 102.82522 225950.4
+# 2    mod6_YDCcr_Cr_CSr_T_F_TF  98.12051 225974.0
+# 3     mod6_YDCcr_Cr_CS_T_F_TF  96.51127 225994.7
+# 4      mod6_YDCcr_C_CS_T_F_TF  92.93128 226010.4
+# 5        mod6_YDC_C_CS_T_F_TF  27.97005 231079.5
 
@@
 ```{r predictions}
-# BKT
-# system.time(
-# mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
-#                s(ydayCumul, bs = 'cr') +
-#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-#                s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-#                s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-#                s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-#                s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-#                s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
-#                s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
-#                data = dataIn, 
-#                method = "REML")
-# )
-
-predGridBKT <- expand.grid(cohortRiverFactor = unique(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr$model$cohortRiverFactor),
-                           seasonRiverFactor = unique(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr$model$seasonRiverFactor),
-                           ydayCumul = seq(250,1000, by = 30),
-                           salmon01 = c(0,1),
-                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
-                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
-                           meanFlowScaledBySeason = 0) %>% #seq(-1.5,1.5, by = 0.5)) %>%
-                          # meanTemperatureFlowScaledBySeason = seq(-1.5,1.5, by = 0.5)) %>%
-                    mutate(cohortFactor = as.numeric(substr(cohortRiverFactor,1,4)),
-                           riverFactor = substring(cohortRiverFactor,6),
-                           season = substring(seasonRiverFactor,1),
-                           #cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
-                           #seasonRiverFactor = paste(season, riverFactor, sep = '_'),
-                           #cohortCentered = cohortOrdered - cohortOffset,
-                           cohort = cohortFactor,
-                           #river = riverFactor,
-                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)
 
@@
 
-yDayAll <- data.frame(y = 250:1000)
-# pick up here
-#### need to do something like this  left_join(yDayAll, ydayCumulSeason, by = c('yDay' = 'minY'))
-
-predGridBKT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
+yDayAll <- data.frame(ydayCumul = 250:1000)
+yDayAll$season <- NA
+
+# fill in season for each day of the year (yDayAll) based on ydayCumulSeason min values
+# need this to get season into all ydayCumul in predGridBKT
+for(i in 1:nrow(yDayAll)) {
+  for(j in 1:(nrow(ydayCumulSeason) - 1)) {
+    if (between(yDayAll$ydayCumul[i], ydayCumulSeason$minY[j], ydayCumulSeason$minY[j + 1])) {
+      yDayAll$season[i] <- ydayCumulSeason$season[j]
+    }
+  }
+}
+
+predGrid <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
                            cohortFactor = unique(dataIn$cohort),
-                           ydayCumul = seq(250,1000, by = 30),
+                           ydayCumul = seq(250,1000, by = 10),
                            salmon01 = c(0,1),
@@
                            meanFlowScaledBySeason = 0) %>%
-                    left_join(., ydayCumulSeason) %>%
+                    left_join(., yDayAll) %>%
                     mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
@@
 
-predBKT <- predict(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, newdata = predGridBKT)#, type = 'lpmatrix')
-
-predGrid2BKT <- bind_cols(predGridBKT, as.data.frame(predBKT)) %>% mutate(pred = predBKT, species = 'bkt')
-
-ggplot(predGrid2BKT %>% filter(salmon01 == 0),
+# predGridBKT2 <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
+#                            
+#                            ydayCumul = seq(250,1000, by = 10),
+#                            salmon01 = c(0,1)) %>%
+#                     left_join(., yDayAll) %>%
+#                     mutate()
+
+pred <- predict(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr, newdata = predGrid)#, type = 'lpmatrix')
+#predBKT2 <- predict(mod1, newdata = predGridBKT2)#, type = 'lpmatrix')
+
+predGrid2 <- bind_cols(predGrid, as.data.frame(pred)) %>% mutate(pred = pred)
+#predGrid2BKT2 <- bind_cols(predGridBKT2, as.data.frame(predBKT2)) %>% mutate(pred = predBKT2, species = 'bkt')
+
+ggplot(predGrid2 %>% filter(salmon01 == 0),
 
@@
 #  guide_legend(title = 'Cohort') +
-  facet_grid(~riverFactor, scales = 'free_y')
-
-```
-
-```{r}
-#earlier runs before setting fixed effect first. kept forf reference
-# BKT 
-#             rowname        df      AIC
-# 1     modCR_SR_c_cS 162.34756 414267.2
-# 2      modCR_S_c_cS 163.43154 414602.7
-# 3        modCR_c_cS 161.63691 414750.3
-# 4  modCR_SR_c_cSR3a  88.21646 415388.9
-# 5  modCR_SR_c_cSR3b 167.32968 416062.3
-# 6    modCR_SR_c_cSR 163.93920 416587.5
-# 7   modCR_SR_c_cSR3 173.36884 416611.4
-# 8         modCR_S_c 157.23995 418645.9
-# 9           modCR_c 158.50186 418665.1
-# 10            modCR 154.36186 419135.1
-# 11           modC_R  63.10879 421258.1
-# 12              mod  17.99204 427309.6
-# 13             modC  48.73277 431440.2
-# 14             modR  22.80483 432214.7
-
-# BNT
-#             rowname       df      AIC
-# 1  modCR_SR_c_cSR3a 89.52290 231873.4
-# 2  modCR_SR_c_cSR3b 89.54674 231873.6
-# 3     modCR_SR_c_cS 84.40505 231935.7
-# 4      modCR_S_c_cS 83.47382 231962.5
-# 5   modCR_SR_c_cSR3 88.15840 232100.8
-# 6        modCR_c_cS 80.92171 232142.0
-# 7    modCR_SR_c_cSR 80.89406 232201.5
-# 8         modCR_S_c 79.53048 232281.7
-# 9           modCR_c 77.80808 232420.2
-# 10            modCR 78.58400 233816.0
-# 11           modC_R 53.21199 234214.3
-# 12              mod 13.99759 238544.8
-# 13             modR 14.97078 243472.0
-# 14             modC 48.73277 431440.2
+  facet_wrap(~riverFactor)#, scales = 'free_y')
 
@@
 #using salmon01 * river because we need diff effects per river - even though AIC is worse for this model
-modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) +
-               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn),
+modBKT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
@@
 
-#plot(modBKT, pages = 8, scale = 0)
-
-predGridBKT <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
-                           ydayCumul = seq(250,1000, by = 30),
+plot(modBKT, pages = 8, scale = 0)
+
+predGridBKT <- expand.grid(riverFactor = unique(modBKT$model$riverFactor), 
+                           cohortFactor = unique(modBKT$model$cohort),
+                           ydayCumul = seq(250,1000, by = 10),
                            salmon01 = c(0,1),
-                           cohortStrengthScaledBKT = seq(-2,2, by = 0.1)) %>%
-                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
-                           riverOrdered = substring(cohortRiverOrdered,6),
-                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
-                           cohortCentered = cohortOrdered - cohortOffset,
-                           cohort = cohortOrdered,
-                           river = riverOrdered)
+                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
+                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
+                           meanFlowScaledBySeason = 0) %>%
+                    left_join(., yDayAll) %>%
+                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
+                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
+                           cohort = cohortFactor,
+                           #river = riverFactor,
+                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)
 
@@
 
-modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
-               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
-               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
+modBNT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")
@@
 
-predGridBNT <- expand.grid(cohortRiverOrdered = unique(modBNT$model$cohortRiverOrdered),
-                           ydayCumul = seq(250,1000, by = 30),
+predGridBNT <- expand.grid(riverFactor = unique(modBNT$model$riverFactor), 
+                           cohortFactor = unique(modBNT$model$cohort),
+                           ydayCumul = seq(250,1000, by = 10),
                            salmon01 = c(0,1),
-                           metaPop01 = c(0,1),
-                           cohortStrengthScaledBKT = seq(-2,2, by = 0.25),
-                           cohortStrengthScaledBNT = seq(-2,2, by = 0.25),
-                           cohortStrengthScaledATS = seq(-2,2, by = 0.25)) %>%
-                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
-                           riverOrdered = substring(cohortRiverOrdered,6),
-                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
-                           cohortCentered = cohortOrdered - cohortOffset,
-                           cohort = cohortOrdered,
-                           river = riverOrdered)
+                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
+                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
+                           meanFlowScaledBySeason = 0) %>%
+                    left_join(., yDayAll) %>%
+                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
+                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
+                           cohort = cohortFactor,
+                           #river = riverFactor,
+                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)
 
@@
 
-predGrid2Both$riverGG <- factor(predGrid2Both$riverOrdered,
+predGrid2Both$riverGG <- factor(predGrid2Both$riverFactor,
                          levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
-predGrid2Both$speciesGG <- factor(predGrid2Both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))
+predGrid2Both$speciesGG <- factor(predGrid2Both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"))
 
@@
 ggplot(predGrid2Both %>% 
-         filter(
-         (cohortStrengthScaledBKT == 0 & salmon01 == 0) | (cohortStrengthScaledBKT == 0 & cohortStrengthScaledBNT == 0 & cohortStrengthScaledATS == 0 & metaPop01 == 1 & salmon01 ==0)),
+         filter(salmon01 ==0
+#         (cohortStrengthScaledBKT == 0 & salmon01 == 0) | (cohortStrengthScaledBKT == 0 & cohortStrengthScaledBNT == 0 & cohortStrengthScaledATS == 0 & metaPop01 == 1 & salmon01 ==0)
+           ),
 #           cohortStrengthScaledBKT == 0, cohortStrengthScaledBNT == 0, cohortStrengthScaledATS == 0, metaPop01 == 1, salmon01 == 0),
@@
 ```
+
+```{r models with season-don't run. gam structure given best model from above}
+
+# BKT, mod0 is best fixed effects model
+
+system.time(
+mod0_YDCcr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCc_r <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) , 
+               data = dataIn, 
+               method = "REML")
+)
+
+AIC(mod0_YDCcr, mod0_YDCc_r) %>% rownames_to_column() %>% arrange(AIC)
+
+system.time(
+mod0_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_Cr_CSr_Ts_Fs_TFs <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = season) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = season) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = season), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
+summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
+plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)
+
+system.time(
+mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_C_CS_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
+summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
+plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)
+
+# take out s()'s
+
+system.time(
+mod0_YDCcr_Cr_CSr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_Cr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               #s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+              # s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod0_YDCcr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+             #  s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+             #  s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+########
+
+AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CSr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_Tsr_Fsr_TFsr,mod0_YDCcr_CSr_Tsr_Fsr_TFsr, mod0_YDCcr_Tsr_Fsr_TFsr, mod6_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
+
+#                           rowname        df      AIC
+# 1  mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr 277.67003 401304.4
+# 2   mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr 281.22421 401407.3
+# 3   mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr 277.09498 401772.0
+# 4     mod0_YDCcr_CSr_Tsr_Fsr_TFsr 279.84632 402037.9
+# 5    mod0_YDCcr_C_CS_Tsr_Fsr_TFsr 269.69828 402073.1
+# 6     mod0_YDCcr_Cr_CSr_Tr_Fr_TFr 207.97536 402187.1
+# 7     mod0_YDCcr_Cr_CSr_Ts_Fs_TFs 190.64645 402464.7
+# 8      mod0_YDCcr_Cr_Tsr_Fsr_TFsr 289.33427 403529.7
+# 9         mod0_YDCcr_Tsr_Fsr_TFsr 283.93242 404139.1
+# 10              mod0_YDCcr_Cr_CSr 184.24038 413605.9
+# 11                     mod0_YDCcr 168.77693 420503.0
+# 12                    mod0_YDCc_r  71.99108 422677.4
+
+summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
+plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)
+
+#AIC(mod6_YDCcr_Cr_CSr_Tsr_Fsr_TFsr)
+#[1] 401367.4
+
+#> AIC(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr)
+#[1] 402517.3
+
+#########
+# models with seasonal effects give a drop in seqson 1 sizes try models without seasonal effects
+# may make sense because season is implicit in yDayCumul
+
+# model without season in fixed
+# this one gives a bump up in sizes for season 4
+system.time(
+mod6_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+# predictions from this models are smooth!
+# use this as the base model
+system.time(
+mod6_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+summary(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr) 
+plot(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr, pages = 8, shade = TRUE, scale = 0)
+
+```
+
+```{r}
+#earlier runs before setting fixed effect first. kept forf reference
+# BKT 
+#             rowname        df      AIC
+# 1     modCR_SR_c_cS 162.34756 414267.2
+# 2      modCR_S_c_cS 163.43154 414602.7
+# 3        modCR_c_cS 161.63691 414750.3
+# 4  modCR_SR_c_cSR3a  88.21646 415388.9
+# 5  modCR_SR_c_cSR3b 167.32968 416062.3
+# 6    modCR_SR_c_cSR 163.93920 416587.5
+# 7   modCR_SR_c_cSR3 173.36884 416611.4
+# 8         modCR_S_c 157.23995 418645.9
+# 9           modCR_c 158.50186 418665.1
+# 10            modCR 154.36186 419135.1
+# 11           modC_R  63.10879 421258.1
+# 12              mod  17.99204 427309.6
+# 13             modC  48.73277 431440.2
+# 14             modR  22.80483 432214.7
+
+# BNT
+#             rowname       df      AIC
+# 1  modCR_SR_c_cSR3a 89.52290 231873.4
+# 2  modCR_SR_c_cSR3b 89.54674 231873.6
+# 3     modCR_SR_c_cS 84.40505 231935.7
+# 4      modCR_S_c_cS 83.47382 231962.5
+# 5   modCR_SR_c_cSR3 88.15840 232100.8
+# 6        modCR_c_cS 80.92171 232142.0
+# 7    modCR_SR_c_cSR 80.89406 232201.5
+# 8         modCR_S_c 79.53048 232281.7
+# 9           modCR_c 77.80808 232420.2
+# 10            modCR 78.58400 233816.0
+# 11           modC_R 53.21199 234214.3
+# 12              mod 13.99759 238544.8
+# 13             modR 14.97078 243472.0
+# 14             modC 48.73277 431440.2
+
+```

### 2019-07-12 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
took T and F out of models

@@
 mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +
-               s(ydayCumul, bs = 'cr'), 
+               s(ydayCumul, bs = 'ps'), 
                data = dataIn, 
@@
 # 3    mod4 11.98667 243594.3
+
+# need to run code at top of 'predictions' block before running this chunk
+predGrid2 <- expand.grid(riverFactor = unique(dataIn$riverFactor),
+                           ydayCumul = seq(250,1000, by = 10),
+                           salmon01 = c(0,1),
+                         season = 2) %>%
+                    #left_join(., yDayAll) %>%
+                    mutate()
+
+pred2 <- predict(mod6, newdata = predGrid2)#, type = 'lpmatrix')
+
+predGrid2 <- bind_cols(predGrid2, as.data.frame(pred2)) %>% mutate(pred = pred2, species = 'bkt')
+
+ggplot(predGrid2 %>% filter(salmon01 == 0),
+
+       aes(ydayCumul, pred)) +
+  #geom_point() + 
+  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
+  geom_line(size = 1.1) +
+  labs(y = 'Predicted body size (mm)',
+       x = 'Days since age-0 Jan 1',
+       color = 'Cohort') +
+  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
+  theme_publication() +
+  theme(legend.position = "right", 
+        legend.direction = "vertical",
+        legend.title = element_text(face = "plain")) +
+  guides(color = guide_colorbar(barheight = 10,
+                                barwidth = 1.5,
+                                ticks = FALSE,
+                                nbin = 100)) +
+#  guide_legend(title = 'Cohort') +
+  facet_wrap(~riverFactor)#, scales = 'free_y')
 
@@
 
+```
+
+```{r models without T and F}
+# Not sure how to get models with season-specific T and F effects to work within season without messing up predictions. # Leave T and F out and focus on densities
+
+mod6_YDCcr_Cr_CSr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor),  # this is conspecific cS only
+               data = dataIn, 
+               method = "REML")
+
+system.time(
+mod6_YDCcr_Cr_CS3r <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDC <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr'),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr_CSBKTr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr_CSBNTr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr_CSATSr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+AIC(mod6_YDCcr_Cr_CSr, mod6_YDCcr_Cr_CS3r, mod6_YDCcr_Cr_CSBKTr, mod6_YDCcr_Cr_CSBNTr, mod6_YDCcr_Cr_CSATSr, mod6_YDCcr_Cr,mod6_YDCcr,mod6_YDC)  %>% rownames_to_column() %>% arrange(AIC)
+
+summary(mod6_YDCcr_Cr_CSr)
+plot(mod6_YDCcr_Cr_CS3r, pages = 7, shade = TRUE, scale = 0)
+
+# bkt
+#                rowname        df      AIC
+# 1    mod6_YDCcr_Cr_CSr 173.85954 413790.2
+# 2   mod6_YDCcr_Cr_CS3r 188.68933 415278.2
+# 3 mod6_YDCcr_Cr_CSBKTr 167.72761 416582.2
+# 4 mod6_YDCcr_Cr_CSBNTr 168.43917 417098.5
+# 5 mod6_YDCcr_Cr_CSATSr 167.84300 417569.6
+# 6        mod6_YDCcr_Cr 160.60559 417764.7
+# 7           mod6_YDCcr 149.80639 420935.0
+# 8             mod6_YDC  14.99337 429769.2
+
+# bnt
+#                rowname       df      AIC
+# 1   mod6_YDCcr_Cr_CS3r 96.91663 231844.1
+# 2    mod6_YDCcr_Cr_CSr 84.99160 231875.4
+# 3 mod6_YDCcr_Cr_CSBKTr 84.39999 232019.6
+# 4 mod6_YDCcr_Cr_CSBNTr 87.25717 232120.2
+# 5 mod6_YDCcr_Cr_CSATSr 85.68893 232203.3
+# 6        mod6_YDCcr_Cr 81.27554 232235.7
+# 7           mod6_YDCcr 75.00492 233885.5
+# 8             mod6_YDC 12.99685 238636.2
 ```
@@
                         salmon01 = c(0,1),
-                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
-                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
-                           meanFlowScaledBySeason = 0) %>%
-                    left_join(., yDayAll) %>%
+                        cohortStrengthScaled = seq(-2.5,2.5, by = 0.25)) %>%
+                    #left_join(., yDayAll) %>%
                     mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
-                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
-                           cohort = cohortFactor,
-                           #river = riverFactor,
-                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)
-
-# predGridBKT2 <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
-#                            
-#                            ydayCumul = seq(250,1000, by = 10),
-#                            salmon01 = c(0,1)) %>%
-#                     left_join(., yDayAll) %>%
-#                     mutate()
-
-pred <- predict(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr, newdata = predGrid)#, type = 'lpmatrix')
-#predBKT2 <- predict(mod1, newdata = predGridBKT2)#, type = 'lpmatrix')
+                           cohort = cohortFactor)
+
+pred <- predict(mod6_YDCcr_Cr_CSr, newdata = predGrid)#, type = 'lpmatrix')
 
 predGrid2 <- bind_cols(predGrid, as.data.frame(pred)) %>% mutate(pred = pred)
-#predGrid2BKT2 <- bind_cols(predGridBKT2, as.data.frame(predBKT2)) %>% mutate(pred = predBKT2, species = 'bkt')
-
-ggplot(predGrid2 %>% filter(salmon01 == 0),
+
+ggplot(predGrid2 %>% filter(salmon01 == 0, cohortStrengthScaled == 0),
 
@@
                s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor), 
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
@@
 
-predGridBKT <- expand.grid(riverFactor = unique(modBKT$model$riverFactor), 
-                           cohortFactor = unique(modBKT$model$cohort),
-                           ydayCumul = seq(250,1000, by = 10),
+predGridBKT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
+                        cohortFactor = unique(dataIn$cohort),
+                        ydayCumul = seq(250,1000, by = 30),
                         salmon01 = c(0,1),
-                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
-                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
-                           meanFlowScaledBySeason = 0) %>%
-                    left_join(., yDayAll) %>%
+                        cohortStrengthScaled = seq(-2.5,2.5, by = 0.25)) %>%
+                    #left_join(., yDayAll) %>%
                     mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
-                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
-                           cohort = cohortFactor,
-                           #river = riverFactor,
-                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)
+                           cohort = cohortFactor)
 
@@
                s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor), 
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")
@@
 
-predGridBNT <- expand.grid(riverFactor = unique(modBNT$model$riverFactor), 
-                           cohortFactor = unique(modBNT$model$cohort),
-                           ydayCumul = seq(250,1000, by = 10),
+predGridBNT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
+                           cohortFactor = unique(dataIn$cohort),
+                           ydayCumul = seq(250,1000, by = 30),
                            salmon01 = c(0,1),
-                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
-                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
-                           meanFlowScaledBySeason = 0) %>%
-                    left_join(., yDayAll) %>%
+                           cohortStrengthScaledBKT = seq(-2.5,2.5, by = 0.5),
+                           cohortStrengthScaledBNT = seq(-2.5,2.5, by = 0.5),
+                           cohortStrengthScaledATS = seq(-2.5,2.5, by = 0.5)) %>%
+                    #left_join(., yDayAll) %>%
                     mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
-                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
-                           cohort = cohortFactor,
-                           #river = riverFactor,
-                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)
+                           cohort = cohortFactor)
 

### 2019-07-12 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
updated plotGams

@@
 summary(modBKT)
-
 plot(modBKT, pages = 8, scale = 0)
 
-predGridBKT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
-                        cohortFactor = unique(dataIn$cohort),
+predGridBKT <- expand.grid(riverFactor = unique(modBKT$model$riverFactor), 
+                           cohortFactor = unique(modBKT$model$cohort),
                            ydayCumul = seq(250,1000, by = 30),
@@
 # Brown trout
-
+# using CSr model becasue delta AIC is not huge btw it and CS3r. Better to have same models for both species
 modBNT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
@@
                s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
-               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor) +
-               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor), 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor), 
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")
@@
 
-predGridBNT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
-                           cohortFactor = unique(dataIn$cohort),
+predGridBNT <- expand.grid(riverFactor = unique(modBNT$model$riverFactor), 
+                           cohortFactor = unique(modBNT$model$cohort),
                            ydayCumul = seq(250,1000, by = 30),
                            salmon01 = c(0,1),
-                           cohortStrengthScaledBKT = seq(-2.5,2.5, by = 0.5),
-                           cohortStrengthScaledBNT = seq(-2.5,2.5, by = 0.5),
-                           cohortStrengthScaledATS = seq(-2.5,2.5, by = 0.5)) %>%
+                           cohortStrengthScaled = seq(-2.5,2.5, by = 0.25)) %>%
                     #left_join(., yDayAll) %>%
@@
 ggplot(predGrid2Both %>% 
-         filter(salmon01 ==0
+         filter(salmon01 ==0, cohortStrengthScaled == 0
 #         (cohortStrengthScaledBKT == 0 & salmon01 == 0) | (cohortStrengthScaledBKT == 0 & cohortStrengthScaledBNT == 0 & cohortStrengthScaledATS == 0 & metaPop01 == 1 & salmon01 ==0)

### 2019-07-12 functions/plotGams.R

@@ -1,14 +1,18 @@
 
 plotGamVarBySpecies <- function(var = 's(cohort)', xAxisLab = 'Cohort', bkt = modBKT, bnt = modBNT){
   require(gratia)
-  bkt <- evaluate_smooth(modBKT, var) %>% mutate(species = 'bkt')
-  bnt <- evaluate_smooth(modBNT, var) %>% mutate(species = 'bnt')
+  BKT <- evaluate_smooth(modBKT, var) %>% mutate(species = 'bkt')
+  BNT <- evaluate_smooth(modBNT, var) %>% mutate(species = 'bnt')
   both <- bind_rows(BKT,BNT) %>% mutate(upperCI = est + se * 1.96, lowerCI = est - se * 1.96)
   
   both$speciesGG <- factor(both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-  
+  both$riverFactor <- factor(both$riverFactor,
+                                  levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
+                                  labels = c("West Brook","Open Large","Open Small","Isolated Small"))  
+  xVar <- substr(var,3,nchar(var)-1)
+  byVar <- as.character(unique(both$by_variable))
   gg <- 
-    ggplot(both, aes(cohort, group = speciesGG)) + 
+    ggplot(both, aes(x = eval(as.name(xVar)), group = speciesGG)) + 
       geom_ribbon(aes( ymin = lowerCI, ymax = upperCI), fill  = 'grey70') +
       geom_line(aes(y = est, linetype = speciesGG), size = 1.5) +
       labs(y = 'Effect size',
@@ -17,7 +21,8 @@ plotGamVarBySpecies <- function(var = 's(cohort)', xAxisLab = 'Cohort', bkt = mo
       theme_publication() +
       theme(legend.position = "right", 
             legend.direction = "vertical",
-            legend.title = element_text(face="plain"))
+            legend.title = element_text(face="plain")) +
+      facet_wrap(~ eval(as.name(byVar)))
   gg
 } 
   # 

### 2019-07-15 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
added figures for MS

@@
 
-# by species, by cohort gridded
-# ggplot(vB %>% filter(cohort %in% 2000:2013), aes(ydayCumul,observedLength, group = species, color = (species))) +
-#   geom_point(size = 0.5, alpha = 0.1) +
-#  # geom_smooth(span = 0.1, size = 1.5, se = F) +
+ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats')), aes(ydayCumul,observedLength)) +
+  geom_point(size = 0.5, alpha = 0.1) +
+ # geom_smooth(span = 0.1, size = 1.5, se = F) +
  # geom_smooth(method='gam', alpha=0.1, formula=y~s(x, k = kValue), se = F) +
-#   facet_wrap( ~ cohort) +
-#   theme_publication()
+  labs(y = 'Observed body size (mm)',
+       x = 'Days since age-0 Jan 1') +
+  facet_wrap(~ speciesGG) +
+  theme_publication()
+
+ggsave('./figures/rawBySpecies.png', width = 11, height = 8, units='in')
+
+# by species (bkt, bnt), by cohort, by river gridded
+ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats')), aes(ydayCumul,observedLength, group = species, color = (species))) +
+  geom_point(size = 0.5, alpha = 0.1) +
+ # geom_smooth(span = 0.1, size = 1.5, se = F) +
+  geom_smooth(method='gam', alpha=0.1, formula=y~s(x, k = kValue), se = F) +
+  facet_wrap( ~ cohort) +
+  theme_publication()
 
@@
   theme_publication()
+
+```
+
+```{r rivers where fish are fish}
+
+vB$riverGG <- factor(vB$riverFactor,
+                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
+                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))
+
+ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats'), !is.na(observedLength)),
+       aes(riverGG, fill = speciesGG)) +
+  geom_bar() +
+  labs(x = 'River', y = 'Number of observations', fill = 'Species') +
+  theme_publication() +
+  scale_fill_grey() +
+  theme(legend.position = "right", 
+    legend.direction = "vertical",
+    legend.title = element_text(face = "plain"))
+
+ggsave('./figures/rawCountsBySpecies.png', width = 11, height = 8, units='in')
+
 ```
@@
 summary(modBKT)
-plot(modBKT, pages = 8, scale = 0)
+#plot(modBKT, pages = 8, scale = 0)
 
@@
                            cohortFactor = unique(modBKT$model$cohort),
-                           ydayCumul = seq(250,1000, by = 30),
+                           ydayCumul = seq(250,1000, by = 14),
                            salmon01 = c(0,1),
@@
 summary(modBNT)
-#plot(modBNT, pages = 8, scale = 0)
+#plot(modBNT,  scale = 0)
 
@@
                            cohortFactor = unique(modBNT$model$cohort),
-                           ydayCumul = seq(250,1000, by = 30),
+                           ydayCumul = seq(250,1000, by = 14),
                            salmon01 = c(0,1),
@@
 
-ggsave('./figures/predLengths.png', width = 11, height = 11, units='in')
-
-# specific splines by species
+ggsave('./figures/predLengths14.png', width = 11, height = 8, units='in')
+
+# specific spline effects by species
 (plotCohort <- plotGamVarBySpecies(var = 's(cohort)', xAxisLab = 'Cohort'))
+ggsave('./figures/cohortEffect.png', width = 11, height = 8, units='in')
 (plotCohortStrength <- plotGamVarBySpecies(var = 's(cohortStrengthScaled)', xAxisLab = 'Cohort Strength'))
-
+ggsave('./figures/cohortStrengthEffect.png', width = 11, height = 8, units='in')
 #
 plot.gam(modBKT, pages = 10)
+
+# plot specific predictions
+
+# across rivers, by species
+ggplot(predGrid2Both %>% 
+         filter(salmon01 ==0, cohortStrengthScaled == 0, cohort == 2006, cohortStrengthScaled == 0),
+       aes(ydayCumul, pred, group = speciesGG, color = speciesGG)) +
+  #geom_point() + 
+  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
+  geom_line(size = 1.1) +
+  labs(y = 'Predicted body size (mm)',
+       x = 'Days since age-0 Jan 1',
+       color = 'Cohort') +
+#  scale_color_viridis_c() +
+  theme_publication() +
+  theme(legend.position = "right", 
+        legend.direction = "vertical",
+        legend.title = element_text(face = "plain")) +
+  facet_wrap(~ riverGG)
+
+ggsave('./figures/ydayCumulAcrossRiverBySpp.png', width = 11, height = 8, units='in')
+
+# across species, by rivers
+ggplot(predGrid2Both %>% 
+         filter(salmon01 == 0, cohortStrengthScaled == 0, cohort == 2006, cohortStrengthScaled == 0),
+       aes(ydayCumul, pred, group = riverGG, color = riverGG)) +
+  #geom_point() + 
+  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
+  geom_line(size = 1.1) +
+  labs(y = 'Predicted body size (mm)',
+       x = 'Days since age-0 Jan 1',
+       color = 'River') +
+  #scale_color_viridis_c() +
+  theme_publication() +
+  theme(legend.position = "right", 
+        legend.direction = "vertical",
+        legend.title = element_text(face = "plain")) +
+  facet_grid(~ speciesGG)
+
+ggsave('./figures/ydayCumulAcrossSppByRiver.png', width = 11, height = 8, units='in')
 

### 2019-07-25 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
update figures

@@
             #!(riverOrdered == 'wb mitchell' & ageInSamples > 8) #low sample sizes
-            )
+            ) %>%
+    filter(!(cohort < 2002 & river %in% c('wb jimmy','wb mitchell','wb obear'))) # added 7/25/2019. Limits tribs to after 2002
 
@@
 
+#ggplot(cohortStrength, aes(cohortStrengthScaled, color = species)) + geom_freqpoly()
+
 vB <- left_join(vB, cohortStrength)
-
-cohortStrengthWide <- spread(cohortStrength, species, cohortStrengthScaled) %>%
+# 
+# cohortStrengthWide <- spread(cohortStrength, species, cohortStrengthScaled) %>%
+#   rename(cohortStrengthScaledBKT = bkt,
+#          cohortStrengthScaledBNT = bnt,
+#          cohortStrengthScaledATS = ats
+#          ) %>%
+#   dplyr::select(-cohortStrength) %>%
+#   ungroup()
+
+cohortStrengthWide <- cohortStrength %>%
+  dplyr::select(-cohortStrength) %>%
+  spread( species, cohortStrengthScaled) %>%
   rename(cohortStrengthScaledBKT = bkt,
@@
          ) %>%
-  dplyr::select(-cohortStrength)
+  ungroup()
 
@@
 
-vB$cohortStrengthScaledBKT <- ifelse(is.na(vB$cohortStrengthScaledBKT), -2.5, vB$cohortStrengthScaledBKT)
-vB$cohortStrengthScaledBNT <- ifelse(is.na(vB$cohortStrengthScaledBNT), -2.5, vB$cohortStrengthScaledBNT)
-vB$cohortStrengthScaledATS <- ifelse(is.na(vB$cohortStrengthScaledATS), -2.5, vB$cohortStrengthScaledATS)
-
-ggplot(cohortStrength, aes(cohort, cohortStrengthScaled, group = species, linetype = species, shape = species, color = species)) +
+vB$cohortStrengthScaledBKT2p5 <- ifelse(is.na(vB$cohortStrengthScaledBKT), -2.5, vB$cohortStrengthScaledBKT)
+vB$cohortStrengthScaledBNT2p5 <- ifelse(is.na(vB$cohortStrengthScaledBNT), -2.5, vB$cohortStrengthScaledBNT)
+vB$cohortStrengthScaledATS2p5 <- ifelse(is.na(vB$cohortStrengthScaledATS), -2.5, vB$cohortStrengthScaledATS)
+
+vB$cohortStrength01BKT <- ifelse(is.na(vB$cohortStrengthScaledBKT), 0, 1)
+vB$cohortStrength01BNT <- ifelse(is.na(vB$cohortStrengthScaledBNT), 0, 1)
+vB$cohortStrength01ATS <- ifelse(is.na(vB$cohortStrengthScaledATS), 0, 1)
+
+cohortStrength$speciesGG <- factor(cohortStrength$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"))
+
+ggplot(cohortStrength %>% filter(cohort > 1999), aes(cohort, cohortStrengthScaled, group = speciesGG, shape = speciesGG, color = speciesGG)) +
   geom_point(size = 2) +
-  geom_line(size = 1.1, color = 'grey40') +
+  geom_line(size = 1.2) +
   labs(x = 'Cohort',
-       y = 'Cohort strength') +
+       y = 'Cohort strength',
+       shape = 'Species', color='Species') +
   theme_publication() +
-  facet_grid(~metaPop01)
+ # scale_fill_grey() +
+  theme(legend.position = "right", 
+    legend.direction = "vertical",
+    legend.title = element_text(face = "plain")) +
+  facet_grid(metaPop01 ~ .)
+
+ggsave('./figures/cohortStrengthByMetapop.png', width = 11, height = 8, units='in')
 
@@
 ```{r dataIn}
-dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn)
+dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn) %>%
+  filter(!(cohort < 2002 & riverFactor %in% c('wb jimmy','wb mitchell','wb obear')))
 ```
@@
 
-```{r mean lengths: from MS_Analysis_3spp_WB.Rmd. For Grenada talk}
-
-# Has mean size changed over time?
-means <- 
-  cd %>% mutate( age = year - cohort ) %>%
-    filter( riverOrdered == 'west brook', ageInSamples %in% 1:10, age < 3, year < 2016 ) %>%
-    group_by(species,ageInSamples,year,season,age) %>%
-    summarize(meanLen = mean(observedLength,na.rm = TRUE),
-              sdLen = sd(observedLength,na.rm = TRUE),
-              meanGR = mean(grLength,na.rm = TRUE),
-              sdGR = sd(grLength,na.rm = TRUE)
-              )
-
-lenSlopes <- means %>%
-  group_by( species, ageInSamples, season, age ) %>%
-  nest()
-
-lengthModel <- function(df){
-  lm( meanLen ~ year, data = df )
-}
-
-grModel <- function(df){
-  lm( meanGR ~ year, data = df )
-}
-
-lenSlopes <- lenSlopes %>%
-  mutate( model = map(lenSlopes$data, lengthModel),
-          modelGR = map(lenSlopes$data, grModel))
-
-###################################
-######### len here, gr below ######
-lenSlopesTidy <- lenSlopes %>%
-  mutate(glance = map(model, broom::tidy)) %>% 
-  unnest(glance, .drop = TRUE) %>%
-  filter(term == 'year') %>%
-  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
-  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
-  mutate( sig = ifelse(p.value < 0.0001,"***",sig))
-
-means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-lenSlopesTidy$seasonGG <- factor(lenSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-lenSlopesTidy$speciesGG <- factor(lenSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanLen, group= speciesGG)) +
-  geom_point( aes(color = speciesGG), size = 4) +
-  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
-  labs( x = "Year", y = "Mean body length (mm)" ) +
-
-  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age<2), 
-            aes(x=2000,y=225, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
-  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age<2), 
-            aes(x=2000,y=195, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
-
-  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), 
-            aes(x=2011.8,y=104, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
-  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), 
-            aes(x=2011.8,y=74, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+
-
-  theme_publication(base_size = 24) +
-  theme(legend.title = element_text("Species"),
-      #  legend.position = "right",
-     #   legend.direction = "vertical"
-        legend.position = "bottom",
-        legend.direction = "horizontal") + 
-  labs(color='Species') +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(seasonGG ~ age, scales='free')
-
-ggsave('./figures/meanSizes.png', width = 11, height = 11, units='in')
-
-# one season/year only
-ggplot( means %>% filter(year >= 2000, species != 'ats', ageInSamples == 8), aes(year,meanLen, group= speciesGG)) +
-  geom_point( aes(color = speciesGG), size = 4) +
-  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
-  labs( x = "Year", y = "Mean body length (mm)" ) +
-  theme_publication(base_size = 24) +
-  theme(legend.title = element_text("Species"),
-      #  legend.position = "right",
-     #   legend.direction = "vertical"
-                legend.position = "bottom",
-        legend.direction = "horizontal") + 
-  labs(color='Species') +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(seasonGG ~ age)
-ggsave('./figures/meanSizesAIS_1.png', width = 11, height = 8, units='in')
-
-# is a model with salmon presence absence better than a year model?
-means$ats01 <- ifelse(means$year <= 2005, 1, 0)
-
-modMeans1 = lm(meanLen ~ year * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
-modMeans2 = lm(meanLen ~ year * factor(ats01) * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
-modMeans3 = lm(meanLen ~ factor(ats01) * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
-modMeans4 = lm(meanLen ~ factor(ats01)*year, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
-AIC(modMeans1, modMeans2, modMeans3, modMeans4)
-
-# one season/year only
-ggplot( means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1), aes(year,meanLen, group= speciesGG)) +
-  geom_point( aes(color = speciesGG), size = 4) +
-  geom_smooth( aes(group = factor(ats01)), color = 'black', se = FALSE, method = 'lm') +
-  labs( x = "Year", y = "Mean body length (mm)" ) +
-  theme_publication(base_size = 24) +
-  theme(legend.title = element_text("Species"),
-      #  legend.position = "right",
-     #   legend.direction = "vertical"
-        legend.position = "bottom",
-        legend.direction = "horizontal") + 
-  labs(color='Species') +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(seasonGG ~ age)
-
-ggsave('./figures/meanSizesAIS_1_ats01.png', width = 11, height = 8, units='in')
-
-# mean increases 
-lenSlopesTidy %>%
-  filter(species != 'ats') %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# mean increases by species
-lenSlopesTidy %>%
-  group_by(species) %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# do increases (slopes) increase over age? Yes
-ggplot(lenSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
-  geom_point(size = 5) +
-  geom_smooth(method='lm', se=FALSE, size = 1.5) +
-  scale_x_continuous(breaks = 1:10, labels = c('0 Fall','0 Winter','1 Spring','1 Summer','1 Fall','1 Winter','2 Spring','2 Summer','2 Fall','2 Winter')) +
-  labs( x = "Age Season", y = "Slope of mean body length over years" ) +
-  theme_publication(base_size = 24) +
-  theme(legend.title = element_text("Species"),
-        legend.position = "bottom",
-        legend.direction = "horizontal",
-        axis.text.x = element_text(angle = 40, hjust = 1)) + 
-  labs(color='Species') +
-  scale_color_grey(start = 0.15, end = 0.75)
-
-ggsave(paste0('figures/meanLengthsOverAge.png'), width = 11, height = 10, units='in')
-
-# species diff in increases in length?
-sppEffect <- (lm(estimate ~ ageInSamples * (species), data = lenSlopesTidy))
-summary(sppEffect)
-anova(sppEffect)
-
-# mean diff in slopes btw bkt and bnt
-lenSlopesTidy %>%
-  filter(species != 'ats') %>%
-  group_by(species) %>%
-  summarise(means = mean(estimate))
-
-#####################################  
-#################################
-# growth rate
-grSlopesTidy <- lenSlopes %>%
-  mutate(glance = map(modelGR, broom::tidy)) %>% 
-  unnest(glance, .drop = TRUE) %>%
-  filter(term == 'year') %>%
-  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
-  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
-  mutate( sig = ifelse(p.value < 0.0001,"***",sig))
-
-##
-means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanGR, group= speciesGG)) +
-  geom_point( aes(color = speciesGG), size = 4) +
-  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
-  labs( x = "Year", y = "Mean growth rate (mm/d)" ) +
-  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age<2), 
-            aes(x=2000,y=0.68, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
-  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age<2), 
-            aes(x=2000,y=0.55, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
-
-  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), 
-            aes(x=2011.5,y=0.64, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
-  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), 
-            aes(x=2011.5,y=0.51, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+
-  theme_publication(base_size = 24) +
-  theme(legend.title = element_text("Species"),
-      #  legend.position = "right",
-     #   legend.direction = "vertical"
-        legend.position = "bottom",
-        legend.direction = "horizontal") + 
-  labs(color='Species') +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(seasonGG~age)
-
-ggsave('./figures/meanGRs.png', width = 11, height = 11, units='in')
-
-# mean increases 
-grSlopesTidy %>%
-  filter(species != 'ats') %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# mean increases by species
-grSlopesTidy %>%
-  group_by(species) %>%
-  summarise( means = mean(estimate),
-             sd = sd(estimate))
-
-# do increases (slopes) increase over age? Yes
-ggplot(grSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
-  geom_point(size = 5) +
-  geom_smooth(method='lm', se=FALSE) +  
-  geom_hline(yintercept = 0) +
-  scale_x_continuous(breaks = 1:10, labels = c('0 Fall','0 Winter','1 Spring','1 Summer','1 Fall','1 Winter','2 Spring','2 Summer','2 Fall','2 Winter')) +
-  labs( x = "Age Season", y = "Slope of mean growth rate over years" ) +
-  theme_publication(base_size = 24) +
-  theme(legend.title = element_text("Species"),
-        legend.position = "bottom",
-        legend.direction = "horizontal",
-        axis.text.x = element_text(angle = 40, hjust = 1)) + 
-  labs(color='Species') +
-  scale_color_grey(start = 0.15, end = 0.75)
-
-ggsave('./figures/meanGRsOverAge.png', width = 11, height = 11, units='in')
-
-summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bkt')))
-summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bnt')))
-
-# species diff in increases in gr?
-sppEffectGR <- (lm(estimate ~ ageInSamples * (species), data = grSlopesTidy))
-summary(sppEffectGR)
-anova(sppEffectGR)
-
-# mean diff in slopes btw bkt and bnt
-grSlopesTidy %>%
-  filter(species != 'ats') %>%
-  group_by(species) %>%
-  summarise(means = mean(estimate))
-
-```
-
-```{r size-dep imm or mort? For Spain talk}
-
-### for losses
-
-cd01 <-  cd %>%
-  filter(knownZ == 1) %>%
-  group_by(tagIndex) %>%
-  mutate(first = min(sampleNumber),
-         firstObs = sampleNumber == first,
-         last = max(sampleNumber),
-         lastObs = sampleNumber == last,
-         notLastObs = !lastObs,
-         nObs = n(),
-         age = year - cohort
-  ) %>%
-  ungroup()
-
-meansByAIS <- cd01 %>%
-  group_by(species, ageInSamples, cohort, river) %>%
-  summarise(meanLen = mean(observedLength, na.rm =T),
-            sdLen = sd(observedLength, na.rm = T),
-            n = n())
-cd01 <- left_join(cd01, meansByAIS) %>%
-  mutate(observedLengthStd = (observedLength - meanLen) / sdLen,
-         cohortStd = cohort - 2007)
-
-#cd %>% select(tagIndex,sampleNumber, firstObs,lastObs, knownZ, observedLength) %>% head(50) %>% as.data.frame()
-cd01Mod <- cd01 %>% filter(species %in% c('bkt', 'bnt'), river == 'west brook', ageInSamples %in% 1:10, cohort %in% 2000:2014)
-
-ggplot(cd01Mod, aes(observedLengthStd, notLastObs*1, color = factor(species), group = factor(species))) +
-  geom_point(alpha = 0.01) +
-  geom_smooth(method = "glm", 
-    method.args = list(family = "binomial"), 
-    se = FALSE) +
- # scale_color_viridis_c() +
-  facet_wrap( ~ cohort, scales = 'free')
-
-#### by species and ageInSamples
-
-sensSlopes <- cd01Mod %>%
-  group_by( species, ageInSamples ) %>%
-  nest()
-
-survModel <- function(df){
-  glm(notLastObs*1 ~ observedLengthStd + cohortStd, family = 'binomial', data = df )
-}
-enterModel <- function(df){
-  glm(firstObs*1 ~ observedLengthStd + cohortStd, family = 'binomial', data = df )
-}
-
-sensSlopes <- sensSlopes %>%
-  mutate(modelSurv = map(sensSlopes$data, survModel),
-         modelEnter = map(sensSlopes$data, enterModel))
-
-sensSlopesTidy <- sensSlopes %>%
-  mutate(surv = map(modelSurv, broom::tidy),
-         enter = map(modelEnter, broom::tidy)) %>% 
-  unnest(surv, enter, .drop = TRUE) %>%
-  filter(term == 'cohortStd') %>% # cohort effect
-  rename(estimateSurv = estimate, estimateEnter = estimate1) %>%
-  select(species,ageInSamples,estimateSurv, estimateEnter) #%>%
-  #gather("type", "estimate", estimateSurv:estimateEnter)
-
-ggplot(sensSlopesTidy, aes(ageInSamples, estimate, color = type)) +
-  geom_point() +
-  geom_line() +
-  facet_wrap(~species)
-
-```
-
-```{r dummy graphs for Spain talk}
-
-sizesForGraph <- data.frame(year = c(2000,2015), surv = c(100, 150))
-ggplot(sizesForGraph, aes(year,surv)) +
-  geom_line(size = 2) +
-  labs(x = 'Year', y = 'Mean body length (mm)') +
-  ylim(80,200) +
-  theme_publication(base_size = 36)
-ggsave('./figures/sizesDummy.png', width = 11, height = 11, units='in')
-
-survForGraph <- data.frame(size = c(-2,2), surv = c(0.25,0.75))
-ggplot(survForGraph, aes(size,surv)) +
-  geom_line(size = 2) +
-  labs(x = 'Standardized length', y = 'Probability of seen again') +
-  theme_publication(base_size = 36)
-ggsave('./figures/pSurvivalDummy.png', width = 11, height = 11, units='in')
-
-firstForGraph <- data.frame(size = c(-2,2), surv = c(0.35,0.65))
-ggplot(firstForGraph, aes(size,surv)) +
-  geom_line(size = 2) +
-  labs(x = 'Standardized length', y = 'Probability of first observation') +
-  theme_publication(base_size = 36)
-ggsave('./figures/pFirstDummy.png', width = 11, height = 11, units='in')
-
-sizesForGraph <- data.frame(year = c(2000,2015), surv = c(100, 150))
-ggplot(sizesForGraph, aes(year,surv)) +
-  geom_line(size = 3) +
-  labs(x = 'Year', y = 'Mean body length (mm)') +
-  ylim(80,200) +
-  theme_publication(base_size = 36)
-ggsave('./figures/sizesDummy.png', width = 11, height = 11, units='in')
-
-#### show interactions
-
-survForGraph <- data.frame(ind = c(1,1,2,2,3,3), size = c(-2,2), surv = c(0.52,0.55,0.51,0.24,0.54,0.82))
-ggplot(survForGraph, aes(size,surv, group = ind)) +
-  geom_line(size = 2) +
-  labs(x = 'Standardized length', y = 'Probability of seen again') +
-  theme_publication(base_size = 36)
-ggsave('./figures/pSurvivalDummy3.png', width = 11, height = 11, units='in')
-
-firstForGraph <- data.frame(ind = c(1,1,2,2,3,3), size = c(-2,2), surv = c(0.42,0.45,0.41,0.14,0.44,0.72))
-ggplot(firstForGraph, aes(size,surv, group = ind)) +
-  geom_line(size = 2) +
-  labs(x = 'Standardized length', y = 'Probability of first observation') +
-  theme_publication(base_size = 36)
-ggsave('./figures/pFirstDummy3.png', width = 11, height = 11, units='in')
-
-```
-
-```{r get betas from growth model for sensitivities}
-load("D:/projects/linkedModels/data/out/allBetas.RData") #loads df 'q'
-qFig1 <- q %>% filter(riverGG == 'WB', species %in% c('bkt','bnt'))
-qFig1$season <- as.numeric(as.character(qFig1$season))
-qFig1$isYOY <- as.numeric(as.character(qFig1$isYOY))
-qFig <- qFig1 %>%
-  mutate( varText = tolower(betaGG)) %>%
-  mutate( varText = ifelse(varText == 'temp*flow','temp:flow', varText)) %>%
-  mutate( varText = ifelse(varText == 'bnt*bkt','bkt:bnt', varText)) %>%
-  select(isYOY,season,species,median,varText)
-qFig$seasonGG <- factor(qFig$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-#get slopes of env variables
-
-load('D:/projects/linkedModels/data/out/flowSlopesTidy.RData')
-flowSlopesTidy2 <- flowSlopesTidy %>% filter(riverOrderedGG == 'West Brook') %>% select(estimate, seasonGG) %>% mutate(varText = 'flow')
-load('D:/projects/linkedModels/data/out/tempSlopesTidy.RData')
-tempSlopesTidy2 <- tempSlopesTidy %>% filter(riverOrderedGG == 'West Brook') %>% select(estimate, seasonGG) %>% mutate(varText = 'temp')
-load('D:/projects/linkedModels/data/out/nSlopesTidy.RData')
-nSlopesTidy2 <- nSlopesTidy %>% select(estimate, season, species) %>% mutate(varText = species)
-nSlopesTidy2$seasonGG <- factor(nSlopesTidy2$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-nSlopesTidy2 <- nSlopesTidy2 %>% select(estimate,seasonGG,varText)
-
-allSlopesTidy <- bind_rows(nSlopesTidy2, tempSlopesTidy2, flowSlopesTidy2)
-
-allSlopesTidyWBetas <- left_join(qFig, allSlopesTidy, by = c('seasonGG', 'varText')) %>% mutate(sens = median * estimate)
-
-ggplot(allSlopesTidyWBetas%>%filter(!is.na(sens), varText != 'ats'), aes(seasonGG, sens, color = varText)) +
-  geom_point() +
-  facet_grid(species + isYOY ~ varText, scales = 'free')
-```
-
-```{r }
-############# 1st attempt at imm em
-mod1 <- glm(notLastObs*1 ~ observedLengthStd * cohortStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
-#mod2 <- glm(notLastObs*1 ~ observedLengthStd * factor(cohortStd) * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
-mod3 <- glm(notLastObs*1 ~ observedLengthStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
-
-AIC(mod1,mod3)
-
-summary(mod1)
-anova(mod1, test="Chi")
-
-pred <- expand.grid(observedLengthStd = seq(-2,2, length.out = 50), cohortStd = unique(cd01Mod$cohortStd), species = c('bkt', 'bnt'), ageInSamples = unique(cd01Mod$ageInSamples))
-
-pred$pred <- predict.glm(mod1, newdata = pred, type = 'response')
-
-u <- cd01Mod %>% distinct(age, season, ageInSamples) %>% arrange(ageInSamples) %>% filter(!(age==2 & ageInSamples == 6))
-pred <- left_join(pred, u)
-
-pred$seasonGG <- factor(pred$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-pred$cohort <- pred$cohortStd + 2007
-pred$speciesGG <- factor(pred$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-ggplot(pred, aes(observedLengthStd, pred, color = (cohortStd), group = (cohortStd))) +
-  geom_point() +
-  scale_color_viridis_c() +
-  scale_y_continuous("Probability of survival") +
-  facet_grid(species ~ ageInSamples, scales = 'free')
-
-ggplot(pred %>% filter(cohortStd %in% 0, age < 3), aes(observedLengthStd, pred, color = speciesGG)) +
-  geom_point(size = 2) +
-  scale_x_continuous("Standardized length") +
-  scale_y_continuous(lim = c(0,1),"Probability of seen again") +
-  theme_publication(base_size = 28) +
-  theme(legend.title = element_text("Species"),
-        legend.position = "bottom",
-        legend.direction = "horizontal"
-        ) + 
-  labs(color='Species') +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(seasonGG ~ age)
-
-ggsave('./figures/pSurvival.png', width = 11, height = 11, units='in')
-
-################################
-# gains - immigration, firstObs
-
-ggplot(cd01Mod, aes(observedLengthStd, firstObs*1, color = factor(species), group = factor(species))) +
-  geom_jitter(alpha = 0.1) +
-  geom_smooth(method = "glm", 
-    method.args = list(family = "binomial"), 
-    se = FALSE) +
- # scale_color_viridis_c() +
-  xlim(c(-2.75,2.75)) +
-  facet_grid(ageInSamples ~ cohort)
-
-ggplot(cd01Mod%>%filter(ageInSamples == 3), aes(observedLengthStd, firstObs*1, color = factor(species), group = factor(species))) +
-  geom_jitter(alpha = 0.1) +
-  geom_smooth(method = "glm", 
-    method.args = list(family = "binomial"), 
-    se = FALSE) +
- # scale_color_viridis_c() +
-  xlim(c(-2.75,2.75)) +
-  facet_wrap( ~ cohort)
-
-modi1 <- glm(firstObs*1 ~ observedLengthStd * cohortStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
-modi2 <- glm(firstObs*1 ~ observedLengthStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
-modi3 <- glm(firstObs*1 ~ observedLengthStd * species , family = 'binomial', data = cd01Mod)
-modi4 <- glm(firstObs*1 ~ observedLengthStd * factor(ageInSamples), family = 'binomial', data = cd01Mod)
-modi5 <- glm(firstObs*1 ~ observedLengthStd, family = 'binomial', data = cd01Mod)
-
-AIC(modi1,modi2,modi3,modi4,modi5) %>% rownames_to_column() %>% arrange(AIC)
-
-summary(modi1)
-anova(modi1, test="Chi")
-###
-predi <- expand.grid(observedLengthStd = seq(-2,2, length.out = 50), cohortStd = unique(cd01Mod$cohortStd), species = c('bkt', 'bnt'), ageInSamples = unique(cd01Mod$ageInSamples))
-
-predi$pred <- predict.glm(modi1, newdata = predi, type = 'response')
-
-u <- cd01Mod %>% distinct(age, season, ageInSamples) %>% arrange(ageInSamples) %>% filter(!(age==2 & ageInSamples == 6))
-predi <- left_join(predi, u)
-
-ggplot(predi, aes(observedLengthStd, pred, color = (cohortStd), group = (cohortStd))) +
-  geom_point() +
-  scale_color_viridis_c() +
-  scale_y_continuous("Probability of first observation") +
-  facet_grid(species ~ ageInSamples, scales = 'free')
-
-predi$seasonGG <- factor(predi$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
-predi$cohort <- predi$cohortStd + 2007
-predi$speciesGG <- factor(predi$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
-
-ggplot(predi %>% filter(cohortStd %in% c(0), species == 'bkt', age < 3), aes(observedLengthStd, pred)) +
-  geom_point() +
-  scale_color_viridis_c() +
-  scale_x_continuous("Standardized length") +
-  scale_y_continuous(lim = c(0,1),"Probability of first observation") +
-  theme_publication() +
-  facet_grid(seasonGG ~ age)
-
-ggplot(predi %>% filter(cohortStd %in% c(0), age < 3), aes(observedLengthStd, pred, color = speciesGG)) +
-  geom_point(size = 2) +
-  scale_x_continuous("Standardized length") +
-  scale_y_continuous(lim = c(0,1),"Probability of first observation") +
-  theme_publication(base_size = 28) +
-  theme(legend.title = element_text("Species"),
-        legend.position = "bottom",
-        legend.direction = "horizontal"
-        ) + 
-  labs(color='Species') +
-  scale_color_grey(start = 0.15, end = 0.75) +
-  facet_grid(seasonGG ~ age)
-
-ggsave('./figures/pFirstObs.png', width = 11, height = 11, units='in')
-```
-
 ```{r gam models - find best lm, before fitting detailed gams - only need to run for model selection}
@@
 #   rowname       df      AIC
-# 1    mod6 14.99337 429769.2
-# 2    mod2 25.99569 432704.3
-# 3    mod4 13.98558 433764.8
+# 1    mod6 14.82321 197082.8
+# 2    mod2 25.99790 197663.0
+# 3    mod4 13.96042 198350.4
 
@@
 #   rowname       df       AIC
-# 1    mod6 12.99685 238636.2
-# 2    mod2 17.99755 242840.1
-# 3    mod4 11.98667 243594.3
+# 1    mod6 12.29048  98814.72
+# 2    mod2 17.99609  99877.98
+# 3    mod4 11.96329 100212.22
 
@@
 
-```{r models with no fixed season effects or categories for splines}
-
-system.time(
-mod6_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod6_YDCcr_Cr_CSr_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod6_YDCcr_Cr_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod6_YDCcr_C_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod6_YDC_C_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(cohort, bs = 'cr', k = kIn) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
-               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
-               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
-               data = dataIn, 
-               method = "REML")
-)
-
-AIC(mod6_YDC_C_CS_T_F_TF,mod6_YDCcr_C_CS_T_F_TF,mod6_YDCcr_Cr_CS_T_F_TF,mod6_YDCcr_Cr_CSr_T_F_TF,mod6_YDCcr_Cr_CSr_Tr_Fr_TFr)  %>% rownames_to_column() %>% arrange(AIC)
-
-summary(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr)
-plot(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr, pages = 8, shade = TRUE, scale = 0)
-
-#BKT
-#                       rowname        df      AIC
-# 1 mod6_YDCcr_Cr_CSr_Tr_Fr_TFr 194.27159 402517.3
-# 2    mod6_YDCcr_Cr_CSr_T_F_TF 178.16547 402696.5
-# 3     mod6_YDCcr_Cr_CS_T_F_TF 170.29804 403228.7
-# 4      mod6_YDCcr_C_CS_T_F_TF 166.30172 404261.2
-# 5        mod6_YDC_C_CS_T_F_TF  29.89126 412576.2
-
-#BNT
-#                       rowname        df      AIC
-# 1 mod6_YDCcr_Cr_CSr_Tr_Fr_TFr 102.82522 225950.4
-# 2    mod6_YDCcr_Cr_CSr_T_F_TF  98.12051 225974.0
-# 3     mod6_YDCcr_Cr_CS_T_F_TF  96.51127 225994.7
-# 4      mod6_YDCcr_C_CS_T_F_TF  92.93128 226010.4
-# 5        mod6_YDC_C_CS_T_F_TF  27.97005 231079.5
-
-```
-
-```{r models without T and F}
-# Not sure how to get models with season-specific T and F effects to work within season without messing up predictions. # Leave T and F out and focus on densities
-
-mod6_YDCcr_Cr_CSr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor),  # this is conspecific cS only
-               data = dataIn, 
-               method = "REML")
-
-system.time(
-mod6_YDCcr_Cr_CS3r <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+```{r remove cohort, focus on spp-specific cohort strengths}
+#sppoke with keith. decided focus on cohrot is less interesting than river effects and density effects
+
+mod1 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
                s(ydayCumul, bs = 'cr') +
@@
                s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
-               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor) +
-               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor),  
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod6_YDCcr_Cr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
+  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+summary(modCS3)
+#plot(modCS3, pages = 8)
+
+mod2 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
+  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+mod3 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor),
+              # s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
+  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+mod4 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
                s(cohort, bs = 'cr', k = kIn, by = riverFactor),
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod6_YDCcr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor),  
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod6_YDC <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr'),  
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod6_YDCcr_Cr_CSBKTr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor),  
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod6_YDCcr_Cr_CSBNTr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor),  
-               data = dataIn, 
-               method = "REML")
-)
-
-system.time(
-mod6_YDCcr_Cr_CSATSr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
-               s(ydayCumul, bs = 'cr') +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
-               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
-               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
-               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor),  
-               data = dataIn, 
-               method = "REML")
-)
-
-AIC(mod6_YDCcr_Cr_CSr, mod6_YDCcr_Cr_CS3r, mod6_YDCcr_Cr_CSBKTr, mod6_YDCcr_Cr_CSBNTr, mod6_YDCcr_Cr_CSATSr, mod6_YDCcr_Cr,mod6_YDCcr,mod6_YDC)  %>% rownames_to_column() %>% arrange(AIC)
-
-summary(mod6_YDCcr_Cr_CSr)
-plot(mod6_YDCcr_Cr_CS3r, pages = 7, shade = TRUE, scale = 0)
-
-# bkt
-#                rowname        df      AIC
-# 1    mod6_YDCcr_Cr_CSr 173.85954 413790.2
-# 2   mod6_YDCcr_Cr_CS3r 188.68933 415278.2
-# 3 mod6_YDCcr_Cr_CSBKTr 167.72761 416582.2
-# 4 mod6_YDCcr_Cr_CSBNTr 168.43917 417098.5
-# 5 mod6_YDCcr_Cr_CSATSr 167.84300 417569.6
-# 6        mod6_YDCcr_Cr 160.60559 417764.7
-# 7           mod6_YDCcr 149.80639 420935.0
-# 8             mod6_YDC  14.99337 429769.2
-
-# bnt
-#                rowname       df      AIC
-# 1   mod6_YDCcr_Cr_CS3r 96.91663 231844.1
-# 2    mod6_YDCcr_Cr_CSr 84.99160 231875.4
-# 3 mod6_YDCcr_Cr_CSBKTr 84.39999 232019.6
-# 4 mod6_YDCcr_Cr_CSBNTr 87.25717 232120.2
-# 5 mod6_YDCcr_Cr_CSATSr 85.68893 232203.3
-# 6        mod6_YDCcr_Cr 81.27554 232235.7
-# 7           mod6_YDCcr 75.00492 233885.5
-# 8             mod6_YDC 12.99685 238636.2
-```
-
-```{r predictions}
-
-ydayCumulSeason <- distinct(dataIn, ydayCumul,season,age) %>% arrange(ydayCumul) %>% mutate(age_season = paste(age,season,sep = '_')) %>%
-  group_by(age_season) %>% summarise(age=min(age),season=min(season),minY=min(ydayCumul),maxY=max(ydayCumul))
-#38 obear fish were captured in january, but get season 4 and are age 1, so have very small ydayCumul -need to manually reset 
-#dataIn %>% filter(age==1,season==4,ydayCumul==367) %>% select(tag,cohort,detectionDate,year,ydayCumul,yday,river,age,season)
-ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '1_4',670,ydayCumulSeason$minY)
-ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '2_4',1035,ydayCumulSeason$minY)
-ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '0_3',0,ydayCumulSeason$minY)
-ydayCumulSeason %>% mutate(minY2 = lead(minY))
-
-yDayAll <- data.frame(ydayCumul = 250:1000)
-yDayAll$season <- NA
-
-# fill in season for each day of the year (yDayAll) based on ydayCumulSeason min values
-# need this to get season into all ydayCumul in predGridBKT
-for(i in 1:nrow(yDayAll)) {
-  for(j in 1:(nrow(ydayCumulSeason) - 1)) {
-    if (between(yDayAll$ydayCumul[i], ydayCumulSeason$minY[j], ydayCumulSeason$minY[j + 1])) {
-      yDayAll$season[i] <- ydayCumulSeason$season[j]
-    }
-  }
-}
-
-predGrid <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
-                        cohortFactor = unique(dataIn$cohort),
-                        ydayCumul = seq(250,1000, by = 10),
-                        salmon01 = c(0,1),
-                        cohortStrengthScaled = seq(-2.5,2.5, by = 0.25)) %>%
-                    #left_join(., yDayAll) %>%
-                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
-                           cohort = cohortFactor)
-
-pred <- predict(mod6_YDCcr_Cr_CSr, newdata = predGrid)#, type = 'lpmatrix')
-
-predGrid2 <- bind_cols(predGrid, as.data.frame(pred)) %>% mutate(pred = pred)
-
-ggplot(predGrid2 %>% filter(salmon01 == 0, cohortStrengthScaled == 0),
-
-       aes(ydayCumul, pred, group = cohort, color = cohort)) +
-  #geom_point() + 
-  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
-  geom_line(size = 1.1) +
-  labs(y = 'Predicted body size (mm)',
-       x = 'Days since age-0 Jan 1',
-       color = 'Cohort') +
-  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
-  theme_publication() +
-  theme(legend.position = "right", 
-        legend.direction = "vertical",
-        legend.title = element_text(face = "plain")) +
-  guides(color = guide_colorbar(barheight = 10,
-                                barwidth = 1.5,
-                                ticks = FALSE,
-                                nbin = 100)) +
-#  guide_legend(title = 'Cohort') +
-  facet_wrap(~riverFactor)#, scales = 'free_y')
-
+              # s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+              # s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
+  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+AIC(mod1,mod2,mod3,mod4) %>% rownames_to_column() %>% arrange(AIC)
 ```
@@
 
-#using salmon01 * river because we need diff effects per river - even though AIC is worse for this model
 modBKT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
@@
                s(cohort, bs = 'cr', k = kIn, by = riverFactor) +
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor), 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor),
+
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
@@
                            salmon01 = c(0,1),
-                           cohortStrengthScaled = seq(-2.5,2.5, by = 0.25)) %>%
+                           cohortStrengthScaledBKT = seq(-2.5,2.5, by = 0.25),
+                           cohortStrengthScaledBNT2p5 = seq(-2.5,2.5, by = 0.25)) %>%
                     #left_join(., yDayAll) %>%
                     mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
-                           cohort = cohortFactor)
+                           cohort = cohortFactor) %>%
+                    filter(!(cohort < 2002 & riverFactor %in% c('wb jimmy','wb mitchell','wb obear')))
 
@@
 # using CSr model becasue delta AIC is not huge btw it and CS3r. Better to have same models for both species
+
 modBNT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
@@
                s(cohort, bs = 'cr', k = kIn, by = riverFactor) +
-               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor), 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor),
+
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")
@@
 summary(modBNT)
-#plot(modBNT,  scale = 0)
+#plot(modBNT,  pages = 8,scale = 0)
 
@@
 
+predGrid2Both <- predGrid2Both %>% filter(!(cohort < 2002 & riverFactor %in% c('wb jimmy','wb mitchell','wb obear')))
+
 save(modBKT,modBNT,predBKT,predBNT,predGrid2Both, file = paste0('./data/out/bestModsBKTBNT.RData'))
@@
 ggplot(predGrid2Both %>% 
-         filter(salmon01 ==0, cohortStrengthScaled == 0
+         filter(salmon01 ==0, cohortStrengthScaledBKT == 0, cohortStrengthScaledBNT2p5 == 0
 #         (cohortStrengthScaledBKT == 0 & salmon01 == 0) | (cohortStrengthScaledBKT == 0 & cohortStrengthScaledBNT == 0 & cohortStrengthScaledATS == 0 & metaPop01 == 1 & salmon01 ==0)
@@
 ggsave('./figures/cohortEffect.png', width = 11, height = 8, units='in')
-(plotCohortStrength <- plotGamVarBySpecies(var = 's(cohortStrengthScaled)', xAxisLab = 'Cohort Strength'))
-ggsave('./figures/cohortStrengthEffect.png', width = 11, height = 8, units='in')
+
+(plotCohortStrength <- plotGamVarBySpecies(var = 's(cohortStrengthScaledBKT)', xAxisLab = 'Brook Trout cohort Strength'))
+ggsave('./figures/cohortStrengthEffectBKT.png', width = 11, height = 8, units='in')
+
+(plotCohortStrengthScaledBNT2p5 <- plotGamVarBySpecies(var = 's(cohortStrengthScaledBNT2p5)', xAxisLab = 'Brown Trout cohort Strength'))
+ggsave('./figures/cohortStrengthEffectBNT.png', width = 11, height = 8, units='in')
+
 #
-plot.gam(modBKT, pages = 10)
+#plot.gam(modBKT, pages = 10)
 
@@
 ggplot(predGrid2Both %>% 
-         filter(salmon01 ==0, cohortStrengthScaled == 0, cohort == 2006, cohortStrengthScaled == 0),
+         filter(salmon01 == 0, cohortStrengthScaledBKT == 0, cohort == 2006, cohortStrengthScaledBNT2p5 == 0),
        aes(ydayCumul, pred, group = speciesGG, color = speciesGG)) +
@@
 ggplot(predGrid2Both %>% 
-         filter(salmon01 == 0, cohortStrengthScaled == 0, cohort == 2006, cohortStrengthScaled == 0),
+         filter(salmon01 == 0, cohortStrengthScaledBKT == 0, cohort == 2006, cohortStrengthScaledBNT2p5 == 0),
        aes(ydayCumul, pred, group = riverGG, color = riverGG)) +
@@
         legend.title = element_text(face = "plain")) +
-  facet_grid(~ speciesGG)
-
-ggsave('./figures/ydayCumulAcrossSppByRiver.png', width = 11, height = 8, units='in')
+  facet_grid(rows = vars(speciesGG))
+
+ggsave('./figures/ydayCumulAcrossSppByRiver.png', width = 8, height = 8, units='in')
 
@@
 
-# cohort effects
-
-ggplot(tmp2, aes(x,y)) + geom_line()
+# river effects 
+# BKT
+plotBKTAll <- plot.gam(modBKT, page = 9)
+plotBKTAll[[16]]$ylab # WB
+
+plotBKTByRiver <- data.frame()
+bktRiver1 <- 16
+for(i in 1:4){
+  ii <- i + bktRiver1 - 1
+  tmp <- data.frame(x = plotBKTAll[[ii]]$x, y = plotBKTAll[[ii]]$fit, se = plotBKTAll[[i]]$se, ylab = plotBKTAll[[ii]]$ylab, riverIndex = i) %>% mutate(species = 'bkt', upperCI = y + se * 1.96, lowerCI = y - se * 1.96)
+  plotBKTByRiver <- rbind(plotBKTByRiver, tmp)
+}
+
+#BNT
+plotBNTAll <- plot.gam(modBNT, page = 6)
+plotBNTAll[[16]]$ylab # WB
+
+plotBNTByRiver <- data.frame()
+bntRiver1 <- 16
+for(i in 1:2){
+  ii <- i + bntRiver1 - 1
+  tmp <- data.frame(x = plotBNTAll[[ii]]$x, y = plotBNTAll[[ii]]$fit, se = plotBNTAll[[i]]$se, ylab = plotBNTAll[[ii]]$ylab, riverIndex = i) %>% mutate(species = 'bnt', upperCI = y + se * 1.96, lowerCI = y - se * 1.96)
+  plotBNTByRiver <- rbind(plotBNTByRiver, tmp)
+}
+
+plotByRiver <- rbind(plotBKTByRiver, plotBNTByRiver)
+
+plotByRiver$speciesGG <- factor(plotByRiver$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+plotByRiver$riverGG <- factor(plotByRiver$riverIndex,
+                         levels=c(1,2,3,4),
+                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))
+
+ggplot(plotByRiver, aes(x)) +
+  geom_line(aes(y = y, color = speciesGG), size = 1.5) +
+  geom_ribbon(aes(ymin = lowerCI, ymax = upperCI, fill = speciesGG), alpha = 0.5, show.legend = FALSE) +
+  scale_color_manual(values = c('grey30', 'grey70')) +
+  scale_fill_manual(values = c('grey30', 'grey70')) +
+  labs(y = 'Predicted body size (mm)',
+       x = 'Days since age-0 Jan 1',
+       color = 'Species') +
+  theme_publication() +
+  theme(legend.position = "right", 
+        legend.direction = "vertical",
+        legend.title = element_text(face="plain")) +
+  facet_wrap(~ riverGG)
+
+ggsave('./figures/ydayCumulAcrossRiversBySpecies.png', width = 11, height = 8, units='in')
 
 appraise(modBKT)
+
+```
+
+```{r mean lengths: from MS_Analysis_3spp_WB.Rmd. For Grenada talk}
+
+# Has mean size changed over time?
+means <- 
+  cd %>% mutate( age = year - cohort ) %>%
+    filter( riverOrdered == 'west brook', ageInSamples %in% 1:10, age < 3, year < 2016 ) %>%
+    group_by(species,ageInSamples,year,season,age) %>%
+    summarize(meanLen = mean(observedLength,na.rm = TRUE),
+              sdLen = sd(observedLength,na.rm = TRUE),
+              meanGR = mean(grLength,na.rm = TRUE),
+              sdGR = sd(grLength,na.rm = TRUE)
+              )
+
+lenSlopes <- means %>%
+  group_by( species, ageInSamples, season, age ) %>%
+  nest()
+
+lengthModel <- function(df){
+  lm( meanLen ~ year, data = df )
+}
+
+grModel <- function(df){
+  lm( meanGR ~ year, data = df )
+}
+
+lenSlopes <- lenSlopes %>%
+  mutate( model = map(lenSlopes$data, lengthModel),
+          modelGR = map(lenSlopes$data, grModel))
+
+###################################
+######### len here, gr below ######
+lenSlopesTidy <- lenSlopes %>%
+  mutate(glance = map(model, broom::tidy)) %>% 
+  unnest(glance, .drop = TRUE) %>%
+  filter(term == 'year') %>%
+  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
+  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
+  mutate( sig = ifelse(p.value < 0.0001,"***",sig))
+
+means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+lenSlopesTidy$seasonGG <- factor(lenSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+lenSlopesTidy$speciesGG <- factor(lenSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanLen, group= speciesGG)) +
+  geom_point( aes(color = speciesGG), size = 4) +
+  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
+  labs( x = "Year", y = "Mean body length (mm)" ) +
+
+  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age<2), 
+            aes(x=2000,y=225, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age<2), 
+            aes(x=2000,y=195, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+
+  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), 
+            aes(x=2011.8,y=104, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), 
+            aes(x=2011.8,y=74, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+
+
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+      #  legend.position = "right",
+     #   legend.direction = "vertical"
+        legend.position = "bottom",
+        legend.direction = "horizontal") + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG ~ age, scales='free')
+
+ggsave('./figures/meanSizes.png', width = 11, height = 11, units='in')
+
+# one season/year only
+ggplot( means %>% filter(year >= 2000, species != 'ats', ageInSamples == 8), aes(year,meanLen, group= speciesGG)) +
+  geom_point( aes(color = speciesGG), size = 4) +
+  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
+  labs( x = "Year", y = "Mean body length (mm)" ) +
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+      #  legend.position = "right",
+     #   legend.direction = "vertical"
+                legend.position = "bottom",
+        legend.direction = "horizontal") + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG ~ age)
+ggsave('./figures/meanSizesAIS_1.png', width = 11, height = 8, units='in')
+
+# is a model with salmon presence absence better than a year model?
+means$ats01 <- ifelse(means$year <= 2005, 1, 0)
+
+modMeans1 = lm(meanLen ~ year * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
+modMeans2 = lm(meanLen ~ year * factor(ats01) * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
+modMeans3 = lm(meanLen ~ factor(ats01) * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
+modMeans4 = lm(meanLen ~ factor(ats01)*year, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
+AIC(modMeans1, modMeans2, modMeans3, modMeans4)
+
+# one season/year only
+ggplot( means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1), aes(year,meanLen, group= speciesGG)) +
+  geom_point( aes(color = speciesGG), size = 4) +
+  geom_smooth( aes(group = factor(ats01)), color = 'black', se = FALSE, method = 'lm') +
+  labs( x = "Year", y = "Mean body length (mm)" ) +
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+      #  legend.position = "right",
+     #   legend.direction = "vertical"
+        legend.position = "bottom",
+        legend.direction = "horizontal") + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG ~ age)
+
+ggsave('./figures/meanSizesAIS_1_ats01.png', width = 11, height = 8, units='in')
+
+# mean increases 
+lenSlopesTidy %>%
+  filter(species != 'ats') %>%
+  summarise( means = mean(estimate),
+             sd = sd(estimate))
+
+# mean increases by species
+lenSlopesTidy %>%
+  group_by(species) %>%
+  summarise( means = mean(estimate),
+             sd = sd(estimate))
+
+# do increases (slopes) increase over age? Yes
+ggplot(lenSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
+  geom_point(size = 5) +
+  geom_smooth(method='lm', se=FALSE, size = 1.5) +
+  scale_x_continuous(breaks = 1:10, labels = c('0 Fall','0 Winter','1 Spring','1 Summer','1 Fall','1 Winter','2 Spring','2 Summer','2 Fall','2 Winter')) +
+  labs( x = "Age Season", y = "Slope of mean body length over years" ) +
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+        legend.position = "bottom",
+        legend.direction = "horizontal",
+        axis.text.x = element_text(angle = 40, hjust = 1)) + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75)
+
+ggsave(paste0('figures/meanLengthsOverAge.png'), width = 11, height = 10, units='in')
+
+# species diff in increases in length?
+sppEffect <- (lm(estimate ~ ageInSamples * (species), data = lenSlopesTidy))
+summary(sppEffect)
+anova(sppEffect)
+
+# mean diff in slopes btw bkt and bnt
+lenSlopesTidy %>%
+  filter(species != 'ats') %>%
+  group_by(species) %>%
+  summarise(means = mean(estimate))
+
+#####################################  
+#################################
+# growth rate
+grSlopesTidy <- lenSlopes %>%
+  mutate(glance = map(modelGR, broom::tidy)) %>% 
+  unnest(glance, .drop = TRUE) %>%
+  filter(term == 'year') %>%
+  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
+  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
+  mutate( sig = ifelse(p.value < 0.0001,"***",sig))
+
+##
+means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanGR, group= speciesGG)) +
+  geom_point( aes(color = speciesGG), size = 4) +
+  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
+  labs( x = "Year", y = "Mean growth rate (mm/d)" ) +
+  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age<2), 
+            aes(x=2000,y=0.68, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age<2), 
+            aes(x=2000,y=0.55, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+
+  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), 
+            aes(x=2011.5,y=0.64, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
+  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), 
+            aes(x=2011.5,y=0.51, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+      #  legend.position = "right",
+     #   legend.direction = "vertical"
+        legend.position = "bottom",
+        legend.direction = "horizontal") + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG~age)
+
+ggsave('./figures/meanGRs.png', width = 11, height = 11, units='in')
+
+# mean increases 
+grSlopesTidy %>%
+  filter(species != 'ats') %>%
+  summarise( means = mean(estimate),
+             sd = sd(estimate))
+
+# mean increases by species
+grSlopesTidy %>%
+  group_by(species) %>%
+  summarise( means = mean(estimate),
+             sd = sd(estimate))
+
+# do increases (slopes) increase over age? Yes
+ggplot(grSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
+  geom_point(size = 5) +
+  geom_smooth(method='lm', se=FALSE) +  
+  geom_hline(yintercept = 0) +
+  scale_x_continuous(breaks = 1:10, labels = c('0 Fall','0 Winter','1 Spring','1 Summer','1 Fall','1 Winter','2 Spring','2 Summer','2 Fall','2 Winter')) +
+  labs( x = "Age Season", y = "Slope of mean growth rate over years" ) +
+  theme_publication(base_size = 24) +
+  theme(legend.title = element_text("Species"),
+        legend.position = "bottom",
+        legend.direction = "horizontal",
+        axis.text.x = element_text(angle = 40, hjust = 1)) + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75)
+
+ggsave('./figures/meanGRsOverAge.png', width = 11, height = 11, units='in')
+
+summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bkt')))
+summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bnt')))
+
+# species diff in increases in gr?
+sppEffectGR <- (lm(estimate ~ ageInSamples * (species), data = grSlopesTidy))
+summary(sppEffectGR)
+anova(sppEffectGR)
+
+# mean diff in slopes btw bkt and bnt
+grSlopesTidy %>%
+  filter(species != 'ats') %>%
+  group_by(species) %>%
+  summarise(means = mean(estimate))
+
+```
+
+```{r size-dep imm or mort? For Spain talk}
+
+### for losses
+
+cd01 <-  cd %>%
+  filter(knownZ == 1) %>%
+  group_by(tagIndex) %>%
+  mutate(first = min(sampleNumber),
+         firstObs = sampleNumber == first,
+         last = max(sampleNumber),
+         lastObs = sampleNumber == last,
+         notLastObs = !lastObs,
+         nObs = n(),
+         age = year - cohort
+  ) %>%
+  ungroup()
+
+meansByAIS <- cd01 %>%
+  group_by(species, ageInSamples, cohort, river) %>%
+  summarise(meanLen = mean(observedLength, na.rm =T),
+            sdLen = sd(observedLength, na.rm = T),
+            n = n())
+cd01 <- left_join(cd01, meansByAIS) %>%
+  mutate(observedLengthStd = (observedLength - meanLen) / sdLen,
+         cohortStd = cohort - 2007)
+
+#cd %>% select(tagIndex,sampleNumber, firstObs,lastObs, knownZ, observedLength) %>% head(50) %>% as.data.frame()
+cd01Mod <- cd01 %>% filter(species %in% c('bkt', 'bnt'), river == 'west brook', ageInSamples %in% 1:10, cohort %in% 2000:2014)
+
+ggplot(cd01Mod, aes(observedLengthStd, notLastObs*1, color = factor(species), group = factor(species))) +
+  geom_point(alpha = 0.01) +
+  geom_smooth(method = "glm", 
+    method.args = list(family = "binomial"), 
+    se = FALSE) +
+ # scale_color_viridis_c() +
+  facet_wrap( ~ cohort, scales = 'free')
+
+#### by species and ageInSamples
+
+sensSlopes <- cd01Mod %>%
+  group_by( species, ageInSamples ) %>%
+  nest()
+
+survModel <- function(df){
+  glm(notLastObs*1 ~ observedLengthStd + cohortStd, family = 'binomial', data = df )
+}
+enterModel <- function(df){
+  glm(firstObs*1 ~ observedLengthStd + cohortStd, family = 'binomial', data = df )
+}
+
+sensSlopes <- sensSlopes %>%
+  mutate(modelSurv = map(sensSlopes$data, survModel),
+         modelEnter = map(sensSlopes$data, enterModel))
+
+sensSlopesTidy <- sensSlopes %>%
+  mutate(surv = map(modelSurv, broom::tidy),
+         enter = map(modelEnter, broom::tidy)) %>% 
+  unnest(surv, enter, .drop = TRUE) %>%
+  filter(term == 'cohortStd') %>% # cohort effect
+  rename(estimateSurv = estimate, estimateEnter = estimate1) %>%
+  select(species,ageInSamples,estimateSurv, estimateEnter) #%>%
+  #gather("type", "estimate", estimateSurv:estimateEnter)
+
+ggplot(sensSlopesTidy, aes(ageInSamples, estimate, color = type)) +
+  geom_point() +
+  geom_line() +
+  facet_wrap(~species)
+
+```
+
+```{r dummy graphs for Spain talk}
+
+sizesForGraph <- data.frame(year = c(2000,2015), surv = c(100, 150))
+ggplot(sizesForGraph, aes(year,surv)) +
+  geom_line(size = 2) +
+  labs(x = 'Year', y = 'Mean body length (mm)') +
+  ylim(80,200) +
+  theme_publication(base_size = 36)
+ggsave('./figures/sizesDummy.png', width = 11, height = 11, units='in')
+
+survForGraph <- data.frame(size = c(-2,2), surv = c(0.25,0.75))
+ggplot(survForGraph, aes(size,surv)) +
+  geom_line(size = 2) +
+  labs(x = 'Standardized length', y = 'Probability of seen again') +
+  theme_publication(base_size = 36)
+ggsave('./figures/pSurvivalDummy.png', width = 11, height = 11, units='in')
+
+firstForGraph <- data.frame(size = c(-2,2), surv = c(0.35,0.65))
+ggplot(firstForGraph, aes(size,surv)) +
+  geom_line(size = 2) +
+  labs(x = 'Standardized length', y = 'Probability of first observation') +
+  theme_publication(base_size = 36)
+ggsave('./figures/pFirstDummy.png', width = 11, height = 11, units='in')
+
+sizesForGraph <- data.frame(year = c(2000,2015), surv = c(100, 150))
+ggplot(sizesForGraph, aes(year,surv)) +
+  geom_line(size = 3) +
+  labs(x = 'Year', y = 'Mean body length (mm)') +
+  ylim(80,200) +
+  theme_publication(base_size = 36)
+ggsave('./figures/sizesDummy.png', width = 11, height = 11, units='in')
+
+#### show interactions
+
+survForGraph <- data.frame(ind = c(1,1,2,2,3,3), size = c(-2,2), surv = c(0.52,0.55,0.51,0.24,0.54,0.82))
+ggplot(survForGraph, aes(size,surv, group = ind)) +
+  geom_line(size = 2) +
+  labs(x = 'Standardized length', y = 'Probability of seen again') +
+  theme_publication(base_size = 36)
+ggsave('./figures/pSurvivalDummy3.png', width = 11, height = 11, units='in')
+
+firstForGraph <- data.frame(ind = c(1,1,2,2,3,3), size = c(-2,2), surv = c(0.42,0.45,0.41,0.14,0.44,0.72))
+ggplot(firstForGraph, aes(size,surv, group = ind)) +
+  geom_line(size = 2) +
+  labs(x = 'Standardized length', y = 'Probability of first observation') +
+  theme_publication(base_size = 36)
+ggsave('./figures/pFirstDummy3.png', width = 11, height = 11, units='in')
+
+```
+
+```{r get betas from growth model for sensitivities}
+load("D:/projects/linkedModels/data/out/allBetas.RData") #loads df 'q'
+qFig1 <- q %>% filter(riverGG == 'WB', species %in% c('bkt','bnt'))
+qFig1$season <- as.numeric(as.character(qFig1$season))
+qFig1$isYOY <- as.numeric(as.character(qFig1$isYOY))
+qFig <- qFig1 %>%
+  mutate( varText = tolower(betaGG)) %>%
+  mutate( varText = ifelse(varText == 'temp*flow','temp:flow', varText)) %>%
+  mutate( varText = ifelse(varText == 'bnt*bkt','bkt:bnt', varText)) %>%
+  select(isYOY,season,species,median,varText)
+qFig$seasonGG <- factor(qFig$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+#get slopes of env variables
+
+load('D:/projects/linkedModels/data/out/flowSlopesTidy.RData')
+flowSlopesTidy2 <- flowSlopesTidy %>% filter(riverOrderedGG == 'West Brook') %>% select(estimate, seasonGG) %>% mutate(varText = 'flow')
+load('D:/projects/linkedModels/data/out/tempSlopesTidy.RData')
+tempSlopesTidy2 <- tempSlopesTidy %>% filter(riverOrderedGG == 'West Brook') %>% select(estimate, seasonGG) %>% mutate(varText = 'temp')
+load('D:/projects/linkedModels/data/out/nSlopesTidy.RData')
+nSlopesTidy2 <- nSlopesTidy %>% select(estimate, season, species) %>% mutate(varText = species)
+nSlopesTidy2$seasonGG <- factor(nSlopesTidy2$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+nSlopesTidy2 <- nSlopesTidy2 %>% select(estimate,seasonGG,varText)
+
+allSlopesTidy <- bind_rows(nSlopesTidy2, tempSlopesTidy2, flowSlopesTidy2)
+
+allSlopesTidyWBetas <- left_join(qFig, allSlopesTidy, by = c('seasonGG', 'varText')) %>% mutate(sens = median * estimate)
+
+ggplot(allSlopesTidyWBetas%>%filter(!is.na(sens), varText != 'ats'), aes(seasonGG, sens, color = varText)) +
+  geom_point() +
+  facet_grid(species + isYOY ~ varText, scales = 'free')
+```
+
+```{r }
+############# 1st attempt at imm em
+mod1 <- glm(notLastObs*1 ~ observedLengthStd * cohortStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+#mod2 <- glm(notLastObs*1 ~ observedLengthStd * factor(cohortStd) * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+mod3 <- glm(notLastObs*1 ~ observedLengthStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+
+AIC(mod1,mod3)
+
+summary(mod1)
+anova(mod1, test="Chi")
+
+pred <- expand.grid(observedLengthStd = seq(-2,2, length.out = 50), cohortStd = unique(cd01Mod$cohortStd), species = c('bkt', 'bnt'), ageInSamples = unique(cd01Mod$ageInSamples))
+
+pred$pred <- predict.glm(mod1, newdata = pred, type = 'response')
+
+u <- cd01Mod %>% distinct(age, season, ageInSamples) %>% arrange(ageInSamples) %>% filter(!(age==2 & ageInSamples == 6))
+pred <- left_join(pred, u)
+
+pred$seasonGG <- factor(pred$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+pred$cohort <- pred$cohortStd + 2007
+pred$speciesGG <- factor(pred$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot(pred, aes(observedLengthStd, pred, color = (cohortStd), group = (cohortStd))) +
+  geom_point() +
+  scale_color_viridis_c() +
+  scale_y_continuous("Probability of survival") +
+  facet_grid(species ~ ageInSamples, scales = 'free')
+
+ggplot(pred %>% filter(cohortStd %in% 0, age < 3), aes(observedLengthStd, pred, color = speciesGG)) +
+  geom_point(size = 2) +
+  scale_x_continuous("Standardized length") +
+  scale_y_continuous(lim = c(0,1),"Probability of seen again") +
+  theme_publication(base_size = 28) +
+  theme(legend.title = element_text("Species"),
+        legend.position = "bottom",
+        legend.direction = "horizontal"
+        ) + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG ~ age)
+
+ggsave('./figures/pSurvival.png', width = 11, height = 11, units='in')
+
+################################
+# gains - immigration, firstObs
+
+ggplot(cd01Mod, aes(observedLengthStd, firstObs*1, color = factor(species), group = factor(species))) +
+  geom_jitter(alpha = 0.1) +
+  geom_smooth(method = "glm", 
+    method.args = list(family = "binomial"), 
+    se = FALSE) +
+ # scale_color_viridis_c() +
+  xlim(c(-2.75,2.75)) +
+  facet_grid(ageInSamples ~ cohort)
+
+ggplot(cd01Mod%>%filter(ageInSamples == 3), aes(observedLengthStd, firstObs*1, color = factor(species), group = factor(species))) +
+  geom_jitter(alpha = 0.1) +
+  geom_smooth(method = "glm", 
+    method.args = list(family = "binomial"), 
+    se = FALSE) +
+ # scale_color_viridis_c() +
+  xlim(c(-2.75,2.75)) +
+  facet_wrap( ~ cohort)
+
+modi1 <- glm(firstObs*1 ~ observedLengthStd * cohortStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+modi2 <- glm(firstObs*1 ~ observedLengthStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+modi3 <- glm(firstObs*1 ~ observedLengthStd * species , family = 'binomial', data = cd01Mod)
+modi4 <- glm(firstObs*1 ~ observedLengthStd * factor(ageInSamples), family = 'binomial', data = cd01Mod)
+modi5 <- glm(firstObs*1 ~ observedLengthStd, family = 'binomial', data = cd01Mod)
+
+AIC(modi1,modi2,modi3,modi4,modi5) %>% rownames_to_column() %>% arrange(AIC)
+
+summary(modi1)
+anova(modi1, test="Chi")
+###
+predi <- expand.grid(observedLengthStd = seq(-2,2, length.out = 50), cohortStd = unique(cd01Mod$cohortStd), species = c('bkt', 'bnt'), ageInSamples = unique(cd01Mod$ageInSamples))
+
+predi$pred <- predict.glm(modi1, newdata = predi, type = 'response')
+
+u <- cd01Mod %>% distinct(age, season, ageInSamples) %>% arrange(ageInSamples) %>% filter(!(age==2 & ageInSamples == 6))
+predi <- left_join(predi, u)
+
+ggplot(predi, aes(observedLengthStd, pred, color = (cohortStd), group = (cohortStd))) +
+  geom_point() +
+  scale_color_viridis_c() +
+  scale_y_continuous("Probability of first observation") +
+  facet_grid(species ~ ageInSamples, scales = 'free')
+
+predi$seasonGG <- factor(predi$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+predi$cohort <- predi$cohortStd + 2007
+predi$speciesGG <- factor(predi$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+ggplot(predi %>% filter(cohortStd %in% c(0), species == 'bkt', age < 3), aes(observedLengthStd, pred)) +
+  geom_point() +
+  scale_color_viridis_c() +
+  scale_x_continuous("Standardized length") +
+  scale_y_continuous(lim = c(0,1),"Probability of first observation") +
+  theme_publication() +
+  facet_grid(seasonGG ~ age)
+
+ggplot(predi %>% filter(cohortStd %in% c(0), age < 3), aes(observedLengthStd, pred, color = speciesGG)) +
+  geom_point(size = 2) +
+  scale_x_continuous("Standardized length") +
+  scale_y_continuous(lim = c(0,1),"Probability of first observation") +
+  theme_publication(base_size = 28) +
+  theme(legend.title = element_text("Species"),
+        legend.position = "bottom",
+        legend.direction = "horizontal"
+        ) + 
+  labs(color='Species') +
+  scale_color_grey(start = 0.15, end = 0.75) +
+  facet_grid(seasonGG ~ age)
+
+ggsave('./figures/pFirstObs.png', width = 11, height = 11, units='in')
+```
+
+```{r models with no fixed season effects or categories for splines}
+
+system.time(
+mod6_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr_CSr_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_C_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDC_C_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(cohort, bs = 'cr', k = kIn) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
+               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
+               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
+               data = dataIn, 
+               method = "REML")
+)
+
+AIC(mod6_YDC_C_CS_T_F_TF,mod6_YDCcr_C_CS_T_F_TF,mod6_YDCcr_Cr_CS_T_F_TF,mod6_YDCcr_Cr_CSr_T_F_TF,mod6_YDCcr_Cr_CSr_Tr_Fr_TFr)  %>% rownames_to_column() %>% arrange(AIC)
+
+summary(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr)
+plot(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr, pages = 8, shade = TRUE, scale = 0)
+
+#BKT
+#                       rowname        df      AIC
+# 1 mod6_YDCcr_Cr_CSr_Tr_Fr_TFr 194.27159 402517.3
+# 2    mod6_YDCcr_Cr_CSr_T_F_TF 178.16547 402696.5
+# 3     mod6_YDCcr_Cr_CS_T_F_TF 170.29804 403228.7
+# 4      mod6_YDCcr_C_CS_T_F_TF 166.30172 404261.2
+# 5        mod6_YDC_C_CS_T_F_TF  29.89126 412576.2
+
+#BNT
+#                       rowname        df      AIC
+# 1 mod6_YDCcr_Cr_CSr_Tr_Fr_TFr 102.82522 225950.4
+# 2    mod6_YDCcr_Cr_CSr_T_F_TF  98.12051 225974.0
+# 3     mod6_YDCcr_Cr_CS_T_F_TF  96.51127 225994.7
+# 4      mod6_YDCcr_C_CS_T_F_TF  92.93128 226010.4
+# 5        mod6_YDC_C_CS_T_F_TF  27.97005 231079.5
+
+```
+
+```{r models without T and F}
+# Not sure how to get models with season-specific T and F effects to work within season without messing up predictions. # Leave T and F out and focus on densities
+
+mod6_YDCcr_Cr_CSr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor),  # this is conspecific cS only
+               data = dataIn, 
+               method = "REML")
+
+system.time(
+mod6_YDCcr_Cr_CS3r <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor) +
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDC <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr'),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr_CSBKTr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr_CSBNTr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn, by = riverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+system.time(
+mod6_YDCcr_Cr_CSATSr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
+               s(cohortStrengthScaledATS, bs = 'cr', k = kIn, by = riverFactor),  
+               data = dataIn, 
+               method = "REML")
+)
+
+AIC(mod6_YDCcr_Cr_CSr, mod6_YDCcr_Cr_CS3r, mod6_YDCcr_Cr_CSBKTr, mod6_YDCcr_Cr_CSBNTr, mod6_YDCcr_Cr_CSATSr, mod6_YDCcr_Cr,mod6_YDCcr,mod6_YDC)  %>% rownames_to_column() %>% arrange(AIC)
+
+summary(mod6_YDCcr_Cr_CSr)
+plot(mod6_YDCcr_Cr_CS3r, pages = 7, shade = TRUE, scale = 0)
+
+# bkt
+#                rowname        df      AIC
+# 1    mod6_YDCcr_Cr_CSr 173.85954 413790.2
+# 2   mod6_YDCcr_Cr_CS3r 188.68933 415278.2
+# 3 mod6_YDCcr_Cr_CSBKTr 167.72761 416582.2
+# 4 mod6_YDCcr_Cr_CSBNTr 168.43917 417098.5
+# 5 mod6_YDCcr_Cr_CSATSr 167.84300 417569.6
+# 6        mod6_YDCcr_Cr 160.60559 417764.7
+# 7           mod6_YDCcr 149.80639 420935.0
+# 8             mod6_YDC  14.99337 429769.2
+
+# bnt
+#                rowname       df      AIC
+# 1   mod6_YDCcr_Cr_CS3r 96.91663 231844.1
+# 2    mod6_YDCcr_Cr_CSr 84.99160 231875.4
+# 3 mod6_YDCcr_Cr_CSBKTr 84.39999 232019.6
+# 4 mod6_YDCcr_Cr_CSBNTr 87.25717 232120.2
+# 5 mod6_YDCcr_Cr_CSATSr 85.68893 232203.3
+# 6        mod6_YDCcr_Cr 81.27554 232235.7
+# 7           mod6_YDCcr 75.00492 233885.5
+# 8             mod6_YDC 12.99685 238636.2
+```
+
+```{r predictions}
+
+ydayCumulSeason <- distinct(dataIn, ydayCumul,season,age) %>% arrange(ydayCumul) %>% mutate(age_season = paste(age,season,sep = '_')) %>%
+  group_by(age_season) %>% summarise(age=min(age),season=min(season),minY=min(ydayCumul),maxY=max(ydayCumul))
+#38 obear fish were captured in january, but get season 4 and are age 1, so have very small ydayCumul -need to manually reset 
+#dataIn %>% filter(age==1,season==4,ydayCumul==367) %>% select(tag,cohort,detectionDate,year,ydayCumul,yday,river,age,season)
+ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '1_4',670,ydayCumulSeason$minY)
+ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '2_4',1035,ydayCumulSeason$minY)
+ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '0_3',0,ydayCumulSeason$minY)
+ydayCumulSeason %>% mutate(minY2 = lead(minY))
+
+yDayAll <- data.frame(ydayCumul = 250:1000)
+yDayAll$season <- NA
+
+# fill in season for each day of the year (yDayAll) based on ydayCumulSeason min values
+# need this to get season into all ydayCumul in predGridBKT
+for(i in 1:nrow(yDayAll)) {
+  for(j in 1:(nrow(ydayCumulSeason) - 1)) {
+    if (between(yDayAll$ydayCumul[i], ydayCumulSeason$minY[j], ydayCumulSeason$minY[j + 1])) {
+      yDayAll$season[i] <- ydayCumulSeason$season[j]
+    }
+  }
+}
+
+predGrid <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
+                        cohortFactor = unique(dataIn$cohort),
+                        ydayCumul = seq(250,1000, by = 10),
+                        salmon01 = c(0,1),
+                        cohortStrengthScaled = seq(-2.5,2.5, by = 0.25)) %>%
+                    #left_join(., yDayAll) %>%
+                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
+                           cohort = cohortFactor)
+
+pred <- predict(mod6_YDCcr_Cr_CSr, newdata = predGrid)#, type = 'lpmatrix')
+
+predGrid2 <- bind_cols(predGrid, as.data.frame(pred)) %>% mutate(pred = pred)
+
+ggplot(predGrid2 %>% filter(salmon01 == 0, cohortStrengthScaled == 0),
+
+       aes(ydayCumul, pred, group = cohort, color = cohort)) +
+  #geom_point() + 
+  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
+  geom_line(size = 1.1) +
+  labs(y = 'Predicted body size (mm)',
+       x = 'Days since age-0 Jan 1',
+       color = 'Cohort') +
+  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
+  theme_publication() +
+  theme(legend.position = "right", 
+        legend.direction = "vertical",
+        legend.title = element_text(face = "plain")) +
+  guides(color = guide_colorbar(barheight = 10,
+                                barwidth = 1.5,
+                                ticks = FALSE,
+                                nbin = 100)) +
+#  guide_legend(title = 'Cohort') +
+  facet_wrap(~riverFactor)#, scales = 'free_y')
 

### 2019-07-25 functions/plotGams.R

@@ -1,8 +1,16 @@
 
 plotGamVarBySpecies <- function(var = 's(cohort)', xAxisLab = 'Cohort', bkt = modBKT, bnt = modBNT){
   require(gratia)
+  
   BKT <- evaluate_smooth(modBKT, var) %>% mutate(species = 'bkt')
+  if(var == 's(cohort)'){
+    BKT <- BKT %>% filter(!(cohort < 2002 & riverFactor %in% c('wb jimmy','wb mitchell','wb obear')))
+  } 
   BNT <- evaluate_smooth(modBNT, var) %>% mutate(species = 'bnt')
+  if(var == 's(cohort)'){
+    BNT <- BNT %>% filter(!(cohort < 2002 & riverFactor %in% c('wb jimmy','wb mitchell','wb obear')))
+  }     
+  
   both <- bind_rows(BKT,BNT) %>% mutate(upperCI = est + se * 1.96, lowerCI = est - se * 1.96)
   
   both$speciesGG <- factor(both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
@@ -11,6 +19,9 @@ plotGamVarBySpecies <- function(var = 's(cohort)', xAxisLab = 'Cohort', bkt = mo
                                   labels = c("West Brook","Open Large","Open Small","Isolated Small"))  
   xVar <- substr(var,3,nchar(var)-1)
   byVar <- as.character(unique(both$by_variable))
+  
+  if(var == 's(cohortStrengthScaledBNT2p5)') both <- filter(both, riverFactor %in% c("West Brook","Open Large"))
+  
   gg <- 
     ggplot(both, aes(x = eval(as.name(xVar)), group = speciesGG)) + 
       geom_ribbon(aes( ymin = lowerCI, ymax = upperCI), fill  = 'grey70') +
@@ -25,6 +36,8 @@ plotGamVarBySpecies <- function(var = 's(cohort)', xAxisLab = 'Cohort', bkt = mo
       facet_wrap(~ eval(as.name(byVar)))
   gg
 } 
+
+
   # 
   # # main spline over ydayCumul
   # plotBKT = plot.gam(modBKT, select = 1, scale = 0)

### 2019-07-25 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
added CS fig

@@
 ggsave('./figures/cohortStrengthByMetapop.png', width = 11, height = 8, units='in')
+
+ggplot(cohortStrengthWide %>% filter(cohort >= 2002), aes(cohortStrengthScaledBKT, cohortStrengthScaledBNT, color=factor(cohort))) +
+  geom_point() +
+  geom_smooth(method = 'lm', se = FALSE)
 
@@
   scale_fill_manual(values = c('grey30', 'grey70')) +
-  labs(y = 'Predicted body size (mm)',
+  labs(y = 'Effect on body size (mm)',
        x = 'Days since age-0 Jan 1',

### 2019-07-26 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
cleaned up figs

@@
 ggplot(cohortStrength %>% filter(cohort > 1999), aes(cohort, cohortStrengthScaled, group = speciesGG, shape = speciesGG, color = speciesGG)) +
-  geom_point(size = 2) +
+  geom_point(size = 3) +
   geom_line(size = 1.2) +
@@
        shape = 'Species', color='Species') +
-  theme_publication() +
- # scale_fill_grey() +
-  theme(legend.position = "right", 
-    legend.direction = "vertical",
-    legend.title = element_text(face = "plain")) +
+  theme_publication(base_size = 20) +
+  scale_color_grey() +
+  theme(legend.position = "none",
+        #legend.position=c(0.8, 0.8),
+    #legend.direction = "vertical",
+    #legend.title = element_text(face = "plain"),
+    strip.background = element_blank(),
+    strip.text.y = element_blank()) +
   facet_grid(metaPop01 ~ .)
@@
 
-ggplot(cohortStrengthWide %>% filter(cohort >= 2002), aes(cohortStrengthScaledBKT, cohortStrengthScaledBNT, color=factor(cohort))) +
-  geom_point() +
-  geom_smooth(method = 'lm', se = FALSE)
+ggplot(cohortStrengthWide %>% filter(cohort >= 2002, metaPop01 == 1), aes(cohortStrengthScaledBKT, cohortStrengthScaledBNT)) +
+  geom_point(size = 4) +
+  geom_smooth(method = 'lm', se = FALSE, color = 'grey30') +
+  theme_publication(base_size = 24) +
+  labs(x = 'Brook Trout cohort strength',
+       y = 'Brown Trout cohort strength') +
+  theme(legend.position = "none")
+
+ggsave('./figures/cohortStrengthBKT_BNT.png', width = 11, height = 8, units='in')
+
+CS_bkt_bnt <- lm(cohortStrengthScaledBNT ~ cohortStrengthScaledBKT, data = cohortStrengthWide %>% filter(cohort >= 2002, metaPop01 == 1))
+
+summary(CS_bkt_bnt)
 
@@
   facet_wrap(~ speciesGG) +
-  theme_publication()
+  theme_publication(base_size = 20)
 
@@
   labs(x = 'River', y = 'Number of observations', fill = 'Species') +
-  theme_publication() +
+  theme_publication(base_size = 24) +
   scale_fill_grey() +
-  theme(legend.position = "right", 
+  theme(#legend.position = "right", 
     legend.direction = "vertical",
-    legend.title = element_text(face = "plain"))
+    legend.title = element_text(face = "plain"),
+    legend.position=c(0.8, 0.8))
 
@@
   scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
-  theme_publication() +
+  theme_publication(base_size = 20) +
   theme(legend.position = "right", 
@@
 (plotCohort <- plotGamVarBySpecies(var = 's(cohort)', xAxisLab = 'Cohort'))
-ggsave('./figures/cohortEffect.png', width = 11, height = 8, units='in')
+ggsave('./figures/cohortEffect.png', width = 8, height = 8, units='in')
 
 (plotCohortStrength <- plotGamVarBySpecies(var = 's(cohortStrengthScaledBKT)', xAxisLab = 'Brook Trout cohort Strength'))
-ggsave('./figures/cohortStrengthEffectBKT.png', width = 11, height = 8, units='in')
+ggsave('./figures/cohortStrengthEffectBKT.png', width = 8, height = 8, units='in')
 
 (plotCohortStrengthScaledBNT2p5 <- plotGamVarBySpecies(var = 's(cohortStrengthScaledBNT2p5)', xAxisLab = 'Brown Trout cohort Strength'))
-ggsave('./figures/cohortStrengthEffectBNT.png', width = 11, height = 8, units='in')
+ggsave('./figures/cohortStrengthEffectBNT.png', width = 8, height = 4.75, units='in')
 
@@
 ggplot(predGrid2Both %>% 
-         filter(salmon01 == 0, cohortStrengthScaledBKT == 0, cohort == 2006, cohortStrengthScaledBNT2p5 == 0),
-       aes(ydayCumul, pred, group = speciesGG, color = speciesGG)) +
+         filter(salmon01 == 0, cohortStrengthScaledBKT == 0, cohort == 2006, cohortStrengthScaledBNT2p5 == 0, riverGG %in% c("West Brook", "Open Large")),
+       aes(ydayCumul, pred, group = speciesGG)) +
   #geom_point() + 
   geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
-  geom_line(size = 1.1) +
+  geom_line(aes(linetype = speciesGG), size = 1.1) +
   labs(y = 'Predicted body size (mm)',
@@
 #  scale_color_viridis_c() +
-  theme_publication() +
-  theme(legend.position = "right", 
+  theme_publication(base_size = 20) +
+  #scale_color_grey() +
+  theme(legend.position = "none", 
         legend.direction = "vertical",
@@
 
-ggsave('./figures/ydayCumulAcrossRiverBySpp.png', width = 11, height = 8, units='in')
+ggsave('./figures/ydayCumulAcrossRiverBySpp.png', width = 8, height = 4.75, units='in')
 
@@
   #scale_color_viridis_c() +
-  theme_publication() +
-  theme(legend.position = "right", 
+  theme_publication(base_size = 20) +
+  theme(legend.position = c(0.125, 0.8), 
         legend.direction = "vertical",
         legend.title = element_text(face = "plain")) +
-  facet_grid(rows = vars(speciesGG))
-
-ggsave('./figures/ydayCumulAcrossSppByRiver.png', width = 8, height = 8, units='in')
+  theme(legend.text=element_text(size=12),
+        legend.title=element_text(size=12)) +
+  facet_grid(cols = vars(speciesGG))
+
+ggsave('./figures/ydayCumulAcrossSppByRiver.png', width = 8, height = 4.75, units='in')
 

### 2019-07-26 functions/plotGams.R

@@ -26,11 +26,11 @@ plotGamVarBySpecies <- function(var = 's(cohort)', xAxisLab = 'Cohort', bkt = mo
     ggplot(both, aes(x = eval(as.name(xVar)), group = speciesGG)) + 
       geom_ribbon(aes( ymin = lowerCI, ymax = upperCI), fill  = 'grey70') +
       geom_line(aes(y = est, linetype = speciesGG), size = 1.5) +
-      labs(y = 'Effect size',
+      labs(y = 'Effect size (mm)',
            x = xAxisLab,
            linetype = 'Species') +
-      theme_publication() +
-      theme(legend.position = "right", 
+      theme_publication(base_size = 20) +
+      theme(legend.position = "none", 
             legend.direction = "vertical",
             legend.title = element_text(face="plain")) +
       facet_wrap(~ eval(as.name(byVar)))

### 2019-09-16 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
added model selection models

@@
 vB$seasonGG <- factor(vB$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
+vB$riverGG <- factor(vB$riverFactor,
+                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
+                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))
 
@@
 vB$meanTemperatureFlowScaledBySeason <- vB$meanTemperatureScaledBySeason * vB$meanFlowScaledBySeason
+
 ```
@@
 
-ggplot(cohortStrength %>% filter(cohort > 1999), aes(cohort, cohortStrengthScaled, group = speciesGG, shape = speciesGG, color = speciesGG)) +
+ggplot(cohortStrength %>% filter(cohort >= 2002), aes(cohort, cohortStrengthScaled, group = speciesGG, shape = speciesGG, color = speciesGG)) +
   geom_point(size = 3) +
@@
        y = 'Brown Trout cohort strength') +
+  xlim(c(-1.4,1.9)) +
   theme(legend.position = "none")
@@
 
+# bkt, metapop comparison
+cohortStrengthWideBKT <- cohortStrengthWide %>%
+  select(-cohortStrengthScaledATS,-cohortStrengthScaledBNT) %>%
+  spread( metaPop01, cohortStrengthScaledBKT) %>%
+  rename(metaPop0 = `0`, metaPop1 = `1`)
+
+ggplot(cohortStrengthWideBKT %>% filter(cohort >= 2002), aes(metaPop1, metaPop0)) +
+  geom_point(size = 4) +
+  geom_smooth(method = 'lm', se = FALSE, color = 'grey30') +
+  theme_publication(base_size = 24) +
+  labs(x = 'Cohort strength connected metapopulation',
+       y = 'Cohort strength isolated population') +
+  xlim(c(-1.4,1.9)) +
+  theme(legend.position = "none")
+
+ggsave('./figures/cohortStrengthBKT.png', width = 11, height = 8, units='in')
+
+CS_metaPopBKT <- lm(metaPop0 ~ metaPop1, data = cohortStrengthWideBKT %>% filter(cohort >= 2002))
+
+summary(CS_metaPopBKT)
+
 ```
@@
 ```{r dataIn}
+write.csv(vB, file = './data/out/vB.csv')
+
 dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn) %>%
@@
 ```{r rivers where fish are fish}
-
-vB$riverGG <- factor(vB$riverFactor,
-                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))
 
@@
 
+## Count Summaries
+#number of observations
+vB %>% 
+  filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats'), !is.na(observedLength)) %>%
+  group_by( speciesGG) %>%
+  summarize(n = n())
+
+vB %>% 
+  filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats'), !is.na(observedLength)) %>%
+  group_by(riverGG, speciesGG, year, season) %>%
+  summarize(n = n()) %>%
+  as.data.frame() %>%
+  filter(season == 4)
+
+#number of individuals
+tags <- vB %>% 
+  filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats'), !is.na(observedLength)) %>%
+  group_by(tag,species,metaPop01) %>%
+  summarize(n = n())
+
+tags %>% group_by(species) %>% summarise(n = n())
+tags %>% group_by(species, metaPop01) %>% summarise(n = n())
 ```
@@
 
+# across years
+ggplot(vB %>% filter(cohort %in% 2000:2013,  species == 'bkt'), aes(observedLength)) +
+ # geom_density(aes(color = factor(cohort)), size = 1.35) +
+    geom_freqpoly(aes(color = factor(cohort)), binwidth = 8, size = 1.35) +
+  theme_publication() +
+  theme(strip.text.x = element_text(size = 8), 
+        strip.background = element_blank(),
+        legend.title = element_text("Species"),
+        legend.position = "right",
+        legend.direction = "vertical") + 
+  labs(color = 'Cohort') +
+  scale_x_continuous("Fork Length (mm)") +
+  scale_y_continuous("Count") +
+  #scale_color_manual(values = c("lightblue", "lightgreen")) +
+  #scale_colour_brewer(palette = "Paired")+
+  facet_grid(riverGG ~ ageInSamples, scales = 'free')
+```
+
+```{r length ageInSamples = 1}
+
+ggplot(vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1), aes(yday,observedLength, color = factor(cohort))) + geom_point() + facet_grid(cohort~river)
+
+ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt'), ageInSamples == 1), aes(observedLength)) +
+  geom_freqpoly(aes(group = cohort, color = cohort), binwidth = 7, size = 1.35) + 
+  theme_publication(base_size = 20) +
+  theme(
+        #strip.text.x = element_text(size = 8), 
+        #strip.background = element_blank(),
+        #legend.title = element_text("Species"),
+        legend.position = "right",
+        legend.direction = "vertical"
+        ) + 
+  labs(color = 'Cohort') +
+#  scale_color_grey(start = 0.15, end = 0.75) +
+  scale_colour_gradient(low = "grey30", high = "grey90") +
+  scale_x_continuous("Fork Length (mm)") +
+  scale_y_continuous("Frequency") +
+  facet_grid(speciesGG ~ riverGG)
+
+ggsave('./figures/AIS1_freqpoly.png', width = 11, height = 8, units='in')
+
+# length models for ageInSamples = 1 to look at variation within and among years/species/rivers
+lenModBKT1 <- lm(observedLength ~  factor(cohort) * factor(riverGG), data = vB %>% filter(cohort %in% 2000:2013, species == 'bkt', ageInSamples == 1))
+lenModBKT2 <- lm(observedLength ~  factor(cohort) + riverGG, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bkt', ageInSamples == 1))
+
+AIC(lenModBKT1,lenModBKT2) %>% rownames_to_column() %>% arrange(AIC)
+summary(lenModBKT1)
+
+anovaBKT <- anova(lenModBKT1)
+lenModBKT1ss <- anovaBKT$"Sum Sq"
+print(cbind(anovaBKT,PctExp=lenModBKT1ss/sum(lenModBKT1ss)*100))
+
+###########################
+meanCVLen <- 
+vB %>% filter(cohort %in% 2000:2013, ageInSamples == 1) %>%
+  group_by(species, riverGG, cohort) %>%
+  summarize(meanL = mean(observedLength),
+            sdL = sd(observedLength))
+
+ggplot(meanCVLen, aes(meanL, sdL, color = factor(species))) +
+  geom_point() +
+  facet_wrap(~riverGG)
+
+#no visual diff among rivers in relationship btw mean and cv in len
+ggplot(meanCVLen, aes(meanL, sdL, color = riverGG)) +
+  geom_point() +
+  geom_smooth(se = FALSE, method = "lm")
+
+ggplot(meanCVLen, aes(meanL, sdL, color = factor(cohort))) +
+  geom_point() +
+  facet_wrap(~riverGG)
+
+#BNT
+lenModBNT1 <- lm(observedLength ~  factor(cohort) * factor(riverGG), data = vB %>% filter(cohort %in% 2000:2013, species == 'bnt', ageInSamples == 1))
+lenModBNT2 <- lm(observedLength ~  factor(cohort) + riverGG, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
+# lenModBNT3 <- lm(observedLength ~ 0 + factor(cohort) * riverGG + yday, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
+# lenModBNT4 <- lm(observedLength ~ 0 + factor(cohort) * riverGG + yday * riverGG, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
+# lenModBNT5 <- lm(observedLength ~ 0 + factor(cohort) * riverGG * yday, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
+
+AIC(lenModBNT1,lenModBNT2) %>% rownames_to_column() %>% arrange(AIC)
+summary(lenModBNT1)
+anovaBNT <- anova(lenModBNT1)
+lenModBNT1ss <- anovaBNT$"Sum Sq"
+print(cbind(anovaBNT,PctExp=lenModBNT1ss/sum(lenModBNT1ss)*100))
+
+# both species together - use this for paper
+
+lenMod1 <- lm(observedLength ~  factor(cohort) * factor(riverGG) * speciesGG, data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt'), ageInSamples == 1))
+lenMod2 <- lm(observedLength ~  (factor(cohort) + factor(riverGG) + speciesGG)^2, data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt'), ageInSamples == 1))
+lenMod3 <- lm(observedLength ~  factor(cohort) + factor(riverGG) + speciesGG, data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt'), ageInSamples == 1))
+
+AIC(lenMod1,lenMod2,lenMod3) %>% rownames_to_column() %>% arrange(AIC)
+
+summary(lenMod1)
+anova1 <- anova(lenMod1)
+lenMod1ss <- anova1$"Sum Sq"
+print(cbind(anova1,PctExp=lenMod1ss/sum(lenMod1ss)*100))
+
+calc.relimp(lenMod3)
+#                         lmg
+# factor(cohort)  0.193090771
+# factor(riverGG) 0.045329273
+# speciesGG       0.000531708
+```
+
+```{r length at ageInSamples1 vs cohort strength}
+
+# NO relationship between cohortStrengthScaled and meanLen for ageInSamples = 1
+
+ais1 <- vB %>% filter(cohort %in% 2000:2013, ageInSamples == 1) %>%
+  group_by(speciesGG, metaPop01, cohort) %>%
+  summarise(meanLen = mean(observedLength),
+            meanYday = mean(yday))
+
+cs_ais1 <- left_join(cohortStrength,ais1)
+
+# does mean size increase with day of year? No, but there are some early samples with yday < 255 that have small meanLen
+ggplot(cs_ais1, aes(meanYday,meanLen, color = species)) +
+  geom_point() +
+  geom_smooth(method = 'lm', se = FALSE) +
+  facet_grid(~ metaPop01)
+summary(lm(meanLen ~ 0 + meanYday * species * factor(metaPop01), data = cs_ais1))
+
+ggplot(cs_ais1 %>% filter(meanYday > 0, species %in% c('bkt','bnt')), aes(cohortStrengthScaled, meanLen)) +
+  geom_point(aes(size = meanYday)) +
+  geom_smooth(method = 'lm', se = FALSE, color = 'black') +
+  theme_publication() +
+  theme(strip.text.x = element_text(size = 8), 
+        strip.background = element_blank(),
+        legend.title = element_text("Species"),
+        legend.position = "right",
+        legend.direction = "vertical") + 
+  labs(color = 'Cohort') +
+  scale_y_continuous("Mean Fork Length (mm)") +
+  scale_x_continuous("Cohort Strength") +
+  facet_grid(speciesGG ~ metaPop01)
+
+ggsave('./figures/cohortStrengh_v_meanLenAIS1.png', width = 11, height = 8, units='in')
+
+summary(lm(meanLen ~  cohortStrengthScaled * factor(metaPop01), data = cs_ais1 %>% filter(species == 'bkt')))
+
+summary(lm(meanLen ~  cohortStrengthScaled, data = cs_ais1 %>% filter(species == 'bnt', metaPop01 == 1)))
 ```
@@
 
-mod2 <- bam(observedLength ~ riverFactor * factor(season) + 
+mod2 <- bam(observedLength ~ riverFactor * factor(salmon01) + 
                s(ydayCumul, bs = 'cr'), 
@@
 mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +
-               s(ydayCumul, bs = 'ps'), 
+               s(ydayCumul, bs = 'cr'), 
                data = dataIn, 
@@
 ```{r remove cohort, focus on spp-specific cohort strengths}
-#sppoke with keith. decided focus on cohrot is less interesting than river effects and density effects
+#sppoke with keith. decided focus on cohort is less interesting than river effects and density effects
 
@@
              data = dataIn, method = "REML")
-summary(modCS3)
+#summary(modCS3)
 #plot(modCS3, pages = 8)
@@
 
-AIC(mod1,mod2,mod3,mod4) %>% rownames_to_column() %>% arrange(AIC)
+mod5 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) ,
+               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+         #      s(cohort, bs = 'cr', k = kIn, by = riverFactor),
+              # s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+              # s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
+  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') +
+               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) ,
+          #     s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+         #      s(cohort, bs = 'cr', k = kIn, by = riverFactor),
+              # s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+              # s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
+  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+mod7 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+               s(ydayCumul, bs = 'cr') ,
+            #   s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) ,
+          #     s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+         #      s(cohort, bs = 'cr', k = kIn, by = riverFactor),
+              # s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+              # s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
+  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+mod8 <- bam(observedLength ~ riverFactor + factor(salmon01) ,              
+             #  s(ydayCumul, bs = 'cr') 
+            #   s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) ,
+          #     s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+         #      s(cohort, bs = 'cr', k = kIn, by = riverFactor),
+              # s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
+              # s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
+  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
+             data = dataIn, method = "REML")
+
+AIC(mod1,mod2,mod3,mod4,mod5,mod6,mod7,mod8) %>% rownames_to_column() %>% arrange(AIC)
 ```
@@
              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
+
+# modBKT2 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
+#                s(ydayCumul, bs = 'cr') +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
+#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
+#                s(cohort, bs = 'cr', k = kIn, by = riverFactor) +
+#                s(cohortStrengthScaledBKT,cohortStrengthScaledBNT2p5, k = kIn, by = riverFactor) ,
+#             
+#              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
+# 
+# AIC(modBKT,modBKT2)
 
@@
 
+# across bkt cohort strengths
+ggplot(predGrid2Both %>% 
+         filter(salmon01 == 0, cohort == 2006, cohortStrengthScaledBNT2p5 == 0),
+       aes(ydayCumul, pred, group = cohortStrengthScaledBKT, color = cohortStrengthScaledBKT)) +
+  #geom_point() + 
+  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
+  geom_line(size = 1.1) +
+  labs(y = 'Predicted body size (mm)',
+       x = 'Days since age-0 Jan 1',
+       color = 'Brook Trout cohort strength') +
+  #scale_color_viridis_c() +
+  theme_publication(base_size = 20) +
+  theme(legend.position = c(0.7, 0.37), 
+        legend.direction = "vertical",
+        legend.title = element_text(face = "plain")) +
+  theme(legend.text = element_text(size = 10),
+        legend.title = element_text(size = 16)) +
+  facet_grid(riverGG ~ speciesGG)
+
+ggsave('./figures/ydayCumulAcrossCSBKTByRiver.png', width = 8, height = 4.75, units='in')
+
+# across bnt cohort strengths
+ggplot(predGrid2Both %>% 
+         filter(salmon01 == 0, cohort == 2006, cohortStrengthScaledBKT == 0,
+                riverGG %in% c('West Brook', 'Open Large')),
+       aes(ydayCumul, pred, group = cohortStrengthScaledBNT2p5, color = cohortStrengthScaledBNT2p5)) +
+  #geom_point() + 
+  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
+  geom_line(size = 1.1) +
+  labs(y = 'Predicted body size (mm)',
+       x = 'Days since age-0 Jan 1',
+       color = 'Brown Trout cohort strength') +
+  #scale_color_viridis_c() +
+  theme_publication(base_size = 20) +
+  theme(legend.position = c(0.7, 0.37), 
+        legend.direction = "vertical",
+        legend.title = element_text(face = "plain")) +
+  theme(legend.text = element_text(size = 10),
+        legend.title = element_text(size = 16)) +
+  facet_grid(riverGG ~ speciesGG)
+
+ggsave('./figures/ydayCumulAcrossCSBNTByRiver.png', width = 8, height = 4.75, units='in')
+
 # main spline over ydayCumul

### 2020-09-22 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
first ms draft

@@
 title: "Body size patterns"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
 editor_options:
@@
 library(itsadug)
+library(gratia)
 library(arm)
@@
 library(GGally)
-library(linkedModels)
+# library(linkedModels)
 library(arrayhelpers)
@@
 
+```{r theme_publication}
+
+theme_publication <- function(base_size=14, base_family="") {
+  library(grid)
+  library(ggthemes)
+  (theme_foundation(base_size=base_size, base_family=base_family)
+    + theme(plot.title = element_text(face = "bold",
+                                      size = rel(1.2), hjust = 0.5),
+            text = element_text(),
+            panel.background = element_rect(colour = NA),
+            plot.background = element_rect(colour = NA),
+            axis.title = element_text(face = "bold",size = rel(1)),
+            axis.title.y = element_text(angle=90,vjust =2),
+            axis.title.x = element_text(vjust = -0.2),
+            axis.text = element_text(),
+            axis.line = element_line(colour="black"),
+            axis.ticks = element_line(),
+            panel.grid.major = element_line(colour="#f0f0f0"),
+            panel.grid.minor = element_blank(),
+            legend.key = element_rect(colour = NA),
+            legend.position = "bottom",
+            legend.direction = "horizontal",
+            legend.key.size= unit(0.2, "cm"),
+            #    legend.margin = unit(0, "cm"),
+            legend.title = element_text(face="italic"),
+            plot.margin=unit(c(10,5,5,5),"mm"),
+            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
+            strip.text = element_text(face="bold"),
+            panel.border = element_rect(colour = "black")
+    ))
+
+}
+```
+
 ```{r gams for body size - data prep}
-speciesIn <- 'bnt'
+speciesIn <- 'bkt'
 depVarIn <- 'observedLength'
-kIn <- 4 # # of knots for the speces and river smooths
+kIn <- 4 # # of knots for the species and river smooths
 
@@
 
+########################################
+# tags that may need cohort fixes
+
+#           tag
+# 1  00088cc1c3
+# 2  00088cc5c8
+# 3  00088d0a10
+# 4  00088d1a28
+# 5  00088d20f5
+# 6  00088d215f
+# 7  00088d23b4
+# 8  0009f6ecc2
+# 9  0009f6edf1
+# 10 0009f6ee11
+# 11 0009f6f33c
+# 12 0009f6f404
+# 13 1bf20e52f0
+# 14 1bf20ebfa7
+########################################
 ```
@@
 
+# temporary fix - added 9/18/19, use this just to remake AIC1... fig
+#vB[vB$tag == '0009f6f33c',]$cohort <- 2014  
+#vB[vB$tag == '0009f6f404',]$cohort <- 2014
+
 dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn) %>%
@@
 Run to here to start
+
+```{r raw data tables}
+
+# get proportion of fish tagged for the first time as age0
+dataAllSpp = vB %>% filter(cohort %in% 2000:2013, enc == 1) %>%
+  filter(!(cohort < 2002 & riverFactor %in% c('wb jimmy','wb mitchell','wb obear')))
+
+AIS <- dataAllSpp %>%
+  group_by(tag) %>%
+  summarise(minAIS = min(ageInSamples))
+
+table(AIS$minAIS)/nrow(AIS)
+
+# counts by season, year, spp
+counts <- dataAllSpp %>%
+  group_by(season, year, species) %>%
+  summarize(n = n())
+
+```
 
@@
 
-```{r length ageInSamples = 1}
+```{r length ageInSamples equals one}
 
@@
 
+################################################
+# compare mean lengths between trout
+meansWide <- meanCVLen %>%
+  select(-sdL) %>%
+  pivot_wider(names_from = species, values_from = meanL) %>%
+  filter(riverGG %in% c('West Brook', 'Open Large'))
+
+ggplot(meansWide, aes(bkt, bnt)) +
+  geom_point() +
+  theme_publication() +
+  scale_x_continuous('Brook Trout mean length (mm)') +
+  scale_y_continuous('Brown Trout mean length (mm)', limits = c(60,90)) +
+  geom_smooth(method = 'lm', color = 'black') +
+  facet_wrap(~riverGG)
+
+ggsave('./figures/AIS1_MeanLengthBKT_BNT.png', width = 11, height = 6, units='in')
+
+#summary(lm(bnt ~ bkt, data = filter(meansWide, riverGG == 'West Brook')))
+#summary(lm(bnt ~ bkt, data = filter(meansWide, riverGG == 'Open Large', bnt < 90)))
+corr(meansWide %>% ungroup() %>% filter(riverGG == 'West Brook') %>% select(bkt, bnt) %>% drop_na())
+corr(meansWide %>% ungroup() %>% filter(riverGG == 'Open Large', bnt < 90) %>% select(bkt, bnt) %>% drop_na())
+#########################
 #BNT
@@
 
+```{r corr btw bkt/bnt mean lengths all samples}
+meanCVLenAll <- 
+#vB %>% filter(cohort %in% 2000:2013) %>% # leave out 2013 cohort - some bad cohort assignments I've asked Matt to check 8/4/2020
+  vB %>% filter(cohort %in% 2000:2012) %>%
+  group_by(species, riverGG, cohort, ageInSamples) %>%
+  summarize(meanL = mean(observedLength),
+            sdL = sd(observedLength))
+
+################################################
+# compare mean lengths between trout
+meansWideAll <- meanCVLenAll %>%
+  select(-sdL) %>%
+  pivot_wider(names_from = species, values_from = meanL) %>%
+  filter(riverGG %in% c('West Brook', 'Open Large'))
+
+# get correlations
+corrMod <- function(df) {
+  df <- df %>% select(bkt, bnt) %>% drop_na()
+  corr(df)
+}
+
+byAIS_River <- meansWideAll %>%
+  group_by(riverGG, ageInSamples) %>%
+  nest()
+
+byAIS_River <- byAIS_River %>%
+  mutate(corr = map(data, corrMod)) %>%
+  unnest() %>%
+  distinct(riverGG, ageInSamples, corr)
+
+meansWideAll$AIS_labels <- factor(meansWideAll$ageInSamples, labels = c('Autumn, 0', 'Winter, 0', 'Spring, 1', 'Summer, 1', 'Autumn, 1', 'Winter, 1', 'Spring, 2', 'Summer, 2', 'Autumn, 2'))
+
+byAIS_River$AIS_labels <- factor(byAIS_River$ageInSamples, labels = c('Autumn, 0', 'Winter, 0', 'Spring, 1', 'Summer, 1', 'Autumn, 1', 'Winter, 1', 'Spring, 2', 'Summer, 2', 'Autumn, 2'))
+
+ggplot(meansWideAll, aes(bkt, bnt)) +
+  geom_point() +
+  theme_publication() +
+  scale_x_continuous('Brook Trout mean length (mm)') +
+  scale_y_continuous('Brown Trout mean length (mm)') +
+  geom_smooth(method = 'lm', se = FALSE, color = 'black') +
+  geom_abline(intercept = 0, slope = 1, color = 'grey') +
+  geom_text(aes(label = round(corr, digits = 2), x = 95, y = 235), size = 3.5, data = byAIS_River) +
+  facet_grid(riverGG ~ AIS_labels)
+
+ggsave('./figures/AISALL_MeanLengthBKT_BNT.png', width = 11, height = 6, units='in')
+
+```
+
 ```{r length at ageInSamples1 vs cohort strength}
@@
 
-cs_ais1 <- left_join(cohortStrength,ais1)
+cohortStrength2 <- cohortStrength
+cohortStrength2$speciesGG <- factor(cohortStrength2$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
+
+cs_ais1 <- left_join(cohortStrength2,ais1)
 
@@
 
+cs_ais1$metaPopLabel <- factor(cs_ais1$metaPop01, label = c("Isolated Large", "MetaPopulation"))
+
 ggplot(cs_ais1 %>% filter(meanYday > 0, species %in% c('bkt','bnt')), aes(cohortStrengthScaled, meanLen)) +
@@
   theme(strip.text.x = element_text(size = 8), 
-        strip.background = element_blank(),
+        #strip.background = element_blank(),
         legend.title = element_text("Species"),
@@
   scale_x_continuous("Cohort Strength") +
-  facet_grid(speciesGG ~ metaPop01)
+  facet_grid(speciesGG ~ metaPopLabel)
 
@@
 save(modBKT,modBNT,predBKT,predBNT,predGrid2Both, file = paste0('./data/out/bestModsBKTBNT.RData'))
+
+appraise(modBKT)
+ggsave('./figures/appraiseBKT.png', width = 11, height = 8, units='in')
+
+appraise(modBNT)
+ggsave('./figures/appraiseBNT.png', width = 11, height = 8, units='in')
 
@@
 
-```{r models with season-don't run. gam structure given best model from above}
+```{r models with season-do not run. gam structure given best model from above}
 

### 2020-11-04 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
added flow/temp means

@@
 
+```{r}
+vB %>%
+  group_by(season, riverOrdered) %>%
+  summarize(meanT = mean(meanTemperature, na.rm = TRUE)) %>%
+  ggplot(aes(season, meanT, color = riverOrdered)) +
+  geom_point()
+
+vB %>%
+  group_by(season) %>%
+  summarize(meanT = mean(meanTemperature, na.rm = TRUE))
+
+vB %>%
+  group_by(season, riverOrdered) %>%
+  summarize(meanF = mean(meanFlow, na.rm = TRUE)) %>%
+  ggplot(aes(season, meanF, color = riverOrdered)) +
+  geom_point()
+
+vB %>%
+  group_by(season) %>%
+  summarize(meanF = mean(meanFlow, na.rm = TRUE))
+```
+
 ```{r rivers where fish are fish}

### 2021-11-16 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
updated figures

@@
                          levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))
+                         #labels = c("West Brook","Open Large","Open Small","Isolated Small"))
+                         labels = c("West Brook","Open Large","Open Small","Isolated Large"))
 
@@
                          levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))
+                         #labels = c("West Brook","Open Large","Open Small","Isolated Small"))
+                         labels = c("West Brook","Open Large","Open Small","Isolated Large"))
+
 predGrid2Both$speciesGG <- factor(predGrid2Both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"))
@@
                          levels=c(1,2,3,4),
-                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))
+                         #labels = c("West Brook","Open Large","Open Small","Isolated Small"))
+                         labels = c("West Brook","Open Large","Open Small","Isolated Large"))
 

### 2021-11-16 functions/plotGams.R

@@ -16,7 +16,8 @@ plotGamVarBySpecies <- function(var = 's(cohort)', xAxisLab = 'Cohort', bkt = mo
   both$speciesGG <- factor(both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
   both$riverFactor <- factor(both$riverFactor,
                                   levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
-                                  labels = c("West Brook","Open Large","Open Small","Isolated Small"))  
+                                 # labels = c("West Brook","Open Large","Open Small","Isolated Small")) 
+                                  labels = c("West Brook","Open Large","Open Small","Isolated Large")) 
   xVar <- substr(var,3,nchar(var)-1)
   byVar <- as.character(unique(both$by_variable))
   
